---
title: "Basic area chart with ggplot2"
description: "A few sentences of description to explain what this page is <a href='https://www.google.com'>gonna be about</a> mate."
author: "[Yan Holtz](https://github.com/holtzy)"
date: "`r format(Sys.time(), '%d %B %Y')`"
mail: "yan.holtz.data@gmail.com"
github: "holtzy"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<!-- Table of content?-->
<div>
<nav class="col-sm-3 col-4" id="myScrollspy">
<ul class="nav nav-pills flex-column">
<li class="nav-item">
<a class="nav-link active" href="#intro">Intro</a>
</li>
<li class="nav-item">
<a class="nav-link" href="#ipsum">R ipsum theme</a>
</li>
<li class="nav-item">
<a class="nav-link" href="#publication">Publication</a>
</li>
<li class="nav-item">
<a class="nav-link" href="#leftright">Left - Right</a>
</li>
</ul>
</nav>
</div>
<!-- ========= -->


<div class="container">




# Intro about data input {#intro}
***

Always start by loading *libraries*. The [tidyverse](https://www.tidyverse.org) is probably the package I use the most. Quite difficult to imagine analysing data without it. It is a good practice to load all the packages in the same time. It allows to know exactly what is needed to reproduce your analysis.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with devtools: install_github("holtzy/epuRate", force=TRUE)
```

Then I explain what is my input data and I load it. Here is how it looks like:
```{r}
# Show it:
library(DT)
datatable(mtcars, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T) )
```

This file has `r nrow(mtcars)` lines and `r ncol(mtcars)` columns. It is ready to be analysed.





# Code top - chart bottom
***

An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Library
library(ggplot2)
 
# mtcars data
head(mtcars)
 
# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl))
```

*Note*: You can read more about this project [here](https://www.data-to-viz.com/story/TwoNumOrdered.html).




# Double chart
***

An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE, out.width=c('33%', '33%', '33%'), fig.show='hold'}
# Library
library(ggplot2)
 
# mtcars data
head(mtcars)
 
# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl))

# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl))

ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl))
```

*Note*: You can read more about this project [here](https://www.data-to-viz.com/story/TwoNumOrdered.html).







# Chart left - explanation right {#leftright}
***

<div class = "row">
  
<div class = "col-md-5">
<br><br><br>An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE}
# Library
library(hrbrthemes)
 
# First type of color
p <- ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl)) +
  theme_ipsum()
```

</div>

<div class = "col-md-7">
```{r, warning=FALSE, echo=FALSE}
p
```
</div>
</div>







