---
title: "Basic area chart with ggplot2"
author: "[Yan Holtz](https://github.com/holtzy)"
date: "`r format(Sys.time(), '%d %B %Y')`"
mail: "yan.holtz.data@gmail.com"
github: "holtzy"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---





<br><br>

> Let's start with a short introduction. Explain briefly what this document is going to talk about. May be add a [useful link](https://github.com/holtzy/epuRate) relative to this project.

# Get the data
***
Always start by loading *libraries*. The [tidyverse](https://www.tidyverse.org) is probably the package I use the most. Quite difficult to imagine analysing data without it. It is a good practice to load all the packages in the same time. It allows to know exactly what is needed to reproduce your analysis.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with devtools: install_github("holtzy/epuRate", force=TRUE)
```

Then I explain what is my input data and I load it. Here is how it looks like:
```{r}
# Show it:
mtcars %>% head(5)
```

This file has `r nrow(mtcars)` lines and `r ncol(mtcars)` columns. It is ready to be analysed.





# A first title
***

An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE, message=FALSE, fig.align="center", fig.height=5, fig.width=10}
# Library
library(ggplot2)
 
# mtcars data
#head(mtcars)
 
# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_violin(aes(fill = cyl))
```

*Note*: You can read more about this project [here](https://www.data-to-viz.com/story/TwoNumOrdered.html).



# Comments
***
Any thoughts on this? Found any mistake? Disagree? Please drop me a word on [twitter](https://twitter.com/R_Graph_Gallery).















