---
title: "Basic template for RGG chart"
descriptionMeta: "How to build histograms showing the distribution of several groups with R and ggplot2."
descriptionTop: "A [histogram](https://www.data-to-viz.com/graph/histogram.html) displays the distribution of a numeric variable. A common task is to compare this distribution through several groups. This document explains how to do so using R and [ggplot2](ggplot2.html)."
sectionText: "Histogram Section"
sectionLink: "histogram.html"
DataToVizText: "About histogram"
DataToVizLink: "data-to-viz.com/graph/histogram.html"
url: "histogram_several_group.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---



```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<!--html_preserve-->
<div>
  <nav class="col-sm-3 col-4" id="myScrollspy">
    <ul class="nav nav-pills flex-column">
      <li class="nav-item">
        <a class="nav-link" href="#intro">Intro</a>
      </li>
      <li class="nav-item">
        <a class="nav-link" href="#topbottom">Top bottom</a>
      </li>
      <li class="nav-item">
        <a class="nav-link active" href="#double">Double</a>
      </li>
      <li class="nav-item">
        <a class="nav-link" href="#leftright">Left - Right</a>
      </li>
    </ul>
  </nav>
</div>
<!--/html_preserve-->

<div class="container">




# Intro about data input {#intro}
***

  Always start by loading *libraries*. The [tidyverse](https://www.tidyverse.org) is probably the package I use the most. Quite difficult to imagine analysing data without it. It is a good practice to load all the packages in the same time. It allows to know exactly what is needed to reproduce your analysis.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rmarkdown)    # You need this library to run this template.
library(epuRate)      # Install with devtools: install_github("holtzy/epuRate", force=TRUE)
```

This file has `r nrow(mtcars)` lines and `r ncol(mtcars)` columns. It is ready to be analysed.





# Code top - chart bottom {#topbottom}
***

An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Library
library(ggplot2)

# mtcars data
head(mtcars)

# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_violin(aes(fill = cyl))
```

*Note*: You can read more about this project [here](https://www.data-to-viz.com/story/TwoNumOrdered.html).




# Double chart {#double}
***

An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE, out.width=c('33%', '33%', '33%'), fig.show='hold'}
# Library
library(ggplot2)

# mtcars data
head(mtcars)

# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_violin(aes(fill = cyl))

# First type of color
ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_violin(aes(fill = cyl))

ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_violin(aes(fill = cyl))
```

*Note*: You can read more about this project [here](https://www.data-to-viz.com/story/TwoNumOrdered.html).







# Chart left - explanation right {#leftright}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">
An `area chart` is really similar to a [line chart](https://www.data-to-viz.com/graph/line.html) and represents the evolution of a numeric variable. Basically, the X axis represents time or an ordered variable, and the Y axis gives the value of anothe.

```{r, warning=FALSE}
# Library
library(hrbrthemes)

# First type of color
p <- ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_violin(aes(fill = cyl)) +
  theme_ipsum()
```

</div>

<div class = "col-md-6  col-sm-12">
```{r, warning=FALSE, echo=FALSE, out.width = "100%", fig.width=5, fig.height=5}
p
```
</div>
</div>



# Insert a table
***

It's kind of a mess to insert a table due to the change of the R Markdown template. A workaround is to save a DT table and insert it:
```{r}
# mtcars data
library(DT)
p = datatable(mtcars, options = list(pageLength = 6, scrollX=T, dom = 't') )
saveWidget(p, file=paste0( getwd(), "/HtmlWidget/table95.html"))
htmltools::includeHTML("HtmlWidget/table95.html")
```























<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
