---
title: "Use mtext() to write text in margin"
descriptionMeta: "This document describes how to use the mtext function to add text in the plot margin. Usefull to add title on a multi chart."
descriptionTop: "This document describes how to use the `mtext()` function to add text in the plot margin. Usefull to add title on a multi chart."
sectionText: "Scatter section"
sectionLink: "scatterplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "1-one-title-for-2-graphs"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">


<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `mtext()` function allows to write text in one of the four margins of the current figure region or one of the outer margins of the device region.

Here, the figure is first split thanks to `par(mfrow..)`. Then, only one title is added and centered using `mtext()`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
#Dummy data 
Ixos <- rnorm(4000,100,30)
Primadur <- Ixos+rnorm(4000 , 0 , 30)
 
#Divide the screen in 1 line and 2 columns
par(
  mfrow=c(1,2), 
  oma = c(0, 0, 2, 0)
) 
 
#Make the margin around each graph a bit smaller
par(mar=c(4,2,2,2))
 
# Histogram and Scatterplot
hist(Ixos,  main="" , breaks=30 , col=rgb(0.3,0.5,1,0.4) , xlab="height" , ylab="nbr of plants")
plot(Ixos , Primadur,  main="" , pch=20 , cex=0.4 , col=rgb(0.3,0.5,1,0.4)  , xlab="primadur" , ylab="Ixos" )
 
#And I add only ONE title :
mtext("Primadur : Distribution and correlation with Ixos", outer = TRUE, cex = 1.5, font=4, col=rgb(0.1,0.3,0.5,0.5) )
```
</div>
</div>



```{r thecode, eval=FALSE}
```




# Customizations
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here is a description of the most common customization:  

- `cex`: circle size
- `xlim` and `ylim`: limits of the X and Y axis
- `pch`: shape of markers. See all [here](6-graph-parameters-reminder.html).
- `xlab` and `ylab`: X and Y axis labels
- `col`: marker color
- `main`: chart title


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
data = data.frame(
  x=seq(1:100) + 0.1*seq(1:100)*sample(c(1:10) , 100 , replace=T),
  y=seq(1:100) + 0.2*seq(1:100)*sample(c(1:10) , 100 , replace=T)
)

# Basic scatterplot
plot(data$x, data$y,
     xlim=c(0,250) , ylim=c(0,250), 
     pch=18, 
     cex=2, 
     col="#69b3a2",
     xlab="value of X", ylab="value of Y",
     main="A simple scatterplot"
     )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
