---
title: "Line chart with error envelop: ggplot2 and geom_ribbon()"
descriptionMeta: "This post explains how to add an error envelop around a line chart using ggplot2 and the geom_ribbon() function."
descriptionTop: "This post explains how to add an error envelop around a [line chart](line-plot.html) using `ggplot2` and the `geom_ribbon()` function."
sectionText: "Line chart Section"
sectionLink: "line-plot.html"
DataToVizText: "About line chart"
DataToVizLink: "data-to-viz.com/graph/line.html"
url: "104-plot-lines-with-error-envelopes-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basic line chart with `ggplot2` and `geom_line()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This graph has been made by [Alastair Sanderson](http://www.alastairsanderson.com/). You can have a look to his gallery [here](http://www.sr.bham.ac.uk/~ajrs/R/r-gallery.html).

It shows mean temperature profiles and their error envelopes, using the [ggplot2](ggplot2.html) package and its `geom_ribbon()` function. 

Note that `geom_ribbon` is used since upper and lower values of the envelop are available in the input data. As an alternative, the [geom_smooth](50-51-52-scatter-plot-with-ggplot2.html) function autamatically draw an error envelop using different statistical models.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=6}
library(ggplot2)
 
# Get the data from the web !
CC <- read.table("http://www.sr.bham.ac.uk/~ajrs/papers/sanderson06/mean_Tprofile-CC.txt" ,  header=TRUE)
nCC <- read.table("http://www.sr.bham.ac.uk/~ajrs/papers/sanderson06/mean_Tprofile-nCC.txt" , header=TRUE)
CC$type <- "Cool core"
nCC$type <- "Non-cool core"
A <- rbind(CC, nCC)
 
 
# Make the plot
ggplot(data=A, aes(x=r.r500, y=sckT, ymin=sckT.lo, ymax=sckT.up, fill=type, linetype=type)) + 
 geom_line() + 
 geom_ribbon(alpha=0.5) + 
 scale_x_log10() + 
 scale_y_log10() + 
 xlab(as.expression(expression( paste("Radius (", R[500], ")") ))) + 
 ylab("Scaled Temperature")

```


</div>
</div>



```{r thecode, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
