---
title: "Correlogram with the corrgram library"
descriptionMeta: "This post explains how to build a correlogram with the corrgram R package. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build a [correlogram](correlogram.html) with the `corrgram` R package. It provides several reproducible examples with explanation and `R` code."
sectionText: "Correlogram section"
sectionLink: "correlogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/correlogram.html"
url: "199-correlation-matrix-with-ggally"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Scatterplot matrix with `ggpairs()`
***

```{r thecode, echo=FALSE, out.width = c("33%","33%","33%") , fig.height=7, fig.show="hold"}
# Corrgram library
library(corrgram)

# mtcars dataset is natively available in R
# head(mtcars)

# First
corrgram(mtcars, order=TRUE, lower.panel=panel.shade, upper.panel=panel.pie, text.panel=panel.txt, main="Car Milage Data in PC2/PC1 Order") 

# Second
corrgram(mtcars, order=TRUE, lower.panel=panel.ellipse, upper.panel=panel.pts, text.panel=panel.txt, diag.panel=panel.minmax, main="Car Milage Data in PC2/PC1 Order") 

# Third
corrgram(mtcars, order=NULL, lower.panel=panel.shade, upper.panel=NULL, text.panel=panel.txt, main="Car Milage Data (unsorted)")
```


The `corrgram` package allows to build [correlogram](correlogram.html). The output allows to check the relationship between each pair of a set of numeric variable.

Relationship can be visualized with different methods:

- `panel.ellipse` to display ellipses
- `panel.shade` for coloured squares
- `panel.pie` for pie charts
- `panel.pts` for scatterplots

```{r thecode, eval=FALSE}
```











# Visualize correlation with `ggcorr()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `ggcorr()` function allows to visualize the correlation of each pair of variable as a square. Note that the `method` argument allows to pick the correlation type you desire.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Quick display of two cabapilities of GGally, to assess the distribution and correlation of variables
library(GGally)

# Create data
data <- data.frame( var1 = 1:100 + rnorm(100,sd=20), v2 = 1:100 + rnorm(100,sd=27), v3 = rep(1, 100) + rnorm(100, sd = 1))
data$v4 = data$var1 ** 2
data$v5 = -(data$var1 ** 2)

# Check correlation between variables
#cor(data)

# Nice visualization of correlations
ggcorr(data, method = c("everything", "pearson"))
```
</div>
</div>



```{r thecode2, eval=FALSE}
```








# Split by group {#category}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to use [ggplot2](ggplot2-package.html) aesthetics on the chart, for instance to color each category.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Quick display of two cabapilities of GGally, to assess the distribution and correlation of variables
library(GGally)

# From the help page:
data(flea)
ggpairs(flea, columns = 2:4, ggplot2::aes(colour=species))
```
</div>
</div>



```{r thecode3, eval=FALSE}
```








# Change plot types
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Change the type of plot used on each part of the [correlogram](correlogram.html). This is done with the `upper` and `lower` argument.</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Quick display of two cabapilities of GGally, to assess the distribution and correlation of variables
library(GGally)

# From the help page:
data(tips, package = "reshape")
ggpairs(
  tips[, c(1, 3, 4, 2)],
  upper = list(continuous = "density", combo = "box_no_facet"),
  lower = list(continuous = "points", combo = "dot_no_facet")
)
```
</div>
</div>



```{r thecode4, eval=FALSE}
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
