---
title: "Add a legend to a base R chart"
descriptionMeta: "This post explains how to add a legend to a chart made with base R, using the legend() function. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to add a legend to a chart made with base R, using the `legend()` function. It provides several reproducible examples with explanation and `R` code."
sectionText: "Connected scatter section"
sectionLink: "connected-scatterplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/connectedscatter.html"
url: "119-add-a-legend-to-a-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">
This page aims to explain how to add a legend to a plot made in base R. It is done using the `legend()` function. The main arguments are:
 
 
- `legend` : names to display
- `bty`: type of box around the legend. See [graph #73](73-box-style-with-the-bty-function.html)
- `horiz` : legend in column or in row

- `col` : symbol color
- `pch`: symbol type. See [graph #6](6-graph-parameters-reminder.html)
- `pt.cex` : symbol size

- `cex` : text size
- `text.col`: text color

- `topright` : legend position: `bottomright`, `bottom`, `bottomleft`, `left`, `topleft`, `top`, `topright`, `right`, `center`.
- `inset` : % (from 0 to 1) to draw the legend away from x and y axis
- You can also give the `X` and `Y` coordinate of the legend: `legend(3, 5, ...)`

Note that an equivalent page exist concerning [legends with ggplot2](239-custom-layout-legend-ggplot2.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Create data:
a=c(1:5)
b=c(5,3,4,5,5)
c=c(4,5,4,3,1)
 
# Make a basic graph
plot( b~a , type="b" , bty="l" , xlab="value of a" , ylab="value of b" , col=rgb(0.2,0.4,0.1,0.7) , lwd=3 , pch=17 , ylim=c(1,5) )
lines(c ~a , col=rgb(0.8,0.4,0.1,0.7) , lwd=3 , pch=19 , type="b" )
 
# Add a legend
legend("bottomleft", 
  legend = c("Group 1", "Group 2"), 
  col = c(rgb(0.2,0.4,0.1,0.7), 
  rgb(0.8,0.4,0.1,0.7)), 
  pch = c(17,19), 
  bty = "n", 
  pt.cex = 2, 
  cex = 1.2, 
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
```
</div>
</div>



```{r thecode, eval=FALSE}
```
















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
