---
title: "Donut chart with ggplot2"
descriptionMeta: "The ggplot2 package allows to build donut chart with R. This post describes how, providing explanation and reproducible code."
descriptionTop: "The ggplot2 package allows to build donut chart with R. This post describes how, providing explanation and reproducible code."
sectionText: "Donut section"
sectionLink: "doughnut-plot.html"
DataToVizText: "Warning"
DataToVizLink: "data-to-viz.com/caveat/pie.html"
url: "128-ring-or-donut-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic doughnut chart with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `ggplot2` package allows to build [donut charts](doughnut-plot.html). Note however that this is possible thanks a hack, since no specific function has been created for this kind of chart. (This is voluntary, to avoid donut charts that are dataviz [bad practice](https://www.data-to-viz.com/caveat/pie.html))

Here is the process:
- input data provides a numeric variable for a set of entities
- absolute numeric values must be translated to proportion
- group positions must be stacked: we're gonna display them one after the other
- `geom_rect()` is used to plot each group as a rectangle
- `coord_polar()` is used to switch from stacked rectangles to a ring
- `xlim()` allows to switch from pie to donut: it adds the empty circle in the middle

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# load library
library(ggplot2)
 
# Create test data.
data <- data.frame(
  category=c("A", "B", "C"),
  count=c(10, 60, 30)
)
 
# Compute percentages
data$fraction = data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax = cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin = c(0, head(data$ymax, n=-1))
 
# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
     geom_rect() +
     coord_polar(theta="y") + # Try to remove that to understand how the chart is built initially
     xlim(c(2, 4)) # Try to remove that to see how to make a pie chart
```
</div>
</div>



```{r thecode, eval=FALSE}
```













# Customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here are a couple of things you can do improve your donut chart style:

- use `theme_void()` to get rid of the unnecessary background, axis, labels and so on.
- use a better color palette
- don't use a legend, add labels to groups directly

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# load library
library(ggplot2)

# Create test data.
data <- data.frame(
  category=c("A", "B", "C"),
  count=c(10, 60, 30)
)
 
# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, "\n value: ", data$count)

# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=6) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```










# Donut thickness
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is important to understand that donut chart are just stacked rectangles that are made circular thanks to `coord_polar`.

Thus, the empty circle that makes it a donut chart is just the space between the initial Y axis and the left part of the rectangle.

- If `xlim` left boundary is big, no empty circle. You get a pie chart
- If `xlim` is low, the ring becomes thinner.

If you don't get it, just plot the chart without `coord_polar()`

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# load library
library(ggplot2)

# Create test data.
data <- data.frame(
  category=c("A", "B", "C"),
  count=c(10, 60, 30)
)
 
# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute a good label
data$label <- paste0(data$category, "\n value: ", data$count)

# Make the plot
ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label, color=category), size=6) + # x here controls label position (inner / outer)
  scale_fill_brewer(palette=3) +
  scale_color_brewer(palette=3) +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() +
  theme(legend.position = "none")
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
