---
title: "Pie plot with basic R"
descriptionMeta: "How to build a pie chart with basic R to visualize the proportion of a set of groups. Several examples with reproducible code provided."
descriptionTop: "A [pie chart](pie-plot.html) is a circle divided into sectors that each represent a proportion of the whole. This page explains how to build one with basic `R` commands."
sectionText: "Pie Chart section"
sectionLink: "pie-plot.html"
DataToVizText: "Why you should not do it"
DataToVizLink: "data-to-viz.com/caveat/pie.html"
url: "131-pie-plot-with-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">


# Most basic pie chart with `pie()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`R` natively offers the `pie()` function that builds pie charts. The input is just a numeric variable, each value providing the value of a group of the piechart.

<u>Important note</u>: pie chart are widely known as a bad way to visualize information. Check [this post](https://www.data-to-viz.com/caveat/pie.html) for reasons and alternatives.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Create Data
Prop <- c(3,7,9,1,2)
 
# Make the default Pie Plot
pie(Prop)
```
</div>
</div>



```{r thecode, eval=FALSE}
```










# Change labels with `labels`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Provide a vector of labels to the `labels` argument to add names to piechart groups:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# You can also custom the labels:
pie(Prop , labels = c("Gr-A","Gr-B","Gr-C","Gr-D","Gr-E"))
```

</div>
</div>



```{r thecode2, eval=FALSE}
```













# Non-circular piechart with `edges`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Decrease the value of the `edges` argument to get angles around your piechart.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# If you give a low value to the "edge" argument, you go from something circular to a shape with edges
pie(Prop , labels = c("Gr-A","Gr-B","Gr-C","Gr-D","Gr-E") , edges=10)
```

</div>
</div>



```{r thecode3, eval=FALSE}
```












# Add stripes with `density`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `density` arguments adds stripes.

You can control the angle of those stripes with `angle`.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# The density arguments adds stripes. You can control the angle of this lines with "angle"
pie(Prop , labels = c("Gr-A","Gr-B","Gr-C","Gr-D","Gr-E") , density=10 , angle=c(20,90,30,10,0))

```

</div>
</div>



```{r thecode4, eval=FALSE}
```











# Color with `col` and `border`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Change group color with `col`, and border color with `border`.

Here, the `RcolorBrewer` package is used to build a nice color palette.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 

# You can change the border of each area with the classical parameters:
pie(Prop , labels = c("Gr-A","Gr-B","Gr-C","Gr-D","Gr-E"), border="white", col=myPalette )
```

</div>
</div>



```{r thecode5, eval=FALSE}
```






<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
