---
title: "Basic Stacked area chart with R"
descriptionMeta: "This post provides the basics concerning stacked area chart with R and ggplot2. It takes into account several input format types and show how to customize the output."
descriptionTop: "This post provides the basics concerning [stacked area chart](stacked-area-graph.html) with R and `ggplot2`. It takes into account several input format types and show how to customize the output."
sectionText: "Stacked area section"
sectionLink: "stacked-area-graph.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/stackedarea.html"
url: "136-stacked-area-chart"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic stacked area with ggplot2
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The data frame used as input to build a stacked area chart requires 3 columns:

- `x`: numeric variable used for the X axis, often it is a time.
- `y`: numeric variable used for the Y axis. What are we looking at?
- `group`: one shape will be done per group.

The chart is built using the `geom_area()` function. 

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=6}
# Packages
library(ggplot2)
library(dplyr)
 
# create data
time <- as.numeric(rep(seq(1,7),each=7))  # x Axis
value <- runif(49, 10, 100)               # y Axis
group <- rep(LETTERS[1:7],times=7)        # group, one shape per group
data <- data.frame(time, value, group)

# stacked area chart
ggplot(data, aes(x=time, y=value, fill=group)) + 
    geom_area()
```
</div>
</div>



```{r thecode, eval=FALSE}
```












# Control stacking order with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The gallery offers a post dedicated to [reordering with ggplot2](267-reorder-a-variable-in-ggplot2.html). This step can be tricky but the code below shows how to:

- give a specific order with the `factor()` function.
- order alphabetically using `sort()`
- order following values at a specific data

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=6}
# Give a specific order:
data$group <- factor(data$group , levels=c("B", "A", "D", "E", "G", "F", "C") )

# Plot again
ggplot(data, aes(x=time, y=value, fill=group)) + 
    geom_area()

# Note: you can also sort levels alphabetically:
myLevels <- levels(data$group)
data$group <- factor(data$group , levels=sort(myLevels) )

# Note: sort followinig values at time = 5
myLevels <- data %>%
  filter(time==6) %>%
  arrange(value)
data$group <- factor(data$group , levels=myLevels$group )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```











# Proportional stacked area chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

In a proportional stacked area graph, the sum of each year is always equal to hundred and value of each group is represented through percentages. 

To make it, you have to calculate these percentages first. This can be done using `dplyr` of with `base R`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=6}
# Compute percentages with dplyr
library(dplyr)
data <- data  %>%
  group_by(time, group) %>%
  summarise(n = sum(value)) %>%
  mutate(percentage = n / sum(n))

# Plot
ggplot(data, aes(x=time, y=percentage, fill=group)) + 
    geom_area(alpha=0.6 , size=1, colour="black")

# Note: compute percentages without dplyr:
my_fun <- function(vec){ 
  as.numeric(vec[2]) / sum(data$value[data$time==vec[1]]) *100 
}
data$percentage <- apply(data , 1 , my_fun)

```
</div>
</div>



```{r thecode3, eval=FALSE}
```












# Color & Style
***

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

Let's improve the chart general appearance:

- usage of the `viridis` color scale
- `theme_ipsum` of the `hrbrthemes` package
- add title with `ggtitle` 


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=5}
# Library
library(viridis)
library(hrbrthemes)

# Plot
ggplot(data, aes(x=time, y=value, fill=group)) + 
    geom_area(alpha=0.6 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) +
    theme_ipsum() + 
    ggtitle("The race between ...")


```
</div>
</div>



```{r thecode4, eval=FALSE}
```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
