---
title: "Basic radar chart"
descriptionMeta: "How to build the most basic radar chart with R and the fmsb library: check several reproducible examples with explanation and R code."
descriptionTop: "How to build the most basic [radar chart](spider-or-radar-chart.html) with `R` and the `fmsb` library: check several reproducible examples with explanation and R code."
sectionText: "Radar section"
sectionLink: "spider-or-radar-chart.html"
DataToVizText: "Warning"
DataToVizLink: "data-to-viz.com/caveat/spider.html"
url: "142-basic-radar-chart"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic radar chart with the `fmsb` package
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Radar charts](spider-or-radar-chart.html) are also called Spider or Web or Polar charts. They are drawn in `R` using the [fmsb library](https://cran.r-project.org/web/packages/fmsb/fmsb.pdf). 

Input data format is very specific. Each row must be an entity. Each column is a quantitative variable. First 2 rows provide the min and the max that will be used for each variable.

Once you have this format, the `radarchart()` function makes all the job for you.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(fmsb)
 
# Create data: note in High school for Jonathan:
data <- as.data.frame(matrix( sample( 2:20 , 10 , replace=T) , ncol=10))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding", "data-viz" , "french" , "physic", "statistic", "sport" )
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each topic to show on the plot!
data <- rbind(rep(20,10) , rep(0,10) , data)
 
# Check your data, it has to look like this!
# head(data)

# The default radar chart 
radarchart(data)
```
</div>
</div>



```{r thecode, eval=FALSE}
```











# Customization
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

The `radarchart()` function offers several options to customize the chart:

- Polygon features:
    - `pcol` &rarr; line color
    - `pfcol` &rarr; fill color
    - `plwd` &rarr; line width 
<br>  
- Grid features: 
    - `cglcol` &rarr; color of the net
    - `cglty` &rarr; net line type (see [possibilities](6-graph-parameters-reminder.html))
    - `axislabcol` &rarr; color of axis labels
    - `caxislabels` &rarr; vector of axis labels to display
    - `cglwd` &rarr; net width
<br>  
- Labels: 
    - `vlcex` &rarr; group labels size
    
    
</div>


<div class = "col-md-7  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(fmsb)
 
# Create data: note in High school for Jonathan:
data <- as.data.frame(matrix( sample( 2:20 , 10 , replace=T) , ncol=10))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding", "data-viz" , "french" , "physic", "statistic", "sport" )
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each topic to show on the plot!
data <- rbind(rep(20,10) , rep(0,10) , data)
 
# Check your data, it has to look like this!
# head(data)

# Custom the radarChart !
radarchart( data  , axistype=1 , 
 
    #custom polygon
    pcol=rgb(0.2,0.5,0.5,0.9) , pfcol=rgb(0.2,0.5,0.5,0.5) , plwd=4 , 
 
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,20,5), cglwd=0.8,
 
    #custom labels
    vlcex=0.8 
    )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```







# Next step
***

This page showed the basics of [polar chart](spider-or-radar-chart.html) with `R` and the `fmsb` package. [Next example](143-spider-chart-with-saveral-individuals.html) describes a few more customization and show how to proceed to visualize several groups on the same chart.

<a class="btn btn-secondary" href="143-spider-chart-with-saveral-individuals.html">Chart #143</a>
<a class="btn btn-secondary" href="spider-or-radar-chart.html">Radar chart section</a>



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
