---
title: "Radar chart with several individuals"
descriptionMeta: "It is possible to display several groups on the same radar chart. This post describes how to draw it with R and the fmsb library. Explanation and R code provided."
descriptionTop: "It is possible to display several groups on the same [radar chart](spider-or-radar-chart.html). This post describes how to draw it with `R` and the `fmsb` library."
sectionText: "Radar section"
sectionLink: "spider-or-radar-chart.html"
DataToVizText: "Warning"
DataToVizLink: "data-to-viz.com/caveat/spider.html"
url: "143-spider-chart-with-saveral-individuals"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Multi-group spider chart, no options
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If you're using the `fmsb` package for the first time, have a look to the [most basic radar chart](142-basic-radar-chart.html) first, it explains how the data must be formatted for the `radarchart()` function.

If you have several individuals, the code looks pretty much the same as shown below.

<u>Note:</u> don't show more that 2 or 3 groups on the same web chart, it would make it unreadable. Read more about [radar chart caveats](https://www.data-to-viz.com/caveat/spider.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(fmsb)
 
# Create data: note in High school for several students
set.seed(99)
data <- as.data.frame(matrix( sample( 0:20 , 15 , replace=F) , ncol=5))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding" )
rownames(data) <- paste("mister" , letters[1:3] , sep="-")
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
data <- rbind(rep(20,5) , rep(0,5) , data)
 
# plot with default options:
radarchart(data)
```
</div>
</div>



```{r thecode, eval=FALSE}
```











# Customization
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

The `radarchart()` function offers several options to customize the chart:

- Polygon features:
    - `pcol` &rarr; line color
    - `pfcol` &rarr; fill color
    - `plwd` &rarr; line width
<br>  
- Grid features:
    - `cglcol` &rarr; color of the net
    - `cglty` &rarr; net line type (see [possibilities](6-graph-parameters-reminder.html))
    - `axislabcol` &rarr; color of axis labels
    - `caxislabels` &rarr; vector of axis labels to display
    - `cglwd` &rarr; net width
<br>  
- Labels:
    - `vlcex` &rarr; group labels size


</div>


<div class = "col-md-7  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(fmsb)
 
# Create data: note in High school for several students
set.seed(99)
data <- as.data.frame(matrix( sample( 0:20 , 15 , replace=F) , ncol=5))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding" )
rownames(data) <- paste("mister" , letters[1:3] , sep="-")
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
data <- rbind(rep(20,5) , rep(0,5) , data)

# Color vector
colors_border=c( rgb(0.2,0.5,0.5,0.9), rgb(0.8,0.2,0.5,0.9) , rgb(0.7,0.5,0.1,0.9) )
colors_in=c( rgb(0.2,0.5,0.5,0.4), rgb(0.8,0.2,0.5,0.4) , rgb(0.7,0.5,0.1,0.4) )

# plot with default options:
radarchart( data  , axistype=1 , 
    #custom polygon
    pcol=colors_border , pfcol=colors_in , plwd=4 , plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,20,5), cglwd=0.8,
    #custom labels
    vlcex=0.8 
    )

# Add a legend
legend(x=0.7, y=1, legend = rownames(data[-c(1,2),]), bty = "n", pch=20 , col=colors_in , text.col = "grey", cex=1.2, pt.cex=3)
```

</div>
</div>



```{r thecode2, eval=FALSE}
```









# About axis limits
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

In the previous examples, axis limits were set in the 2 first rows of the input dataset. 

If you do not specify these values, the axis limits will be computed automatically, as shown below.
</div>


<div class = "col-md-7  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(fmsb)
 
# Create data: note in High school for several students
set.seed(99)
data <- as.data.frame(matrix( sample( 0:20 , 15 , replace=F) , ncol=5))
colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding" )
rownames(data) <- paste("mister" , letters[1:3] , sep="-")
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
data <- rbind(rep(20,5) , rep(0,5) , data)
 
# Set graphic colors
library(RColorBrewer)
coul <- brewer.pal(3, "BuPu")
colors_border <- coul
library(scales)
colors_in <- alpha(coul,0.3)

# If you remove the 2 first lines, the function compute the max and min of each variable with the available data:
radarchart( data[-c(1,2),]  , axistype=0 , maxmin=F,
    #custom polygon
    pcol=colors_border , pfcol=colors_in , plwd=4 , plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="black", cglwd=0.8, 
    #custom labels
    vlcex=0.8 
    )

# Add a legend
legend(x=0.7, y=1, legend = rownames(data[-c(1,2),]), bty = "n", pch=20 , col=colors_in , text.col = "grey", cex=1.2, pt.cex=3)
```

</div>
</div>



```{r thecode3, eval=FALSE}
```









# Warning
***

There is a lot of criticism going around spider chart. Before using it in a project, you probably want to [learn more about it](https://www.data-to-viz.com/caveat/spider.html).


<a class="btn btn-secondary" href="https://www.data-to-viz.com/caveat/spider.html">Why it's dangerous</a>
<a class="btn btn-secondary" href="spider-or-radar-chart.html">Radar chart section</a>



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
