---
title: "Interactive area chart with R and plotly"
descriptionMeta: "The plotly package allows to build interactive charts directly from R. Here is a application to area chart, with explanation and reproducible code."
descriptionTop: "The `plotly` package allows to build interactive charts directly from `R`. Here is a application to [area chart](area-chart.html), using both the `plot_ly()` and `ggplotly()` functions."
sectionText: "Area chart Section"
sectionLink: "area-chart.html"
DataToVizText: "About area chart"
DataToVizLink: "data-to-viz.com/graph/area.html"
url: "163-interactive-area-chart-plotly"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Using the `ggplotly()` function
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

This post follows the previous [basic area chart](164-area-chart-ggplot2.html) built with [ggplot2](ggplot2-package.html).

The idea is to turn the chart interactive:

- You can zoom by selecting an area of interest
- Hover the line to get exact time and value
- Export to png
- Slide axis
- Double click to re-initialize.

This is done thanks to the `ggplotly()` function of the [plotly](plotly.html) package that turn any `ggplot2` chart object interactive. 

</div>


<div class = "col-md-7  col-sm-12">

<center><iframe src="HtmlWidget/plotly-line-chart.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>




```{r}
# Libraries
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# Usual area chart
p <- data %>%
  ggplot( aes(x=date, y=value)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("bitcoin price ($)") +
    theme_ipsum()

# Turn it interactive with ggplotly
p <- ggplotly(p)
p

# save the widget
#library(htmlwidgets)
#saveWidget(p, file="HtmlWidget/plotly-line-chart.html")
```















# Using `plot_ly()`
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

The `ggplotly()` function above takes as input a ggplot2 chart and turn it interactive.

But the `plotly` package also allows to build interactive charts using its own function: `plot_ly()`.

<u>Pro</u>: it will allows a greater control of chart style. See the whole [API](https://plot.ly/r/reference/) for customization.

<u>Con</u>: you have to learn a new API, although you probably already know `ggplot2`
</div>


<div class = "col-md-7  col-sm-12">

<center><iframe src="HtmlWidget/plotly-area-chart.html" height="600" width="700" style="border:none;"></iframe></center>

</div>
</div>




```{r}
# library
library(plotly)
 
# Create data
var1 <- seq(1,8)
var2 <- c(0,1,4,1,8,7,5,4)
var3 <- c(7,8,4,2,1,2,0,1)

# Area chart with 2 groups
p <- plot_ly(x = var1, y = var2, type="scatter", mode="markers", fill = "tozeroy")
p <- add_trace(p, x = var1, y = var3, type="scatter", mode="markers", fill = "tonexty")
p

# save the widget
# library(htmlwidgets)
#saveWidget(p, file="HtmlWidget/plotly-area-chart.html")
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
