---
title: "Open and Plot Shapefiles in R"
descriptionMeta: "Shapefiles, a popular format for storing geospatial information, are easily handled in R using the sf package. This guide will demonstrate how to read shapefiles into R and visualize them using either base R or the ggplot2 package."
descriptionTop: "**Shapefiles**, a popular format for storing **geospatial information**, are easily handled in **R** using the `sf` package. This guide will demonstrate **how to read shapefiles** into R and visualize them using either **base R** or the [ggplot2](ggplot2-package.html) package."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "168-load-a-shape-file-into-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container" style="padding-top: 80px">


# About shapefiles
***

If you did not find the geospatial data you need in existing R packages (see the [map section](map.html)), you need to find this information elsewhere on the web. 

Usually, you will find it as a *shape file* format. This format is composed by **several files** that you need to **keep together** in the same folder.


# Find and download a shapefile.
***

You need to dig the internet to find the shape file you are interested in. For instance, [this URL](http://thematicmapping.org/downloads/world_borders.php) will redirect you to a zipped shape file containing the world boundaries.

You can download it and unzip it with R:

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip", destfile = "DATA/world_shape_file.zip")
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
unzip("DATA/world_shape_file.zip", junkpaths = FALSE)
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
```










# Read it with `sf`
***

The `sf` package offers the `st_read()` and the `read_sf()` functions that allows to **read shapefile** using the following syntax. 

As a result you get a geospatial object (`my_sf` here) that contains all the information we need for further mapping. Please try the following command to understand how this object works:

- `print(my_sf)`: tells you the max and min coordinates, the kind of projection in use and the the **first few rows** of the data associated with the regions.
- `nrow(my_sf)`: how many regions you have

```{r, warning=FALSE, message=FALSE}
# Read this shape file with the sf library.
library(sf)
my_sf <- read_sf("TM_WORLD_BORDERS_SIMPL-0.3.shp")

# -- > Now you have a sf object (simple feature data frame). You can start doing maps!
```



# Plot it with `base R`
***

The basic `plot()` function knows how to plot a geospatial object. Thus you just need to pass it `my_sf` and add a couple of **options to customize** the output.


```{r thecode4, out.width = "100%", fig.height=4}
# Basic plot of this shape file:
par(mar = c(0, 0, 0, 0))
plot(st_geometry(my_sf), col = "#f2f2f2", bg = "skyblue", lwd = 0.25, border = 0)
```















# Plot it with `ggplot2`
***


It is totally possible (and advised) to build the map with [ggplot2](ggplot2-package.html), using the `geom_sf()` function as described below.


```{r thecode5, out.width = "100%", fig.height=4}
# Plot it
library(ggplot2)
ggplot(my_sf) +
  geom_sf(fill = "#69b3a2", color = "white") +
  theme_void()
```




# Going further
***

This post explains how to read and plot a shapefile in R.

You might be interested in how to [customize the map](169-170-basic-manipulation-of-shapefiles.html) (select regions, change colors, etc.) and more generally in the [map section](map.html).










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
