---
title: "Open and Plot Shapefiles in R"
descriptionMeta: "Shapefiles are a common way to store geospatial data. This post explains how to read it with R and the rgdal package, and how to plot it in base R or with ggplot2."
descriptionTop: "Shapefiles are a common way to store geospatial data. This post explains how to read it with `R` and the `rgdal` package, and how to plot it in base `R` or with `ggplot2`."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "168-load-a-shape-file-into-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">




If you did not find the geospatial data you need in existing R packages (see the [map section](maps.html)), you need to find this information elsewhere on the web. 

Usually, you will find it as a *shape file* format. This format is composed by several files that you need to keep together in the same folder.

<u>Note</u>: if you found a `.geoJSON` file, read [this post](325-background-map-from-geojson-format-in-r.html) instead.


# Find and download a shapefile.
***

You need to dig the internet to find the shape file you are interested in. For instance, [this URL](http://thematicmapping.org/downloads/world_borders.php) will redirect you to a zipped shape file containing the worl boundaries.

You can download it and unzip it with R:

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="DATA/world_shape_file.zip")
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
system("unzip DATA/world_shape_file.zip")
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
 
```










# Read it with `rgdal`
***

The `rgdal` package offers the `readOGR()` function that allows to read shapefile using the following syntax. 

As a result you get a geospatial object (`my_spdf` here) that contains all the information we need for further mapping. Please try th following command to understand how this object works:

- `summary(my_spdf)`: tells you the max and min coordinates, the kind of projection in use
- `length(my_spdf)`: how many regions you have
- `head(my_spdf@data)`: the firs few rows of the data slot associated with the regions

```{r, warning=FALSE, message=FALSE}
# Read this shape file with the rgdal library. 
library(rgdal)
my_spdf <- readOGR( 
  dsn= paste0(getwd(),"/DATA/world_shape_file/") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# -- > Now you have a Spdf object (spatial polygon data frame). You can start doing maps!
```

# Plot it with `base R`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The basic `plot()` function knows how to plot a geospatial object. Thus you just need to pass it `my_spdf` and add a couple of options to customize the output.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Basic plot of this shape file:
par(mar=c(0,0,0,0))
plot(my_spdf, col="#f2f2f2", bg="skyblue", lwd=0.25, border=0 )
```
</div>
</div>

```{r thecode4, eval=FALSE}
```













# Plot it with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally possible (and advised imo) to build the map with [ggplot2](ggplot2-package.html). However, ggplot2 takes as input data frames, not geospatial data.

`my_spdf` thus needs to be transformed using the `tidy()` function of the `broom` package. The `region` argument of this function expect one of the column name if the @data slot. It will be the region name in the new dataframe.

Once the data frame is created, it is plotted using the `geom_polygon()` function as described below.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=4}
# 'fortify' the data to get a dataframe format required by ggplot2
library(broom)
spdf_fortified <- tidy(my_spdf, region = "NAME")

# Plot it
library(ggplot2)
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="#69b3a2", color="white") +
  theme_void() 
```
</div>
</div>

```{r thecode5, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
