---
title: "Geospatial data manipulation in R"
descriptionMeta: "Map data in R are stored in a specialized geospatial format. This post explores key manipulations you might need, including selecting zones, simplifying borders, and more."
descriptionTop: "[**Map data**](map.html) in `R` is stored in a specialized **geospatial format**: shapefiles. This post explores key manipulations you might need, including **selecting zones**, **simplifying borders**, and more.<br>For an introduction to shapefiles, check [this post](168-load-a-shape-file-into-r.html)."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "169-170-basic-manipulation-of-shapefiles"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container">



# Get a geospatial object
***

The region boundaries required to make maps are usually stored in **geospatial objects**. Those objects can come from [shapefiles](168-load-a-shape-file-into-r.html), [geojson files](325-background-map-from-geojson-format-in-r.html) or provided in a [R package](278-the-maps-library.html). See the [map](map.html) section for possibilities.

Let's get a geospatial object from a shape file available [here](http://thematicmapping.org/downloads/world_borders.php). This step is extensively described in [this post](168-load-a-shape-file-into-r.html) in case you're not familiar with it.

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip",
  destfile = "DATA/world_shape_file.zip"
)
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
system("unzip DATA/world_shape_file.zip")
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
```

<br>

And let's load it in R

```{r}
# Read this shape file with the sf library.
library(sf)
my_sf <- read_sf(
  file.path(getwd(), "/DATA/world_shape_file/TM_WORLD_BORDERS_SIMPL-0.3.shp")
)

# -- > Now you have a sf object (simple feature data frame). You can start doing maps!
```











# Select a region {#select}
***


You can filter the geospatial object to plot only a subset of the regions. The following code keeps only Africa and plot it.

```{r thecode4, out.width = "100%", fig.height=4}
# Keep only data concerning Africa
africa <- my_sf[my_sf$REGION == 2, ]

# Plot africa
par(mar = c(0, 0, 0, 0))
plot(st_geometry(africa),
  xlim = c(-20, 60), ylim = c(-40, 35),
  col = "steelblue", lwd = 0.5
)
```









# Simplify the geospatial object {#simplify}
***


It's a common task to *simplify* the geospatial object. Basically, it decreases the border precision which results in a lighter object that will be plotted faster.

The `rmapshaper` package offers the `ms_simplify()` function to makes the simplification. Play with the `keep` argument to control simplification rate.



```{r thecode6, out.width = "100%", fig.height=4}
# Simplification with rmapshaper
library("rmapshaper")
africaSimple <- ms_simplify(africa, keep = 0.01, keep_shapes = TRUE)

# Plot it
par(mar = c(0, 0, 0, 0))
plot(st_geometry(africaSimple),
  xlim = c(-20, 60), ylim = c(-40, 35),
  col = "#59b2a3", lwd = 0.5
)
```















# Compute region centroid  {#centroid}
***

Another common task is to compute the centroid of each region to add labels. This is doable using the `st_centroid()` function of the `sf` package.


```{r thecode7, out.width = "100%", fig.height=4}
# The st_centroid function computes the centroid of each region:
# st_centroid(africa, of_largest_polygon = TRUE)

# select big countries only
africaBig <- africa[which(africa$AREA > 75000), ]

centroids <- st_centroid(africaBig, of_largest_polygon = TRUE)
# Small manipulation to add coordinates as columns
centers <- cbind(centroids, st_coordinates(centroids))

# Show it on the map
par(mar = c(0, 0, 0, 0))
plot(st_geometry(africa), xlim = c(-20, 60), ylim = c(-40, 35), lwd = 0.5)
text(centers$X, centers$Y, centers$FIPS, cex = .9, col = "#69b3a2")
```



# Going further
***

This post explains how to manipulate geospatial objects in R.

You might be interested in creating a [choropleth map](choropleth-map.html) or a [bubble map](bubble-map.html) with this object.










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
