---
title: "Geospatial data manipulation in R"
descriptionMeta: "Map data are stored in a very specific geospatial format in R. This post describes the most common manipulations you may have to apply: selecting zone, simplifying the borders, and more."
descriptionTop: "[Map data](maps.html) are stored in a very specific geospatial format in `R`. This post describes the most common manipulations you may have to apply: selecting zone, simplifying the borders, and more."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "169-170-basic-manipulation-of-shapefiles"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Get a geospatial object
***

The region boundaries required to make maps are usually stored in geospatial objects. Those objects can come from [shapefiles](168-load-a-shape-file-into-r.html), [geojson files](325-background-map-from-geojson-format-in-r.html) or provided in a [R package](278-the-maps-library.html). See the [map](maps.html) section for possibilities.

Let's get a geospatial object from a shape file available [here](http://thematicmapping.org/downloads/world_borders.php). This step is extensively described in [this post](168-load-a-shape-file-into-r.html) in case you're not familiar with it.

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="DATA/world_shape_file.zip")
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
system("unzip DATA/world_shape_file.zip")
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
```

<br>

And let's load it in R

```{r}
# Read this shape file with the rgdal library. 
library(rgdal)
my_spdf <- readOGR( 
  dsn= paste0(getwd(),"/DATA/world_shape_file/") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# -- > Now you have a Spdf object (spatial polygon data frame). You can start doing maps!
```











# Select a region {#select}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

You can filter the geospatial object to plot only a subset of the regions. The following code keeps only Africa and plot it.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Keep only data concerning Africa
africa <- my_spdf[my_spdf@data$REGION==2 , ]

# Plot africa
par(mar=c(0,0,0,0))
plot(africa , xlim=c(-20,60) , ylim=c(-40,35), col="steelblue", lwd=0.5 )
```

</div>
</div>

```{r thecode4, eval=FALSE}
```









# Simplify the geospatial object {#simplify}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It's a common task to *simplify* the geospatial object. Basically, it decreases the border precision which results in a lighter object that will be plotted faster.

The `rgeos` package offers the `gSimplify()` function to makes the simplification. Play with the `tol` argument to control simplification rate.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode6, echo=FALSE, out.width = "100%", fig.height=4}
# Simplification with rgeos
library("rgeos")
africaSimple <- gSimplify(africa, tol = 4, topologyPreserve = TRUE)

# Plot it
par(mar=c(0,0,0,0))
plot(africaSimple , xlim=c(-20,60) , ylim=c(-40,35), col="#59b2a3", lwd=0.5 )
```

</div>
</div>

```{r thecode6, eval=FALSE}
```
















# Compute region centroid  {#centroid}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Another common task is to compute the centroid of each region to add labels. This is doable using the `gCentroid()` function of the `rgeos` package.

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode7, echo=FALSE, out.width = "100%", fig.height=4}
# Load the rgeos library
library(rgeos)

# The gCentroid function computes the centroid of each region:
# gCentroid(africa, byid=TRUE)

# select big countries only
africaBig <- africa[which(africa@data$AREA>75000), ]

# Small manipulation to put it in a dataframe:
centers <- cbind.data.frame(data.frame(gCentroid(africaBig, byid=TRUE), id=africaBig@data$FIPS))

# Show it on the map?
par(mar=c(0,0,0,0))
plot(africa , xlim=c(-20,60) , ylim=c(-40,35), lwd=0.5 )
text(centers$x, centers$y, centers$id, cex=.9, col="#69b3a2")
```

</div>
</div>

```{r thecode7, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
