---
title: "Basic choropleth map with the Cartography package"
descriptionMeta: "The cartography package is a great alternative to make maps with R. This post explains how to build a basic choropleth map using it, with explanation and reproducible code."
descriptionTop: "The `cartography` package is a great alternative to make maps with R. This post explains how to build a basic [choropleth](choropleth-map.html) map using it, with explanation and reproducible code."
sectionText: "Choropleth section"
sectionLink: "choropleth-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/choropleth.html"
url: "175-choropleth-map-cartography-pkg"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container" style="padding-top: 80px">



# Data provided in the library
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `cartography` package is a very good alternative to the [other tools](bubble-map.html) presented in the gallery for building maps.

It notably allows to:

- display a legend with embedded circles
- add data sources on the bottom of the chart
- show north / south


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, fig.show="hold", fig.height=6, out.width="100%"}
# Use the cartography library to do the choropleth map
library(cartography)

# Load data
data(nuts2006)

# Build a choropleth
choroLayer(spdf = nuts2.spdf, df = nuts2.df, var = "pop2008", legend.pos = "right")
title("Population in 2008")
```
</div>
</div>



```{r thecode, eval=FALSE}
```













# Data provided in a shape file
***

<!-- warning message about the deprecated rgdal package -->

<div class = "alert alert-warning">
  <strong>Warning!</strong> The <a href='package/rgdal.html'>rgdal</a> package is deprecated. You should use the <a href='package/sf.html'>sf</a> package instead.
</div>

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `cartography` package is a very good alternative to the [other tools](bubble-map.html) presented in the gallery for building maps.

It notably allows to:

- display a legend with embedded circles
- add data sources on the bottom of the chart
- show north / south


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = c("100%"), fig.height=6}
# Download the shape file from the web and unzip it:
# download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="world_shape_file.zip")
# system("unzip world_shape_file.zip")

# Load it as a geospatial object in R
library(rgdal)
my_spdf <- readOGR(dsn = "/Users/yan.holtz/Desktop/R-graph-gallery/world_shape_file/", layer = "TM_WORLD_BORDERS_SIMPL-0.3", verbose = FALSE)
africa <- my_spdf[my_spdf@data$REGION == 2, ]

africa@data$POP2005 <- as.numeric(africa@data$POP2005)

# Use the cartography library to do the choropleth map
library(cartography)
choroLayer(spdf = africa, df = africa@data, var = "POP2005")
title("Number of people living in Africa in 2005")
```
</div>
</div>



```{r thecode2, eval=FALSE}
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
