---
title: "Most basic background map with R and Leaflet"
descriptionMeta: "This post is a very basic introduction to the leaflet R package. It shows how to display a background map using default parameters, with reproducible code provided."
descriptionTop: "This post is a very basic introduction to the `leaflet` R package. It shows how to display a [background map](map.html) using default parameters, with reproducible code provided."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "179-show-a-map-with-leaflet-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This page explains how to display a [background map](map.html) with the R `leaflet` package. 

First initiate the map with the `leaflet()` function. Then add tiles with `addTiles()`. Note the use of the `%>%` operator to "pipe" functions. Here you can see how to get the same result using or not this operator. 

By default, you get the map beside. See next charts to learn how to [zoom on a zone](179-2-select-a-zone-in-leaflet-map.html), change [background style](180-change-background-in-leaflet-map.html).

</div>


<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/backgroundMapBasic.html" height="260" width="100%" style="border:none; max-width: 500px"></iframe></center>

</div>
</div>

<br><br>

```{r, eval=FALSE}
# Load the library
library(leaflet)
 
# Note: if you do not already installed it, install it with:
# install.packages("leaflet")

# Initialize the leaflet map with the leaflet() function
m <- leaflet()
# Then we Add default OpenStreetMap map tiles
m <- addTiles(m)
m
 
# Same stuff but using the %>% operator
m <- leaflet() %>% 
  addTiles()
m

# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/backgroundMapBasic.html"))
```






</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
