---
title: "Change background tile with leaflet and R"
descriptionMeta: "This post shows how to change the background tile used by the leaflet R package. It provides an overview of the most common tiles with their associated code."
descriptionTop: "This post shows how to change the background tile used by the `leaflet` R package. It provides an overview of the most common options with their associated code."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "180-change-background-in-leaflet-map"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" >


# Loading a tile
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This post follows the previous posts on `leaflet`: most [basic map](179-show-a-map-with-leaflet-r.html), and zooming on a [specific area](179-2-select-a-zone-in-leaflet-map.html). 

`Leaflet` offers several tiles to customize the background used on a map. 

A tile can be loaded thanks to the `addProviderTiles()` function.

</div>


<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/backgroundMapTile.html" height="400" width="100%" style="border:none; max-width: 500px"></iframe></center>

</div>
</div>

<br><br>

```{r, eval=FALSE}
# Load the library
library(leaflet)

# Note: if you do not already installed it, install it with:
# install.packages("leaflet")

# Background 1: NASA
m <- leaflet() %>% 
   addTiles() %>% 
   setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>% 
   addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")
m
 
# Background 2: World Imagery
m <- leaflet() %>% 
   addTiles() %>% 
   setView( lng = 2.34, lat = 48.85, zoom = 3 ) %>% 
   addProviderTiles("Esri.WorldImagery")
m

# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/backgroundMapTile.html", width="1000px"))
```




# List of tiles
***

Many tiles are made available. A complete list can be found [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html). Most famous are probably:

- Nasa: `NASAGIBS.ViirsEarthAtNight2012`
- Google map: `Esri.WorldImagery`
- Gray: `Esri.WorldGrayCanvas`
- Terrain: `Esri.WorldTerrain`
- Topo Map: `Esri.WorldTopoMap`












</div>






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
