---
title: "Change and customize circles and rectangles on a leaflet map"
descriptionMeta: "How to change and customize circles and rectangles on a leaflet map. Three examples are provided: basic circle, custom circle and circle with size that change when you zoom. Step by step explanation with reproducible code."
descriptionTop: "Learn how to change and customize circles and rectangles on a `leaflet` map. Three examples are provided: **basic** circle, **custom** circle and circle with **size that change** when you zoom."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "182-add-circles-rectangles-on-leaflet-map"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(
   warning=FALSE,
   message=FALSE,
   fig.align='center'
)
```

<div class="container" >




# List of tiles
***

Let's start with basic circle (default propose by `leaflet`). Please note that since we use here the `addCircleMarkers()` function, the **size of the circle** are set in pixels. It means that markers will keep always the **same size**, even when you zoom in or out! 

See below in order to find how to change the size of the circle, **dynamically or not**.

```{r}
#library
library(leaflet)
 
# Create 20 markers (Random points)
data = data.frame(
   long=sample(seq(-150,150),20),
   lat=sample(seq(-50,50),20),
   val=round(rnorm(20),2),
   name=paste("point",letters[1:20],sep="_")
) 
 
# Show a circle at each position
m = leaflet(data = data) %>%
   addTiles() %>%
   addCircleMarkers(~long, ~lat , popup = ~as.character(name))
m

# save the widget
# library(htmlwidgets)
# saveWidget(m, file="HtmlWidget/leaflet-182-add-circles.html", selfcontained = F)
```

<center>
   <iframe src="HtmlWidget/leaflet-182-add-circles.html" width="100%" height="400"></iframe>
</center>


# Custom circles
***

It is possible to custom the circles. Really helpful since it permits to change it **following the value of a variable** from the dataset. Here, the size of the circle is defined by the `val`

```{r}
 
# Create 20 markers (Random points)
data = data.frame(
   long=sample(seq(-150,150),20),
   lat=sample(seq(-50,50),20),
   val=round(rnorm(20),2),
   name=paste("point",letters[1:20],sep="_")
) 
 
# Show a CUSTOM circle at each position. Size defined in Pixel. Size does not change when you zoom
m=leaflet(data = data) %>%
   addTiles() %>%
   addCircleMarkers(
      ~long, ~lat, 
      radius=~val*14 , 
      color=~ifelse(data$val>0 , "red", "orange"),
      stroke = TRUE, 
      fillOpacity = 0.2,
      popup = ~as.character(name)
   ) 
m

# save the widget
# library(htmlwidgets)
# saveWidget(m, file="HtmlWidget/leaflet-182-add-circles-2.html", selfcontained = F)
```

<center>
   <iframe src="HtmlWidget/leaflet-182-add-circles-2.html" width="100%" height="400"></iframe>
</center>


# Custom circles with size that change when you zoom
***

You can also add circles with the `addCircle()` function. This time, the radius of the circle is **defined in meters**. It means that its **size will change** when you zoom on the map!

```{r}
#library
library(leaflet)
 
# Create 20 markers (Random points)
data = data.frame(
   long=sample(seq(-150,150),20),
   lat=sample(seq(-50,50),20),
   val=round(rnorm(20),2),
   name=paste("point",letters[1:20],sep="_")
) 
 
# Show a CUSTOM circle at each position. Size in meters --> change when you zoom.
m = leaflet(data = data) %>%
   addTiles() %>%
   addCircles(
      ~long, ~lat, 
      radius=~val*1000000 , 
      color=~ifelse(data$val>0 , "red", "orange"),
      stroke = TRUE, 
      fillOpacity = 0.2,
      popup = ~as.character(name)
   ) %>% 
  setView(lng = 166.45, lat = 21, zoom = 2)
m

# save the widget
# library(htmlwidgets)
# saveWidget(m, file="HtmlWidget/leaflet-182-add-circles-3.html", selfcontained = F)
```

<center>
   <iframe src="HtmlWidget/leaflet-182-add-circles-3.html" width="100%" height="400"></iframe>
</center>

# Add rectangles
***

Lastly, how to **add a rectangle** with the `addRectangles()` function. You have to give the **2 coordinates of 2 opposites** corners of the rectangle.

```{r}
#library
library(leaflet)
 
# Create 20 markers (Random points)
data = data.frame(
   long=sample(seq(-150,150),20),
   lat=sample(seq(-50,50),20),
   val=round(rnorm(20),2),
   name=paste("point",letters[1:20],sep="_")
) 
 
# Show a rectangle
m = leaflet() %>% addTiles() %>%  
  addRectangles(
    lng1=-118.456554, lat1=34.078039,
    lng2=-118.436383, lat2=34.062717,
    fillColor = "transparent"
  )
m

# save the widget
# library(htmlwidgets)
# saveWidget(m, file="HtmlWidget/leaflet-182-add-rectangles.html", selfcontained = F)
```

<center>
   <iframe src="HtmlWidget/leaflet-182-add-rectangles.html" width="100%" height="400"></iframe>
</center>


# Going further
***

You might be interested in how to [change the tile style](180-change-background-in-leaflet-map.html) and the [map section](map.html) of the leaflet map.

</div> 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
