---
title: "Interactive choropleth map with R and leaflet"
descriptionMeta: "This post explains how to build an interactive choropleth map with R and the leaflet package."
descriptionTop: "This post explains how to build an interactive [choropleth map](choropleth-map.html) with R and the `leaflet` package. A tooltip is available for each region on hover, and the map is zoomable."
sectionText: "Choropleth section"
sectionLink: "choropleth-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/choropleth.html"
url: "183-choropleth-map-with-leaflet"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container" style="padding-top: 50px" >



`leaflet` is a R package allowing to build **interactive maps**. If you're not familiar to it, have a look to this [leaflet introduction](179-show-a-map-with-leaflet-r.html). This post is a step-by-step tutorial leading to the following [choropleth map](choropleth-map.html).

Here is an example of what you can do with `leaflet` (try to zoom in and out!):

<br>

<center><iframe src="./HtmlWidget/choroplethLeaflet5.html" height="400" width="100%" style="border:none;"></iframe></center>

<br>










# Find, Download and Load geospatial data
***
<u>Note</u>: this step is described in detail [here](168-load-a-shape-file-into-r.html). Read it if you are not familiar with geospatial data management in R.

The region boundaries required to make maps are usually stored in **geospatial objects**. Those objects can come from [shapefiles](168-load-a-shape-file-into-r.html), [geojson files](325-background-map-from-geojson-format-in-r.html) or provided in a [R package](278-the-maps-library.html). See the [background map](map.html) section for possibilities.

This tutorial uses a geospatial object stored in a shape file available [here](http://thematicmapping.org/downloads/world_borders.php). Start by downloading the file:

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip", destfile = "DATA/world_shape_file.zip")
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
system("unzip DATA/world_shape_file.zip")
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
```

<br>

And load it in R

```{r}
# Read this shape file with the sf library.
library(sf)
world_sf <- read_sf(paste0(
  getwd(), "/DATA/world_shape_file/",
  "TM_WORLD_BORDERS_SIMPL-0.3.shp"
))

# Clean the data object
library(dplyr)
world_sf <- world_sf %>%
  mutate(POP2005 = ifelse(POP2005 == 0, NA, round(POP2005 / 1000000, 2)))

# -- > Now you have a sf object (simple feature data frame). You can start doing maps!
```








# Default choropleth
***

It is now possible to draw a first [choropleth map](choropleth-map.html). Here are the main steps to follow:

- create a **color palette** thanks to the `colorNumeric()` function
- make the **background map** with `leaflet()`, `addTiles()` and `setView()`
- use `addPolygons()` to add the **shape of all country**, with a color representing the number of people in the country. 

The resulting map is quite disappointing: China and India having very numerous population, all the variation between other countries gets hard to observe on the map.

<center><iframe src="./HtmlWidget/choroplethLeaflet1.html" height="400" width="100%" style="border:none;"></iframe></center>


```{r, eval=FALSE}
# Library
library(leaflet)

# Create a color palette for the map:
mypalette <- colorNumeric(
  palette = "viridis", domain = world_sf$POP2005,
  na.color = "transparent"
)
mypalette(c(45, 43))

# Basic choropleth with leaflet?
m <- leaflet(world_sf) %>%
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 2) %>%
  addPolygons(fillColor = ~ mypalette(POP2005), stroke = FALSE)

m

# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/choroplethLeaflet1.html"))
```










# Visualize the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

In a choropleth map, each region has a color that represents the value of a numeric variable (population here).

It is a good practice to check the **distribution** of this variable to understand what kind of color scale should be used. Using a [histogram](histogram.html) is often a good option for that:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# load ggplot2
library(ggplot2)

# Distribution of the population per country?
ggplot(world_sf, aes(x = POP2005)) +
  geom_histogram(bins = 20, fill = "#69b3a2", color = "white") +
  xlab("Population (M)") +
  theme_bw()
```

</div>
</div>

```{r thecode, eval=FALSE}
```













# Change color scale
***

There are several ways to translate a numeric variable to a palette of color. Leaflet offers 3 options: 

- quantile with `colorQuantile`
- numeric with `colorNumeric` 
- bin with `colorBin`

Results can be very different and the best option usually depends on the distribution of your input data.


Quantile
<center><iframe src="./HtmlWidget/choroplethLeaflet2.html" height="200" width="100%" style="border:none;"></iframe></center>

Numeric
<center><iframe src="./HtmlWidget/choroplethLeaflet3.html" height="200" width="100%" style="border:none;"></iframe></center>

Bins
<center><iframe src="./HtmlWidget/choroplethLeaflet4.html" height="200" width="100%" style="border:none;"></iframe></center>


```{r, eval=FALSE}
# Color by quantile
m <- leaflet(world_sf) %>%
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 2) %>%
  addPolygons(
    stroke = FALSE, fillOpacity = 0.5,
    smoothFactor = 0.5, color = ~ colorQuantile("YlOrRd", POP2005)(POP2005)
  )
m

# # save the widget in a html file if needed.
# htmlwidgets::saveWidget(m, file=paste0( getwd(), "/HtmlWidget/choroplethLeaflet2.html"))

# Numeric palette
m <- leaflet(world_sf) %>%
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 2) %>%
  addPolygons(
    stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5,
    color = ~ colorNumeric("YlOrRd", POP2005)(POP2005)
  )
m


# htmlwidgets::saveWidget(m, file=paste0( getwd(), "/HtmlWidget/choroplethLeaflet3.html"))

# Bin
m <- leaflet(world_sf) %>%
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 2) %>%
  addPolygons(
    stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5,
    color = ~ colorBin("YlOrRd", POP2005)(POP2005)
  )
m

# htmlwidgets::saveWidget(m, file=paste0( getwd(), "/HtmlWidget/choroplethLeaflet4.html"))
```












# Customizied leaflet choropleth map
***

In order to get a **quality** [choropleth map](choropleth-map.html), there are several improvements we need to apply:

- Add a legend with `addLegend()`
- Change the **color scale**: binning is a good option here since it **avoids to put too much weight** on China and India
- Add a tooltip with `labelOptions`. When you hover a specific region, a box appears with custom text. This is very handy to add additional information to your map.

Here is the result and the code:


<center><iframe src="./HtmlWidget/choroplethLeaflet5.html" height="400" width="100%" style="border:none;"></iframe></center>


```{r, eval=FALSE}
# Create a color palette with handmade bins.
library(RColorBrewer)
mybins <- c(0, 10, 20, 50, 100, 500, Inf)
mypalette <- colorBin(
  palette = "YlOrBr", domain = world_sf$POP2005,
  na.color = "transparent", bins = mybins
)

# Prepare the text for tooltips:
mytext <- paste(
  "Country: ", world_sf$NAME, "<br/>",
  "Area: ", world_sf$AREA, "<br/>",
  "Population: ", round(world_sf$POP2005, 2),
  sep = ""
) %>%
  lapply(htmltools::HTML)

# Final Map
m <- leaflet(world_sf) %>%
  addTiles() %>%
  setView(lat = 10, lng = 0, zoom = 2) %>%
  addPolygons(
    fillColor = ~ mypalette(POP2005),
    stroke = TRUE,
    fillOpacity = 0.9,
    color = "white",
    weight = 0.3,
    label = mytext,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = mypalette, values = ~POP2005, opacity = 0.9,
    title = "Population (M)", position = "bottomleft"
  )

m

# save the widget in a html file if needed.
# htmlwidgets::saveWidget(m, file=paste0( getwd(), "/HtmlWidget/choroplethLeaflet5.html"))
```
  





# Going further
***

This post explains how to build a basic [choropleth map](choropleth-map.html) with R and the leaflet package.

You might be interested in how to create a [choropleth map in ggplot2](327-chloropleth-map-from-geojson-with-ggplot2.html), and more generally in the [choropleth section](choropleth-map.html).

</div>


```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
