---
title: "Interactive bubble map with R and leaflet"
descriptionMeta: "This post explains how to build a bubble map with R and the leaflet package. It shows a set of earthquakes and their intensity in the Pacific Ocean. Reproducible code is provided."
descriptionTop: "This post explains how to build a [bubble map](bubble-map.html) with R and the `leaflet` package. It shows a set of earthquakes and their intensity in the Pacific Ocean."
sectionText: "Bubble map section"
sectionLink: "bubble-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/bubblemap.html"
url: "19-map-leafletr"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 50px" >



`leaflet` is a R package allowing to build interactive maps. This post follows the [leaflet introduction](179-show-a-map-with-leaflet-r.html) and explain how to build a [bubble map](bubble-map.html).

The input required to build such a map is a list of GPS coordinates, with a numeric value attributed to each location to control the bubble size and color.

Markers are added using the `addCircleMarkers()` function, and bubble color is mapped to a color scale build with binning.

<br>

<center><iframe src="./HtmlWidget/bubblemapQuakes.html" height="400" width="100%" style="border:none;"></iframe></center>

<br>

```{r, eval=FALSE}
# Library
library(leaflet)

# load example data (Fiji Earthquakes) + keep only 100 first lines
data(quakes)
quakes <-  head(quakes, 100)

# Create a color palette with handmade bins.
mybins <- seq(4, 6.5, by=0.5)
mypalette <- colorBin( palette="YlOrBr", domain=quakes$mag, na.color="transparent", bins=mybins)

# Prepare the text for the tooltip:
mytext <- paste(
   "Depth: ", quakes$depth, "<br/>", 
   "Stations: ", quakes$stations, "<br/>", 
   "Magnitude: ", quakes$mag, sep="") %>%
  lapply(htmltools::HTML)

# Final Map
m <- leaflet(quakes) %>% 
  addTiles()  %>% 
  setView( lat=-27, lng=170 , zoom=4) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(~long, ~lat, 
    fillColor = ~mypalette(mag), fillOpacity = 0.7, color="white", radius=8, stroke=FALSE,
    label = mytext,
    labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "13px", direction = "auto")
  ) %>%
  addLegend( pal=mypalette, values=~mag, opacity=0.9, title = "Magnitude", position = "bottomright" )

m 

# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/bubblemapQuakes.html"))
```











</div>






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
