---
title: "Manage dates data with base R"
descriptionMeta: "This post explains how to deal with date data in base R. It takes a connected scatterplot as an example and display several options to deal with dates."
descriptionTop: "This post explains how to deal with date data in base R. It takes a [connected scatterplot](connected-scatterplot.html) as an example and display several options to deal with dates."
sectionText: "Time series section"
sectionLink: "time-series.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/connectedscatter.html"
url: "191-manage-date-data"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Important note about the `lubridate()` library.
***
I strongly advise to have a look to the `lubridate()` library. It allows to easily manipulate the date format, and is very powerfull in conjunction with [ggplot2](ggplot2-package.html). Have a look to the [time series section](time-series.html) of the gallery.


# Is your date recognized as a date?
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

R offers a special data type for dates. It is important to use it since it will make the creation of charts lot easier. 

The `str()` function allows to check the type of each column. In the example beside, the `date` column is recognized as a `factor`

</div>


<div class = "col-md-6  col-sm-12">

```{r}
# Create data
set.seed(124)
date <- paste(   "2015/03/" , sample(seq(1,31),6) , sep="")
value <- sample(seq(1,100) , 6)
data <- data.frame(date,value)

# Date and time are recognized as factor:
str(data)
```


</div>
</div>








# Why it matters
***


<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The issue is that your plot is gonna be very disapointing if the date is not recognized properly, as shown beside.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
set.seed(124)
date <- paste(   "2015/03/" , sample(seq(1,31),6) , sep="")
value <- sample(seq(1,100) , 6)
data <- data.frame(date,value)

# Date and time are recognized as factor:
#str(data)

# So ploting them works bad --> wrong order, date without value are not represented, 
plot(data$value~data$date, type="b")
```


</div>
</div>



```{r thecode2, eval=FALSE}
```






# Switch to date format
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

You can use the `as.Date()` function to specify that a column is at the date format. Now, with a bit of customization, we can get a nice [connected scatterplot](connected-scatterplot.html) from our data:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
set.seed(124)
date <- paste(   "2015/03/" , sample(seq(1,31),6) , sep="")
value <- sample(seq(1,100) , 6)
data <- data.frame(date,value)

# Let's change the date to the "date" format:
data$date <- as.Date(data$date)
 
# So we can sort the table:
data <- data[order(data$date) , ]
 
# Easy to make it better now:
plot(data$value~data$date , type="b" , lwd=3 , col=rgb(0.1,0.7,0.1,0.8) , ylab="value of ..." , xlab="date" , bty="l" , pch=20 , cex=4)
abline(h=seq(0,100,10) , col="grey", lwd=0.8)

```


</div>
</div>



```{r thecode3, eval=FALSE}
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
