---
title: "Available theme for ggplot2"
descriptionMeta: "A set of pre-built theme are available to quickly customize a ggplot2 chart. This post aims to give an overview of the available themes, provided by ggplot2 itself or by other packages."
descriptionTop: "A set of pre-built theme are available to quickly customize a `ggplot2` chart. This post aims to give an overview of the available themes, provided by `ggplot2` itself or by other packages."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "192-ggplot-themes"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Themes provided by `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# library
library(ggplot2)
library(gridExtra)
 
# create data
set.seed(123)
var=rnorm(1000)
 
# Without theme
plot1 <- qplot(var , fill=I(rgb(0.1,0.2,0.4,0.6)) )
 
# With themes
plot2 = plot1+theme_bw()+annotate("text", x = -1.9, y = 75, label = "bw()" , col="orange" , size=4)
plot3 = plot1+theme_classic()+annotate("text", x = -1.9, y = 75, label = "classic()" , col="orange" , size=4)
plot4 = plot1+theme_gray()+annotate("text", x = -1.9, y = 75, label = "gray()" , col="orange" , size=4)
plot5 = plot1+theme_linedraw()+annotate("text", x = -1.9, y = 75, label = "linedraw()" , col="orange" , size=4)
plot6 = plot1+theme_dark()+annotate("text", x = -1.9, y = 75, label = "dark()" , col="orange" , size=4)
plot7 = plot1+theme_get()+annotate("text", x = -1.9, y = 75, label = "get()" , col="orange" , size=4)
plot8 = plot1+theme_minimal()+annotate("text", x = -1.9, y = 75, label = "minimal()" , col="orange" , size=4)
 
 
# Arrange and display the plots into a 2x1 grid
grid.arrange(plot1,plot2,plot3,plot4, ncol=2)
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
grid.arrange(plot5,plot6,plot7,plot8, ncol=2)
```

</div>

</div>

Two main types of grid exist with `ggplot2`: major and minor. They are controled thanks to the `panel.grid.major` and `panel.grid.minor` options.

Once more, you can add the options `.y` or `.x` at the end of the function name to control one orientation only.

Features are wrapped in an `element_line()` function. Specifying `element_blank()` will simply removing the grid.




```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```





# Code used for the `ggplot2` page.
***

Here is the code used for the examples displayed on the [ggplot2 page](ggplot2-package.html). It shows how to used most of the common packages related to `ggplot2` themes.

[See how themes look like](ggplot2-package.html#themes).
```{r, eval=FALSE}
# library
library(ggplot2)
 
# My margini
mytheme <- theme(
      plot.margin=unit(rep(1.3,4),"cm")
      ) 

# Plot
p <- ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species, shape=Species)) + 
    geom_point(size=6, alpha=0.6, show.legend = FALSE)

setwd("~/Desktop/R-graph-gallery/img/graph")
png("192-ggplot-theme_default.png") ;  p + mytheme ; dev.off()
png("192-ggplot-theme_bw.png") ;  p + theme_bw() + mytheme ; dev.off()
png("192-ggplot-theme_minimal.png") ;  p + theme_minimal() + mytheme ; dev.off()
png("192-ggplot-theme_classic.png") ;  p + theme_classic() + mytheme ; dev.off()
png("192-ggplot-theme_gray.png") ;  p + theme_gray() + mytheme ; dev.off()

#library(ggthemes)
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_economist.png") ;  p + theme_economist() + mytheme ; dev.off()
png("192-ggplot-theme_fivethirtyeight.png") ;  p + theme_fivethirtyeight() + mytheme ; dev.off()
png("192-ggplot-theme_tufte.png") ;  p + theme_tufte() + mytheme ; dev.off()
png("192-ggplot-theme_gdocs.png") ;  p + theme_gdocs() + mytheme ; dev.off()
png("192-ggplot-theme_wsj.png") ;  p + theme_wsj() + mytheme ; dev.off()
png("192-ggplot-theme_calc.png") ;  p + theme_calc() + scale_colour_calc() + mytheme ; dev.off()
png("192-ggplot-theme_hc.png") ;  p + theme_hc() + scale_colour_hc() + mytheme ; dev.off()

# library(hrbrthemes)
png("192-ggplot-theme_ipsum.png") ;  p + theme_ipsum() + scale_color_ipsum() + mytheme ; dev.off()

# library(egg)
png("192-ggplot-theme_article.png") ;  p + theme_article() + mytheme ; dev.off()

# library(ggpubr)
png("192-ggplot-theme_pubclean.png") ;  p + theme_pubclean() + mytheme ; dev.off()
# library(bigstatsr)
png("192-ggplot-theme_bigstatsr.png") ;  p + theme_bigstatsr() + mytheme ; dev.off()


png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()
png("192-ggplot-theme_excel.png") ;  p + theme_excel() + mytheme ; dev.off()

theme_bigstatsr()

```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
