---
title: "Barplot for likert type items"
descriptionMeta: "This post shows how to use the likert R package. It allows to build 0-centered stacked barplot  to study likert type items."
descriptionTop: "This post shows how to use the likert R package. It allows to build 0-centered stacked barplot  to study likert type items."
sectionText: "Stacked bar section"
sectionLink: "stacked-barplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "202-barplot-for-likert-type-items"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">






<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Likert](https://github.com/jbryer/likert) is an R package designed to help analyzing and visualizing Likert type items. It has been developped by Jason Bryer and Kim Speerschneider.

This barplot comes from the [demo page](http://jason.bryer.org/likert/) and has been sent by [Carlos Ortega](https://www.linkedin.com/in/carlosortegafernandez).

 It allows to analyse the reading attitudes from a panel of people. Each line represents a question. The barplot explains the feeling of people concerning this question.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(likert) 
 
# Use a provided dataset
data(pisaitems) 
items28 <- pisaitems[, substr(names(pisaitems), 1, 5) == "ST24Q"] 
 
# Build plot
p <- likert(items28) 
plot(p)
```
</div>
</div>



```{r thecode, eval=FALSE}
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
