---
title: "Basic R barplot customization"
descriptionMeta: "The barplot() function allows to build a barplot in base R. Learn how to customize the chart: color, bar width, orientation and more."
descriptionTop: "The `barplot()` function allows to build a [barplot](barplot.html) in base R. Learn how to customize the chart: color, bar width, orientation and more."
sectionText: "Barchart section"
sectionLink: "barplot.html"
DataToVizText: "Barplot tips"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "209-the-options-of-barplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Start basic: the `barplot()` function
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the most basic [barplot](barplot.html) you can build with R and the barplot() funtion. It was described in [graph #208](208-basic-barplot.html).

This post describes how to custom this basic barplot.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# The most basic barplot you can do:
barplot(height=data$value, names=data$name)
```
</div>
</div>



```{r thecode, eval=FALSE}
```





# Custom color {#color}
***

```{r thecode2, echo=FALSE, out.width = c("33%","33%","33%"), fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Uniform color
barplot(height=data$value, names=data$name, col=rgb(0.2,0.4,0.6,0.6) )

# Specific color for each bar? Use a well known palette
library(RColorBrewer)
coul <- brewer.pal(5, "Set2") 
barplot(height=data$value, names=data$name, col=coul )

# Change border color
barplot(height=data$value, names=data$name, border="#69b3a2", col="white" )
```
Here are 2 examples showing how to custom the barplot color:

- uniform color with `col`, asking one color only
- using a palette coming from `RColorBrewer`
- change border color with the `border` argument

```{r thecode2, eval=FALSE}
```








# Title, Axis label, Custom limits {#title}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Usual customizations with `xlab`, `ylab`, `main` and `ylim`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Uniform color
barplot(height=data$value, names=data$name, 
        col=rgb(0.8,0.1,0.1,0.6),
        xlab="categories", 
        ylab="values", 
        main="My title", 
        ylim=c(0,40)
        )

```
</div>
</div>



```{r thecode4, eval=FALSE}
```


















# Horizontal barplot {#horiz}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Usual customizations with `xlab`, `ylab`, `main` and `ylim`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Uniform color
barplot(height=data$value, names=data$name, 
        col="#69b3a2",
        horiz=T, las=1
        )

```
</div>
</div>



```{r thecode5, eval=FALSE}
```














# Bar width & space between bars {#width}
***

```{r thecode6, echo=FALSE, out.width = c("50%","50%"), fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Control space:
barplot(height=data$value, names=data$name, col=rgb(0.2,0.4,0.6,0.6), space=c(0.1,0.2,3,1.5,0.3) )
 
# Control width:
barplot(height=data$value, names=data$name, col=rgb(0.2,0.4,0.6,0.6), width=c(0.1,0.2,3,1.5,0.3) )
```

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to control the space between bars and the width of the bars using `space` and `width`. 

Can be usefull to represent the number of value behind each bar.

</div>


<div class = "col-md-6  col-sm-12">


</div>
</div>



```{r thecode6, eval=FALSE}
```













# Barplot texture {#texture}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Change bar texture with the `density` and `angle` arguments.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode7, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# barplot
barplot( height=data$value, names=data$name , density=c(5,10,20,30,7) , angle=c(0,45,90,11,36) , col="brown"   )

```
</div>
</div>



```{r thecode7, eval=FALSE}
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
