---
title: "Advanced R barplot customization"
descriptionMeta: "Take your basic R barplot to the next step: modify axis, label orientation, margins, and more."
descriptionTop: "Take your basic R [barplot](barplot.html) to the next step: modify axis, label orientation, margins, and more."
sectionText: "Barchart section"
sectionLink: "barplot.html"
DataToVizText: "Barplot tips"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "210-custom-barplot-layout"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Start basic: the `barplot()` function
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Graph #208](208-basic-barplot.html) describes the most simple barchart you can do with `R` and the `barplot()` function. [Graph #209](209-the-options-of-barplot) shows the basic options of `barplot()`. 

Let's recall how to build a basic [barplot](barplot.html):

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# The most basic barplot you can do:
barplot(height=data$value, names=data$name, col="#69b3a2")
```
</div>
</div>



```{r thecode, eval=FALSE}
```





# Axis labels orientation with `las` {#las}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `las` argument allows to change the orientation of the axis labels:  

- `0`: always parallel to the axis
- `1`: always horizontal
- `2`: always perpendicular to the axis
- `3`: always vertical. 

This is specially helpful for horizontal bar chart.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# The most basic barplot you can do:
barplot(height=data$value, names=data$name, col="#69b3a2", horiz=T , las=1)
```
</div>
</div>



```{r thecode2, eval=FALSE}
```








# Change group labels with `names.arg` {#labels}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Change the group names using the `names.arg` argument. The vector you provide must be the same length as the number of categories.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Uniform color
barplot(height=data$value, names.arg=c("group1","group2","group3","group4","group5"), col="#69b3a2")

```
</div>
</div>



```{r thecode4, eval=FALSE}
```

















# Axis labels & axis title style {#width}
***

```{r thecode6, echo=FALSE, out.width = c("50%","50%"), fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Customize labels (left)
barplot(height=data$value, names=data$name, 
        names.arg=c("group1","group2","group3","group4","group5"), 
        font.axis=2, 
        col.axis="orange", 
        cex.axis=1.5 
        )

# Customize title (right)
barplot(height=data$value, names=data$name, 
        xlab="category", 
        font.lab=2, 
        col.lab="orange", 
        cex.lab=2  
        )
```

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Customize the labels:  

- `font.axis`: font: 1: normal, 2: bold, 3: italic, 4: bold italic
- `col.axis`: color
- `cex.axis`: size

Customize axis title:

- `font.lab`
- `col.lab`
- `cex.lab`

</div>


<div class = "col-md-6  col-sm-12">


</div>
</div>



```{r thecode6, eval=FALSE}
```
















# Increase margin size {#margin}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If your group names are long, you need to:

- rotate them to avoid overlapping. This is done with `las`
- increase bottom margin size using the `mar` parameter of the `par()` function. Four values are provided: bottom, left, top, right respectively.

<u>Note</u>: prefer a [horizontal barplot](209-the-options-of-barplot.html#horiz) in this case.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# create dummy data
data <- data.frame(
  name=letters[1:5],
  value=sample(seq(4,15),5)
)

# Increase margin size
par(mar=c(11,4,4,4))

# Uniform color
barplot(height=data$value,
        col="#69b3a2",
        names.arg=c("very long group name 1","very long group name 2","very long group name 3","very long group name 4","very long group name 5"), 
        las=2 
)

```
</div>
</div>



```{r thecode5, eval=FALSE}
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
