---
title: "Basic barplot with ggplot2"
descriptionMeta: "This post explains how to draw barplots with R and ggplot2, using the geom_bar() function. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to draw barplots with `R` and [ggplot2](ggplot2.html), using the `geom_bar()` function. It starts with the most basic example and describes a few possible customizations."
sectionText: "Barchart section"
sectionLink: "barplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "218-basic-barplots-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic barplot with `geom_bar()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the most basic barplot you can build using the [ggplot2](ggplot2.html) package. It follows those steps:

- always start by calling the `ggplot()` function.
- then specify the `data` object. It has to be a data frame. And it needs one numeric and one categorical variable.
- then come thes aesthetics, set in the `aes()` function: set the categoric variable for the X axis, use the numeric for the Y axis
- finally call `geom_bar()`. You have to specify `stat="identity"` for this kind of dataset.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create data
data <- data.frame(
  name=c("A","B","C","D","E") ,  
  value=c(3,12,5,18,45)
  )

# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# Control bar color {#color}
***

```{r thecode2, echo=FALSE, out.width = c("20%","20%","20%","20%","20%"), fig.height=7, fig.show="hold"}
# Libraries
library(ggplot2)

# 1: uniform color. Color is for the border, fill is for the inside
ggplot(mtcars, aes(x=as.factor(cyl) )) +
  geom_bar(color="blue", fill=rgb(0.1,0.4,0.5,0.7) )
 
# 2: Using Hue
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) + 
  geom_bar( ) +
  scale_fill_hue(c = 40) +
  theme(legend.position="none")
 
# 3: Using RColorBrewer
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) + 
  geom_bar( ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position="none")

 
# 4: Using greyscale:
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) + 
  geom_bar( ) +
  scale_fill_grey(start = 0.25, end = 0.75) +
  theme(legend.position="none")

 
# 5: Set manualy
ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(cyl) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red", "green", "blue") ) +
  theme(legend.position="none")
```


Here are a few different methods to control bar colors. Note that using a legend in this case is not necessary since names are already displayed on the X axis. You can remove it with `theme(legend.position="none")`.

```{r thecode2, eval=FALSE}
```










# Horizontal barplot with `coord_flip()`  {#horiz}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It often makes sense to turn your [barplot](barplot.html) horizontal. Indeed, it makes the group labels much easier to read. 

Fortunately, the `coord_flip()` function makes it a breeze.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create data
data <- data.frame(
  name=c("A","B","C","D","E") ,  
  value=c(3,12,5,18,45)
  )

# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity") +
  coord_flip()
```
</div>
</div>



```{r thecode3, eval=FALSE}
```











# Control bar width with `width`  {#width}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `width` argument of the `geom_bar()` function allows to control the bar width. It ranges between 0 and 1, 1 being full width.

See how this can be used to make bar charts with [variable width](81-barplot-with-variable-width.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create data
data <- data.frame(
  name=c("A","B","C","D","E") ,  
  value=c(3,12,5,18,45)
  )

# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_bar(stat = "identity", width=0.2) 
```
</div>
</div>



```{r thecode4, eval=FALSE}
```




# What's next?
***

This post was an overview of ggplot2 [barplots](barplot.html), showing the basic options of `geom_barplot()`. Visit the barplot section for more:

- how to [reorder](267-reorder-a-variable-in-ggplot2.html) your barplot
- how to use [variable bar width](81-barplot-with-variable-width.html)
- what about [error bars](4-barplot-with-error-bar.html)
- circular [barplots](circular-barplot.html)





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
