---
title: "Basic histogram with ggplot2"
descriptionMeta: "How to build a basic histogram with R and ggplot2: basic examples with explanations and code."
descriptionTop: "A [histogram](https://www.data-to-viz.com/graph/histogram.html) is a representation of the distribution of a numeric variable. This document explains how to build it with R and the [ggplot2 package](ggplot2-package.html). You can find more examples in the [histogram section](histogram.html."
sectionText: "Histogram Section"
sectionLink: "histogram.html"
DataToVizText: "About histogram"
DataToVizLink: "data-to-viz.com/graph/histogram.html"
url: "220-basic-ggplot2-histogram.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```


<div class="container">

 
# Basic histogram with `geom_histogram` 
***

<div class = "row">
  
<div class = "col-md-6  col-sm-12 align-self-center">

It is relatively straightforward to build a histogram with `ggplot2` thanks to the `geom_histogram()` function. Only one numeric variable is needed in the input. Note that a warning message is triggered with this code: we need to take care of the bin width as explained in the next section.

```{r}
# library
library(ggplot2)
 
# dataset:
data=data.frame(value=rnorm(100))

# basic histogram
p <- ggplot(data, aes(x=value)) + 
  geom_histogram()

#p
```

</div>

<div class = "col-md-6  col-sm-12 align-self-center">
```{r, out.width=c('100%'), echo=FALSE}
p
```

</div>
</div>










# Control bin size with `binwidth` {#binSize}
***

<div class = "row">
  
<div class = "col-md-6  col-sm-12 align-self-center">

A histogram takes as input a numeric variable and cuts it into several bins. Playing with the bin size is a very important step, since its value can have a big impact on the histogram appearance and thus on the message you're trying to convey. This concept is explained in depth in [data-to-viz](https://www.data-to-viz.com/caveat/bin_size.html).

Ggplot2 makes it a breeze to change the bin size thanks to the `binwidth` argument of the `geom_histogram` function. See below the impact it can have on the output.

</div>

<div class = "col-md-6  col-sm-12 align-self-center">
```{r}
# Libraries
library(tidyverse)
library(hrbrthemes)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/1_OneNum.csv", header=TRUE)

# plot
p <- data %>%
  filter( price<300 ) %>%
  ggplot( aes(x=price)) +
    geom_histogram( binwidth=3, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("Bin size = 3") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=15)
    )
#p
```
</div>
</div>


```{r, out.width=c('50%', '50%','50%','50%'), fig.show="hold", echo=FALSE}
p

data %>%
  filter( price<300 ) %>%
  ggplot( aes(x=price)) +
    geom_histogram( binwidth=15, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("Bin size = 15") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=15)
    )

data %>%
  filter( price<300 ) %>%
  ggplot( aes(x=price)) +
    geom_histogram( binwidth=30, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("Bin size = 30") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=15)
    )

data %>%
  filter( price<300 ) %>%
  ggplot( aes(x=price)) +
    geom_histogram( binwidth=100, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("Bin size = 100") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=15)
    )
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```







