---
title: "Faceting with `ggplot2`"
descriptionMeta: "This post describes all the available options to use small multiples with R and ggplot2. it shows how to efficiently split the chart window by row, column or both to show every group of the dataset separately."
descriptionTop: "This post describes all the available options to use small multiples with R and ggplot2. it shows how to efficiently split the chart window by row, column or both to show every group of the dataset separately."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "223-faceting-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# What is small multiple
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Faceting is the process that split the chart window in several small parts (a grid), and display a similar chart in each section. Each section usually shows the same graph for a specific group of the dataset. The result is usually called small multiple.

It is a very [powerful technique](https://www.data-to-viz.com/graph/area.html) in data visualization, and a major advantage of using `ggplot2` is that it offers functions allowing to build it in a single line of code. Here is an example showing the evolution of a few baby names in the US. ([source](https://www.data-to-viz.com/graph/area.html))
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(babynames)
library(viridis)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# Load dataset from github
don <- babynames %>% 
  filter(name %in% c("Ashley", "Amanda", "Mary", "Deborah",   "Dorothy", "Betty", "Helen", "Jennifer", "Shirley")) %>%
  filter(sex=="F")

# Plot
don %>%
  ggplot( aes(x=year, y=n, group=name, fill=name)) +
    geom_area() +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Popularity of American names in the previous 30 years") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=13)
    ) +
    facet_wrap(~name, scale="free_y")
```
</div>

</div>



```{r thecode, eval=FALSE}
```
















# Faceting with `facet_wrap()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# library & datset
library(ggplot2)
head(mtcars)
 
# Split by columns (default)
ggplot( mtcars , aes(x=mpg, y=wt, color=as.factor(cyl) )) + 
  geom_point(size=3) +  
  facet_wrap(~cyl) +
  theme(legend.position="none")
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Split by row
ggplot( mtcars , aes(x=mpg, y=wt, color=as.factor(cyl)  )) + 
  geom_point(size=3) +  
  facet_wrap(~cyl , dir="v")  +
  theme(legend.position="none")
```

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3c, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Add label at the bottom
ggplot( mtcars , aes(x=mpg, y=wt, color=as.factor(cyl)  )) + 
  geom_point(size=3) +  
  facet_wrap(~cyl , strip.position="bottom") +
  theme(legend.position="none")
```

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3d, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Use same scales for all
ggplot( mtcars , aes(x=mpg, y=wt, color=as.factor(cyl)  )) + 
  geom_point(size=3) +  
  facet_wrap(~cyl , scales="free" ) +
  theme(legend.position="none")
```

</div>
</div>



`facet_wrap()` is the most common function for faceting with `ggplot2`. It builds a new chart for each level of a categorical variable. You can add the charts horizontally (graph1) or vertically (graph2, using `dir="v"`). Note that if the number of group is big enough, `ggplot2` will automatically display charts on several rows/columns.

The grey bar showing the related level can be placed on top or on the bottom thanks to the `strip.position` option (graph3). 

Last but not least, you can choose whether every graph have the same scale or not with the `scales` option (graph4).




```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```
```{r thecode3c, eval=FALSE}
```
```{r thecode3d, eval=FALSE}
```




















# `facet_grid()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`facet_grid()` is the second function allowing to build small multiples with `ggplot2`. It allows to build one chart for each combinations of 2 categorical variables. One variable will be used for rows, the other for columns.

The two variables must be given around a `~`, the first being displayed as row, the second as column. The `facet-grid()` function also accepts the `scales` argument described above.



</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# Faceted ggplot2 using facet_grid():
ggplot( mtcars , aes(x=mpg, y=wt )) + 
  geom_point() +
  facet_grid( cyl ~ gear)
```
</div>

</div>



```{r thecode5, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
