---
title: "Introduction to the circlize package"
descriptionMeta: "This post is an introduction to the circlize package: the ultimate way to build circular charts with R. It shows how to initialize a circular section and fill it with a scatterplot."
descriptionTop: "This post is an introduction to the `circlize` package: the ultimate way to build [circular charts](chord-diagram.html) with R. It shows how to initialize a circular section and fill it with a [scatterplot](scatterplot.html)."
sectionText: "Chord section"
sectionLink: "chord-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/chord.html"
url: "224-basic-circular-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">


In R, circular plots are made using the `circlize` package. Circular plots are composed by several regions (8 here), each representing a level of a factor.  Three steps  are required to build a circular plot:




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">


- *Step 1*: Initialize the chart with `circos.initialize()`. Provide the factor vector, and the numeric values to use for the X axis. The circle will be split in as many zone as the number of levels present in your factor. Each region will be as long as the coresponding x axis.

- *Step2*: Build the regions with `circos.trackPlotRegion()`. You have to specify the factors once again, and tell what to use for the Y axis if needed.

- *Step3*: Add a chart in each region. Here `circos.trackPoints()` is used to build a scatterplot. See [chart #226](226-plot-types-for-circular-plot.html) for other chart types.



</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)
 
# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000), 
    y = runif(1000)
    )
 
# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )
 
# Step 2: Build the regions. 
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis()
    })
 
# Step 3: Add points
circos.trackPoints(data$factor, data$x, data$y, col = "blue", pch = 16, cex = 0.5) 
```
</div>
</div>



```{r thecode, eval=FALSE}
```


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
