---
title: "Circular chart customization for the circlize R package"
descriptionMeta: "How to apply basic customizations to a circular chart made with the R circlize package. Learn how to change track width, background color, start angle and more."
descriptionTop: "This post follows the previous [introduction](224-basic-circular-plot.html) to the `circlize` package. It shows how to apply basic customizations to a circular chart: how to change track width, background color, start angle and more."
sectionText: "Chord section"
sectionLink: "chord-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/chord.html"
url: "225-circular-plot-custom-a-track"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Most basic circular chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Let's remind how to build a basic circular chart with the `circlize` package. 

This is extensively described in [chart #224](224-basic-circular-plot.html).


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

# Step 2: Build the regions.
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis()
    })

# Step 3: Add points
circos.trackPoints(data$factor, data$x, data$y)
```
</div>
</div>



```{r thecode, eval=FALSE}
```













# Customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Customization can happen at 3 different levels:

- initialization: use the usual `par()` and the specific `circos.par()` functions for general stuff.
- in `circos.axis()` to customize axis
- in circos.trackPoints()` to customize the chart shapes.

Note that most of the parameters are consistent with base R. Have a look to the [scatterplot](scatterplot.html) section of the gallery for more customization.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# General Customization:
par(
  mar = c(1, 1, 1, 1),           # Margin around chart
  bg = rgb(0.4,0.1,0.7,0.05)     # background color
) 
circos.par("track.height" = 0.6) # track hight, 0.6 = 60% of total height

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

# Step2: Build regions. 
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis(
        h="top",                   # x axis on the inner or outer part of the track?
        labels=TRUE,               # show the labels of the axis?
        major.tick=TRUE,           # show ticks?
        labels.cex=0.5,            # labels size (higher=bigger)
        labels.font=1,             # labels font (1, 2, 3 , 4)
        direction="outside",       # ticks point to the outside or inside of the circle ?
        minor.ticks=4,             # Number of minor (=small) ticks
        major.tick.percentage=0.1, # The size of the ticks in percentage of the track height
        lwd=2                      # thickness of ticks and x axis.
        )
    })

# Step 3: Add points
circos.trackPoints(data$factor, data$x, data$y, col = "#69b3a2", pch = 16, cex = 0.5)
```

</div>
</div>



```{r thecode2, eval=FALSE}
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
