---
title: "Available circular chart types with circlize"
descriptionMeta: "The circlize package allows to build circular charts with R. Several chart types are offered: bar, scatter, line, abline and more. This post shows how to build them."
descriptionTop: "The `circlize` package allows to build [circular charts](chord-diagram.html) with R. Several chart types are offered: bar, scatter, line, abline and more. This post shows how to build them."
sectionText: "Chord section"
sectionLink: "chord-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/chord.html"
url: "226-plot-types-for-circular-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Circular Scatterplot
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Circular scatterplot has already been extensively described in chart [#224](224-basic-circular-plot.html) and [#225](225-circular-plot-custom-a-track.html). 

Here is a reminder:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)
circos.par("track.height" = 0.4)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

# Step 2: Build the regions.
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis()
    })

# Step 3: Add points
circos.trackPoints(data$factor, data$x, data$y, col="#69b3a2")
```
</div>
</div>



```{r thecode, eval=FALSE}
```















# Circular Line chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to switch to line chart using the `circos.trackLines()` function. Visit the [line chart](line-plot.html) section of the gallery to learn how to customize that kind of chart.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)
circos.par("track.height" = 0.4)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

# Step 2: Build the regions.
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis()
    })

# Step 3: Add points
circos.trackLines(data$factor, data$x[order(data$x)], data$y[order(data$x)], col = rgb(0.1,0.5,0.8,0.3), lwd=2)

```
</div>
</div>



```{r thecode2, eval=FALSE}
```











# Vertical ablines
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `circos.trackLines()` function can also be used to display vertical ablines.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)
circos.par("track.height" = 0.4)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

# Step 2: Build the regions.
circos.trackPlotRegion(factors = data$factor, y = data$y, panel.fun = function(x, y) {
    circos.axis()
    })

# Step 3: Add points
circos.trackLines(data$factor, data$x[order(data$x)], data$y[order(data$x)], col = rgb(0.1,0.5,0.8,0.3), lwd=2, type="h")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```














# Circular histogram
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Note that each plot type must be coherent with what you specified in the `circos.trackPlotRegion` function.

You have to specify an Y axis for a scatterplot as seen before. But not for a histogram that is built with `circos.trackHist()`.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Upload library
library(circlize)
circos.par("track.height" = 0.4)

# Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000),
    y = runif(1000)
    )

# Step1: Initialise the chart giving factor and x-axis.
circos.initialize( factors=data$factor, x=data$x )

circos.trackHist(data$factor, data$x, bg.col = "white", col = "#69b3a2")
```
</div>
</div>



```{r thecode4, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
