---
title: "Circular chart with several tracks"
descriptionMeta: "The circlize package allows to build circular charts with several tracks. This post describes how, providing explanation and reproducible code."
descriptionTop: "The `circlize` package allows to build [circular charts](chord-diagram.html) with several tracks. This post describes how, providing explanation and reproducible code.."
sectionText: "Chord section"
sectionLink: "chord-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/chord.html"
url: "227-add-several-tracks"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If you read the graph [#224](224-basic-circular-plot.html), and [#225](225-circular-plot-custom-a-track.html) you know how to make circular plots and customize them. 

It is possible to add several tracks on your circular plot: just repeat the step 2 (building regions with `circos.trackPlotRegion`) and 3 (adding dots with `circos.trackPoints` for instance). 

Note that you can custom the width of each track using `track.height`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
#library
library(circlize)
 
#Create data
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000), 
    y = runif(1000)
    )
 
#Initialize the plot.
par(mar = c(1, 1, 1, 1) ) 
circos.initialize(factors = data$factor, x = data$x )
 

# Build the regions of track #1 
circos.trackPlotRegion(factors = data$factor, y=data$y, panel.fun = function(x, y) {
    circos.axis(labels.cex=0.5, labels.font=1, lwd=0.8)
    })
# --> Add a scatterplot on it:
circos.trackPoints(data$factor, data$x, data$y, col = rgb(0.1,0.5,0.8,0.3), pch=20)
 
 
# Build the regions of track #2:
circos.trackPlotRegion(factors = data$factor, y=data$y, panel.fun = function(x, y) {
    circos.axis(labels=FALSE, major.tick=FALSE)
    })
# --> Add a scatterplot on it
circos.trackPoints(data$factor, data$x, data$y, col = rgb(0.9,0.5,0.8,0.3), pch=20, cex=2)
 
 
# Add track #3 --> don't forget you can custom the height of tracks!
circos.par("track.height" = 0.4)
circos.trackPlotRegion(factors = data$factor, y=data$y, panel.fun = function(x, y) {
    circos.axis(labels=FALSE, major.tick=FALSE)
    })
circos.trackLines(data$factor, data$x, data$y, col = rgb(0.9,0.5,0.1,0.3), pch=20, cex=2, type="h")
# and continue as long as needed!
```
</div>
</div>



```{r thecode, eval=FALSE}
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
