---
title: "Introduction to chord diagram"
descriptionMeta: "The circlize package allows to build chord diagrams, where arcs between sections are built to show flows."
descriptionTop: "The `circlize` package allows to build [chord diagrams](chord-diagram.html), where arcs between sections are built to show flows."
sectionText: "Chord section"
sectionLink: "chord-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/chord.html"
url: "228-add-links-between-regions"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Chart [#224](224-basic-circular-plot.html), and [#225](225-circular-plot-custom-a-track.html) introduced the `circlize` package and its ability to build circular charts.

It is possible to add connections between tracks with `circos.links()`. 

See [chart #122](122-a-circular-plot-with-the-circlize-package.html) for a customized version, and [chart #123](123-circular-plot-circlize-package-2.html) for an automatized version.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(circlize)
 
# Create data
set.seed(123)
data = data.frame(
    factor = sample(letters[1:8], 1000, replace = TRUE),
    x = rnorm(1000), 
    y = runif(1000)
    )
 
# Initialize the plot.
par(mar = c(1, 1, 1, 1) ) 
circos.initialize(factors = data$factor, x = data$x )
 
# Build the regions of track #1
circos.trackPlotRegion(factors = data$factor, y=data$y , bg.col = rgb(0.1,0.1,seq(0,1,0.1),0.4) , bg.border = NA)
 
# Add a link between a point and another
circos.link("a", 0, "b", 0, h = 0.4)
 
# Add a link between a point and a zone
circos.link("e", 0, "g", c(-1,1), col = "green", lwd = 2, lty = 2, border="black" )
 
# Add a link between a zone and another
circos.link("c", c(-0.5, 0.5), "d", c(-0.5,0.5), col = "red", border = "blue", h = 0.2)
```
</div>
</div>



```{r thecode, eval=FALSE}
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
