---
title: "Add color to specific groups of a boxplot"
descriptionMeta: "How to build a boxplot with basic R, with custom color for each group."
descriptionTop: "A [boxplot](boxplot.html) summarizes the distribution of a continuous variable for one or several groups. This example illustrates how to build it with basic R, coloring each group with a specific color."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "23-add-colors-to-specific-groups-of-a-boxplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [boxplot](boxplot.html) summarizes the distribution of a numeric variable for one or several groups.

It can be usefull to add colors to specific groups to highlight them. For exemple, positive and negative controls are likely to be in different colors.  

The easiest way is to give a vector (`myColor` here) of colors when you call the `boxplot()` function. Use `ifelse` statements to add the color you want to a specific name.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
#Create data
names <- c(rep("Maestro", 20) , rep("Presto", 20) , 
      rep("Nerak", 20), rep("Eskimo", 20), rep("Nairobi", 20), rep("Artiko", 20))
value <- c(  sample(3:10, 20 , replace=T) , sample(2:5, 20 , replace=T) , 
      sample(6:10, 20 , replace=T), sample(6:10, 20 , replace=T) , 
      sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) )
data <- data.frame(names,value)

# Prepare a vector of colors with specific color for Nairobi and Eskimo
myColors <- ifelse(levels(data$names)=="Nairobi" , rgb(0.1,0.1,0.7,0.5) , 
              ifelse(levels(data$names)=="Eskimo", rgb(0.8,0.1,0.3,0.6),
              "grey90" ) )

# Build the plot
boxplot(data$value ~ data$names , 
    col=myColors , 
    ylab="disease" , xlab="- variety -")
 
# Add a legend
legend("bottomleft", legend = c("Positiv control","Negativ control") , 
    col = c(rgb(0.1,0.1,0.7,0.5) , rgb(0.8,0.1,0.3,0.6)) , bty = "n", pch=20 , pt.cex = 3, cex = 1, horiz = FALSE, inset = c(0.03, 0.1))

```
</div>
</div>



```{r, eval=FALSE}
#Create data
names <- c(rep("Maestro", 20) , rep("Presto", 20) , 
      rep("Nerak", 20), rep("Eskimo", 20), rep("Nairobi", 20), rep("Artiko", 20))
value <- c(  sample(3:10, 20 , replace=T) , sample(2:5, 20 , replace=T) , 
      sample(6:10, 20 , replace=T), sample(6:10, 20 , replace=T) , 
      sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) )
data <- data.frame(names,value)

# Prepare a vector of colors with specific color for Nairobi and Eskimo
myColors <- ifelse(levels(data$names)=="Nairobi" , rgb(0.1,0.1,0.7,0.5) , 
              ifelse(levels(data$names)=="Eskimo", rgb(0.8,0.1,0.3,0.6),
              "grey90" ) )

# Build the plot
boxplot(data$value ~ data$names , 
    col=myColors , 
    ylab="disease" , xlab="- variety -")
 
# Add a legend
legend("bottomleft", legend = c("Positiv control","Negativ control") , 
    col = c(rgb(0.1,0.1,0.7,0.5) , rgb(0.8,0.1,0.3,0.6)) , bty = "n", pch=20 , pt.cex = 3, cex = 1, horiz = FALSE, inset = c(0.03, 0.1))

```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
