---
title: "How to annotate a plot in ggplot2"
descriptionMeta: "Once your chart is done, annotating it is a crucial step to make it more insightful. This post will guide you through the best practices using R and ggplot2."
descriptionTop: "Once your chart is done, annotating it is a crucial step to make it more insightful. This post will guide you through the best practices using `R` and `ggplot2`."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Why annotating?"
DataToVizLink: "data-to-viz.com/caveat/annotation.html"
url: "233-add-annotations-on-ggplot2-chart"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Adding text with `geom_text()` or `geom_label()`
***

<div class = "row">


<div class = "col-md-6  col-sm-12">

```{r thecodeA, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)

# basic graph
p <- ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point()
 
# a data frame with all the annotation info
annotation <- data.frame(
   x = c(2,4.5),
   y = c(20,25),
   label = c("label 1", "label 2")
)

# Add text
p + geom_text(data=annotation, aes( x=x, y=y, label=label),                 , 
           color="orange", 
           size=7 , angle=45, fontface="bold" )

# Note: possible to shorten with annotate:
# p +
#  annotate("text", x = c(2,4.5), y = c(20,25), 
#           label = c("label 1", "label 2") , color="orange", 
#           size=7 , angle=45, fontface="bold")

```
</div>



<div class = "col-md-6  col-sm-12">

```{r thecodeB, echo=FALSE, out.width = "100%", fig.height=7}
# Right chart: using labels
p + geom_label(data=annotation, aes( x=x, y=y, label=label),                 , 
           color="orange", 
           size=7 , angle=45, fontface="bold" )
```
</div>

</div>

Text is the most common kind of annotation. It allows to give more information on the most important part of the chart.

Using `ggplot2`, 2 main functions are available for that kind of annotation:

- `geom_text` to add a simple piece of text
- `geom_label` to add a label: framed text

Note that the `annotate()` function is a good alternative that can reduces the code length for simple cases.


```{r thecodeA, eval=FALSE}
```

```{r thecodeB, eval=FALSE}
```













# Add shapes with `annotate()`
***

<div class = "row">


<div class = "col-md-4  col-sm-12">

```{r thecode1A, echo=FALSE, out.width = "100%", fig.height=7}
# Add rectangles
p + annotate("rect", xmin=c(2,4), xmax=c(3,5), ymin=c(20,10) , ymax=c(30,20), alpha=0.2, color="blue", fill="blue")
```

</div>


<div class = "col-md-4  col-sm-12">

```{r thecode1B, echo=FALSE, out.width = "100%", fig.height=7}
# Add segments
p + annotate("segment", x = 1, xend = 3, y = 25, yend = 15, colour = "purple", size=3, alpha=0.6)
```

</div>


<div class = "col-md-4  col-sm-12">

```{r thecode1C, echo=FALSE, out.width = "100%", fig.height=7}
# Add arrow
p + annotate("segment", x = 2, xend = 4, y = 15, yend = 25, colour = "pink", size=3, alpha=0.6, arrow=arrow())
```

</div>


</div>

The `annotate()` function allows to add all kind of shape on a `ggplot2` chart. The first argument will control what kind is used: `rect` or `segment` for rectangle, segment or arrow.

```{r thecode1A, eval=FALSE}
```

```{r thecode1B, eval=FALSE}
```

```{r thecode1C, eval=FALSE}
```









# Add ablines with `geom_hline()` and `geom_vline()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

An abline is a segment that goes from one chart extremity to the other. `ggplot2` offers the `geom_hline()` and `geom_vline()` functions that are dedicated to it.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
p + 
  # horizontal
  geom_hline(yintercept=25, color="orange", size=1) + 
  # vertical
  geom_vline(xintercept=3, color="orange", size=1)
```

</div>
</div>



```{r thecode2, eval=FALSE}
```








# Add a point and a range with.. `pointrange()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Last kind of annotation, add a dot and a segment directly with `pointrange()`.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Add point and range
p + annotate("pointrange", x = 3.5, y = 20, ymin = 12, ymax = 28,colour = "orange", size = 1.5, alpha=0.4)
```

</div>
</div>



```{r thecode3, eval=FALSE}
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
