---
title: "Most basic treemap with R"
descriptionMeta: "This post explains how to build a very basic treemap with R. It uses the treemap package, provides reproducible code and explain how input data must be formatted."
descriptionTop: "This post explains how to build a very basic [treemap](treemap.html) with `R`. It uses the `treemap` package, provides reproducible code and explains how input data must be formatted."
sectionText: "Treemap section"
sectionLink: "treemap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/treemap.html"
url: "234-a-very-basic-treemap"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic treemap
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the most basic [treemap](treemap.html) you can do. The input dataset is simple: we just have 3 groups, and each has a value which we map to an area.

It allows to learn the syntax of the `treemap` library: you need to provide at least a dataset (`data`), the column that provides groups (`index`), and the column that gives the size of each group (`vSize`).

See graph [#235](235-treemap-with-subgroups.html) to learn how to add subgroups, and graph [#236](236-custom-your-treemap.html) to customize the chart appearance.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(treemap)
 
# Create data
group <- c("group-1","group-2","group-3")
value <- c(13,5,22)
data <- data.frame(group,value)
 
# treemap
treemap(data,
            index="group",
            vSize="value",
            type="index"
            )
```
</div>
</div>



```{r thecode, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
