---
title: "Customize your R treemap"
descriptionMeta: "How to customize your treemap built with R? Learn how to control borders, labels, and more. Several examples with reproducible code provided."
descriptionTop: "How to customize your [treemap](treemap.html) built with `R`? Learn how to control borders, labels, and more. Several examples with reproducible code provided."
sectionText: "Treemap section"
sectionLink: "treemap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/treemap.html"
url: "236-custom-your-treemap"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">


#Labels
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This page aims to explain how to customize R treemaps. Make sure you already understood how to build a [basic treemap](234-a-very-basic-treemap.html) with R.

The first step is to control label appearance. All the options are explained in the code below. Note that you can apply a different feature to each level of the treemap, for example using white for group labels, and orange for subgroup labels.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(treemap)
 
# Create data
group <- c(rep("group-1",4),rep("group-2",2),rep("group-3",3))
subgroup <- paste("subgroup" , c(1,2,3,4,1,2,1,2,3), sep="-")
value <- c(13,5,22,12,11,7,3,1,23)
data <- data.frame(group,subgroup,value)
 
# Custom labels:
treemap(data, index=c("group","subgroup"),     vSize="value", type="index",
 
    fontsize.labels=c(15,12),                # size of labels. Give the size per level of aggregation: size for group, size for subgroup, sub-subgroups...
    fontcolor.labels=c("white","orange"),    # Color of labels
    fontface.labels=c(2,1),                  # Font of labels: 1,2,3,4 for normal, bold, italic, bold-italic...
    bg.labels=c("transparent"),              # Background color of labels
    align.labels=list(
        c("center", "center"), 
        c("right", "bottom")
        ),                                   # Where to place labels in the rectangle?
    overlap.labels=0.5,                      # number between 0 and 1 that determines the tolerance of the overlap between labels. 0 means that labels of lower levels are not printed if higher level labels overlap, 1  means that labels are always printed. In-between values, for instance the default value .5, means that lower level labels are printed if other labels do not overlap with more than .5  times their area size.
    inflate.labels=F,                        # If true, labels are bigger when rectangle is bigger.
 
)
```
</div>
</div>



```{r thecode, eval=FALSE}
```










#Borders
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

You can control the border:

- color with `border.col`
- width with `border.lwds`

Remember that you can still provide a vector to each option: it gives the value for groups, subgroups and so on.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Custom borders:
treemap(data, index=c("group","subgroup"), vSize="value", type="index",
 
    border.col=c("black","white"),             # Color of borders of groups, of subgroups, of subsubgroups ....
    border.lwds=c(7,2)                         # Width of colors
        
    )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```














#General features
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A few other arguments for more general customization. The `palette` arguments accepts any palette from `RColorBrewer`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# General features:
treemap(data, index=c("group","subgroup"), vSize="value", 
 
    type="index",                            # How you color the treemap. type help(treemap) for more info
    palette = "Set1",                        # Select your color palette from the RColorBrewer presets or make your own.
    title="My Treemap",                      # Customize your title
    fontsize.title=12,                       # Size of the title
 
) 
```
</div>
</div>



```{r thecode3, eval=FALSE}
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
