---
title: "Interactive Treemap in R"
descriptionMeta: "With a big amount of data, a treemap can get cluttered and unreadable. Interactivity allows to keep a clean an insightful figure. Learn how to in R."
descriptionTop: "With a big amount of data, a [treemap](treemap.html) can get cluttered and unreadable. Interactivity allows to keep a clean an insightful figure. This post shows how to build an interactive treemap with `R` and the `d3treeR` package"
sectionText: "Treemap section"
sectionLink: "treemap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/treemap.html"
url: "237-interactive-treemap"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">





<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

This post follows the previous chart [#234](234-a-very-basic-treemap.html), [#235](235-treemap-with-subgroups.html) and [#236](236-custom-your-treemap.html) that describe how to build and customize [treemaps](treemap.html) with the `treemap` package.

The idea is to turn the chart interactive: you can now click on a group to zoom in and show its subgroups. Click on the group name on top to unzoom and come back to the previous state.

This is done thanks to the `d3treeR` package:

</div>


<div class = "col-md-7  col-sm-12">

<center><iframe src="HtmlWidget/interactiveTreemap.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>




```{r, eval=FALSE}
# library
library(treemap)
library(d3treeR)
 
# dataset
group <- c(rep("group-1",4),rep("group-2",2),rep("group-3",3))
subgroup <- paste("subgroup" , c(1,2,3,4,1,2,1,2,3), sep="-")
value <- c(13,5,22,12,11,7,3,1,23)
data <- data.frame(group,subgroup,value)
 
# basic treemap
p <- treemap(data,
            index=c("group","subgroup"),
            vSize="value",
            type="index",
            palette = "Set2",
            bg.labels=c("white"),
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )            
 
# make it interactive ("rootname" becomes the title of the plot):
inter <- d3tree2( p ,  rootname = "General" )

# save the widget
# library(htmlwidgets)
# saveWidget(inter, file=paste0( getwd(), "/HtmlWidget/interactiveTreemap.html"))
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
