---
title: "Axis manipulation with R and ggplot2"
descriptionMeta: "This post describes all the available options to customize chart axis with R and ggplot2. It shows how to control the axis itself, its label, title, position and more."
descriptionTop: "This post describes all the available options to customize chart axis with `R` and `ggplot2`. It shows how to control the axis itself, its label, title, position and more."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "238-custom-layout-axis-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Default `ggplot2` axis
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Let's start with a very basic `ggplot2` [scatterplot](scatterplot.html). The axis usually looks very good with default option as you can see here.

Basically two main functions will allow to customize it:

- `theme()` to change the axis appearance
- `scale_x_` and `scale_y_` to change the axis type

Let's see how to use them



</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Very basic chart
basic <- ggplot( mtcars , aes(x=mpg, y=wt)) + 
  geom_point()
basic
```
</div>

</div>



```{r thecode, eval=FALSE}
```
















# Set axis title and limits with `xlab()` and `xlim()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Two basic options that are used in almost every charts are `xlab()` and `xlim()` to control the axis title and the axis limits respectively.

<u>Note </u>: it is possible to specify only the lower or upper bound of a limit. For instance, try `xlim(0,NA)`
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
basic+
    xlab("mpg value") + 
    xlim(0,50)
```
</div>

</div>



```{r thecode2, eval=FALSE}
```











# Customize axis title appearance: `axis.title`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# Left -> both axis are modified
basic + theme(axis.title = element_text( angle = 90, color="red", size=15, face=3)) # face = title location
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Right -> only the x axis is modified
basic + theme(axis.title.x = element_text( angle = 90, color="red", size=15, face=3))
```

</div>

</div>


The `theme()` function allows to customize all parts of the `ggplot2` chart. The `axis.title.` controls the axis title appearance. Since it is text, features are wrapped in a `element_text()` function. The code below shows how to change the most common features:

```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```











# Customize axis labels: `axis.text`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Using pretty much the same process, the `axis.text()` function allows to control text label features. Once more, using `axis.text.x()` would modify the x axis only.

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
basic + 
  theme(axis.text = element_text( 
    angle = 90, 
    color="blue", 
    size=15, 
    face=3)
  )
```
</div>

</div>


```{r thecode4, eval=FALSE}
```












# Customize axis ticks and axis line with `axis.ticks` and `axis.line`
***

<div class = "row">

<div class = "col-md-4  col-sm-12 ">

```{r thecode5a, echo=FALSE, out.width = "100%", fig.height=7, fig.align="center"}
# chart 1: ticks
basic + theme(
   axis.ticks = element_line(size = 2, color="red") , 
   axis.ticks.length = unit(.5, "cm")
)
```

</div>

<div class = "col-md-4  col-sm-12">

```{r thecode5b, echo=FALSE, out.width = "100%", fig.height=7,  fig.align="center"}
# chart 2: axis lines
basic + theme(axis.line = element_line(size = 3, colour = "green", linetype=2))
```

</div>

<div class = "col-md-4  col-sm-12">

```{r thecode5c, echo=FALSE, out.width = "100%", fig.height=7,  fig.align="center"}
# chart 3: combination
ggplot( mtcars , aes(x=mpg, y=wt)) + geom_point() +
    theme(
        axis.title = element_text( color="red", size=15, face=2),
        axis.line = element_line(size = 3, colour = "green", linetype=2),
        axis.text = element_text( angle = 90, color="blue", size=15, face=2)
        )
```

</div>

</div>



The `axis.ticks()` function controls the ticks appearance. `axis.line()` controles the axis line. Both of them are lines, so options are wrapped in a `element_line()` statement. 

`linetype` controls the type of line in use, see the [ggplot2 section](ggplot2-package.html) for more.

```{r thecode5a, eval=FALSE}
```
```{r thecode5b, eval=FALSE}
```
```{r thecode5c, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
