---
title: "Building a nice legend with R and ggplot2"
descriptionMeta: "This post describes all the available options to customize the chart legend with R and ggplot2. It shows how to control the title, text, location, symbols and more."
descriptionTop: "This post describes all the available options to customize the chart legend with `R` and `ggplot2`. It shows how to control the title, text, location, symbols and more."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "239-custom-layout-legend-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">





# Default legend with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

By default, `ggplot2` will automatically build a legend on your chart as soon as a shape feature is mapped to a variable in `aes()` part of the `ggplot()` call. So if you use `color`, `shape` or `alpha`, a legend will be available.

Here is an example based on the `mtcars` dataset. This post is gonna show how to use the `theme()` function to apply all type of customization on this default legend.

<u>Note </u>: this post is strongly inspired by the doc you get typiinig `??ggplot2::theme`, give it a go!

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Very basic chart
basic <- ggplot(mtcars, aes(wt, mpg, colour = factor(cyl), shape = factor(vs) )) +
       geom_point()
basic
```
</div>

</div>



```{r thecode, eval=FALSE}
```
















# Change legend title with `labs()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `labs()` function allows to change the legend titles. You can specify one title per section of the legend, i.e. per aesthetics in use in the chart.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
basic+
  labs(
         colour = "name1",
         shape = "name2"
        )
```
</div>

</div>



```{r thecode2, eval=FALSE}
```











# Get rid of the legend: `guides()` and `theme()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# Left -> get rid of one part of the legend
basic + guides(shape=FALSE)
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Right -> only the x axis is modified
basic + theme(legend.position = "none")
```

</div>

</div>


It is possible to remove a specific part or the whole legend thanks to the `theme()` and the `guides()` function. See code below:

```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```














# Control legend position with `legend.position`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode4a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# Left -> legend around the plot
basic + theme(legend.position = "bottom")
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode4b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Right -> inside the plot area
basic + theme(
    legend.position = c(.95, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
    )
```

</div>

</div>


You can place the legend literally anywhere. 

To put it around the chart, use the `legend.position` option and specify `top`, `right`, `bottom`, or `left`. To put it inside the plot area, specify a vector of length 2, both values going between 0 and 1 and giving the x and y coordinates.

Note: the command `legend.justification` sets the corner that the position refers to.

```{r thecode4a, eval=FALSE}
```
```{r thecode4b, eval=FALSE}
```










# Legend appearance
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode5a, echo=FALSE, out.width = "100%", fig.height=7, fig.align="center"}
# custom box around legend
basic + theme(
    legend.box.background = element_rect(color="red", size=2),
    legend.box.margin = margin(116, 6, 6, 6)
)
```

</div>

<div class = "col-md-6  col-sm-12 ">

```{r thecode5b, echo=FALSE, out.width = "100%", fig.height=7,  fig.align="center"}
# custom the key
basic + theme(legend.key = element_rect(fill = "white", colour = "black"))
```

</div>

<div class = "col-md-6  col-sm-12 ">

```{r thecode5c, echo=FALSE, out.width = "100%", fig.height=7,  fig.align="center"}
# custom the text
basic + theme(legend.text = element_text(size = 8, colour = "red"))
```

</div>

<div class = "col-md-6  col-sm-12 ">

```{r thecode5d, echo=FALSE, out.width = "100%", fig.height=7,  fig.align="center"}
# custom the title
basic + theme(legend.title = element_text(face = "bold"))
```

</div>

</div>

Here are 4 examples showing how to customize the legend main features:

- box with `legend.box.`: it is a rectangle that frames the legend. Give features with `element_rect()`
- key with `legend.key`: the key is the part showing the symbols. Note that symbols will automatically be the ones used on the chart
- text with `legend.text`: here you can control the color, the size of the right part of the legend
- title with `legend.title`.

```{r thecode5a, eval=FALSE}
```
```{r thecode5b, eval=FALSE}
```
```{r thecode5c, eval=FALSE}
```
```{r thecode5d, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
