---
title: "Boxplot with variable width"
descriptionMeta: "How to make a boxplot with width proportional to category sample size in base R: code and explanation."
descriptionTop: "This examples demonstrates how to build a [boxplot](boxplot.html) with variable width. It is useful to indicate what sample size is hidden behind each box. It is a base R implementation, see [here](266-ggplot2-boxplot-with-variable-width.html) for a ggplot2 version."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "24-boxplot-with-variable-width"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

When the sample size behind each category is highly variable, it can be great to represent it through the box widths.

First calculate the proportion of each level using the `table()` function. Using these proportions will make the box twice bigger if a level is twice more represented. Then give these proportions to the `width` argument when you call the `boxplot()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
names <- c(rep("A", 20) , rep("B", 8) , rep("C", 30), rep("D", 80))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 8 , replace=T), 
       sample(1:7, 30 , replace=T), sample(3:8, 80 , replace=T) )
data <- data.frame(names,value)
 
# Calculate proportion of each level
proportion <- table(data$names)/nrow(data)
 
#Draw the boxplot, with the width proportionnal to the occurence !
boxplot(data$value ~ data$names , width=proportion , col=c("orange" , "seagreen"))
```
</div>
</div>



```{r, eval=FALSE}
# Dummy data
names <- c(rep("A", 20) , rep("B", 8) , rep("C", 30), rep("D", 80))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 8 , replace=T), 
       sample(1:7, 30 , replace=T), sample(3:8, 80 , replace=T) )
data <- data.frame(names,value)
 
 
# Calculate proportion of each level
proportion <- table(data$names)/nrow(data)
 
#Draw the boxplot, with the width proportionnal to the occurence !
boxplot(data$value ~ data$names , width=proportion , col=c("orange" , "seagreen"))
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
