---
title: "Background manipulation with R and ggplot2"
descriptionMeta: "This post describes all the available options to customize chart background with R and ggplot2. It shows how to control the background color and the minor and major gridlines."
descriptionTop: "This post describes all the available options to customize chart background with `R` and `ggplot2`. It shows how to control the background color and the minor and major gridlines."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "240-custom-layout-background-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Default `ggplot2` background
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Let's start with a very basic `ggplot2` [scatterplot](scatterplot.html). By default, `ggplot2` offers a grey background with white major and minor gridlines.

It is possible to change that thanks to the `theme()` function. Keep reading to learn how!


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Very basic chart
basic <- ggplot( mtcars , aes(x=mpg, y=wt)) +
  geom_point()
basic
```
</div>

</div>



```{r thecode, eval=FALSE}
```
















# Background color: `plot.background` and `panel.background`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Two options of the `theme()` functions are available to control the map background color. `plot_background` controls the color of the whole chart. `panel.background` controls the part between the axis.

Both are rectangles, their features is thus specified through an `element_rect()` function.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
basic + theme(
    plot.background = element_rect(fill = "green"), 
    panel.background = element_rect(fill = "red", colour="blue")
    )
```
</div>

</div>



```{r thecode2, eval=FALSE}
```











# Customize the grid: `panel.grid.major` and `panel.grid.minor`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=6, fig.align="center"}
# Modify for both axis
basic + theme(
    panel.grid.major = element_line(colour = "red", linetype = "dotted"),
    panel.grid.minor = element_line(colour = "blue", size = 2)
    )
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=6,  fig.align="center"}
# Modify y axis only (remove)
basic + theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
)
```

</div>

</div>

Two main types of grid exist with `ggplot2`: major and minor. They are controled thanks to the `panel.grid.major` and `panel.grid.minor` options.

Once more, you can add the options `.y` or `.x` at the end of the function name to control one orientation only.

Features are wrapped in an `element_line()` function. Specifying `element_blank()` will simply removing the grid.




```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
