---
title: "Customize small multiple appearance with ggplot2"
descriptionMeta: "ggplot2 makes it a breeze to build small multiples. This post shows how to customize the layout, notably using the strip options."
descriptionTop: "[ggplot2](ggplot2-package.html) makes it a breeze to build small multiples. This post shows how to customize the layout, notably using the strip options."
sectionText: "ggplot2 Section"
sectionLink: "ggplot2-package.html"
DataToVizText: "About Scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "241-custom-layout-strip-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 40px">






This page aims to provide useful R code to customize the strips of a `ggplot2` plots when using faceting. For other ggplot2 customization, visit the [dedicated page](ggplot2-package.html).

Here we want to modify non-data components, which is often done trough the `theme()` command. This page is strongly inspired from the help page of ggplot2 (`?theme`). Also, do not hesitate to visit the very strong [ggplot2 documentation](http://docs.ggplot2.org/0.9.3/theme.html) for more information.

Chart 1 is a really basic plot relying on the `mtcars` dataset. The 3 following charts show how to customize `strip` background, text, and how to modify the space between sections.


<div class = "row">

<div class = "col-md-6  col-sm-12">

```{r thecodeA, echo=FALSE, out.width = "100%", fig.height=7}

library(ggplot2)
 
# basic chart
basic=ggplot(mtcars, aes(wt, mpg)) +
    geom_point() +
    facet_wrap(~ cyl)
basic
 
```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeB, echo=FALSE, out.width = "100%", fig.height=7}

# Change background
basic + theme(strip.background = element_rect(colour = "red", fill = alpha("blue",0.2) ))

```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeC, echo=FALSE, out.width = "100%", fig.height=7}

# Change the text 
basic + theme(strip.text.x = element_text(colour = "red", face = "bold", size=10, angle=30))
 

```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeD, echo=FALSE, out.width = "100%", fig.height=7}

# Change the space between parts:
basic + theme(panel.spacing = unit(4, "lines"))
```
</div>
</div>




```{r thecodeA, eval=FALSE}
```

```{r thecodeB, eval=FALSE}
```

```{r thecodeC, eval=FALSE}
```

```{r thecodeD, eval=FALSE}
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
