---
title: "Use leaflet control widget"
descriptionMeta: "This post explains how to build a bubble map with R and the leaflet package. It shows a set of earthquakes and their intensity in the Pacific Ocean. Reproducible code is provided."
descriptionTop: "This post explains how to build a [bubble map](bubble-map.html) with R and the `leaflet` package. It shows a set of earthquakes and their intensity in the Pacific Ocean."
sectionText: "Bubble map section"
sectionLink: "bubble-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/bubblemap.html"
url: "242-use-leaflet-control-widget"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 50px" >



`leaflet` is a R package allowing to build interactive maps. This post follows the [leaflet introduction](179-show-a-map-with-leaflet-r.html) and explain how to build a [bubble map](bubble-map.html).

The input required to build such a map is a list of GPS coordinates, with a numeric value attributed to each location to control the bubble size and color.

Markers are added using the `addCircleMarkers()` function, and bubble color is mapped to a color scale build with binning.

<br>

<center><iframe src="./HtmlWidget/bubblemapControl.html" height="400" width="100%" style="border:none;"></iframe></center>

<br>

```{r, eval=FALSE}
# Load libraries
library(leaflet)

# Make data with several positions
data_red <- data.frame(LONG=42+rnorm(10), LAT=23+rnorm(10), PLACE=paste("Red_place_",seq(1,10)))
data_blue <- data.frame(LONG=42+rnorm(10), LAT=23+rnorm(10), PLACE=paste("Blue_place_",seq(1,10)))

# Initialize the leaflet map:
m <- leaflet() %>% 
  setView(lng=42, lat=23, zoom=6 ) %>%

  # Add two tiles
  addProviderTiles("Esri.WorldImagery", group="background 1") %>%
  addTiles(options = providerTileOptions(noWrap = TRUE), group="background 2") %>%

  # Add 2 marker groups
  addCircleMarkers(data=data_red, lng=~LONG , lat=~LAT, radius=8 , color="black",
                   fillColor="red", stroke = TRUE, fillOpacity = 0.8, group="Red") %>%
  addCircleMarkers(data=data_blue, lng=~LONG , lat=~LAT, radius=8 , color="black",
                   fillColor="blue", stroke = TRUE, fillOpacity = 0.8, group="Blue") %>%

  # Add the control widget
  addLayersControl(overlayGroups = c("Red","Blue") , baseGroups = c("background 1","background 2"), 
                   options = layersControlOptions(collapsed = FALSE))

m

# save the widget in a html file if needed.
# library(htmlwidgets)
# saveWidget(m, file=paste0( getwd(), "/HtmlWidget/bubblemapControl.html"))
```











</div>






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
