---
title: "Network graph layouts with R and igraph"
descriptionMeta: "This post describes how to apply different layouts to a network diagram using the igraph R library. It gives reproducible code showing how to use the offered algorithm."
descriptionTop: "This post describes how to apply different layouts to a [network diagram](network.html) using the `igraph` R library. It gives reproducible code showing how to use the offered algorithms."
sectionText: "Network section"
sectionLink: "network.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/network.html"
url: "247-network-chart-layouts"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px;">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Network layouts are algorithms that return coordinates for each node in a [network](network.html). The `igraph` library offers several built-in layouts, and a sample of them is presented here. 

Usually, algorithms try to minimize edge crossing and prevent overlap. Edges can have uniform length or not.

Choose your layout using the `layout` argument. Type `help(layout)` to see all the possibilities.

In practice, the `fruchterman reingold` algorythm is often used.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode8, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(igraph)
 
# Create data
data <- matrix(sample(0:1, 400, replace=TRUE, prob=c(0.8,0.2)), nrow=20)
network <- graph_from_adjacency_matrix(data , mode='undirected', diag=F )
 
# When ploting, we can use different layouts:
par(mfrow=c(2,2), mar=c(1,1,1,1))
plot(network, layout=layout.sphere, main="sphere")
plot(network, layout=layout.circle, main="circle")
plot(network, layout=layout.random, main="random")
plot(network, layout=layout.fruchterman.reingold, main="fruchterman.reingold")
 
# See the complete list with
# help(layout)
```
</div>
</div>



```{r thecode8, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
