---
title: "Map variable to node and links feature in R igraph network"
descriptionMeta: "The igraph package is the best way to build network diagram with R. This post shows how to map a variable to node or link features, allowing to add more insight to the chart."
descriptionTop: "The `igraph` package is the best way to build [network diagram](network.html) with R. This post shows how to map a variable to node or link features, allowing to add more insight to the chart."
sectionText: "Network section"
sectionLink: "network.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/network.html"
url: "249-igraph-network-map-a-color"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px;">



# Map variable to node features
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Once you know how to make a [basic network graph](257-input-formats-for-network-charts.html) and how to customize its general features, you probably want to map the features according to another variable.

Here we consider a network with 10 people. Each is either adult, old or young and we want one specific color for each category.

The dataset is composed by 2 data frames.  `Links` provides the links between people.  `Nodes` gives features concerning people. What we need is to transform the  `carac` column into a vector of 3 colors, and provide this vector to the plot. The 3 colors are picked up in an `Rcolorbrewer` palette as described in [graph #39](39-pie-plot-with-rcolorbrewer-palette.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode8, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(igraph)
 
# create data:
links <- data.frame(
    source=c("A","A", "A", "A", "A","J", "B", "B", "C", "C", "D","I"),
    target=c("B","B", "C", "D", "J","A","E", "F", "G", "H", "I","I"),
    importance=(sample(1:4, 12, replace=T))
    )
nodes <- data.frame(
    name=LETTERS[1:10],
    carac=c( rep("young",3),rep("adult",2), rep("old",5))
    )
 
# Turn it into igraph object
network <- graph_from_data_frame(d=links, vertices=nodes, directed=F) 
 
# Make a palette of 3 colors
library(RColorBrewer)
coul  <- brewer.pal(3, "Set1") 
 
# Create a vector of color
my_color <- coul[as.numeric(as.factor(V(network)$carac))]
 
# Make the plot
plot(network, vertex.color=my_color)
 
# Add a legend
legend("bottomleft", legend=levels(as.factor(V(network)$carac))  , col = coul , bty = "n", pch=20 , pt.cex = 3, cex = 1.5, text.col=coul , horiz = FALSE, inset = c(0.1, 0.1))

```
</div>
</div>



```{r thecode8, eval=FALSE}
```















# Map variable to link features
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Following the same principle, it is possible to map other variables to other parameters.

Here is an example where we map the importance of the nodes to the edge width. (There is an  `importance` column in the  `links` data frame)

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode1, echo=FALSE, out.width = "100%", fig.height=7}
# Check
#E(network)$importance

# Plot
plot(network, vertex.color=my_color, edge.width=E(network)$importance*2 )
legend("bottomleft", legend=levels(as.factor(V(network)$carac))  , col = coul , bty = "n", pch=20 , pt.cex = 3, cex = 1.5, text.col=coul , horiz = FALSE, inset = c(0.1, 0.1))
```
</div>
</div>



```{r thecode1, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
