---
title: "Clustering result visualization with network diagram"
descriptionMeta: "This post explains how to compute a correlation matrix and display the result as a network chart using R and the igraph package."
descriptionTop: "This post explains how to compute a correlation matrix and display the result as a [network](network.html) chart using R and the `igraph` package."
sectionText: "Network section"
sectionLink: "network.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/network.html"
url: "250-correlation-network-with-igraph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Compute the correlation matrix
***

Consider a dataset composed by entities (usually in rows) and features (usually in columns). 

It is possible to compute a correlation matrix from it. It is a square matrix showing the relationship between each pair of entity. It can be computed using correlation (`cor()`) or euclidean distance (`dist()`).

Let's apply it to the `mtcars` dataset that is natively provided by R.

```{r thecode1, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(igraph)
 
# data
# head(mtcars)
 
# Make a correlation matrix:
mat <- cor(t(mtcars[,c(1,3:6)]))
```



```{r thecode1, eval=FALSE}
```












# Basic network diagram
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A correlation matrix can be visualized as a network diagram. Each entity of the dataset will be a node. And 2 nodes will be connected if their correlation or distance reach a threshold (`0.995` here).

To make a `graph` object from the correlation matrix, use the `graph_from_adjacency_matrix()` function of the `igraph` package. If you're not familiar with `igraph`, the [network](network.html) section is full of examples to get you started.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Keep only high correlations
mat[mat<0.995] <- 0
 
# Make an Igraph object from this matrix:
network <- graph_from_adjacency_matrix( mat, weighted=T, mode="undirected", diag=F)

# Basic chart
plot(network)
```
</div>
</div>



```{r thecode2, eval=FALSE}
```














# Customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The hardest part of the job has been done. The chart just requires a bit of polishing for a better output:

- customize node, link, label and background features as you like
- map the node feature to a variable (`cyl` here). It gives an additional layer of information, allowing to compare the network structure with a potential expected organization.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# color palette
library(RColorBrewer)
coul <- brewer.pal(nlevels(as.factor(mtcars$cyl)), "Set2")

# Map the color to cylinders
my_color <- coul[as.numeric(as.factor(mtcars$cyl))]

# plot
par(bg="grey13", mar=c(0,0,0,0))
set.seed(4)
plot(network, 
    vertex.size=12,
    vertex.color=my_color, 
    vertex.label.cex=0.7,
    vertex.label.color="white",
    vertex.frame.color="transparent"
    )

# title and legend
text(0,0,"mtcars network",col="white", cex=1.5)
legend(x=-0.2, y=-0.12, 
       legend=paste( levels(as.factor(mtcars$cyl)), " cylinders", sep=""), 
       col = coul , 
       bty = "n", pch=20 , pt.cex = 2, cex = 1,
       text.col="white" , horiz = F)
```
</div>
</div>



```{r thecode3, eval=FALSE}
```











# Customize link features
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Last but not least, control edges with arguments starting with `edge.`.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
plot(network,
    edge.color=rep(c("red","pink"),5),           # Edge color
    edge.width=seq(1,10),                        # Edge width, defaults to 1
    edge.arrow.size=1,                           # Arrow size, defaults to 1
    edge.arrow.width=1,                          # Arrow width, defaults to 1
    edge.lty=c("solid")                           # Line type, could be 0 or “blank”, 1 or “solid”, 2 or “dashed”, 3 or “dotted”, 4 or “dotdash”, 5 or “longdash”, 6 or “twodash”
    #edge.curved=c(rep(0,5), rep(1,5))            # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
    )
```
</div>
</div>



```{r thecode4, eval=FALSE}
```











# All customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Of course, you can use all the options described above all together on the same chart, for a high level of customization.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
par(bg="black")

plot(network,

    # === vertex
    vertex.color = rgb(0.8,0.4,0.3,0.8),          # Node color
    vertex.frame.color = "white",                 # Node border color
    vertex.shape="circle",                        # One of “none”, “circle”, “square”, “csquare”, “rectangle” “crectangle”, “vrectangle”, “pie”, “raster”, or “sphere”
    vertex.size=14,                               # Size of the node (default is 15)
    vertex.size2=NA,                              # The second size of the node (e.g. for a rectangle)

    # === vertex label
    vertex.label=LETTERS[1:10],                   # Character vector used to label the nodes
    vertex.label.color="white",
    vertex.label.family="Times",                  # Font family of the label (e.g.“Times”, “Helvetica”)
    vertex.label.font=2,                          # Font: 1 plain, 2 bold, 3, italic, 4 bold italic, 5 symbol
    vertex.label.cex=1,                           # Font size (multiplication factor, device-dependent)
    vertex.label.dist=0,                          # Distance between the label and the vertex
    vertex.label.degree=0 ,                       # The position of the label in relation to the vertex (use pi)

    # === Edge
    edge.color="white",                           # Edge color
    edge.width=4,                                 # Edge width, defaults to 1
    edge.arrow.size=1,                            # Arrow size, defaults to 1
    edge.arrow.width=1,                           # Arrow width, defaults to 1
    edge.lty="solid",                             # Line type, could be 0 or “blank”, 1 or “solid”, 2 or “dashed”, 3 or “dotted”, 4 or “dotdash”, 5 or “longdash”, 6 or “twodash”
    edge.curved=0.3    ,                          # Edge curvature, range 0-1 (FALSE sets it to 0, TRUE to 0.5)
    )
```
</div>
</div>



```{r thecode5, eval=FALSE}
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
