---
title: "Network with node size based on edge number"
descriptionMeta: "This post explains how to build a network diagram where node size is proportionnal to its number of connection with other nodes. It uses R and the igraph package."
descriptionTop: "This post explains how to build a [network diagram](network.html) where node size is proportionnal to its number of connection with other nodes. It uses R and the `igraph` package."
sectionText: "Network section"
sectionLink: "network.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/network.html"
url: "251-network-with-node-size-based-on-edges-number"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">





# Basic network diagram
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common task to make nodes bigger when they are heavily connected with other nodes. Indeed, it means they have an importance in the network and thus deserves to be highlighted.

The `degree()` function of the `igraph` package allows to compute the number of connection per node. It is possible to pass its result to the `vertex.size` argument of the `plot()` function to get the targeted result.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(igraph)
 
# create data:
links=data.frame(
    source=c("A","A", "A", "A", "A","J", "B", "B", "C", "C", "D","I"),
    target=c("B","B", "C", "D", "J","A","E", "F", "G", "H", "I","I")
    )
 
# Turn it into igraph object
network <- graph_from_data_frame(d=links, directed=F) 
 
# Count the number of degree for each node:
deg <- degree(network, mode="all")
 
# Plot
plot(network, vertex.size=deg*6, vertex.color=rgb(0.1,0.7,0.8,0.5) )
```

</div>
</div>



```{r thecode2, eval=FALSE}
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
