---
title: "Add text over boxplot in basic R"
descriptionMeta: "How to make a boxplot with category sample size written on top of each box in base R: code and explanation."
descriptionTop: "This examples demonstrates how to build a [boxplot](boxplot.html) with sample size written on top of each box. It is useful to indicate what sample size is hidden behind each box. Basic R implementation."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "26-add-text-over-a-boxplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The first challenge here is to recover the position of the top part of each box. This is done by saving the `boxplot()` result in an object (called `boundaries` here). Now, typing `boundaries$stats` gives a dataframe with all information concerning boxes.

Then, it is possible to use the `text` function to add labels on top of each box. This function takes 3 inputs:  

- x axis positions of the labels. In our case, it will be 1,2,3,4 for 4 boxes.
- y axis positions, available in the `boundaries$stats` object.
- text of the labels : the number of value per group or whatever else.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
names <- c(rep("A", 20) , rep("B", 8) , rep("C", 30), rep("D", 80))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 8 , replace=T), 
       sample(1:7, 30 , replace=T), sample(3:8, 80 , replace=T) )
data <- data.frame(names,value)
 
# Draw the boxplot. Note result is also stored in a object called boundaries
boundaries <- boxplot(data$value ~ data$names , col="#69b3a2" , ylim=c(1,11))
# Now you can type boundaries$stats to get the boundaries of the boxes

# Add sample size on top
nbGroup <- nlevels(data$names)
text( 
  x=c(1:nbGroup), 
  y=boundaries$stats[nrow(boundaries$stats),] + 0.5, 
  paste("n = ",table(data$names),sep="")  
)
```
</div>
</div>



```{r, eval=FALSE}
# Dummy data
names <- c(rep("A", 20) , rep("B", 8) , rep("C", 30), rep("D", 80))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 8 , replace=T), 
       sample(1:7, 30 , replace=T), sample(3:8, 80 , replace=T) )
data <- data.frame(names,value)
 
# Draw the boxplot. Note result is also stored in a object called boundaries
boundaries <- boxplot(data$value ~ data$names , col="#69b3a2" , ylim=c(1,11))
# Now you can type boundaries$stats to get the boundaries of the boxes

# Add sample size on top
nbGroup <- nlevels(data$names)
text( 
  x=c(1:nbGroup), 
  y=boundaries$stats[nrow(boundaries$stats),] + 0.5, 
  paste("n = ",table(data$names),sep="")  
)
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
