---
title: "Multiple ggplot2 charts on a single page"
descriptionMeta: "This post shows how to use the gridExtra library to combine several ggplot2 charts on the same figure. Several examples are provided, illustrating several ways to split the graphing window."
descriptionTop: "This post shows how to use the `gridExtra` library to combine several `ggplot2` charts on the same figure. Several examples are provided, illustrating several ways to split the graphing window."
sectionText: "ggplot2 Section"
sectionLink: "ggplot2-package.html"
DataToVizText: "About Scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "261-multiple-graphs-on-same-page"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 40px">




Mixing multiple graphs on the same page is a common practice. It allows to summarize a lot of information on the same figure, and is for instance widely used for scientific publication.

The gridExtra package makes it a breeze. It offers the `grid.arrange()` function that does exactly that. Its `nrow` argument allows to specify how to arrange the layout. 

For more complex layout, the `arrangeGrob()` functions allows to do some nesting. Here are 4 examples to illustrate how `gridExtra` works:

<div class = "row">

<div class = "col-md-6  col-sm-12">

```{r thecodeA, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# libraries
library(ggplot2)
library(gridExtra)
 
# Make 3 simple graphics:
g1 <- ggplot(mtcars, aes(x=qsec)) + geom_density(fill="slateblue")
g2 <- ggplot(mtcars, aes(x=drat, y=qsec, color=cyl)) + geom_point(size=5) + theme(legend.position="none")
g3 <- ggplot(mtcars, aes(x=factor(cyl), y=qsec, fill=cyl)) + geom_boxplot() + theme(legend.position="none")
g4 <- ggplot(mtcars , aes(x=factor(cyl), fill=factor(cyl))) +  geom_bar()
 
# Plots
grid.arrange(g2, arrangeGrob(g3, g4, ncol=2), nrow = 2)
```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeB, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}

grid.arrange(g1, g2, g3, nrow = 3)

```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeC, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}

grid.arrange(g2, arrangeGrob(g3, g4, ncol=2), nrow = 1)


```
</div>

<div class = "col-md-6  col-sm-12">

```{r thecodeD, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
grid.arrange(g2, arrangeGrob(g3, g4, nrow=2), nrow = 1)
```
</div>
</div>




```{r thecodeA, eval=FALSE}
```

```{r thecodeB, eval=FALSE}
```

```{r thecodeC, eval=FALSE}
```

```{r thecodeD, eval=FALSE}
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
