---
title: "Control ggplot2 boxplot colors"
descriptionMeta: "A focus on the different ways to color a ggplot2 boxplot. Also describes how to highlight a specific group."
descriptionTop: "A [boxplot](boxplot.html) summarizes the distribution of a continuous variable. Different color scales can be apply to it, and this post describes how to do so using the ggplot2 library. It is notably described how to highlight a specific group of interest."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "264-control-ggplot2-boxplot-colors"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" >


# General color customization
***


```{r, echo=FALSE, out.width=c('50%','50%','50%','50%'), fig.show='hold'}
# library
library(ggplot2)

# The mtcars dataset is natively available in R
#head(mpg)
 
# Top Left: Set a unique color with fill, colour, and alpha
ggplot(mpg, aes(x=class, y=hwy)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2) +
    xlab("") + 
    theme(plot.margin=unit(c(2,2,2,2),"cm"))
 
# Top Right: Set a different color for each group with default color palette
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot() +
    theme(legend.position="none") +
    xlab("") + 
    theme(plot.margin=unit(c(2,2,2,2),"cm"))

# Bottom Left: Use a palette from colorBrewer
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot() +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Set3") +
    xlab("") + 
    theme(plot.margin=unit(c(2,2,2,2),"cm"))
 
# Bottom Right
library(viridis)
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot(alpha=0.9) +
    theme(legend.position="none") +
    scale_fill_viridis(discrete=TRUE) +
    xlab("") + 
    theme(plot.margin=unit(c(2,2,2,2),"cm"))
```

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

These for examples illustrate the most common color scales used in [boxplot](boxplot.html). 

Note the use of [RcolorBrewer](package/rColorBrewer.html) and viridis to automatically generate nice color palette.
</div>


<div class = "col-md-6  col-sm-12">

```{r, eval=FALSE}
# library
library(ggplot2)
 
# The mtcars dataset is natively available in R
#head(mpg)
 
# Top Left: Set a unique color with fill, colour, and alpha
ggplot(mpg, aes(x=class, y=hwy)) + 
    geom_boxplot(color="red", fill="orange", alpha=0.2)
 
# Top Right: Set a different color for each group
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none")

# Bottom Left
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="BuPu")
 
# Bottom Right
ggplot(mpg, aes(x=class, y=hwy, fill=class)) + 
    geom_boxplot(alpha=0.3) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2")
```
</div>
</div>








# Highlighting a group {#highlight}
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Highlighting the main message conveid by your chart is an important step in dataviz. If your story focuses on a specific group, you should highlight it in your boxplot.

To do so, first create a new column with `mutate` where you store the binary information: highlight ot not. Then just provide this column to the `fill` argument of ggplot2 and eventually custom the appearance of the highlighted group with `scale_fill_manual` and `scale_alpha_manual`.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Work with the natively available mpg dataset
mpg %>% 
  
  # Add a column called 'type': do we want to highlight the group or not?
  mutate( type=ifelse(class=="subcompact","Highlighted","Normal")) %>%
  
  # Build the boxplot. In the 'fill' argument, give this column
  ggplot( aes(x=class, y=hwy, fill=type, alpha=type)) + 
    geom_boxplot() +
    scale_fill_manual(values=c("#69b3a2", "grey")) +
    scale_alpha_manual(values=c(1,0.1)) +
    theme_ipsum() +
    theme(legend.position = "none")  +
    xlab("")

```
</div>
</div>


```{r eval=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Work with the natively available mpg dataset
mpg %>% 
  
  # Add a column called 'type': do we want to highlight the group or not?
  mutate( type=ifelse(class=="subcompact","Highlighted","Normal")) %>%
  
  # Build the boxplot. In the 'fill' argument, give this column
  ggplot( aes(x=class, y=hwy, fill=type, alpha=type)) + 
    geom_boxplot() +
    scale_fill_manual(values=c("#69b3a2", "grey")) +
    scale_alpha_manual(values=c(1,0.1)) +
    theme_ipsum() +
    theme(legend.position = "none") +
    xlab("")
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
