---
title: "Grouped boxplot with ggplot2"
descriptionMeta: "How to build a grouped boxplot with the ggplot2 R package: code and explanation"
descriptionTop: "A [boxplot](boxplot.html) summarizes the distribution of a continuous variable for several categories. If categories are organized in groups and subgroups, it is possible to build a grouped boxplot. Here is an example with R and ggplot2."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "265-grouped-boxplot-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" >


# Grouped boxplot
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A grouped boxplot is a [boxplot](boxplot.html) where categories are organized in groups and subgroups. 

Here we visualize the distribution of 7 groups (called A to G) and 2 subgroups (called low and high). Note that the group must be called in the `X` argument of `ggplot2`. The subgroup is called in the `fill` argument.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# create a data frame
variety=rep(LETTERS[1:7], each=40)
treatment=rep(c("high","low"),each=20)
note=seq(1:280)+sample(1:150, 280, replace=T)
data=data.frame(variety, treatment ,  note)
 
# grouped boxplot
ggplot(data, aes(x=variety, y=note, fill=treatment)) + 
    geom_boxplot()

```
</div>
</div>


```{r eval=FALSE}
# library
library(ggplot2)
 
# create a data frame
variety=rep(LETTERS[1:7], each=40)
treatment=rep(c("high","low"),each=20)
note=seq(1:280)+sample(1:150, 280, replace=T)
data=data.frame(variety, treatment ,  note)
 
# grouped boxplot
ggplot(data, aes(x=variety, y=note, fill=treatment)) + 
    geom_boxplot()
```



# Using small multiple
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">
Note that an alternative to grouped boxplot is to use [faceting](223-faceting-with-ggplot2.html): each subgroup (left) or each group (right) is represented in a distinct panel.
</div>


<div class = "col-md-6  col-sm-12">

```{r}
# One box per treatment
p1 <- ggplot(data, aes(x=variety, y=note, fill=treatment)) + 
    geom_boxplot() +
    facet_wrap(~treatment)
# one box per variety
p2 <- ggplot(data, aes(x=variety, y=note, fill=treatment)) + 
    geom_boxplot() +
    facet_wrap(~variety, scale="free")

```
</div>
</div>
<br><br>

```{r, echo=FALSE, out.width=c('50%', '50%'),  fig.show='hold'}
p1
p2
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
