---
title: "Ggplot2 boxplot with variable width"
descriptionMeta: "Boxplots hide the category sample sizes. One way to tackle this issue is to build boxplot with width proportionnal to sample size. Here is how to do it with R and ggplot2."
descriptionTop: "[Boxplots](boxplot.html) hide the category sample sizes. One way to tackle this issue is to build boxplot with width proportionnal to sample size. Here is how to do it with R and [ggplot2](ggplot2.html)."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "266-ggplot2-boxplot-with-variable-width.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```


<div class="container" style="padding-top: 50px">



# Grouped boxplot
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Boxplot](boxplot.html) are often [critized](https://www.data-to-viz.com/caveat/boxplot.html) for hiding the underlying distribution of each category. Since individual data points are hidden, it is also impossible to know what sample size is available for each category.

In this example, box widths are proportional to sample size thanks to the `varwidth` option. On top of that, the exact sample size is added to the X axis labels for more accuracy.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# create data
names <- c(rep("A", 20) , rep("B", 5) , rep("C", 30), rep("D", 100))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 5 , replace=T), sample(1:7, 30 , replace=T), sample(3:8, 100 , replace=T) )
data <- data.frame(names,value)
 
# prepare a special xlab with the number of obs for each group
my_xlab <- paste(levels(data$names),"\n(N=",table(data$names),")",sep="")
 
# plot
ggplot(data, aes(x=names, y=value, fill=names)) +
    geom_boxplot(varwidth = TRUE, alpha=0.2) +
    theme(legend.position="none") +
    scale_x_discrete(labels=my_xlab)
```
</div>
</div>


```{r eval=FALSE}
# library
library(ggplot2)
 
# create data
names <- c(rep("A", 20) , rep("B", 5) , rep("C", 30), rep("D", 100))
value <- c( sample(2:5, 20 , replace=T) , sample(4:10, 5 , replace=T), sample(1:7, 30 , replace=T), sample(3:8, 100 , replace=T) )
data <- data.frame(names,value)
 
# prepare a special xlab with the number of obs for each group
my_xlab <- paste(levels(data$names),"\n(N=",table(data$names),")",sep="")
 
# plot
ggplot(data, aes(x=names, y=value, fill=names)) +
    geom_boxplot(varwidth = TRUE, alpha=0.2) +
    theme(legend.position="none") +
    scale_x_discrete(labels=my_xlab)
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
