---
title: "ggplot2 boxplot from continuous variable"
descriptionMeta: "This post explains how to build a boxplot with ggplot2 where categories are actually bins of a numeric variable."
descriptionTop: "A [boxplot](boxplot.html) summarizes the distribution of a continuous variable. This post explains how to build a boxplot with [ggplot2](ggplot2) where categories are actually bins of a numeric variable. It is sometimes useful to study the relationship between 2 nnumeric variables."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "268-ggplot2-boxplot-from-continuous-variable"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Let's say we want to study the relationship between 2 numeric variables. It is possible to cut on of them in different bins, and to use the created groups to build a [boxplot](boxplot.html).

Here, the numeric variable called `carat` from the `diamonds` dataset in cut in 0.5 length bins thanks to the `cut_width` function. Then, we just need to provide the newly created variable to the X axis of `ggplot2`.
```{r, echo=T}
# library
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Start with the diamonds dataset, natively available in R:
p <- diamonds %>%
  
  # Add a new column called 'bin': cut the initial 'carat' in bins
  mutate( bin=cut_width(carat, width=0.5, boundary=0) ) %>%
  
  # plot
  ggplot( aes(x=bin, y=price) ) +
    geom_boxplot(fill="#69b3a2") +
    theme_ipsum() +
    xlab("Carat")
```

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
p
```
</div>
</div>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
