---
title: "Heatmap with lattice and levelplot()"
descriptionMeta: "This document explains how to use the levelplot() function of the lattice R package to build heatmaps."
descriptionTop: "This document explains how to use the `levelplot()` function of the `lattice` R package to build [heatmaps](heatmap.html)."
sectionText: "Heatmap section"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/heatmap.html"
url: "27-levelplot-with-lattice"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basis use of levelplot()
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `plot()` function of R allows to build a [scatterplot](scatterplot.html). Both numeric variables of the input dataframe must be specified in the `x` and `y` argument.


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
data = data.frame(
  x=seq(1:100) + 0.1*seq(1:100)*sample(c(1:10) , 100 , replace=T),
  y=seq(1:100) + 0.2*seq(1:100)*sample(c(1:10) , 100 , replace=T)
)

# Basic scatterplot
plot(x=data$x, y=data$y)
```
</div>
</div>



```{r thecode, eval=FALSE}
```




# Customizations
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here is a description of the most common customization:  

- `cex`: circle size
- `xlim` and `ylim`: limits of the X and Y axis
- `pch`: shape of markers. See all [here](6-graph-parameters-reminder.html).
- `xlab` and `ylab`: X and Y axis labels
- `col`: marker color
- `main`: chart title


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
data = data.frame(
  x=seq(1:100) + 0.1*seq(1:100)*sample(c(1:10) , 100 , replace=T),
  y=seq(1:100) + 0.2*seq(1:100)*sample(c(1:10) , 100 , replace=T)
)

# Basic scatterplot
plot(data$x, data$y,
     xlim=c(0,250) , ylim=c(0,250), 
     pch=18, 
     cex=2, 
     col="#69b3a2",
     xlab="value of X", ylab="value of Y",
     main="A simple scatterplot"
     )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
