---
title: "Heatmap with lattice and levelplot()"
descriptionMeta: "This document explains how to use the levelplot() function of the lattice R package to build heatmaps."
descriptionTop: "This document explains how to use the `levelplot()` function of the `lattice` R package to build [heatmaps](heatmap.html)."
sectionText: "Heatmap section"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/heatmap.html"
url: "27-levelplot-with-lattice"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basis use of levelplot()
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `lattice` package allows to build [heatmaps](heatmap.html) thanks to the `levelplot()` function. 

<u>Input data</u>: here input is a data frame with 3 columns prividing the X and Y coordinate of the cell and its value. (<u>Long</u> format).
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load the lattice package
library("lattice")
 
# Dummy data
x <- seq(1,10, length.out=20)
y <- seq(1,10, length.out=20)
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)

## Try it out
levelplot(Z ~ X*Y, data=data  ,xlab="X",
          main="")
```
</div>
</div>



```{r thecode, eval=FALSE}
```






# From wide input matrix
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Previous example of this document was based on a data frame at the long format. Here, a square matrix is used instead. It is the second format understood by the `levelplot()` function.

<u>Note</u>: here row and column order isn't respected in the heatmap.

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Load the library
library("lattice")
 
# Dummy data
data <- matrix(runif(100, 0, 5) , 10 , 10)
colnames(data) <- letters[c(1:10)]
rownames(data) <- paste( rep("row",10) , c(1:10) , sep=" ")
 
# plot it flipping the axis
levelplot(data)
```
</div>
</div>



```{r thecode3, eval=FALSE}
```








# Flip and reorder axis
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `t()` function of R allows to transpose the input matrix, and thus to flip X and Y coordinates.

Moreover, you can reverse matrix order as shown below to reverse order in the heatmap as well. Now the heatmap is organized exactly as the input matrix.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Load the library
library("lattice")
 
# Dummy data
data <- matrix(runif(100, 0, 5) , 10 , 10)
colnames(data) <- letters[c(1:10)]
rownames(data) <- paste( rep("row",10) , c(1:10) , sep=" ")
 
# plot it flipping the axis
levelplot( t(data[c(nrow(data):1) , ]),
           col.regions=heat.colors(100))
```
</div>
</div>



```{r thecode2, eval=FALSE}
```












# Custom colors
***


```{r thecode4, echo=FALSE, out.width = c("30%","30%","30%"), fig.height=9, fig.show="hold"}
# Lattice package
require(lattice)

# The volcano dataset is provided, it looks like that:
#head(volcano)

# 1: native palette from R
levelplot(volcano, col.regions = terrain.colors(100)) # try cm.colors() or terrain.colors()

# 2: Rcolorbrewer palette
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "PiYG"))(25)
levelplot(volcano, col.regions = coul) # try cm.colors() or terrain.colors()

# 3: Viridis
library(viridisLite)
coul <- viridis(100)
levelplot(volcano, col.regions = coul) 
#levelplot(volcano, col.regions = magma(100)) 
```



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

There are several ways to custom the color palette:  

- native palettes of R: `terrain.color()`, `rainbow()`, `heat.colors()`, `topo.colors()` or `cm.colors()`

- palettes of `RColorBrewer`. See list of available palettes [here](package/rColorBrewer.html).
- palettes of `Viridis`: viridis, magma, inferno, plasma.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, fig.width=5, fig.height=3, fig.align="center"}
levelplot(volcano, col.regions = magma(100)) 
```


</div>
</div>

```{r thecode4, eval=FALSE}
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
