---
title: "Basic scatterplot with R and ggplot2"
descriptionMeta: "This post provides reproducible code and explanation for the most basic scatterplot you can build with R and ggplot2."
descriptionTop: "This post provides reproducible code and explanation for the most basic [scatterplot](scatterplot.html) you can build with `R` and [ggplot2](ggplot2-package.html)."
sectionText: "Scatter Section"
sectionLink: "scatterplot.html"
DataToVizText: "About Scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "272-basic-scatterplot-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [scatterplot](scatterplot.html) displays the values of two variables along two axes. It shows the relationship between them, eventually revealing a correlation.

Here the relationship between Sepal width and Sepal length of several plants is shown.

It illustrates the basic utilization of `ggplot2` for scatterplots:  

- 1 - provide a dataframe
- 2 - tell which variable to show on x and y axis
- 3 - add a `geom_point()` to show points.


</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# The iris dataset is provided natively by R
#head(iris)
 
# basic scatterplot
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point()
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggplot2)
 
# The iris dataset is provided natively by R
#head(iris)
 
# basic scatterplot
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point()
```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
