---
title: "Custom ggplot2 scatterplot"
descriptionMeta: "This post is dedicated to customization you can apply to a scatterplot built with ggplot2."
descriptionTop: "This post is dedicated to customization you can apply to a [scatterplot](scatterplot.html) built with [ggplot2](ggplot2.html)"
sectionText: "Scatter Section"
sectionLink: "scatterplot.html"
DataToVizText: "About Scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "273-custom-your-scatterplot-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This post follows the previous [basic scatterplot](272-basic-scatterplot-with-ggplot2.html) with `ggplot2`. It shows the kind of customization you can apply to circles thanks to the `geom_point()` options:  

- `color`: the stroke color, the circle outline
- `stroke`: the stroke width
- `fill`: color of the circle inner part
- `shape`: shape of the marker. See list in the [ggplot2 section](ggplot2.html)
- `alpha`: circle transparency, [0->1], 0 is fully transparent
- `color`: the stroke color, the circle outline
- `size`: circle size

<u>Note:</u> these options will be uniform among markers if you put it in the `geom_point()` call. You can also map them to a variable if put inside the `aes()` part of the code.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
        )
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggplot2)
 
# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
        )
```








# Using `theme_ipsum`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Note that applying the `theme_ipsum` of the `hrbrthemes` package is always a good option.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
library(hrbrthemes)

# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="black",
        fill="#69b3a2",
        shape=22,
        alpha=0.5,
        size=6,
        stroke = 1
        ) +
    theme_ipsum()
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggplot2)
library(hrbrthemes)

# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="black",
        fill="#69b3a2",
        shape=22,
        alpha=0.5,
        size=6,
        stroke = 1
        ) +
    theme_ipsum()
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
