---
title: "Map a variable to marker feature in ggplot2 scatterplot"
descriptionMeta: "Ggplot2 allows to easily map a variable to marker features. This post explaines how it works through several examples, with explanation and code."
descriptionTop: "`ggplot2` allows to easily map a variable to marker features of a [scatterplot](scatterplot.html). This post explaines how it works through several examples, with explanation and code."
sectionText: "Scatter Section"
sectionLink: "scatterplot.html"
DataToVizText: "About Scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "274-map-a-variable-to-ggplot2-scatterplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Basic example
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here is the magick of [ggplot2](ggplot2): the ability to map a variable to marker features. Here, the marker `color` depends on its value in the field called `Species` in the input data frame.

Note that the legend is built automatically.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# load ggplot2
library(ggplot2)
library(hrbrthemes)

# mtcars dataset is natively available in R
# head(mtcars)
 
# A basic scatterplot with color depending on Species
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + 
    geom_point(size=6) +
    theme_ipsum()
```
</div>
</div>









# Works with any aesthetics
***

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

You can map variables to any marker features. For instance, specie is represente below using transparency (left), shape (middle) and size (right).

</div>


<div class = "col-md-8  col-sm-12">

```{r eval=FALSE}
# Transparency
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, alpha=Species)) + 
    geom_point(size=6, color="#69b3a2") +
    ggtitle("Transparency") +
    theme_ipsum()

# Shape
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, shape=Species)) + 
    geom_point(size=6, color="lightblue") +
    ggtitle("Shape") +
    theme_ipsum()

# Size
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, size=Petal.Width)) + 
    geom_point(color="darkred") +
    ggtitle("Size") +
    theme_ipsum()
```
</div>
</div>

```{r echo=FALSE, out.width = c('33%', '33%', '33%'), fig.height=7, fig.show='hold'}
# Transparency
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, alpha=Species)) + 
    geom_point(size=6, color="#69b3a2") +
    ggtitle("Transparency") +
    theme_ipsum()

# Shape
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, shape=Species)) + 
    geom_point(size=6, color="lightblue") +
    ggtitle("Shape") +
    theme_ipsum()

# Size
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, size=Petal.Width)) + 
    geom_point(color="darkred") +
    ggtitle("Size") +
    theme_ipsum()
```



# Mapping to several features
***


Last but not least, note that you can map one or several variables to one or several features. Here, shape, transparency, size and color all depends on the marker `Species` value.


```{r thecode3, out.width = "100%", fig.height=7}
# load ggplot2
library(ggplot2)
library(hrbrthemes)

# A basic scatterplot with color depending on Species
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, shape=Species, alpha=Species, size=Species, color=Species)) + 
    geom_point() +
    theme_ipsum()
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
