---
title: "Add text labels with ggplot2"
descriptionMeta: "This document is dedicated to text annotation with ggplot2. It provides several examples with reproducible code showing how to use function like geom_label, geom_text."
descriptionTop: "This document is dedicated to text annotation with [ggplot2](ggplot2-package.html). It provides several examples with reproducible code showing how to use function like `geom_label` and `geom_text`."
sectionText: "More annotation with ggplot2"
sectionLink: "233-add-annotations-on-ggplot2-chart.html"
DataToVizText: "Annotation, why?"
DataToVizLink: "data-to-viz.com/caveat/annotation.html"
url: "275-add-text-labels-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Adding text with `geom_text()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This example demonstrates how to use `geom_text()` to add text as markers. It works pretty much the same as `geom_point()`, but add text instead of circles. A few arguments must be provided:  

- `label`: what text you want to display
- `nudge_x` and `nudge_y`: shifts the text along X and Y axis
- `check_overlap` tries to avoid text overlap. Note that a package called `ggrepel` extends this concept further

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# 1/ add text with geom_text, use nudge to nudge the text
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_text(
    label=rownames(data), 
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )
```
</div>
</div>



```{r thecode, eval=FALSE}
```







# Add labels with `geom_label()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`geom_label()` works pretty much the same way as `geom_text()`. However, text is wrapped in a rectangle that you can customize (see next example).
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# 1/ add text with geom_text, use nudge to nudge the text
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_label(
    label=rownames(data), 
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```





# Add one text label only
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Of course, you don't have to label all dots on the chart. You can also add a piece of text on a specific position. Since we're here, note that you can custom the annotation of `geom_label` with `label.padding`, `label.size`, `color` and `fill` as described below:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
 
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# Add one annotation
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_label(
    label="Look at this!", 
    x=4.1,
    y=20,
    label.padding = unit(0.55, "lines"), # Rectangle size around label
    label.size = 0.35,
    color = "black",
    fill="#69b3a2"
  )
```
</div>
</div>



```{r thecode3, eval=FALSE}
```






# Add labels for a selection of marker
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Last but not least, you can also select a group of marker and annotate them only. Here, only car with `mpg` > 20 and `wt` > 3 are annotated thanks to a data filtering in the `geom_label()` call.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
library(dplyr)
library(tibble)

# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)

# Change data rownames as a real column called 'carName'
data <- data %>%
  rownames_to_column(var="carName")
  
# Plot
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + 
  geom_label( 
    data=data %>% filter(mpg>20 & wt>3), # Filter data first
    aes(label=carName)
  )
```
</div>
</div>



```{r thecode4, eval=FALSE}
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
