---
title: "Marginal distribution with ggplot2 and ggExtra"
descriptionMeta: "This post explains how to add marginal distributions to the X and Y axis of a ggplot2 scatterplot. It can be done using histogram, boxplot or density plot using the ggExtra library."
descriptionTop: "This post explains how to add marginal distributions to the X and Y axis of a `ggplot2` [scatterplot](scatterplot.html). It can be done using [histogram](histogram.html), [boxplot](boxplot.html) or [density plot](density.html) using the `ggExtra` library."
sectionText: "Scatter section"
sectionLink: "scatterplot.html"
DataToVizText: "About scatter"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "277-marginal-histogram-for-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Basic use of `ggMarginal()`
***
```{r thecode, echo=FALSE, out.width = c('33%', '33%', '33%'), fig.show="hold", fig.height=7}
# library
library(ggplot2)
library(ggExtra)
 
# The mtcars dataset is proposed in R
head(mtcars)
 
# classic plot :
p <- ggplot(mtcars, aes(x=wt, y=mpg, color=cyl, size=cyl)) +
      geom_point() +
      theme(legend.position="none")
 
# with marginal histogram
p1 <- ggMarginal(p, type="histogram")
 
# marginal density
p2 <- ggMarginal(p, type="density")
 
# marginal boxplot
p3 <- ggMarginal(p, type="boxplot")
```

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p1
```
</div>

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p2
```
</div>

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p3
```
</div>
</div>

Here are 3 examples of marginal distribution added on X and Y axis of a scatterplot. The `ggExtra` library makes it a breeze thanks to the `ggMarginal()` function. Three main types of distribution are available: [histogram](histogram.html), [density](density.html) and [boxplot](boxplot.html).

```{r thecode, eval=FALSE}
```











# More customization
***
```{r thecode2, echo=FALSE, out.width = c('33%', '33%', '33%'), fig.show="hold", fig.height=7}
# library
library(ggplot2)
library(ggExtra)
 
# The mtcars dataset is proposed in R
head(mtcars)
 
# classic plot :
p <- ggplot(mtcars, aes(x=wt, y=mpg, color=cyl, size=cyl)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=10)
 
# Custom marginal plots:
p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)
```

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p1
```
</div>

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p2
```
</div>

<div class = "col-md-4  col-sm-12 align-self-center">
```{r, out.width='100%', echo=FALSE, fig.height=7}
p3
```
</div>
</div>

Three additional examples to show possible customization:  

- change marginal plot size with `size`
- custom marginal plot appearance with all usual parameters
- show only one marginal plot with `margins = 'x'` or `margins = 'y'`
```{r thecode2, eval=FALSE}
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
