---
title: "The `Maps`, `Mapdata` and `Oz` packages to get most common boundaries"
descriptionMeta: "The maps, mapdata and oz packages provide the boundaries of the most common world regions like the US, Europe, New Zealand and more. This post describes how to use them to avoid struggling finding input data."
descriptionTop: "The [maps](#maps), [mapdata](#mapdata) and [oz](#oz) packages provide the boundaries of the most common world regions like the US, Europe, New Zealand and more. This post describes how to use them to avoid struggling finding input data."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "278-the-maps-library"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# `maps` package {#maps}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The [maps](https://cran.r-project.org/web/packages/maps/maps.pdf) package is the best source of geospatial data in R. 

The whole list of offered data is available typing: `help(package='maps')`.

It includes: 

- World: `world`, `world.cities`, `lakes`
- US: states, `county`, `state`, `usa` 
- France: `france`
- Italy: `italy`
- New zealand: `nz` 

The code below shows how to use one of this geospatial object.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# Load library
library(maps)

# Check all available geospatial objects:
# help(package='maps')

# Map of the world:
map('world',col="grey", fill=TRUE, bg="white", lwd=0.05, mar=rep(0,4),border=0, ylim=c(-80,80) )
```
</div>
</div>



```{r thecode, eval=FALSE}
```











# `mapdata` package {#mapdata}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The [mapdata](https://cran.r-project.org/web/packages/mapdata/mapdata.pdf) package extends the `maps` package with more geospatial datasets:

- `china`
- `japan`
- Other world versions like pacific Centric (`world2Hires`)

See complete list with `help(package='mapdata')`

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=4}
# Load library
library(mapdata)

# Check all available geospatial objects:
# help(package='mapdata')

# Map of Japan:
map('japan',col="black", lwd=1, mar=rep(0,4) )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```









# `oz` package {#oz}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The [oz](https://github.com/cran/oz) package provides several geospatial object concerning Australia. 

Plot the general country using the `oz()` function, or states using `nsw()`, `qld()` and so on..

Type `help(package='oz')` for documentation.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=4}
# Load library
library(oz)

# Check all available geospatial objects:
# help(package='oz')

# Map of Australia
par(mar=rep(0,4))
oz( states=TRUE, col="#69b3a2")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```






<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
