---
title: "Title manipulation with R and ggplot2"
descriptionMeta: "This post describes all the available options to customize the chart title with R and ggplot2. It shows how to control its color, its position, and more."
descriptionTop: "This post describes all the available options to customize the chart title with `R` and `ggplot2`. It shows how to control its color, its position, and more."
sectionText: "ggplot2 section"
sectionLink: "ggplot2-package.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "289-control-ggplot2-title"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Default `ggplot2` title
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to add a title to a ggplot2 chart using the `ggtitle()` function. It works as follow:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library & data
library(ggplot2)
data <- data.frame(x=rnorm(100))
 
# Basic plot with title
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue Histogram") +
  theme_minimal()
```
</div>

</div>



```{r thecode, eval=FALSE}
```
















# Title on several lines
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common need to set the title on several lines. To add a break and skip to a second line, just add a `\n` in the text.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# title on several lines
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue \nHistogram") +
  theme_minimal() 
```
</div>

</div>



```{r thecode2, eval=FALSE}
```











# Title appearance and position
***

<div class = "row">

<div class = "col-md-6  col-sm-12 ">

```{r thecode3a, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Custom title appearance
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.5) +
  ggtitle("A blue Histogram") +
  theme_minimal() +
  theme(
    plot.title=element_text(family='', face='bold', colour='purple', size=26)
  )
 
```

</div>

<div class = "col-md-6  col-sm-12">

```{r thecode3b, echo=FALSE, out.width = "80%", fig.height=7,  fig.align="center"}
# Adjust the position of title
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.8) +
  ggtitle("With title on the right") +
  theme_minimal() +
  theme(
    plot.title=element_text( hjust=1, vjust=0.5, face='bold')
  )
```

</div>

</div>

Here are 2 tricks to control text appearance and its position. Both features are controled thanks to the `plot.title` argument of the `theme()` function. Appearance can be controlled with option such as `family`, `size` or `color`, when position is controlled with `hjust` and `vjust`.


```{r thecode3a, eval=FALSE}
```
```{r thecode3b, eval=FALSE}
```













# Customize a specific word only
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If you want to highlight a specific set of words in the title, it is doable thanks to the `expression()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# Custom a few word of the title only:
my_title <- expression(paste("A ", italic(bold("blue")), " histogram"))
ggplot( data=data, aes(x=x)) + 
  geom_histogram(fill="skyblue", alpha=0.8) +
  ggtitle(my_title) +
  theme_minimal()
```
</div>

</div>



```{r thecode5, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
