---
title: "Most basic dendrogram for clustering with R"
descriptionMeta: "Clustering allows to group samples by similarity and can its result can be visualized as a dendrogram. This post describes a basic usage of the hclust function and builds a dendrogram from its output."
descriptionTop: "Clustering allows to group samples by similarity and can its result can be visualized as a [dendrogram](dendrogram.html). This post describes a basic usage of the `hclust()` function and builds a dendrogram from its output."
sectionText: "Dendrogram section"
sectionLink: "dendrogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/dendrogram.html"
url: "29-basic-dendrogram"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic dendrogram with R
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

&rarr; Input dataset is a `matrix` where each row is a sample, and each column is a variable. Keep in mind you can transpose a matrix using the `t()` function if needed.

&rarr; Clustering is performed on a square matrix (sample x sample) that provides the distance between samples. It can be computed using the `dist()` or the `cor()` function depending on the question your asking

&rarr; The `hclust()` function is used to perform the hierarchical clustering

&rarr; Its output can be visualized directly with the `plot()` function. See possible [customization](31-custom-colors-in-dendrogram.html).


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode8, echo=FALSE, out.width = "100%", fig.height=7}
# Dataset 
data <- matrix( sample(seq(1,2000),200), ncol = 10 )
rownames(data) <- paste0("sample_" , seq(1,20))
colnames(data) <- paste0("variable",seq(1,10))

# Euclidean distance
dist <- dist(data[ , c(4:8)] , diag=TRUE)

# Hierarchical Clustering with hclust
hc <- hclust(dist)

# Plot the result
plot(hc)
```
</div>
</div>



```{r thecode8, eval=FALSE}
```

<br><br>




# Hierarchical clustering principle:
***

- Take distances between objects. 
- Seek the smallest distance between 2 objects. 
- Aggregate the 2 objects in a cluster. 
- Replace them with their barycenter. 
&rarr; Again until having only one cluster containing every points. 

There are several ways to calculate the distance between 2 clusters ( using the max between 2 points of the clusters, or the mean, or the min, or ward (default) ).







# Zoom on a group
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to zoom on a specific part of the tree. Select the group of interest using the `[[..]]` operator:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=5}
# store the dedrogram in an object
dhc <- as.dendrogram(hc)

# set the margin
par(mar=c(4,4,2,2))

# Plot the Second group
plot(dhc[[2]] , main= "zoom on a part of the dendrogram")
```

</div>
</div>



```{r thecode, eval=FALSE}
```

<br><br>







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
