---
title: "Basic ridgeline plot"
descriptionMeta: "The ridgeline plot allows to study the distribution of a numeric variable for several groups. Learn how to build one with R and the ggridges library."
descriptionTop: "The [ridgeline plot](ridgeline-plot.html) allows to study the distribution of a numeric variable for several groups. This document explains how to build it with `R` and the `ggridges` library."
sectionText: "Ridgeline Section"
sectionLink: "ridgeline-plot.html"
DataToVizText: "What is it?"
DataToVizLink: "data-to-viz.com/graph/ridgeline.html"
url: "294-basic-ridgeline-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [Ridgelineplot](https://www.data-to-viz.com/graph/ridgeline.html) (formerly called Joyplot) allows to study the distribution of a numeric variable for several groups. In this example, we check the distribution of diamond prices according to their quality.
 
This graph is made using the `ggridges` library, which is a `ggplot2` extension and thus respect the syntax of the grammar of graphic. We specify the `price` column for the X axis and the `cut` column for the Y axis. Adding `fill=cut` allows to use one colour per category and display them as separate groups.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggridges)
library(ggplot2)
 
# Diamonds dataset is provided by R natively
#head(diamonds)
 
# basic example
ggplot(diamonds, aes(x = price, y = cut, fill = cut)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggridges)
library(ggplot2)
 
# Diamonds dataset is provided by R natively
#head(diamonds)
 
# basic example
ggplot(diamonds, aes(x = price, y = cut, fill = cut)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```


# Shape Variation {#shape}
***
<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to represent the density with different aspects. For instance, using `stat="binline"` makes a histogram like shape to represent each distribution.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggridges)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0)) %>%
  filter(text %in% c("Almost Certainly","Very Good Chance","We Believe","Likely","About Even", "Little Chance", "Chances Are Slight", "Almost No Chance"))

# Plot
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, stat="binline", binwidth=5) +
    scale_x_continuous(breaks=seq(0,100,25)) +
    theme_ridges() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggridges)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0)) %>%
  filter(text %in% c("Almost Certainly","Very Good Chance","We Believe","Likely","About Even", "Little Chance", "Chances Are Slight", "Almost No Chance"))

# Plot
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
    theme_ridges() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```



# Color relative to numeric value {#color}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to set color depending on the numeric variable instead of the categoric one. (code from the [ridgeline R package](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html) by [Claus O. Wilke](https://github.com/clauswilke) )

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```
</div>
</div>



```{r, eval=FALSE}
# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
