---
title: "Most basic circular barplot"
descriptionMeta: "A circular barplot is a barplot where bars are displayed along a circle instead of a line. Learn how to build a basic circular barplot with R and ggplot2."
descriptionTop: "A circular barplot is a [barplot](barplot.html) where bars are displayed along a circle instead of a line. This post explains how to build a basic version with `R` and [ggplot2](ggplot2.html). It provides the reproducible code and explain how the `coord_polar()` function works."
sectionText: "Circular bar section"
sectionLink: "circular-barplot.html"
DataToVizText: "Warning"
DataToVizLink: "data-to-viz.com/graph/circularbarplot.html"
url: "295-basic-circular-barplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [circular barplot](circular-barplot.html) is a [barplot](barplot.html) where bars are displayed along a circle instead of a line. 
 
The input dataset is the same than for a barplot: we need one numeric value per group (one group = one bar). (See more explanation in the [barplot](barplot.html) section).
 
Basically, the method is the same than to do a classic barplot. At the end, we call `coord_polar()` to make the chart circular. Note that the `ylim()` argument is really important. If it starts at 0, the bars will start from the centre of the circle. If you provide a negative value, a white circle space will appear!
 
This chart is not really insightful, go to the [next example](296-add-labels-to-circular-barplot.html) to learn how to add labels!

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(tidyverse)
 
# Create dataset
data <- data.frame(
  id=seq(1,60),
  individual=paste( "Mister ", seq(1,60), sep=""),
  value=sample( seq(10,100), 60, replace=T)
)
 
# Make the plot
p <- ggplot(data, aes(x=as.factor(id), y=value)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  # This add the bars with a blue color
  geom_bar(stat="identity", fill=alpha("blue", 0.3)) +
  
  # Limits of the plot = very important. The negative value controls the size of the inner circle, the positive one is useful to add size over each bar
  ylim(-100,120) +
  
  # Custom the theme: no axis title and no cartesian grid
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-2,4), "cm")     # This remove unnecessary margin around plot
  ) +
  
  # This makes the coordinate polar instead of cartesian.
  coord_polar(start = 0)
p
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# What's next?
***

This chart is not really insightful, go to the [next example](296-add-labels-to-circular-barplot.html) to learn how to add labels!

<a class="btn btn-secondary" href="296-add-labels-to-circular-barplot">Add labels</a>
<a class="btn btn-secondary" href="circular-barplot.html">See all examples</a>




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
