---
title: "Add labels to circular barplot"
descriptionMeta: "This post explains how to add labels on a circular barchart, on top of each bar. It uses R and ggplot2, and reproducible code is provided."
descriptionTop: "This post explains how to add labels on a [ggplot2](ggplot2-package.html) circular barchart, on top of each bar. It follows the previous most basic circular barchart."
sectionText: "Circular bar section"
sectionLink: "circular-barplot.html"
DataToVizText: "Warning"
DataToVizLink: "data-to-viz.com/graph/circularbarplot.html"
url: "296-add-labels-to-circular-barplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The [chart #295](295-basic-circular-barplot.html) explains how to make a basic [circular barplot](circular-barplot.html). The next step is to add labels to each bar, to give insight to the graphic.
 
Here I suggest a method to add label at the top of each bar, using the same angle that the central part of the bar. In the code below, a short section creates a dataframe with the feature of each label, that we can then call in `geom_text()`. 

Note that labels are always in an angle that allows to read them easily, what requires a 180 degrees flip for some of them.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(tidyverse)
 
# Create dataset
data <- data.frame(
  id=seq(1,60),
  individual=paste( "Mister ", seq(1,60), sep=""),
  value=sample( seq(10,100), 60, replace=T)
)
 
# ----- This section prepare a dataframe for labels ---- #
# Get the name and the y position of each label
label_data <- data
 
# calculate the ANGLE of the labels
number_of_bar <- nrow(label_data)
angle <-  90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
label_data$hjust<-ifelse( angle < -90, 1, 0)
 
# flip angle BY to make them readable
label_data$angle<-ifelse(angle < -90, angle+180, angle)
# ----- ------------------------------------------- ---- #
 
 
# Start the plot
p <- ggplot(data, aes(x=as.factor(id), y=value)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  # This add the bars with a blue color
  geom_bar(stat="identity", fill=alpha("skyblue", 0.7)) +
  
  # Limits of the plot = very important. The negative value controls the size of the inner circle, the positive one is useful to add size over each bar
  ylim(-100,120) +
  
  # Custom the theme: no axis title and no cartesian grid
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm")      # Adjust the margin to make in sort labels are not truncated!
  ) +
  
  # This makes the coordinate polar instead of cartesian.
  coord_polar(start = 0) +
  
  # Add the labels, using the label_data dataframe that we have created before
  geom_text(data=label_data, aes(x=id, y=value+10, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 
 
p
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# What's next?
***

Have a look to [graph #297](297-circular-barplot-with-groups.html) for a grouped barplot, or check the circular barplot [section](circular-barplot.html) for more examples.

<a class="btn btn-secondary" href="297-circular-barplot-with-groups.html">Grouped barplot</a>
<a class="btn btn-secondary" href="circular-barplot.html">See all examples</a>




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
