---
title: "2d density plot with ggplot2"
descriptionMeta: "This post introduces the concept of 2d density chart and explains how to build it with R and ggplot2. 2d histograms, hexbin charts, 2d distributions and others are considered."
descriptionTop: "This post introduces the concept of 2d density chart and explains how to build it with R and ggplot2. 2d histograms, hexbin charts, 2d distributions and others are considered."
sectionText: "2d density section"
sectionLink: "2d-density-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/density2d.html"
url: "2d-density-plot-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# The issue with `geom_point()`
***


<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [2d density plot](2d-density-chart.html) is useful to study the relationship between 2 numeric variables if you have a huge number of points. 

To avoid overlapping (as in the scatterplot beside), it divides the plot area in a multitude of small fragment and represents the number of points in this fragment.

There are several types of 2d density plots. Each has its proper [ggplot2](ggplot2.html) function. This post describes all of them.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(tidyverse)
 
# Data
a <- data.frame( x=rnorm(20000, 10, 1.9), y=rnorm(20000, 10, 1.2) )
b <- data.frame( x=rnorm(20000, 14.5, 1.9), y=rnorm(20000, 14.5, 1.9) )
c <- data.frame( x=rnorm(20000, 9.5, 1.9), y=rnorm(20000, 15.5, 1.9) )
data <- rbind(a,b,c)
 
 
# Basic scatterplot
ggplot(data, aes(x=x, y=y) ) +
  geom_point()
```
</div>
</div>



```{r thecode, eval=FALSE}
```










<br><br>



# 2d Histogram with `geom_bin2d()` {#hist}
***

```{r thecode2, echo=FALSE, out.width = c("40%", "40%"), fig.show="hold"}
# 2d histogram with default option
ggplot(data, aes(x=x, y=y) ) +
  geom_bin2d() +
  theme_bw()
 
# Bin size control + color palette
ggplot(data, aes(x=x, y=y) ) +
  geom_bin2d(bins = 70) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
```

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

For 2d histogram, the plot area is divided in a multitude of squares. (It is a 2d version of the classic [histogram](histogram.html)). It is called using the `geom_bin_2d()` function. This function offers a `bins` argument that controls the number of bins you want to display.

<u>Note</u>: If you're not convinced about the importance of the `bins` option, read [this](https://www.data-to-viz.com/caveat/bin_size.html).

</div>


<div class = "col-md-6  col-sm-12">


```{r thecode2, eval=FALSE}
```



</div></div>




<br><br>


# Hexbin chart with `geom_hex()` {#hex}
***

```{r thecode3, echo=FALSE, out.width = c("40%", "40%"), fig.show="hold"}
# Hexbin chart with default option
ggplot(data, aes(x=x, y=y) ) +
  geom_hex() +
  theme_bw()
 
# Bin size control + color palette
ggplot(data, aes(x=x, y=y) ) +
  geom_hex(bins = 70) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()
```

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Another alternative is to divide the plot area in a multitude of hexagons: it is thus called a hexbin chart, and is made using the `geom_hex()` function. 

This function provides the `bins` argument as well, to control the number of division per axis.

</div>


<div class = "col-md-6  col-sm-12">


```{r thecode3, eval=FALSE}
```

</div>
</div>










<br><br>



# 2d distribution with `geom_density_2d` or `stat_density_2d`  {#distr}
***

```{r thecode4, echo=FALSE, out.width = c("50%", "50%", "50%", "50%"), fig.show="hold", fig.height=7}
# Show the contour only
ggplot(data, aes(x=x, y=y) ) +
  geom_density_2d()
 
# Show the area only
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon")
 
# Area + contour
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")
 
# Using raster
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
```


As you can plot a [density chart](density.html) instead of a [histogram](histogram.html), it is possible to compute a 2d density and represent it. Several possibilities are offered by `ggplot2`: you can show the contour of the distribution, or the area, or use the `raster` function:


```{r thecode4, eval=FALSE}
```









<br><br>


# Customize the color palette
***

```{r thecode5, echo=FALSE, out.width = c("33%","33%","33%"), fig.show="hold"}
# Call the palette with a number
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette=4, direction=-1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
 
# The direction argument allows to reverse the palette
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette=4, direction=1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
 
# You can also call the palette using a name.
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette= "Spectral", direction=1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
```

Whatever you use a 2d histogram, a hexbin chart or a 2d distribution, you can and should custom the colour of your chart. Here is a suggestion using the `scale_fill_distiller()` function. You can see other methods in the [ggplot2 section](ggplot2.html) of the gallery.

```{r thecode5, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
