---
title: "Most basic lollipop plot"
descriptionMeta: "How to build a very basic lollipop chart with R and ggplot2. Several examples illustrating how to deal with diverse input format."
descriptionTop: "How to build a very basic [lollipop chart](lollipop-plot.html) with `R` and [ggplot2](ggplot2-package.html). This post how to use `geom_point()` and `geom_segment()` based on different input formats."
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "300-basic-lollipop-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# From 2 numeric variables
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [lollipop plot](lollipop-plot.html) is very close from both [scatterplots](scatterplot.html) and [barplots](barplot.html). 

Thus, 2 types of input format work to build it:

- 2 numeric values like for a scatterplot
- one numeric and one categorical variable like for the barplot.

In any case, a lollipop is built using `geom_point()` for the circle, and `geom_segment()` for the stem.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(x=seq(1,30), y=abs(rnorm(30)))
 
# Plot
ggplot(data, aes(x=x, y=y)) +
  geom_point() + 
  geom_segment( aes(x=x, xend=x, y=0, yend=y))
```

</div>
</div>



```{r thecode, eval=FALSE}
```











# From 1 numeric and 1 categorical variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The code works pretty much the same way, but it is important to note that the X axis can represent a categorical variable as well. In this case, the lollipop chart is a good replacement of the [barplot](barplot.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(
  x=LETTERS[1:26], 
  y=abs(rnorm(26))
)
 
# Plot
ggplot(data, aes(x=x, y=y)) +
  geom_point() + 
  geom_segment( aes(x=x, xend=x, y=0, yend=y))
```

</div>
</div>



```{r thecode2, eval=FALSE}
```




# What's next
***

The lollipop chart is one of my favourite. There is so much to do with it and it is under-utilized in favor of barplot. Visit the [dedicated section](lollipop-plot.html) for more examples produced with `R`, or [data-to-viz](https://www.data-to-viz.com/graph/lollipop.html) to learn about the available variations and caveats to avoid.

<a class="btn btn-secondary" href="lollipop-plot.html">Lollipop section</a>
<a class="btn btn-secondary" href="https://www.data-to-viz.com/graph/lollipop.html">Data to Viz</a>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
