---
title: "Custom lollipop chart"
descriptionMeta: "A lollipop chart is constituted of a circle (made with geom_point) and a segment (made by geom_segment). This page explains how to customize the chart appearance with R and ggplot2."
descriptionTop: "A [lollipop chart](lollipop-plot.html) is constituted of a circle (made with `geom_point()`) and a segment (made with `geom_segment()`). This page explains how to customize the chart appearance with `R` and `ggplot2`."
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "301-custom-lollipop-chart"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Marker
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [lollipop plot](lollipop-plot.html) is constituted of a marker and a stem. You can customize the marker as usual with [ggplot2](ggplot2-package.html):

- `size`, `color`
- `alpha` &rarr; transparency
- `shape` &rarr; see list of available shape [here](6-graph-parameters-reminder.html)
- `stroke` and `fill` &rarr; only for shapes that have stroke, like the `21`

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(tidyverse)
 
# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)
 
# plot
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y)) +
  geom_point( size=5, color="red", fill=alpha("orange", 0.3), alpha=0.7, shape=21, stroke=2) 
```

</div>
</div>



```{r thecode, eval=FALSE}
```











# Stem {#stem}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The stem is built using `geom_segment()` and can be customized as well:

- `size`, `color`
- `linetype` &rarr; can be an integer (see [list](6-graph-parameters-reminder.html)), a word like `dotted`, `dashed`, `dotdash` and more (type `help(linetype)`)

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)

# Plot
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y) , size=1, color="blue", linetype="dotdash" ) +
  geom_point()
```

</div>
</div>



```{r thecode2, eval=FALSE}
```





# General appearance with `theme()`  {#general}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

As usual, you can customize the general appearance of the chart using the `theme()` function.

<u>Note</u>: another solution is to use the pre-built `theme_ipsum()` offered in the `hrbrthemes` package.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)

# Plot
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y), color="grey") +
  geom_point( color="orange", size=4) +
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y")
```

</div>
</div>



```{r thecode3, eval=FALSE}
```













# Horizontal version  {#horiz}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is pretty straightforward to flip the chart using the `coord_flip()` function.

It makes sense to do so if you have long labels &rarr; they will be much easier to read.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)

# Horizontal version
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y), color="skyblue") +
  geom_point( color="blue", size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )
```

</div>
</div>



```{r thecode4, eval=FALSE}
```







# Baseline  {#baseline}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Lastly, you can easily change the baseline of the chart. It gives more insight to the figure if there is a specific threshold in the data that interests you. 

You just have to change the `y` argument in the `geom_segment()` call.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)

# Create data
data <- data.frame(
  x=LETTERS[1:26],
  y=abs(rnorm(26))
)

# Change baseline
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=1, yend=y), color="grey") +
  geom_point( color="orange", size=4) +
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  xlab("") +
  ylab("Value of Y")

```

</div>
</div>



```{r thecode5, eval=FALSE}
```




# What's next
***

The lollipop chart is one of my favourite. There is so much to do with it and it is under-utilized in favor of barplot. Visit the [dedicated section](lollipop-plot.html) for more examples produced with `R`, or [data-to-viz](https://www.data-to-viz.com/graph/lollipop.html) to learn about the available variations and caveats to avoid.

<a class="btn btn-secondary" href="lollipop-plot.html">Lollipop section</a>
<a class="btn btn-secondary" href="https://www.data-to-viz.com/graph/lollipop.html">Data to Viz</a>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
