---
title: "Lollipop chart with conditional color"
descriptionMeta: "If your lollipop plot goes on both side of an interesting threshold, you probably want to change the color of its components conditionally. Here is how using R and ggplot2"
descriptionTop: "If your [lollipop plot ](lollipop-plot.html) goes on both side of an interesting threshold, you probably want to change the color of its components conditionally. Here is how using `R` and [ggplot2](ggplot2.html)"
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "302-lollipop-chart-with-conditional-color"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Marker
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here is the process to use conditional color on your `ggplot2` chart:

- add a new column to your dataframe specifying if you are over or under the threshold (use an `ifelse` statement)
- give this column to the `color` aesthetic


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
library(dplyr)

# Create data (this takes more sense with a numerical X axis)
x <- seq(0, 2*pi, length.out=100)
data <- data.frame(
  x=x, 
  y=sin(x) + rnorm(100, sd=0.2)
)
 
# Add a column with your condition for the color
data <- data %>% 
  mutate(mycolor = ifelse(y>0, "type1", "type2"))
 
# plot
ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y, color=mycolor), size=1.3, alpha=0.9) +
  theme_light() +
  theme(
    legend.position = "none",
    panel.border = element_blank(),
  ) +
  xlab("") +
  ylab("Value of Y")
```

</div>
</div>



```{r thecode, eval=FALSE}
```












# What's next
***

The lollipop chart is one of my favourite. There is so much to do with it and it is under-utilized in favor of barplot. Visit the [dedicated section](lollipop-plot.html) for more examples produced with `R`, or [data-to-viz](https://www.data-to-viz.com/graph/lollipop.html) to learn about the available variations and caveats to avoid.

<a class="btn btn-secondary" href="lollipop-plot.html">Lollipop section</a>
<a class="btn btn-secondary" href="https://www.data-to-viz.com/graph/lollipop.html">Data to Viz</a>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
