---
title: "Lollipop chart with 2 groups"
descriptionMeta: "A lollipop chart can be used to compare 2 categories, linking them with a segment to stress out their difference. This post explains how to do it with R and ggplot2."
descriptionTop: "A [lollipop chart](lollipop-plot.html) can be used to compare 2 categories, linking them with a segment to stress out their difference. This post explains how to do it with `R` and [ggplot2](ggplot2-package.html)."
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "303-lollipop-plot-with-2-values"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Lollipop plots](lollipop-plot.html) can be very appropriate when it comes to compare 2 values for several entities. 

For each entity, one point is drawn for each variable, with a different color. Their difference is then highlighted using a segment. This type of visualisation is also called [Cleveland dot plots](https://www.data-to-viz.com/graph/lollipop.html). 

As usual, it is advised to [order your individuals](267-reorder-a-variable-in-ggplot2.html) by mean, median, or group difference to give even more insight to the figure. 


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Create data
value1 <- abs(rnorm(26))*2
data <- data.frame(
  x=LETTERS[1:26], 
  value1=value1, 
  value2=value1+1+rnorm(26, sd=1) 
)
 
# Reorder data using average? Learn more about reordering in chart #267
data <- data %>% 
  rowwise() %>% 
  mutate( mymean = mean(c(value1,value2) )) %>% 
  arrange(mymean) %>% 
  mutate(x=factor(x, x))
 
# Plot
ggplot(data) +
  geom_segment( aes(x=x, xend=x, y=value1, yend=value2), color="grey") +
  geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_ipsum() +
  theme(
    legend.position = "none",
  ) +
  xlab("") +
  ylab("Value of Y")
```

</div>
</div>



```{r thecode, eval=FALSE}
```





# What's next
***

The lollipop chart is one of my favourite. There is so much to do with it and it is under-utilized in favor of barplot. Visit the [dedicated section](lollipop-plot.html) for more examples produced with `R`, or [data-to-viz](https://www.data-to-viz.com/graph/lollipop.html) to learn about the available variations and caveats to avoid.

<a class="btn btn-secondary" href="lollipop-plot.html">Lollipop section</a>
<a class="btn btn-secondary" href="https://www.data-to-viz.com/graph/lollipop.html">Data to Viz</a>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
