---
title: "Highlight a group in a lollipop chart"
descriptionMeta: "This post describes how to build a lollipop chart with R and ggplot2. It shows how to highlight one or several groups of interest to convey your message more efficiently."
descriptionTop: "This post describes how to build a [lollipop chart](lollipop-plot.html) with `R` and `ggplot2`. It shows how to highlight one or several groups of interest to convey your message more efficiently."
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "304-highlight-a-group-in-lollipop"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Annotation is key](https://www.data-to-viz.com/caveat/annotation.html) in data visualization: it allows the reader to focus on the main message you want to convey.

If one or a few groups specifically interest you, it is a good practice to highlight them on the plot. Your reader will understand quicker what the story behind the chart is.

To do so, you can use an `ifelse` statement to change `size`, `color`, `alpha` or any other aesthetics. Moreover, it is even more insightful to add text annotation directly on the chart. 

Learn more about ggplot2 annotation [here](https://www.r-graph-gallery.com/233-add-annotations-on-ggplot2-chart/) and more generally in the [ggplot2](ggplot2.html) section.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Create data
set.seed(1000)
data <- data.frame(
  x=LETTERS[1:26], 
  y=abs(rnorm(26))
)
 
# Reorder the data
data <- data %>%
  arrange(y) %>%
  mutate(x=factor(x,x))
  
# Plot
p <- ggplot(data, aes(x=x, y=y)) +
  geom_segment( aes(x=x, xend=x, y=0, yend=y ), color=ifelse(data$x %in% c("A","D"), "orange", "grey"), size=ifelse(data$x %in% c("A","D"), 1.3, 0.7) ) +
  geom_point( color=ifelse(data$x %in% c("A","D"), "orange", "grey"), size=ifelse(data$x %in% c("A","D"), 5, 2) ) +
  theme_ipsum() +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  xlab("") +
  ylab("Value of Y") +
  ggtitle("How did groups A and D perform?")

# Add annotation
p + annotate("text", x=grep("D", data$x), y=data$y[which(data$x=="D")]*1.2, 
           label="Group D is very impressive", 
           color="orange", size=4 , angle=0, fontface="bold", hjust=0) + 
  
    annotate("text", x = grep("A", data$x), y = data$y[which(data$x=="A")]*1.2, 
           label = paste("Group A is not too bad\n (val=",data$y[which(data$x=="A")] %>% round(2),")",sep="" ) , 
           color="orange", size=4 , angle=0, fontface="bold", hjust=0) 
  

```

</div>
</div>



```{r thecode, eval=FALSE}
```





# What's next
***

The lollipop chart is one of my favourite. There is so much to do with it and it is under-utilized in favor of barplot. Visit the [dedicated section](lollipop-plot.html) for more examples produced with `R`, or [data-to-viz](https://www.data-to-viz.com/graph/lollipop.html) to learn about the available variations and caveats to avoid.

<a class="btn btn-secondary" href="lollipop-plot.html">Lollipop section</a>
<a class="btn btn-secondary" href="https://www.data-to-viz.com/graph/lollipop.html">Data to Viz</a>


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
