---
title: "Basic circle packing with one level"
descriptionMeta: "This page aims to describe how to build a basic circle packing chart with only one level of hierarchy. It uses the packcircle package for circle position, and ggplot2 for drawing."
descriptionTop: "This page aims to describe how to build a basic [circle packing](circle-packing.html) chart with only one level of hierarchy. It uses the `packcircle` package for circle position, and `ggplot2` for drawing."
sectionText: "Circle Packing section"
sectionLink: "circle-packing.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/circularpacking.html"
url: "305-basic-circle-packing-with-one-level"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This page aims to describe how to build a basic [circle packing](circle-packing.html) chart with only one level of hierarchy. Basically, you just represent each entity or individual of your dataset with a circle, its size depending on a provided value.
 
It is like a [barplot](barplot.html), but you use circle size instead of bar length. It is close to a [bubble plot](bubble-chart.html), but X and Y positions do not mean anything. It is a circle version of a [treemap](treemap.html).
 
Calculating the arrangement of dots is not a trivial problem. The `packcircles` library solves it and output coordinates of every points of the circle edges. 

Finally, ggplot2 allows to draw shapes thanks to `geom_polygon()`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(packcircles)
library(ggplot2)
 
# Create data
data <- data.frame(group=paste("Group", letters[1:20]), value=sample(seq(1,100),20)) 
 
# Generate the layout. This function return a dataframe with one line per bubble. 
# It gives its center (x and y) and its radius, proportional of the value
packing <- circleProgressiveLayout(data$value, sizetype='area')
 
# We can add these packing information to the initial data frame
data <- cbind(data, packing)
 
# Check that radius is proportional to value. We don't want a linear relationship, since it is the AREA that must be proportionnal to the value
# plot(data$radius, data$value)
 
# The next step is to go from one center + a radius to the coordinates of a circle that
# is drawn by a multitude of straight lines.
dat.gg <- circleLayoutVertices(packing, npoints=50)
 
# Make the plot
ggplot() + 
  
  # Make the bubbles
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=as.factor(id)), colour = "black", alpha = 0.6) +
  
  # Add text in the center of each bubble + control its size
  geom_text(data = data, aes(x, y, size=value, label = group)) +
  scale_size_continuous(range = c(1,4)) +
  
  # General theme:
  theme_void() + 
  theme(legend.position="none") +
  coord_equal()
```
</div>
</div>



```{r thecode, eval=FALSE}
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
