---
title: "Circle packing customization with R"
descriptionMeta: "This page is dedicated to one level circle packing customization with R. It notably shows how to use different color palettes and provides reproducible code snippets."
descriptionTop: "This page is dedicated to one level [circle packing](circle-packing.html) customization with R. It notably shows how to use different color palettes and provides reproducible code snippets."
sectionText: "Circle Packing section"
sectionLink: "circle-packing.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/circularpacking.html"
url: "306-custom-circle-packing-with-one-level"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" >




# Using the `viridis` color scale
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This chart follows the previous most [basic circle packing](305-basic-circle-packing-with-one-level) post. 

It shows how to use the awesome `viridis` package to build color scales, a very good alternative to the usual `colorBrewer`.

Note that `magma` is used here, but you could use the same code with `inferno` or viridis` instead.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# libraries
library(packcircles)
library(ggplot2)
library(viridis)
 
# Create data
data <- data.frame(group=paste("Group", letters[1:20]), value=sample(seq(1,100),20)) 
 
# Generate the layout. sizetype can be area or radius, following your preference on what to be proportional to value.
packing <- circleProgressiveLayout(data$value, sizetype='area')
data <- cbind(data, packing)
dat.gg <- circleLayoutVertices(packing, npoints=50)
 
# Basic color customization
ggplot() + 
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=as.factor(id)), colour = "black", alpha = 0.6) +
  scale_fill_manual(values = magma(nrow(data))) +
  geom_text(data = data, aes(x, y, size=value, label = group)) +
  scale_size_continuous(range = c(1,4)) +
  theme_void() + 
  theme(legend.position="none") +
  coord_equal()
 
```
</div>
</div>



```{r thecode, eval=FALSE}
```











# Map color to bubble value
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common task to make the bubble color being lighter or darker according to its value.

This is possible by passing the focus variable to the dataframe that is read by `ggplot2`, and specifying it in tha `aes()`. 

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# First I need to add the 'value' of each group to dat.gg.
# Here I repeat each value 51 times since I create my polygons with 50 lines
dat.gg$value <- rep(data$value, each=51)

# Plot
ggplot() + 
  
  # Make the bubbles
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=value), colour = "black", alpha = 0.6) +
  scale_fill_distiller(palette = "BuPu", direction = 1 ) +
 
  # Add text in the center of each bubble + control its size
  geom_text(data = data, aes(x, y, size=value, label = group)) +
  scale_size_continuous(range = c(1,4)) +
 
  # General theme:
  theme_void()  + 
  theme(legend.position="none") + 
  coord_equal()
 
```
</div>
</div>



```{r thecode2, eval=FALSE}
```












# Background customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Change the background thanks to the `theme()` function and its `plot.background()` argument.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
ggplot() + 
  
  # Make the bubbles
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=value), colour = "grey", alpha = 0.6, size=.5) +
  scale_fill_distiller(palette = "Spectral", direction = 1 ) +
  
  # Add text in the center of each bubble + control its size
  geom_label(data = data, aes(x, y, size=value, label = group)) +
  scale_size_continuous(range = c(1,4)) +
  
  # General theme:
  theme_void()  + 
  theme(
    legend.position="none",
    plot.background = element_rect(fill="black"),
    plot.title = element_text(color="white") 
  ) + 
  coord_equal() +
  ggtitle("A custom circle packing with\nblack background")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
