---
title: "Circle packing customization with R"
descriptionMeta: "This posts explains how to build a one level circle packing chart with space between bubbles, improving the general appearance."
descriptionTop: "This posts explains how to build a one level [circle packing](circle-packing.html) chart with space between bubbles, improving the general appearance."
sectionText: "Circle Packing section"
sectionLink: "circle-packing.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/circularpacking.html"
url: "307-add-space-in-circle-packing"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This chart is just a customisation of the [chart #305](305-basic-circle-packing-with-one-level.html) which describes the basic process to make a one level circle packing chart. I personally like to add a bit of space between each circle.
 
Basically, all you have to do is to reduce the `radius` size in your data once this one has been calculated. Just multiply it by a number under 0, and it will decrease the circle size.
 
If you have been so far, you probably want to check the [interactive version](308-interactive-circle-packing.html) of the chart !

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# libraries
library(packcircles)
library(ggplot2)
library(viridis)

# Create data
data <- data.frame(group=paste("Group", letters[1:20]), value=sample(seq(1,100),20)) 

# Generate the layout
packing <- circleProgressiveLayout(data$value, sizetype='area')
packing$radius <- 0.95*packing$radius
data <- cbind(data, packing)
dat.gg <- circleLayoutVertices(packing, npoints=50)

# Plot 
ggplot() + 
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=id), colour = "black", alpha = 0.6) +
  scale_fill_viridis() +
  geom_text(data = data, aes(x, y, size=value, label = group), color="black") +
  theme_void() + 
  theme(legend.position="none")+ 
  coord_equal()
```
</div>
</div>



```{r thecode, eval=FALSE}
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
