---
title: "Interactive circle packing with R"
descriptionMeta: "This post describes how to build an interactive circle packing chart with R and the ggiraph package. It allows to hover bubbles to get additionnal information."
descriptionTop: "This post describes how to build an interactive [circle packing](circle-packing.html) chart with R and the `ggiraph` package. It allows to hover bubbles to get additionnal information."
sectionText: "Circle Packing section"
sectionLink: "circle-packing.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/circularpacking.html"
url: "308-interactive-circle-packing"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 50px">



<center><iframe src="HtmlWidget/circular_packing_interactive.html" height="640" width="100%" style="border:none;"></iframe></center>

This chart follows posts [#305](305-basic-circle-packing-with-one-level.html) and [#306](306-custom-circle-packing-with-one-level.html) that explains how to build a static version of circle packing, and how to customize it. 

This interactive version is very close to the static one. It uses the `ggiraph` library to transform the `ggplot2` code in something interactive. The steps are quite easy: 

- First you need to prepare a column in the data frame with the text you want to display while hovering. 
- Second, you need to change the geometries to use the interactive geometries of ggiraph. 

Check the code below:

```{r, eval=FALSE}
# libraries
library(packcircles)
library(ggplot2)
library(viridis)
library(ggiraph)

# Create data
data <- data.frame(group=paste("Group_", sample(letters, 70, replace=T), sample(letters, 70, replace=T), sample(letters, 70, replace=T), sep="" ), value=sample(seq(1,70),70)) 

# Add a column with the text you want to display for each bubble:
data$text <- paste("name: ",data$group, "\n", "value:", data$value, "\n", "You can add a story here!")

# Generate the layout
packing <- circleProgressiveLayout(data$value, sizetype='area')
data <- cbind(data, packing)
dat.gg <- circleLayoutVertices(packing, npoints=50)

# Make the plot with a few differences compared to the static version:
p <- ggplot() + 
  geom_polygon_interactive(data = dat.gg, aes(x, y, group = id, fill=id, tooltip = data$text[id], data_id = id), colour = "black", alpha = 0.6) +
  scale_fill_viridis() +
  geom_text(data = data, aes(x, y, label = gsub("Group_", "", group)), size=2, color="black") +
  theme_void() + 
  theme(legend.position="none", plot.margin=unit(c(0,0,0,0),"cm") ) + 
  coord_equal()

# Turn it interactive
widg <- ggiraph(ggobj = p, width_svg = 7, height_svg = 7)
# widg

# save the widget
# library(htmlwidgets)
# saveWidget(widg, file=paste0( getwd(), "/HtmlWidget/circular_packing_interactive.html"))
```





</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
