---
title: "Add labels to Hierarchical Edge Bundling"
descriptionMeta: "This post describes how to add labels on a hierarchical edge bundling chart. It explains how to proceed, with reproducible R code."
descriptionTop: "This post describes how to add labels on a [hierarchical edge bundling](hierarchical-edge-bundling.html) chart. It explains how to proceed, with reproducible `R` code, using the `ggraph` package."
sectionText: "More examples"
sectionLink: "hierarchical-edge-bundling.html"
DataToVizText: "Data to viz"
DataToVizLink: "data-to-viz.com/caveat/overplotting.html"
url: "311-add-labels-to-hierarchical-edge-bundling"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">






# Without Labels
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Previous post of the [hierarchical edge bundling](hierarchical-edge-bundling.html) section explained:

- how to build a very [basic version](309-intro-to-hierarchical-edge-bundling.html).
- how to [customize](310-custom-hierarchical-edge-bundling.html) connection and node features

Let's remind how to prepare the data for the `ggraph` library.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer)
 
# create a data frame giving the hierarchical structure of your individuals
set.seed(1234)
d1 <- data.frame(from="origin", to=paste("group", seq(1,10), sep=""))
d2 <- data.frame(from=rep(d1$to, each=10), to=paste("subgroup", seq(1,100), sep="_"))
edges <- rbind(d1, d2)
 
# create a dataframe with connection between leaves (individuals)
all_leaves <- paste("subgroup", seq(1,100), sep="_")
connect <- rbind( 
  data.frame( from=sample(all_leaves, 100, replace=T) , to=sample(all_leaves, 100, replace=T)), 
  data.frame( from=sample(head(all_leaves), 30, replace=T) , to=sample( tail(all_leaves), 30, replace=T)), 
  data.frame( from=sample(all_leaves[25:30], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)), 
  data.frame( from=sample(all_leaves[75:80], 30, replace=T) , to=sample( all_leaves[55:60], 30, replace=T)) )
connect$value <- runif(nrow(connect))
 
# create a vertices data.frame. One line per object of our hierarchy
vertices  <-  data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) , 
  value = runif(111)
) 
# Let's add a column with the group of each name. It will be useful later to color points
vertices$group  <-  edges$from[ match( vertices$name, edges$to ) ]
```

</div>
</div>



```{r thecode, eval=FALSE}
```






# Create the labels
***

Next step: computing the label features that will be displayed all around the circle, next to the nodes:

- angle &rarr; vertical on top and botton, horizontal on the side, and so on.
- flip it  &rarr;  labels on the left hand side must be 180° flipped to be readable
- alignment  &rarr; if labels are flipped, they must be right aligned

Those information are computed and added to the `vertices` data frame.

```{r }
#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels
vertices$id <- NA
myleaves <- which(is.na( match(vertices$name, edges$from) ))
nleaves <- length(myleaves)
vertices$id[ myleaves ] <- seq(1:nleaves)
vertices$angle <- 90 - 360 * vertices$id / nleaves
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices$hjust <- ifelse( vertices$angle < -90, 1, 0)
 
# flip angle BY to make them readable
vertices$angle <- ifelse(vertices$angle < -90, vertices$angle+180, vertices$angle)
```



# Plot the labels
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Now that label features have been computed, we just need to display it on the chart using the `geom_node_text()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Create a graph object
mygraph <- igraph::graph_from_data_frame( edges, vertices=vertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( connect$from, vertices$name)
to  <-  match( connect$to, vertices$name)
 
# Basic usual argument
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_node_point(aes(filter = leaf, x = x*1.05, y=y*1.05)) +
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, colour="skyblue", width=0.9) +
  geom_node_text(aes(x = x*1.1, y=y*1.1, filter = leaf, label=name, angle = angle, hjust=hjust), size=1.5, alpha=1) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.2, 1.2), y = c(-1.2, 1.2))
```

</div>
</div>



```{r thecode2, eval=FALSE}
```




# With Customization
***


To get the final figure, it is necessary to add customization described in [graph #310](310-custom-hierarchical-edge-bundling.html):

- control node size, color and transparency.
- control connection color

```{r, fig.align="center", fig.width=7, fig.height=7}
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, width=0.9, aes(colour=..index..)) +
  scale_edge_colour_distiller(palette = "RdPu") +
  
  geom_node_text(aes(x = x*1.15, y=y*1.15, filter = leaf, label=name, angle = angle, hjust=hjust, colour=group), size=2, alpha=1) +
  
  geom_node_point(aes(filter = leaf, x = x*1.07, y=y*1.07, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(9,"Paired") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
