---
title: "Time series visualization with the dygraphs package"
descriptionMeta: "This post describes the options offered by the dygraphs package for interactive time series visualization with R. It shows the different chart types available and how to customize them."
descriptionTop: "This post describes the options offered by the `dygraphs` package for interactive [time series](time-series.html) visualization with R. It shows the different chart types available and how to customize them."
sectionText: "Time series section"
sectionLink: "time-series.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/connectedscatter.html"
url: "317-time-series-with-the-dygraphs-library"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



The `dygraphs` R library is my favorite tool to plot [time series](time-series.html). The [chart #316](316-possible-inputs-for-the-dygraphs-library.html) describes extensively its basic utilisation, notably concerning the required input format. 
This page aims to describe the chart types that this library offers. Remember you can zoom and hover on every following chart.




# Connected scatterplot
***
Most of the chart types described in this post are called using the `dyOptions()` function. For connected scatterplots, use `drawPoints = TRUE`. Note that the gallery offers a [whole section](connected-scatterplot.html) on connected scatterplot.

<center><iframe src="HtmlWidget/dygraphs317-1.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
# Library
library(dygraphs)
library(xts)          # To make the convertion data-frame / xts format
 
# Create data 
data <- data.frame(
  time=seq(from=Sys.Date()-40, to=Sys.Date(), by=1 ), 
  value=runif(41)
)

# Double check time is at the date format
str(data$time)

# Switch to XTS format
data <- xts(x = data$value, order.by = data$time)
 
# Default = line plot --> See chart #316
 
# Add points
p <- dygraph(data) %>%
  dyOptions( drawPoints = TRUE, pointSize = 4 )
p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-1.html"))
```








# Area chart
***

Area chart are built thanks to the `fillGraph = TRUE` option. See the [area chart](area-graph.html) section of the gallery.

<center><iframe src="HtmlWidget/dygraphs317-2.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
p <- dygraph(data) %>%
  dyOptions( fillGraph=TRUE )
p

# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-2.html"))
```








# Step chart
***

The step chart is made using the .. `stepPlot` option! Use it in conjunction with `fillGraph` to fill the area below the curve.


<center><iframe src="HtmlWidget/dygraphs317-3.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
p <- dygraph(data) %>%
  dyOptions( stepPlot=TRUE, fillGraph=TRUE)
p

# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-3.html"))
```









# Lollipop plot
***

Called using the `stemPlot` option. See the [lollipop plot](lollipop-plot.html) section of the gallery for more.

<center><iframe src="HtmlWidget/dygraphs317-4.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
p <- dygraph(data) %>%
  dyOptions( stemPlot=TRUE)
p

# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-4.html"))
```










# Candlestick chart
***

The candlestick chart represents 4 series and is widely used in finance. `dygraphs` offers the `dyCandlestick()` function that allows to build them in minutes.

<center><iframe src="HtmlWidget/dygraphs317-5.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
# Create data (needs 4 data points per date stamp)
trend <- sin(seq(1,41))+runif(41)
data <- data.frame(
  time=seq(from=Sys.Date()-40, to=Sys.Date(), by=1 ), 
  value1=trend, 
  value2=trend+rnorm(41), 
  value3=trend+rnorm(41), 
  value4=trend+rnorm(41) 
)

# switch to xts format
data <- xts(x = data[,-1], order.by = data$time)

# Plot it
p <- dygraph(data) %>%
  dyCandlestick()
p

# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-5.html"))
```











# Line chart with interval
***

This is very handy to represent confidence interval around your time series. Don with `dySeries()` that takes 3 columns as input: trend and upper and lower limits of the confidence interval.


<center><iframe src="HtmlWidget/dygraphs317-6.html" height="300" width="100%" style="border:none;"></iframe></center>

```{r eval=FALSE}
# Create data
trend <- sin(seq(1,41))+runif(41)
data <- data.frame(
  time=seq(from=Sys.Date()-40, to=Sys.Date(), by=1 ), 
  trend=trend, 
  max=trend+abs(rnorm(41)), 
  min=trend-abs(rnorm(41, sd=1))
)

# switch to xts format
data <- xts(x = data[,-1], order.by = data$time)

# Plot
p <- dygraph(data) %>%
  dySeries(c("min", "trend", "max"))
p

# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dygraphs317-6.html"))
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
