---
title: "Bubble plot with ggplot2"
descriptionMeta: "This post explains how to build a bubble chart with R and ggplot2. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build a [bubble chart](bubble-chart.html) with R and [ggplot2](ggplot2-package.html). It provides several reproducible examples with explanation and R code."
sectionText: "Bubble section"
sectionLink: "bubble-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/bubble.html"
url: "320-the-basis-of-bubble-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic bubble chart with `geom_point()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [bubble plot](bubble-chart.html) is a [scatterplot](scatterplot.html) where a third dimension is added: the value of an additional numeric variable is represented through the size of the dots. (source: [data-to-viz](https://www.data-to-viz.com/graph/bubble.html)).

With [ggplot2](ggplot2-package.html), bubble chart are built thanks to the `geom_point()` function. At least three variable must be provided to `aes()`: x, y and size. The legend will automatically be built by ggplot2.

Here, the relationship between life expectancy (`y`) and gdp per capita (`x`) of world countries is represented. The population of each country is represented through circle size.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)

# The dataset is provided in the gapminder library
library(gapminder)
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)

# Most basic bubble plot
ggplot(data, aes(x=gdpPercap, y=lifeExp, size = pop)) +
    geom_point(alpha=0.7)
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# Control circle size with `scale_size()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The first thing we need to improve on the previous chart is the bubble size. `scale_size()` allows to set the size of the smallest and the biggest circles using the `range` argument. Note that you can customize the legend name with `name`.

<u>Note</u>: circles often overlap. To avoid having big circles on top of the chart you have to reorder your dataset first, as in the code below.

<u>ToDo</u>: give more details about how to map a numeric variable to circle size. Use of `scale_radius`, `scale_size` and `scale_size_area`. See [this post](https://www.data-to-viz.com/caveat/radius_or_area.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)

# The dataset is provided in the gapminder library
library(gapminder)
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)

# Most basic bubble plot
data %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size = pop)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 24), name="Population (M)")
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










# Add a fourth dimension: color
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If you have one more variable in your dataset, why not showing it using circle color? Here, the continent of each country is used to control circle color:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)

# The dataset is provided in the gapminder library
library(gapminder)
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)

# Most basic bubble plot
data %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size=pop, color=continent)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 24), name="Population (M)")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```











# Make it pretty
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A few classic improvement:

- use of the `viridis` package for nice color palette
- use of `theme_ipsum()` of the `hrbrthemes` package
- custom axis titles with `xlab` and `ylab`
- add stroke to circle: change `shape` to 21 and specify `color` (stroke) and `fill`

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

# The dataset is provided in the gapminder library
library(gapminder)
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)

# Most basic bubble plot
data %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size=pop, fill=continent)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    scale_size(range = c(.1, 24), name="Population (M)") +
    scale_fill_viridis(discrete=TRUE, guide=FALSE, option="A") +
    theme_ipsum() +
    theme(legend.position="bottom") +
    ylab("Life Expectancy") +
    xlab("Gdp per Capita") +
    theme(legend.position = "none")
```
</div>
</div>



```{r thecode4, eval=FALSE}
```




# What's next?
***

This chart could be much better with annotation, or with an interactive version. Don't you think it is frustrating not to be able to access country name? 

Visit the [bubble chart section](bubble-chart.html) to see it.






<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
