---
title: "Most basic Sankey Diagram"
descriptionMeta: "Sankey Diagram can be built in R using the networkD3 package. This posts displays basic example, focusing on the different input formats that can be used."
descriptionTop: "[Sankey Diagram](sankey-diagram.html) can be built in `R` using the `networkD3` package. This posts displays basic example, focusing on the different input formats that can be used."
sectionText: "Sankey section"
sectionLink: "sankey-diagram.html"
DataToVizText: "About Sankey"
DataToVizLink: "data-to-viz.com/graph/sankey.html"
url: "321-introduction-to-interactive-sankey-diagram-2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">




A [Sankey diagram](sankey-diagram.html) represents flows, i.e. weigthed connections going from one node to another. Input data can be stored in 2 different formats:

- connection data frame (3 columns)
- incidence matrix (square matrix)

This post describes how to build a basic Sankey diagram from these 2 types of input.



# From connection data frame
***

<div class="row">

<div class = "col-md-5  col-sm-12 align-self-center">

A connection data frame lists all the connections one by one in a data frame. 

Usually you have a `source` and a `target` column. You can add a third column that gives further information for each connection, like the value of the flow. 

This is the format you need to use the `networkD3` library. Let’s build a connection data frame and represent it as a Sankey diagram:

</div>

<div class = "col-md-7  col-sm-12">

<center><iframe src="HtmlWidget/sankeyBasic1.html" height="440" width="100%" style="border:none;"></iframe></center>

</div>

</div>

```{r, eval=FALSE}
# Library
library(networkD3)
library(dplyr)
 
# A connection data frame is a list of flows with intensity for each flow
links <- data.frame(
  source=c("group_A","group_A", "group_B", "group_C", "group_C", "group_E"), 
  target=c("group_C","group_D", "group_E", "group_F", "group_G", "group_H"), 
  value=c(2,3, 2, 3, 1, 3)
  )
 
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
 
# Make the Network
p <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", 
              sinksRight=FALSE)
p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/sankeyBasic1.html"))
```









# From incidence matrix
***

<div class="row">

<div class = "col-md-5  col-sm-12 align-self-center">

An incidence matrix is square or rectangle. 

Row and column names are node names. The item in row x and column y represents the flow between x and y. In the Sankey diagram we represent all flows that are over 0. 

Since the `networkD3` library expects a connection data frame, we will fist convert the dataset, and then re-use the code from above.

</div>

<div class = "col-md-7  col-sm-12">

<center><iframe src="HtmlWidget/sankeyBasic2.html" height="440" width="100%" style="border:none;"></iframe></center>

</div>

</div>

```{r, eval=FALSE}
# Library
library(networkD3)
library(dplyr)
 
# Create an incidence matrix. Usually the flow goes from the row names to the column names.
# Remember that our connection are directed since we are working with a flow.
set.seed(1)
data <- matrix(sample( seq(0,40), 49, replace=T ), 7, 7)
data[data < 35] <- 0
colnames(data) = rownames(data) = c("group_A", "group_B", "group_C", "group_D", "group_E", "group_F", "group_G")

# Transform it to connection data frame with tidyr from the tidyverse:
links <- data %>% 
  as.data.frame() %>% 
  rownames_to_column(var="source") %>% 
  gather(key="target", value="value", -1) %>%
  filter(value != 0)
 
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$source), as.character(links$target)) %>% 
    unique()
  )
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
 
# Make the Network
p <- sankeyNetwork(Links = links, Nodes = nodes,
                     Source = "IDsource", Target = "IDtarget",
                     Value = "value", NodeID = "name", 
                     sinksRight=FALSE)

p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/sankeyBasic2.html"))
```







</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
