---
title: "Open and Plot Geojson files in R"
descriptionMeta: "Geojson files are a common way to store geospatial data. This post explains how to read it with R and the sf package, and how to plot it in base R or with ggplot2."
descriptionTop: "Geojson files are a common way to store geospatial data. This post explains how to read it with `R` and the `sf` package, and how to plot it in base R or with `ggplot2`."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "325-background-map-from-geojson-format-in-r.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dpi = 300)
```

<div class="container" style="padding-top: 100px">







If you did not find the geospatial data you need in existing R packages (see the [map section](map.html)), you need to find this information elsewhere on the web.

It will often be stored as a `.geoJSON` format. This post explains how to read it.









# Find and download a `.geoJSON` file
***

You need to dig the internet to find the geoJSON file you are interested in. For instance, [this URL](https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson) provides a file containing french region boundaries.

You can load it in R with:

```{r get_geojson, cache=TRUE}
# Download to a temporary file
tmp_geojson <- tempfile(fileext = ".geojson")
download.file(
  "https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",
  tmp_geojson
)

# Let's read the downloaded geoJson file with the sf library:
library(sf)
my_sf <- read_sf(tmp_geojson)
```

<br>

That's it! You now have a geospatial object called `my_sf`. I strongly advise to read [this post](169-170-basic-manipulation-of-shapefiles) to learn how to manipulate it.

Just in case, here is how to plot it in base R and with `ggplot2`.







# Plot it with `base R`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The basic `plot()` function knows how to plot a geospatial object. Thus you just need to pass it `my_sf` and add a couple of options to customize the output.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Select only the region #6 as you would do with a data frame
my_sf_region_6 <- my_sf[substr(my_sf$code, 1, 2) == "06", ]

par(mar = c(0, 0, 0, 0))
plot(my_sf_region_6$geometry, col = "grey")
```

</div>
</div>

```{r thecode4, eval=FALSE}
```













# Plot it with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally possible (and advised IMO) to build the map with [ggplot2](ggplot2-package.html), using the `geom_sf()` function as described below.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=4}
# Plot it
library(ggplot2)
ggplot(my_sf_region_6) +
  geom_sf(fill = "#69b3a2", color = "white") +
  theme_void()
```
</div>
</div>

```{r thecode5, eval=FALSE}
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
