---
title: "Open and Plot Geojson files in R"
descriptionMeta: "Geojson files are a common way to store geospatial data. This post explains how to read it with R and the geojsonio package, and how to plot it in base R or with ggplot2."
descriptionTop: "Geojson files are a common way to store geospatial data. This post explains how to read it with `R` and the `geojsonio` package, and how to plot it in base R or with `ggplot2`."
sectionText: "Background map section"
sectionLink: "map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "325-background-map-from-geojson-format-in-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">







If you did not find the geospatial data you need in existing R packages (see the [map section](maps.html)), you need to find this information elsewhere on the web. 

It will often be stored as a `.geomJSON` format. This post explains how to read it.

<u>Note</u>: if you found a `shapefile`, read [this post](168-load-a-shape-file-into-r.html) instead.










# Find and download a `.geoJSON` file
***

You need to dig the internet to find the geoJSON file you are interested in. For instance, [this URL](https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson) provides a file containing french region boundaries.

You can load it in R with:

```{r, cache=TRUE}
# Let's read the jeoJson file that is stored on the web with the geojsonio library:
library(geojsonio)
spdf <- geojson_read("https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",  what = "sp")
```

<br>

That's it! You now have a geospatial object called `spdf`. I strongly advise to read [this post](169-170-basic-manipulation-of-shapefiles) to learn how to manipulate it.

Just in case, here is how to plot it in base R and with `ggplot2`.







# Plot it with `base R`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The basic `plot()` function knows how to plot a geospatial object. Thus you just need to pass it `spdf` and add a couple of options to customize the output.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Select only the region #6
spdf@data$mystate = substr( spdf@data$code, 1, 2)
spdf_region_6 = spdf[ spdf@data$mystate == "06" , ]

# plot the selected are with sp
library(sp)
par(mar=c(0,0,0,0))
plot(spdf_region_6, col="grey")
```

</div>
</div>

```{r thecode4, eval=FALSE}
```













# Plot it with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally possible (and advised imo) to build the map with [ggplot2](ggplot2-package.html). However, ggplot2 takes as input data frames, not geospatial data.

The geospatial object thus needs to be transformed using the `tidy()` function of the `broom` package.

Once the data frame is created, it is plotted using the `geom_polygon()` function as described below.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=4}
# 'fortify' the data to get a dataframe format required by ggplot2
library(broom)
spdf_fortified <- tidy(spdf_region_6)

# Plot it
library(ggplot2)
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="#69b3a2", color="white") +
  theme_void() +
  coord_map()
```
</div>
</div>

```{r thecode5, eval=FALSE}
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
