---
title: "Choropleth map with R and ggplot2"
descriptionMeta: "How to build a choropleth map with R and the ggplot2 package. This post shows how to load geospatial data in R, merge region features and build the map. Reproducible code is provided."
descriptionTop: "How to build a [choropleth map](choropleth-map.html) with `R` and the [ggplot2](ggplot2-package.html) package. This post shows how to **load geospatial data** in R, **merge region** features and **build the map**. Reproducible code is provided."
sectionText: "Choropleth section"
sectionLink: "choropleth-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/choropleth.html"
url: "327-chloropleth-map-from-geojson-with-ggplot2"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container" style="padding-top: 100px">






Two inputs are needed to build a [choropleth map](choropleth-map.html):

- A **geospatial object** providing region boundaries (city districts of the south of France in this example). Data are available at the geoJSON format [here](https://github.com/gregoiredavid/france-geojson), and this post explains in detail how to read and represent geoJSON format with R.
- A **numeric variable** that we use to color each geographical unit. Here we will use the number of restaurant per city. The data has been found [here](https://www.insee.fr/fr/accueil).









# Find and download a `.geoJSON` file
***


This step has been extensively describe in [chart #325](325-background-map-from-geojson-format-in-r.html). The `sf` library allows to read this type of format in R. For plotting it with [ggplot2](ggplot2-package.html), the `geom_sf()` function allows to **represent this type of object**.

```{r thecode, cache=TRUE}
# Geospatial data available at the geojson format
tmp_geojson <- tempfile(fileext = ".geojson")

download.file(
  "https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",
  tmp_geojson
)
library(sf)
my_sf <- read_sf(tmp_geojson)

# Since it is a bit too much data, I select only a subset of it:
my_sf <- my_sf[substr(my_sf$code, 1, 2) %in% c(
  "06", "83",
  "13", "30", "34", "11", "66"
), ]
```


```{r thecode, eval=FALSE}
```











# Basic background map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We now have a geospatial object called `my_sf`. This object could be plotted as is using the `plot()` function as explained [here](choropleth-map-in-r.html).

On [ggplot2](ggplot2-package.html) we can use `geom_sf()` to plot the shape.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode10, echo=FALSE, out.width = "100%", fig.height=4}
library(ggplot2)
ggplot(my_sf) +
  geom_sf(fill = "white", color = "black", linewidth = 0.3) +
  theme_void()
```

</div>
</div>

```{r thecode10, eval=FALSE}
```













# Read the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The number of restaurant per city district has been found on the internet and a clean version is stored on the gallery website. It is thus easy to read it with `read.table()`. Before doing a [choropleth map](choropleth-map.html), it is a good practice to check the distribution of your variable.

Here, we have a *long tail* distribution: a few cities have a lot of restaurant. Thus we will probably need to apply a **log scale to our color palette**. It will avoid that all the variation is absorbed by these high values.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=4}
# read data
data <- read.table("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/data_on_french_states.csv",
  header = T, sep = ";"
)

head(data)

# Distribution of the number of restaurant?
library(dplyr)
data %>%
  ggplot(aes(x = nb_equip)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "black") +
  scale_x_log10()
```

</div>
</div>

```{r thecode2, eval=FALSE}
```













# Merge geospatial and numeric data
***

This is a key step in [choropleth map](choropleth-map.html): your 2 inputs must have a id in common to make the link between them!


```{r thecode5, echo=TRUE, out.width = "100%", fig.height=4}
# Make the merge
my_sf_merged <- my_sf %>%
  left_join(data, by = c("code" = "depcom")) %>%
  # Note that if the number of restaurant is NA, it is in fact 0
  mutate(nb_equip = ifelse(is.na(nb_equip), 0.01, nb_equip))
```











# Read the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We can make a first basic choropleth map. We just need to add `fill = our_value` in the aesthetic of our polygons.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=4}
ggplot(my_sf_merged) +
  geom_sf(aes(fill = nb_equip)) +
  theme_void()
```

</div>
</div>

```{r thecode3, eval=FALSE}
```











# Customized choropleth map with R and ggplot2
***

We need to change the **color palette**, improve the **legend**, use a log **scale transformation** for the colorscale, change **background** and add **titles** and explanation.

Here is the code to do that, and the final result!

```{r thecode9, echo=TRUE, out.width = "100%", fig.height=4}
p <- ggplot(my_sf_merged) +
  geom_sf(aes(fill = nb_equip), linewidth = 0, alpha = 0.9) +
  theme_void() +
  scale_fill_viridis_c(
    trans = "log", breaks = c(1, 5, 10, 20, 50, 100),
    name = "Number of restaurant",
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(12, units = "mm"),
      label.position = "bottom",
      title.position = "top",
      nrow = 1
    )
  ) +
  labs(
    title = "South of France Restaurant concentration",
    subtitle = "Number of restaurant per city district",
    caption = "Data: INSEE | Creation: Yan Holtz | r-graph-gallery.com"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    ),
    plot.caption = element_text(
      size = 10,
      color = "#4e4d47",
      margin = margin(
        b = 0.3, r = -99, t = 0.3,
        unit = "cm"
      )
    ),
    legend.position = c(0.7, 0.09)
  )

p
```






# Going further
***

This post explains how to build a [choropleth map](choropleth-map.html) with R and [ggplot2](ggplot2-package.html).

You might be interested in how to create an [interactive choropleth map](183-choropleth-map-with-leaflet.html), and more generally in the [choropleth section](choropleth-map.html) of the gallery.








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
