---
title: "Choropleth map with R and ggplot2"
descriptionMeta: "This post describes how to build a choropleth map with R and the ggplot2 package. It shows how to load geospatial data in R, merge region features and build the map. Reproducible code is provided."
descriptionTop: "This post describes how to build a [choropleth map](choropleth-map.html) with `R` and the `ggplot2` package. It shows how to load geospatial data in R, merge region features and build the map. Reproducible code is provided."
sectionText: "Choropleth section"
sectionLink: "choropleth-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/choropleth.html"
url: "327-chloropleth-map-from-geojson-with-ggplot2"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">






Two inputs are needed to build a choropleth map:

- A geospatial object providing region boundaries (city districts of the south of France in this example). Data are available at the geoJSON format [here](https://github.com/gregoiredavid/france-geojson), and this post explains in detail how to read and represent geoJSON format with R.
- A numeric variable that we use to color each geographical unit. Here we will use the number of restaurant per city. The data has been found [here](https://www.insee.fr/fr/accueil). (Thanks [@joelgombin](https://twitter.com/joelgombin)).









# Find and download a `.geoJSON` file
***


This step has been extensively describe in [chart #325](325-background-map-from-geojson-format-in-r/). The geojsonio library allows to read this type of format in R. To plot it with ggplot2, we first need to transform it to a data frame using the tidy function of the broom library. Then, the geom_polygon() function allows to represent this type of object !

```{r thecode, cache=TRUE}
# Geospatial data available at the geojson format
library(geojsonio)
spdf <- geojson_read("https://raw.githubusercontent.com/gregoiredavid/france-geojson/master/communes.geojson",  what = "sp")

# Since it is a bit to much data, I select only a subset of it:
spdf <- spdf[ substr(spdf@data$code,1,2)  %in% c("06", "83", "13", "30", "34", "11", "66") , ]
```


```{r thecode, eval=FALSE}
```











# Basic background map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We now have a geospatial object called `spdf`. This object could be plotted as is using the `plot()` function as explained [here](choropleth-map-in-r.html).

However, an additionnal step is required to plot it with `ggplot2` that expects a data frame as input. It is possible to make the convertion using the `tidy` function of the `broom` package as shown below.

Finally, `geom_polygon` is used to plot the shape.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode10, echo=FALSE, out.width = "100%", fig.height=4}
#plot(spdf)

# I need to fortify the data AND keep trace of the commune code! (Takes ~2 minutes)
library(broom)
spdf_fortified <- tidy(spdf, region = "code")

# Now I can plot this shape easily as described before:
library(ggplot2)
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="white", color="grey") +
  theme_void() +
  coord_map()
```

</div>
</div>

```{r thecode10, eval=FALSE}
```













# Read the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The number of restaurant per city district has been found on the internet and a clean version is stored on the gallery website. It is thus easy to read it with read.table. Before doing a choropleth map, it is a good practice to check the distribution of your variable.

Here, we have a ‘long tail’ distribution: a few cities have a lot of restaurant. Thus we will probably need to apply a log scale to our color palette. It will avoid that all the variation is absorbed by these high values.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=4}
# read data
data <- read.table("https://www.r-graph-gallery.com/wp-content/uploads/2017/12/data_on_french_states.csv", header=T, sep=";")
head(data)

# Distribution of the number of restaurant?
library(dplyr)
data %>%
  ggplot( aes(x=nb_equip)) +
    geom_histogram(bins=20, fill='skyblue', color='#69b3a2') + scale_x_log10()
```

</div>
</div>

```{r thecode2, eval=FALSE}
```













# Merge geospatial and numeric data
***

This is a key step in choropleth map: your 2 inputs must have a id in common to make the link between them!


```{r thecode5, echo=FALSE, out.width = "100%", fig.height=4}
# Make the merge
spdf_fortified = spdf_fortified %>%
  left_join(. , data, by=c("id"="depcom"))

# Note that if the number of restaurant is NA, it is in fact 0
spdf_fortified$nb_equip[ is.na(spdf_fortified$nb_equip)] = 0.001
```











# Read the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We can make a first basic choropleth map. We just need to add fill = our value in the aesthetic of our polygons.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=4}
ggplot() +
  geom_polygon(data = spdf_fortified, aes(fill = nb_equip, x = long, y = lat, group = group)) +
  theme_void() +
  coord_map()
```

</div>
</div>

```{r thecode3, eval=FALSE}
```











# Customized choropleth map with R and ggplot2
***

There is a bit of work to do to get a descent figure. We need to change the color palette, improve the legend, use a log scale transformation for the colorscale, change background and add titles and explanation. Here is the code to do that, and the final result!


```{r thecode9, echo=true, out.width = "100%", fig.height=4}
library(viridis)
p <- ggplot() +
  geom_polygon(data = spdf_fortified, aes(fill = nb_equip, x = long, y = lat, group = group) , size=0, alpha=0.9) +
  theme_void() +
  scale_fill_viridis(trans = "log", breaks=c(1,5,10,20,50,100), name="Number of restaurant", guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) ) +
  labs(
    title = "South of France Restaurant concentration",
    subtitle = "Number of restaurant per city district",
    caption = "Data: INSEE | Creation: Yan Holtz | r-graph-gallery.com"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),

    plot.title = element_text(size= 22, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    plot.subtitle = element_text(size= 17, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.43, l = 2, unit = "cm")),
    plot.caption = element_text( size=12, color = "#4e4d47", margin = margin(b = 0.3, r=-99, unit = "cm") ),

    legend.position = c(0.7, 0.09)
  ) +
  coord_map()
p
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
