---
title: "Hexbin map in R: an example with US states"
descriptionMeta: "This post describes how to build a hexbin map, with input data available at geojson format. It provides a ste-by-step explanation and reproducible code."
descriptionTop: "This post describes how to build a [hexbin map](hexbin-map.html). It is based on a `geojson` file providing the boundaries of us states as hexagons."
sectionText: "Hexbin map section"
sectionLink: "hexbin-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/hexbinmap.html"
url: "328-hexbin-map-of-the-usa"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basic hexbin map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The first step is to build a basic [hexbin map](hexbin-map.html) of the US. Note that the gallery dedicates a [whole section](hexbin-map.html) to this kind of map.

Hexagones boundaries are provided [here](https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map). You have to download it at the `geojson` format and load it in R thanks to the `geojson_read()` function. You get a geospatial object that you can plot using the `plot()` function. This is widely explained in the [background map](map.html) section of the gallery.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# library
library(tidyverse)
library(geojsonio)
library(RColorBrewer)
library(rgdal)

# Download the Hexagones boundaries at geojson format here: https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map.

# Load this file. (Note: I stored in a folder called DATA)
spdf <- geojson_read("DATA/us_states_hexgrid.geojson.json",  what = "sp")

# Bit of reformating
spdf@data = spdf@data %>%
  mutate(google_name = gsub(" \\(United States\\)", "", google_name))

# Show it
plot(spdf)
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# `ggplot2` and state name
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally doable to plot this geospatial object using `ggplot2` and its `geom_polygon()` function, but we first need to `fortify` it using the `broom` package.

Moreover, the `rgeos` package is used here to compute the centroid of each region thanks to the `gCentroid` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=4}

# I need to 'fortify' the data to be able to show it with ggplot2 (we need a data frame format)
library(broom)
spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")

# Calculate the centroid of each hexagon to add the label:
library(rgeos)
centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))
 
# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="skyblue", color="white") +
  geom_text(data=centers, aes(x=x, y=y, label=id)) +
  theme_void() +
  coord_map()
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










# Basic choropleth
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Now you probably want to adjust the color of each hexagon, according to the value of a specific variable (we call it a [choropleth map](chloropleth-map.html)). 

In this post I suggest to represent the number of wedding per thousand people. The data have been found [here](https://www.cdc.gov/nchs/nvss/marriage-divorce.html), and stored on a clean format [here](https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/State_mariage_rate.csv). 

Let’s start by loading this information and represent its distribution:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Load mariage data
data <- read.table("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/State_mariage_rate.csv", header=T, sep=",", na.strings="---")
 
# Distribution of the marriage rate?
data %>% 
  ggplot( aes(x=y_2015)) + 
    geom_histogram(bins=20, fill='#69b3a2', color='white') + 
    scale_x_continuous(breaks = seq(1,30))
```
</div>
</div>



```{r thecode3, eval=FALSE}
```








<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Most of the state have between 5 and 10 weddings per 1000 inhabitants, but there are 2 outliers with high values (16 and 32).

Let’s represent this information on a map. We have a column with the state id in both the geospatial and the numerical datasets. So we can merge both information and plot it.

Note the use of the `trans = "log"` option in the color scale to decrease the effect of the 2 outliers.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Merge geospatial and numerical information
spdf_fortified <- spdf_fortified %>%
  left_join(. , data, by=c("id"="state")) 
 
# Make a first chloropleth map
ggplot() +
  geom_polygon(data = spdf_fortified, aes(fill =  y_2015, x = long, y = lat, group = group)) +
  scale_fill_gradient(trans = "log") +
  theme_void() +
  coord_map()
```
</div>
</div>



```{r thecode4, eval=FALSE}
```

















# Customized hexbin choropleth map
***

Here is a final version after applying a few customization:

- Use handmade binning for the colorscale with `scale_fill_manual`
- Use `viridis` for the color palette
- Add custom legend and title
- Change background color


```{r thecode5, echo=FALSE, out.width = "80%", fig.height=4, fig.align="center"}
# Prepare binning
spdf_fortified$bin <- cut( spdf_fortified$y_2015 , breaks=c(seq(5,10), Inf), labels=c("5-6", "6-7", "7-8", "8-9", "9-10", "10+" ), include.lowest = TRUE )
 
# Prepare a color scale coming from the viridis color palette
library(viridis)
my_palette <- rev(magma(8))[c(-1,-8)]
 
# plot
ggplot() +
  geom_polygon(data = spdf_fortified, aes(fill = bin, x = long, y = lat, group = group) , size=0, alpha=0.9) +
  geom_text(data=centers, aes(x=x, y=y, label=id), color="white", size=3, alpha=0.6) +
  theme_void() +
  scale_fill_manual( 
    values=my_palette, 
    name="Wedding per 1000 people in 2015", 
    guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) 
  ) +
  ggtitle( "A map of marriage rates, state by state" ) +
  theme(
    legend.position = c(0.5, 0.9),
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(size= 22, hjust=0.5, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
  )
```



```{r thecode5, eval=FALSE}
```



















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
