---
title: "Hexbin map in R: an example with US states"
descriptionMeta: "This post describes how to build a hexbin map, with input data available at geojson format. It provides a ste-by-step explanation and reproducible code."
descriptionTop: "This post describes how to build a [hexbin map](hexbin-map.html). It is based on a `geojson` file providing the boundaries of us states as hexagons."
sectionText: "Hexbin map section"
sectionLink: "hexbin-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/hexbinmap.html"
url: "328-hexbin-map-of-the-usa"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basic hexbin map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The first step is to build a basic [hexbin map](hexbin-map.html) of the US. Note that the gallery dedicates a [whole section](hexbin-map.html) to this kind of map.

Hexagones boundaries are provided [here](https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map). You have to download it at the `geojson` format and load it in R thanks to the `geojson_read()` function. You get a geospatial object that you can plot using the `plot()` function. This is widely explained in the [background map](map.html) section of the gallery.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(tidyverse)
library(geojsonio)
library(RColorBrewer)
library(rgdal)

# Download the Hexagones boundaries at geojson format here: https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map.

# Load this file. (Note: I stored in a folder called DATA)
spdf <- geojson_read("DATA/us_states_hexgrid.geojson.json",  what = "sp")

# Bit of reformating
spdf@data = spdf@data %>%
  mutate(google_name = gsub(" \\(United States\\)", "", google_name))

# Show it
plot(spdf)
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# `ggplot2` and state name
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally doable to plot this geospatial object using `ggplot2` and its `geom_polygon()` function, but we first need to `fortify` it using the `broom` package.

Moreover, the `rgeos` package is used here to compute the centroid of each region thanks to the `gCentroid` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}

# I need to 'fortify' the data to be able to show it with ggplot2 (we need a data frame format)
library(broom)
spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")

# Calculate the centroid of each hexagon to add the label:
library(rgeos)
centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))
 
# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="skyblue", color="white") +
  geom_text(data=centers, aes(x=x, y=y, label=id)) +
  theme_void() +
  coord_map()
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










# Basic choropleth
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Now you probably want to adjust the color of each hexagon, according to the value of a specific variable (we call it a [choropleth map](chloropleth-map.html)). 

In this post I propose to represent the number of wedding per thousand people. The data have been found here, and stored in the gallery. Let’s start by loading this information and represent its distribution:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}

# I need to 'fortify' the data to be able to show it with ggplot2 (we need a data frame format)
library(broom)
spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")

# Calculate the centroid of each hexagon to add the label:
library(rgeos)
centers <- cbind.data.frame(data.frame(gCentroid(spdf, byid=TRUE), id=spdf@data$iso3166_2))
 
# Now I can plot this shape easily as described before:
ggplot() +
  geom_polygon(data = spdf_fortified, aes( x = long, y = lat, group = group), fill="skyblue", color="white") +
  geom_text(data=centers, aes(x=x, y=y, label=id)) +
  theme_void() +
  coord_map()
```
</div>
</div>



```{r thecode2, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
