---
title: "Hexbin map in R: an example with US states"
descriptionMeta: "This post describes how to build a hexbin map, with input data available at geojson format. It provides a step-by-step explanation and reproducible code."
descriptionTop: "This post describes how to build a [hexbin map](hexbin-map.html). It is based on a `geojson` file providing the **boundaries of US states** as hexagons."
sectionText: "Hexbin map section"
sectionLink: "hexbin-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/hexbinmap.html"
url: "328-hexbin-map-of-the-usa"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container">




# Basic hexbin map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The first step is to build a basic [hexbin map](hexbin-map.html) of the US. Note that the gallery dedicates a [whole section](hexbin-map.html) to this kind of map.

Hexagon boundaries are provided [here](https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map). You have to download it at the `geojson` format and load it in R thanks to the `st_read() / read_sf()` function. You get a geospatial object that you can plot using the `plot()` function. This is widely explained in the [background map](map.html) section of the gallery.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# library
library(tidyverse)
library(sf)
library(RColorBrewer)

# Download the Hexagon boundaries at geojson format here: https://team.carto.com/u/andrew/tables/andrew.us_states_hexgrid/public/map.

# Load this file. (Note: I stored in a folder called DATA)
my_sf <- read_sf("DATA/us_states_hexgrid.geojson.json")

# Bit of reformatting
my_sf <- my_sf %>%
  mutate(google_name = gsub(" \\(United States\\)", "", google_name))

# Show it
plot(st_geometry(my_sf))
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# `ggplot2` and state name
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally doable to plot this geospatial object using `ggplot2` and its `geom_sf()` and `geom_sf_text()` functions.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=4}
ggplot(my_sf) +
  geom_sf(fill = "skyblue", color = "white") +
  geom_sf_text(aes(label = iso3166_2)) +
  theme_void()
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










# Basic choropleth
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Now you probably want to **adjust the color of each hexagone**, according to the value of a specific variable (we call it a [choropleth map](chloropleth-map.html)). 

The data is about the number of wedding per thousand people, and you can **load it directly** using the following code.

Let's start by loading this information and represent its distribution:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Load marriage data
data <- read.table("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/State_mariage_rate.csv",
  header = T, sep = ",", na.strings = "---"
)

# Distribution of the marriage rate?
data %>%
  ggplot(aes(x = y_2015)) +
  geom_histogram(bins = 20, fill = "#69b3a2", color = "white") +
  scale_x_continuous(breaks = seq(1, 30))
```
</div>
</div>



```{r thecode3, eval=FALSE}
```








<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Most of the state have between 5 and 10 weddings per 1000 inhabitants, but there are **2 outliers** with high values (16 and 32).

Note the use of the `trans = "log"` option in the color scale to decrease the effect of the 2 outliers.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=4}
# Merge geospatial and numerical information
my_sf_wed <- my_sf %>%
  left_join(data, by = c("google_name" = "state"))

# Make a first choropleth map
ggplot(my_sf_wed) +
  geom_sf(aes(fill = y_2015)) +
  scale_fill_gradient(trans = "log") +
  theme_void()
```
</div>
</div>



```{r thecode4, eval=FALSE}
```

















# Customized hexbin choropleth map
***

Here is a final version after applying a few customization:

- Use **handmade binning** for the colorscale with `scale_fill_manual`
- Use `viridis` for the color palette
- Add custom **legend and title**
- Change **background color**


```{r thecode5, echo=FALSE, out.width = "80%", fig.height=4, fig.align="center"}
# Prepare binning
my_sf_wed$bin <- cut(my_sf_wed$y_2015,
  breaks = c(seq(5, 10), Inf),
  labels = c("5-6", "6-7", "7-8", "8-9", "9-10", "10+"),
  include.lowest = TRUE
)

# Prepare a color scale coming from the viridis color palette
library(viridis)
my_palette <- rev(magma(8))[c(-1, -8)]

# plot
ggplot(my_sf_wed) +
  geom_sf(aes(fill = bin), linewidth = 0, alpha = 0.9) +
  geom_sf_text(aes(label = iso3166_2), color = "white", size = 3, alpha = 0.6) +
  theme_void() +
  scale_fill_manual(
    values = my_palette,
    name = "Wedding per 1000 people in 2015",
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(12, units = "mm"),
      label.position = "bottom", title.position = "top", nrow = 1
    )
  ) +
  ggtitle("A map of marriage rates, state by state") +
  theme(
    legend.position = c(0.5, 0.9),
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 22, hjust = 0.5, color = "#4e4d47",
      margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")
    )
  )
```



```{r thecode5, eval=FALSE}
```










# Going further
***


This post explains how to build a hexbin map of the US.

You might be interested in how to [combine hexbin map and cartogram](332-hexbin-chloropleth-cartogram.html), and more generally by the [hexbin map section](hexbin-map.html) of the gallery.








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
