---
title: "Hexbin map for 2D density visualization"
descriptionMeta: "It is possible to apply 2d density visualization methods on map to study the geographical distribution of a variable. The two most famous techniques beeing Hexbin maps and 2d histogram maps. This post describes how to build it with `R` and `ggplot2`."
descriptionTop: "It is possible to apply [2d density](2d-density-chart.html) visualization methods on map to study the geographical distribution of a variable. The two most famous techniques beeing [Hexbin maps](hexbin-map.html) and 2d histogram maps. This post describes how to build it with `R` and `ggplot2`."
sectionText: "Hexbin map section"
sectionLink: "hexbin-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/hexbinmap.html"
url: "329-hexbin-map-for-distribution"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">


Several techniques exist to visualize the distribution of two variables in the same time ([2D distribution](2d-density-chart.html)). The R graph gallery dedicates a whole [section](2d-density-chart.html) to it.

The idea is to count the number of observation within a particular area of the 2D space and represent this count by a color. This can method can be applied to maps using hexagones or squares, resulting in hexbin maps and 2d histogram maps.

In this post a list of [GPS coordinates](https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/17_ListGPSCoordinates.csv) is used as input data. It comes from a [project](https://www.data-to-viz.com/story/GPSCoordWithoutValue.html) that harvested and geocoded a list of 200k tweets talking about [#surf](https://www.data-to-viz.com/story/GPSCoordWithoutValue.html).














# 2d histogram maps
***

For 2d histogram maps the globe is split in several squares, the number of tweet per square is counted, and a color is attributed to each square. 

- `ggplot2` offers the `geom_bin2d()` function that does all the calculation for us and plot the squares.
- the `geom_polygon()` function is used to show the world map in the background.
- `annotate()` is used to add a title.
- the `guide` option of `scale_fill_viridis` is used to create a nice legend.


```{r thecode, echo=FALSE, out.width = "100%", fig.height=4, cache=TRUE}
# Libraries
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(mapdata)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/17_ListGPSCoordinates.csv", sep=",", header=T)

# Get the world polygon
world <- map_data("world")

# plot
ggplot(data, aes(x=homelon, y=homelat)) + 
    geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
    geom_bin2d(bins=100) +
    ggplot2::annotate("text", x = 175, y = 80, label="Where people tweet about #Surf", colour = "black", size=4, alpha=1, hjust=1) +
    ggplot2::annotate("segment", x = 100, xend = 175, y = 73, yend = 73, colour = "black", size=0.2, alpha=1) +
    theme_void() +
    ylim(-70, 80) +
    scale_fill_viridis(
      trans = "log", 
      breaks = c(1,7,54,403,3000),
      name="Tweet # recorded in 8 months", 
      guide = guide_legend( keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) 
    )  +
    ggtitle( "" ) +
    theme(
      legend.position = c(0.8, 0.09),
      legend.title=element_text(color="black", size=8),
      text = element_text(color = "#22211d"),
      plot.title = element_text(size= 13, hjust=0.1, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    ) 
```


```{r thecode, eval=FALSE}
```
















# Hexbin maps
***

Hexbin maps are done using pretty much the same code. 

Here, `geom_hex()` is used to create the hexagones. Note the `bins` option allowing to control the bin size, and thus the hexagone size on the map.

```{r thecode2, echo=FALSE, out.width = "80%", fig.height=5, fig.align="center", cache=TRUE}
# Libraries
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(mapdata)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/17_ListGPSCoordinates.csv", sep=",", header=T)

# plot
data %>%
  filter(homecontinent=='Europe') %>%
  ggplot( aes(x=homelon, y=homelat)) + 
    geom_hex(bins=59) +
    ggplot2::annotate("text", x = -27, y = 72, label="Where people tweet about #Surf", colour = "black", size=5, alpha=1, hjust=0) +
    ggplot2::annotate("segment", x = -27, xend = 10, y = 70, yend = 70, colour = "black", size=0.2, alpha=1) +
    theme_void() +
    xlim(-30, 70) +
    ylim(24, 72) +
    scale_fill_viridis(
      option="B",
      trans = "log", 
      breaks = c(1,7,54,403,3000),
      name="Tweet # recorded in 8 months", 
      guide = guide_legend( keyheight = unit(2.5, units = "mm"), keywidth=unit(10, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) 
    )  +
    ggtitle( "" ) +
    theme(
      legend.position = c(0.8, 0.09),
      legend.title=element_text(color="black", size=8),
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_rect(fill = "#f5f5f2", color = NA),
      plot.title = element_text(size= 13, hjust=0.1, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    ) 

```


```{r thecode2, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
