---
title: "Basic Cartogram with R"
descriptionMeta: "This post descriibes how to build a very simple cartogram with R and the cartogram package. It is a step a step approach leading to a choropleth map with distorted region size."
descriptionTop: "This post descriibes how to build a very simple [cartogram](cartogram.html) with `R` and the `cartogram` package. It is a step a step approach leading to a choropleth map with distorted region size."
sectionText: "Cartogram section"
sectionLink: "cartogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/cartogram.html"
url: "331-basic-cartogram"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = "ragg_png", dpi = 300)
```

<div class="container">




# Very basic map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [cartogram](cartogram.html) is a  map in which the geometry of regions is **distorted** in order to convey the information of an alternate variable. In this post, we are going to draw a map of Africa where the size of each country is **distorted proportionally to its population**.

First of all, you need to understand what a geospatial object is, and how to plot it with `R`. See the [background map section](map.html) of the gallery!

Let's get a geospatial object from a shape file available [here](http://thematicmapping.org/downloads/world_borders.php). This step is extensively described in [this post](168-load-a-shape-file-into-r.html) in case you're not familiar with it.


We can plot the boundaries using the `plot()` function:

</div>


<div class = "col-md-6  col-sm-12">



```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Get the shape file of Africa, see how on
# https://r-graph-gallery.com/168-load-a-shape-file-into-r.html

# I stored the data on a DATA folder and read it from there
library(sf)
wrld_simpl <- read_sf("DATA/world_shape_file/TM_WORLD_BORDERS_SIMPL-0.3.shp")
afr <- wrld_simpl[wrld_simpl$REGION == 2, ]

# We can visualize the region's boundaries with the plot function
plot(st_geometry(afr))
```
</div>
</div>



```{r thecode, eval=FALSE}
```













 
# Distort country size with `cartogram` {#carto}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The geospatial object has several information for each region. It notably gives the population of each country in 2005. 

We can thus use the `cartogram` library to **distort the size of each country**, proportionally to this column. The new geospatial object we get can be draw with the same `plot` function!

See how Nigeria is now bigger?


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# We work with the cartogram library
library(cartogram)

# construct a cartogram using the population in 2005

# need first to "change" the projection to Mercator (AKA Google Maps): EPSG: 3857
afr_merc <- st_transform(afr, 3857)

afr_cartogram <- cartogram_cont(afr_merc, "POP2005", itermax = 5)

# And back to the previous projection
afr_cartogram <- st_transform(afr_cartogram, st_crs(afr))

# This is a new geospatial object, we can visualise it!
plot(st_geometry(afr_cartogram))
```
</div>
</div>



```{r thecode2, eval=FALSE}
```










# Cartogram and choropleth {#choro}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Cartogram are very often used in combination with [choropleth map](https://www.r-graph-gallery.com/chloropleth-map.html). Since we have a spatial object and a numeric value associated with each region, it is possible to color each region according to its value. 

Let's use [ggplot2](ggplot2-package.html) to add some **color**, **title**, **legend**, background and so on. We now have a nice cartogram choropleth map of Africa!

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# It is a new geospatial object: we can use all the usual techniques on it! Let's start with a basic ggplot2 choropleth map:
library(tidyverse)
ggplot(afr_cartogram) +
  geom_sf(aes(fill = POP2005), linewidth = 0, alpha = 0.9) +
  theme_void()
```
</div>
</div>



```{r thecode3, eval=FALSE}
```











# Make it pretty  {#clean}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Same thing with a tiny bit of **customization**. See more explanation in the [choropleth](choropleth.html) an [ggplot2](ggplot2-package.html) sections.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# As seen before, we can do better with a bit of customization
ggplot(afr_cartogram) +
  geom_sf(aes(fill = POP2005 / 1000000), linewidth = 0, alpha = 0.9) +
  theme_void() +
  scale_fill_viridis_c(
    name = "Population (M)",
    breaks = c(1, 50, 100, 140),
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(12, units = "mm"),
      label.position = "bottom",
      title.position = "top", nrow = 1
    )
  ) +
  labs(title = "Africa 2005 Population") +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f4", color = NA),
    panel.background = element_rect(fill = "#f5f5f4", color = NA),
    legend.background = element_rect(fill = "#f5f5f4", color = NA),
    plot.title = element_text(
      size = 22, hjust = 0.5,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    legend.position = c(0.2, 0.26)
  )
```
</div>
</div>



```{r thecode4, eval=FALSE}
```


# Going further
***

This post explains how to build a very simple cartogram with `R` and the `cartogram` package.

You might be interested by how to [combine cartogram and hexbin map](332-hexbin-chloropleth-cartogram.html), and more generally by the [cartogram section](cartogram.html) of the gallery.


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
