---
title: "Dendrogram customization with R and ggraph"
descriptionMeta: "This post follows the previous introduction to ggraph and dendrogram. It shows how to customize the dendrogram: layout, edge style, node features and more."
descriptionTop: "This post follows the previous [introduction](334-basic-dendrogram-with-ggraph.html) to `ggraph` and dendrogram. It shows how to customize the dendrogram: layout, edge style, node features and more."
sectionText: "Dendrogram section"
sectionLink: "dendrogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/dendrogram.html"
url: "335-custom-ggraph-dendrogram"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px;">




Start by creating a dataset and a graph object using the `igraph` package.

```{r}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
theme_set(theme_void())
 
# data: edge list
d1 <- data.frame(from="origin", to=paste("group", seq(1,7), sep=""))
d2 <- data.frame(from=rep(d1$to, each=7), to=paste("subgroup", seq(1,49), sep="_"))
edges <- rbind(d1, d2)
 
# We can add a second data frame with information for each node!
name <- unique(c(as.character(edges$from), as.character(edges$to)))
vertices <- data.frame(
  name=name,
  group=c( rep(NA,8) ,  rep( paste("group", seq(1,7), sep=""), each=7)),
  cluster=sample(letters[1:4], length(name), replace=T),
  value=sample(seq(10,30), length(name), replace=T)
)
 
# Create a graph object
mygraph <- graph_from_data_frame( edges, vertices=vertices)
````




# Circular or linear `layout`
***

First of all, you can use a linear or a circular representation using the circular option thanks to the `layout` argument of `ggraph`. 

<u>Note</u>: a customized version of the circular dendrogram is available [here](339-circular-dendrogram-with-ggraph.html), with more node features and labels.

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Left
ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_diagonal() 
```
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode1, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Right
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_edge_diagonal()
```


</div>
</div>



```{r thecode, eval=FALSE}
```


```{r thecode1, eval=FALSE}
```












# Edge style
***

Then you can choose between different styles for your edges. The `ggraph` package comes with 2 main functions: `geom_edge_link` and `geom_edge_diagonal`.

Note that the most usual "elbow" representation is not implemented for hierarchical data yet.

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode2, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Left
ggraph(mygraph, layout = 'dendrogram') + 
  geom_edge_link()
```
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Right
ggraph(mygraph, layout = 'dendrogram') + 
  geom_edge_diagonal()
```


</div>
</div>



```{r thecode2, eval=FALSE}
```


```{r thecode3, eval=FALSE}
```













# Labels and Nodes
***

You probably want to add labels to give more insight to your tree. And eventually nodes. This can be done using `the geom_node_text` and `geom_node_point` respectively. 

<u>Note</u>: the label addition is a bit more tricky for circular dendrogram, a solution is suggested in [graph #339](339-circular-dendrogram-with-ggraph.html).

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode4, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Left
ggraph(mygraph, layout = 'dendrogram') + 
  geom_edge_diagonal() +
  geom_node_text(aes( label=name, filter=leaf) , angle=90 , hjust=1, nudge_y = -0.01) +
  ylim(-.4, NA)
```
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "80%", fig.height=7, fig.align="center"}
# Right
ggraph(mygraph, layout = 'dendrogram') + 
  geom_edge_diagonal() +
  geom_node_text(aes( label=name, filter=leaf) , angle=90 , hjust=1, nudge_y = -0.04) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) +
  ylim(-.5, NA)
```


</div>
</div>



```{r thecode4, eval=FALSE}
```


```{r thecode5, eval=FALSE}
```






# Customize aesthetics
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common task to add color or shapes to your dendrogram. It allows to show more clearly the organization of the dataset.

`ggraph` works the same way as `ggplot2`. In the aesthetics part of each component, you can use a column of your initial data frame to be mapped to a shape, a color, a size or other..

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode8, echo=FALSE, out.width = "100%", fig.height=7}
ggraph(mygraph, layout = 'dendrogram') + 
  geom_edge_diagonal() +
  geom_node_text(aes( label=name, filter=leaf, color=group) , angle=90 , hjust=1, nudge_y=-0.1) +
  geom_node_point(aes(filter=leaf, size=value, color=group) , alpha=0.6) +
  ylim(-.6, NA) +
  theme(legend.position="none")
```
</div>
</div>



```{r thecode8, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
