---
title: "Interactive dendrogram with R"
descriptionMeta: "This post describes how to use the CollapsibleTree package to build an interactive tree diagram. Explanation and reproducible code provided"
descriptionTop: "This post describes how to use the `CollapsibleTree` package to build an interactive [tree diagram](dendrogram.html). Explanation and reproducible code provided"
sectionText: "Dendrogram section"
sectionLink: "dendrogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/dendrogram.html"
url: "336-interactive-dendrogram-with-collapsibletree"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 50px">



<center><iframe src="HtmlWidget/dendrogram_interactive.html" height="640" width="100%" style="border:none;"></iframe></center>


The `collapsibletree` package is the best option to build interactive [dendrogram](dendrogram.html) with R.

On the chart above, you can click a node to reveal the following branch that is currently collapsed.

The input must be a data frame that stores the hierarchical information. Basically, each row describes a complete path from the root to the leaf. In this example, the `warpbreaks` dataset has 3 columns: `wool`, `tension` and `breaks`.

```{r, eval=FALSE}
# Load library
# install.packages("collapsibleTree")
library(collapsibleTree) 
 
# input data must be a nested data frame:
head(warpbreaks)
 
# Represent this tree:
p <- collapsibleTree( warpbreaks, c("wool", "tension", "breaks"))
p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/dendrogram_interactive.html"))
```





</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
