---
title: "Circular dendrogram with R and ggraph"
descriptionMeta: "This post shows how to build a customized circular dendrogram using R and the ggraph package. It provides explanation and reproducible code."
descriptionTop: "This post shows how to build a customized circular [dendrogram](dendrogram.html) using R and the `ggraph` package. It provides explanation and reproducible code."
sectionText: "Dendrogram section"
sectionLink: "dendrogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/dendrogram.html"
url: "339-circular-dendrogram-with-ggraph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px;">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The circular dendrogram of the `ggraph` library deserves its own page because it can be a bit tricky to adjust the labels. Indeed they need to have a good angle, be flipped upside down on the left part of the chart, and their alignment needs to be adjusted as well.

The following piece of code should allow you to display them correctly as shown beside. Note that the [graph #334](334-basic-dendrogram-with-ggraph.html) gives the basis of dendrogram with ggraph, and that [graph #335](335-custom-ggraph-dendrogram.html) presents a few possible customizations.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode8, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer) 
# create a data frame giving the hierarchical structure of your individuals
d1=data.frame(from="origin", to=paste("group", seq(1,10), sep=""))
d2=data.frame(from=rep(d1$to, each=10), to=paste("subgroup", seq(1,100), sep="_"))
edges=rbind(d1, d2)
 
# create a vertices data.frame. One line per object of our hierarchy
vertices = data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) , 
  value = runif(111)
) 
# Let's add a column with the group of each name. It will be useful later to color points
vertices$group = edges$from[ match( vertices$name, edges$to ) ]
 
 
#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels
vertices$id=NA
myleaves=which(is.na( match(vertices$name, edges$from) ))
nleaves=length(myleaves)
vertices$id[ myleaves ] = seq(1:nleaves)
vertices$angle= 90 - 360 * vertices$id / nleaves
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices$hjust<-ifelse( vertices$angle < -90, 1, 0)
 
# flip angle BY to make them readable
vertices$angle<-ifelse(vertices$angle < -90, vertices$angle+180, vertices$angle)
 
# Create a graph object
mygraph <- graph_from_data_frame( edges, vertices=vertices )
 
# Make the plot
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_edge_diagonal(colour="grey") +
  scale_edge_colour_distiller(palette = "RdPu") +
  geom_node_text(aes(x = x*1.15, y=y*1.15, filter = leaf, label=name, angle = angle, hjust=hjust, colour=group), size=2.7, alpha=1) +
  geom_node_point(aes(filter = leaf, x = x*1.07, y=y*1.07, colour=group, size=value, alpha=0.2)) +
  scale_colour_manual(values= rep( brewer.pal(9,"Paired") , 30)) +
  scale_size_continuous( range = c(0.1,10) ) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.3, 1.3), y = c(-1.3, 1.3))
```
</div>
</div>



```{r thecode8, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
