---
title: "Customized dendrogram with R and the dendextend package"
descriptionMeta: "Dendextend allows to reach the next step in term of dendrogram. Check many examples with explanation and reproducible code."
descriptionTop: "The `dendextend` package allows to apply all kinds of customization to a dendrogram: coloring nodes, labels, putting several tree face to face and more."
sectionText: "Dendrogram section"
sectionLink: "dendrogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/dendrogram.html"
url: "340-custom-your-dendrogram-with-dendextend"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Basic dendrogram
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

First of all, let's remind how to build a basic [dendrogram](dendrogram.html) with `R`:

- input dataset is a dataframe with individuals in row, and features in column
- `dist()` is used to compute distance between sample
- `hclust()` performs the hierarchical clustering
- the `plot()` function can plot the output directly as a tree



</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(tidyverse)
 
# Data
head(mtcars)
 
# Clusterisation using 3 variables
mtcars %>% 
  select(mpg, cyl, disp) %>% 
  dist() %>% 
  hclust() %>% 
  as.dendrogram() -> dend
 
# Plot
par(mar=c(7,3,1,1))  # Increase bottom margin to have the complete label
plot(dend)
```
</div>
</div>



```{r thecode, eval=FALSE}
```









# The `set()` function
***

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

```{r thecodeA, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(dendextend)

# Chart (left)
dend %>% 
  # Custom branches
  set("branches_col", "grey") %>% set("branches_lwd", 3) %>%
  # Custom labels
  set("labels_col", "orange") %>% set("labels_cex", 0.8) %>%
  plot()
```

</div>


<div class = "col-md-4  col-sm-12 align-self-center">

```{r thecodeB, echo=FALSE, out.width = "100%", fig.height=7}
# Middle
dend %>% 
  set("nodes_pch", 19)  %>% 
  set("nodes_cex", 0.7) %>% 
  set("nodes_col", "orange") %>% 
  plot()
```

</div>


<div class = "col-md-4  col-sm-12 align-self-center">

```{r thecodeC, echo=FALSE, out.width = "100%", fig.height=7}
# right
dend %>% 
  set("leaves_pch", 19)  %>% 
  set("leaves_cex", 0.7) %>% 
  set("leaves_col", "skyblue") %>% 
  plot()

```

</div>

</div>

The `set()` function of `dendextend` allows to modify the attribute of a specific part of the tree. 

You can customize the `cex`, `lwd`, `col`, `lty` for `branches` and `labels` for example. You can also custom the nodes or the leaf. The code below illustrates this concept:


```{r thecodeA, eval=FALSE}
```
```{r thecodeB, eval=FALSE}
```
```{r thecodeC, eval=FALSE}
```


















# Highlight clusters
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode2A, echo=FALSE, out.width = "100%", fig.height=7}
# Color in function of the cluster
par(mar=c(1,1,1,7))
dend %>%
  set("labels_col", value = c("skyblue", "orange", "grey"), k=3) %>%
  set("branches_k_color", value = c("skyblue", "orange", "grey"), k = 3) %>%
  plot(horiz=TRUE, axes=FALSE)
abline(v = 350, lty = 2)
```

</div>


<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode2B, echo=FALSE, out.width = "100%", fig.height=7}
# Highlight a cluster with rectangle
par(mar=c(9,1,1,1))
dend %>%
  set("labels_col", value = c("skyblue", "orange", "grey"), k=3) %>%
  set("branches_k_color", value = c("skyblue", "orange", "grey"), k = 3) %>%
  plot(axes=FALSE)
rect.dendrogram( dend, k=3, lty = 5, lwd = 0, x=1, col=rgb(0.1, 0.2, 0.4, 0.1) ) 

```

</div>



</div>

The `dendextend` library has some good functionalities to highlight the tree clusters. 

You can color branches and label following their cluster attribution, specifying the number of cluster you want. The `rect.dendrogram()` function even allows to highlight one or several specific clusters with a rectangle.


```{r thecode2A, eval=FALSE}
```
```{r thecode2B, eval=FALSE}
```














# Comparing with an expected clustering
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common task to compare the cluster you get with an expected distribution. 

In the `mtcars` dataset we used to build our dendrogram, there is an `am` column that is a binary variable. We can check if this variable is consistent with the cluster we got using the `colored_bars()` function.


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Create a vector of colors, darkgreen if am is 0, green if 1.
my_colors <- ifelse(mtcars$am==0, "forestgreen", "green")
 
# Make the dendrogram
par(mar=c(10,1,1,1))
dend %>%
  set("labels_col", value = c("skyblue", "orange", "grey"), k=3) %>%
  set("branches_k_color", value = c("skyblue", "orange", "grey"), k = 3) %>%
  set("leaves_pch", 19)  %>% 
  set("nodes_cex", 0.7) %>% 
  plot(axes=FALSE)
 
# Add the colored bar
colored_bars(colors = my_colors, dend = dend, rowLabels = "am")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```
















# Comparing 2 dendrograms with `tanglegram()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is possible to compare 2 dendrograms using the `tanglegram()` function.

Here it illustrates a very important concept: when you calculate your distance matrix and when you run your hierarchical clustering algorithm, you cannot simply use the default options without thinking about what you’re doing. Have a look to the differences between 2 different methods of clusterisation.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Make 2 dendrograms, using 2 different clustering methods
d1 <- USArrests %>% dist() %>% hclust( method="average" ) %>% as.dendrogram()
d2 <- USArrests %>% dist() %>% hclust( method="complete" ) %>% as.dendrogram()
 
# Custom these kendo, and place them in a list
dl <- dendlist(
  d1 %>% 
    set("labels_col", value = c("skyblue", "orange", "grey"), k=3) %>%
    set("branches_lty", 1) %>%
    set("branches_k_color", value = c("skyblue", "orange", "grey"), k = 3),
  d2 %>% 
    set("labels_col", value = c("skyblue", "orange", "grey"), k=3) %>%
    set("branches_lty", 1) %>%
    set("branches_k_color", value = c("skyblue", "orange", "grey"), k = 3)
)
 
# Plot them together
tanglegram(dl, 
           common_subtrees_color_lines = FALSE, highlight_distinct_edges  = TRUE, highlight_branches_lwd=FALSE, 
           margin_inner=7,
           lwd=2
)
```
</div>
</div>



```{r thecode4, eval=FALSE}
```














<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
