---
title: "Stacked barplot with negative values with ggplot2"
descriptionMeta: "This post explains how to build a stacked barplot with R and ggplot2. It considers a dataset that includes negative values to see how the package behave."
descriptionTop: "This post explains how to build a stacked barplot with R and [ggplot2](https://www.r-graph-gallery.com/ggplot2-package.html). It focus on a dataset that includes negative values"
sectionText: "Barchart section"
sectionLink: "barplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "341-stacked-barplot-with-negative-values"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Dataset
***

The dataset used in this post has 5 columns:

  - the first column is called `x` and actually provides the month of the year, from January to December
  - the columns 2 to 5 provide a numeric value for 4 employees of a company respectively. Let's say it's the amount of money they brought or spend each month.

The dataset is stored on [github](https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/stacked_barplot_negative_values.csv) and can be loaded as follow:

```{r out.width = "100%"}

# Load the dataset that is stored on the web
data <- read.table("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/stacked_barplot_negative_values.csv", header=T, sep=",")

```

<br/>

This is an overview of how it looks like:

```{r}
library(knitr)
kable(head(data, 4))
```


<br/>

To work with `ggplot2` and more generally with the `tidyverse`, it's necessary to transform it to a "long" or "tidy" format. This is totally doable thanks to the `tidyr` package:

```{r}
# Load the package
library(tidyr)
library(dplyr)

# transform the format
data_long <- gather(data, group, value, groupA:groupD) %>%
  arrange(factor(x, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sept", "Oct"))) %>% 
  mutate(x=factor(x, levels=unique(x)))

# And that's the result!
kable(head(data_long, 4))
```


Note: sorting the data is annoying and is described more in depth [here](https://r-graph-gallery.com/267-reorder-a-variable-in-ggplot2.html)



















# Stacked barchart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It's now pretty straightforward to build the stacked barplot. Pretty much the same guideline that this [specific post](https://r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html) are applied. It's important to notice that the negative values are automatically located below the 0 baseline, without any additional work required.


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
 # library
library(ggplot2)

# plot
ggplot(data_long, aes(fill=group, y=value, x=x)) + 
  geom_bar(position="stack", stat="identity")
```
</div>
</div>



```{r thecode2, eval=FALSE}
```






# Stacked barchart polish
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

As always, a few polish steps to make the chart looks better with proper axis titles, better theme and a classic `viridis` color palette.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, warning=FALSE, comment=FALSE, out.width = "100%", fig.height=7}
 # library
library(ggplot2)
library(hrbrthemes)
library(viridis)

# plot
ggplot(data_long, aes(fill=group, y=value, x=x)) + 
  geom_bar(position="stack", stat="identity") + 
  scale_fill_viridis(discrete=TRUE, name="") +
  theme_ipsum() +
  ylab("Money input") + 
  xlab("Month")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```

































<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
