---
title: "Stacked barplot with negative values with ggplot2"
descriptionMeta: "This post explains how to build a stacked barplot with R and ggplot2. It considers a dataset that includes negative values to see how the package behave."
descriptionTop: "This post explains how to build a stacked barplot with R and [ggplot2](https://www.r-graph-gallery.com/ggplot2-package.html). It focus on a dataset that includes negative values"
sectionText: "Barchart section"
sectionLink: "barplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/barplot.html"
url: "341-stacked-barplot-with-negative-values"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Grouped barchart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A grouped barplot display a numeric value for a set of entities split in groups and subgroups. Before trying to build one, check how to make a [basic barplot](218-basic-barplots-with-ggplot2.html) with `R` and `ggplot2`.

A few explanation about the code below:

- input dataset must provide 3 columns: the numeric value (`value`), and 2 categorical variables for the group (`specie`) and the subgroup (`condition`) levels.
- in the `aes()` call, x is the group (`specie`), and the subgroup (`condition`) is given to the `fill` argument.
- in the `geom_bar()` call, `position="dodge"` must be specified to have the bars one beside each other.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
 # library
library(ggplot2)
library(tidyr)

# Load the dataset that is stored on the web

# Grouped
ggplot(data, aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="dodge", stat="identity")
 
```
</div>
</div>



```{r thecode, eval=FALSE}
```
























# Stacked barchart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A stacked barplot is very similar to the grouped barplot above. The subgroups are just displayed on top of each other, not beside.

The only thing to change to get this figure is to switch the `position` argument to `stack`.



</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
 # library
library(ggplot2)
 
# create a dataset
specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
 
# Stacked
ggplot(data, aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="stack", stat="identity")
 
```
</div>
</div>



```{r thecode2, eval=FALSE}
```




























# Percent stacked barchart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Once more, there is not much to do to switch to a percent stacked barplot. Just switch to `position="fill"`. Now, the percentage of each subgroup is represented, allowing to study the evolution of their proportion in the whole.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
 # library
library(ggplot2)
 
# create a dataset
specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
 
# Stacked + percent
ggplot(data, aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="fill", stat="identity")
 
```
</div>
</div>



```{r thecode3, eval=FALSE}
```












# Grouped barchart customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

As usual, some customization are often necessary to make the chart look better and personnal. Let's:

- add a [title](ggplot2-package.html)
- use a [theme](ggplot2-package.html)
- change color palette. See more [here](ggplot2-color.html).
- customize axis titles

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
library(viridis)
library(hrbrthemes)

# create a dataset
specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
 
# Small multiple
ggplot(data, aes(fill=condition, y=value, x=specie)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_viridis(discrete = T) +
    ggtitle("Studying 4 species..") +
    theme_ipsum() +
    xlab("")
```
</div>
</div>



```{r thecode4, eval=FALSE}
```
















# Small multiple
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Small multiple can be used as an alternative of stacking or grouping. It is straightforward to make thanks to the `facet_wrap()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode5, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(ggplot2)
library(viridis)
library(hrbrthemes)

# create a dataset
specie <- c(rep("sorgho" , 3) , rep("poacee" , 3) , rep("banana" , 3) , rep("triticum" , 3) )
condition <- rep(c("normal" , "stress" , "Nitrogen") , 4)
value <- abs(rnorm(12 , 0 , 15))
data <- data.frame(specie,condition,value)
 
# Graph
ggplot(data, aes(fill=condition, y=value, x=condition)) + 
    geom_bar(position="dodge", stat="identity") +
    scale_fill_viridis(discrete = T, option = "E") +
    ggtitle("Studying 4 species..") +
    facet_wrap(~specie) +
    theme_ipsum() +
    theme(legend.position="none") +
    xlab("")

```
</div>
</div>



```{r thecode5, eval=FALSE}
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
