---
title: "Introduction to waffle charts with R"
descriptionMeta: "This post explains how to build a basic waffle chart with R and the waffle package. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build a basic waffle chart with R and the [waffle](https://github.com/hrbrmstr/waffle) package. It provides several reproducible examples with explanation and R code."
sectionText: "Waffle section"
sectionLink: "waffle-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/waffle.html"
url: "342-introduction-to-waffle-plot-with-R"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---



```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# The `waffle` package
***

The [waffle](https://github.com/hrbrmstr/waffle) package is the best way to build a waffle chart in R. You can read more about this package on its [github page](https://github.com/hrbrmstr/waffle). Something important to note is that you can't install it directly from the CRAN. Instead, use this command line:

```{r eval=FALSE}
install.packages("waffle", repos = "https://cinc.rud.is")
```

<br/>



# Data input: vector
***

<div class = "row">

<div class = "col-md-6 col-sm-12 align-self-center">

The waffle function accepts 2 kinds of input: a vector or a data frame.


Let's start with a first example based on a vector input. The vector is a list of values. Each value will be represented as a set of rectangles. The `rows` argument controls how many rows will be displayed on the chart.

By default, categories will be labeled A, B, C.. and so on. We will see later how to customize this.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load the library
library(waffle)

# Vector
x <- c(30, 25, 20, 5)

# Waffle chart
waffle(x, rows = 8)
```
</div>
</div>

```{r thecode, eval=FALSE}
```


# Data input: data frame
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We can also build a [waffle plot](waffle-chart.html) from a data frame. 

Here a simple example with pizza topings. 

The data frame requires 2 columns: the first one provides the labels, the second the quantities of each group.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# install.packages("waffle", repos = "https://cinc.rud.is")
library(waffle)

# Creating a simple data frame
mypizza <- data.frame(
  ingredients = c("Cheese","Tomatoes","Mushroom","Pepperoni"),
  vals = c(30, 25, 20, 5)
)

waffle(mypizza)
```
</div>
</div>



```{r thecode2, eval=FALSE}
```



# Waffle chart with customization
***



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The waffle function allows a lot of customization to make the graphic more pleasant. In this example:

- a title is added (`title` attribute)
- the width of the grid is controled with `size`
- some custom colors are picked to match the ingredient (`colors`)
- the legend is put at the bottom of the chart (`legend_pos`)

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
library(waffle)

# Creating a simple data frame
mypizza <- data.frame(
  ingredients = c("Cheese","Tomatoes","Mushroom","Pepperoni"),
  vals = c(30, 25, 20, 5)
)

waffle(mypizza, 
   rows = 7, # Numbers of rows
   size = 1, # width of the separator between blocks
   colors = c("#E8D630", "#FF0000", "#582900","#A91101"), # Colors of each group
   legend_pos = "bottom", # Position of the legend
   title = "Most popular toppings on a pizza" # Title
) 
```
</div>
</div>



```{r thecode3, eval=FALSE}
```



# Conclusion
***

That's a very quick introduction to waffle charts with R using the `waffle` package. There is much more you can do using this package, so feel free to visit the [waffle section](waffle-chart.html) of the gallery tp browse more examples.

<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```




