---
title: "Customize your titles in your tables with gt table"
descriptionMeta: "This post explains how to build customize the title and subtitle in a table with the gt package."
descriptionTop: "This post explains how to customize the title and subtitle in a table with the gt package. It provides several reproducible examples with explanation and R code."
sectionText: "Table section"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/waffle.html"
url: "350-custom-title-in-gt-table.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---



```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# The `gt` package
***

The [gt](https://r-graph-gallery.com/package/gt.html) package is an excellent way to **create and customize nice table output** in R. You can read more about this package on its [github page](https://github.com/rstudio/gt). You can install it directly from the CRAN by running the following:

```{r eval=FALSE}
install.packages("gt")
```

<br/>


# Default gt table
***

<div class = "row">

Tables from `gt` are sober but highly customizable

<div class = "col-md-12  col-sm-12 align-self-center">

```{r thecode, echo=TRUE, fig.height=7, out.width="100%"}
library(gt)
library(dplyr)

# dataset
data = data.frame(
  Country = c("USA", "China", "India"),
  Capitals = c("Washington D.C.", "Beijing", "New Delhi"),
  Population = c(331, 1441, 1393),
  GDP = c(21.43, 14.34, 2.87)
)

# create and display the gt table (equivalent to "gt(data)")
data %>%
  gt()
```
</div>
</div>






# Add a title in markdown
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The most practical and intuitive way to create a gt table is to use it in **combination** of the `dplyr` package.

We define a simple dataset and pass it to the `gt()` function. 

Thanks to the `tab_header()` function, we can super easily add a title. And with the `md()` function we can write it in **markdown**.

```{r thecode2, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_header(title = md("Some **title**"))
```
</div>
</div>






# Add a title in HTML
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Thanks to the `tab_header()` function, we can super easily add title. And with the `html()` function we can write the text in **HTML**.


```{r thecode3, echo=TRUE, fig.height=7, out.width="100%"}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_header(title = html("<span style='color:red;'>A red title</span>"))
```
</div>
</div>




# Combine title and subtitle
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Still with the `tab_header()` function, we can combine title and subtitle by just adding a subtitle argument:


```{r thecode4, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_header(title = html("<span style='color:red;'>A <strong>red</strong> title</span>"),
               subtitle = md("This text will be *below the title* and is written in `markdown`"))
```
</div>
</div>






# More complex example
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In this example, we'll add an image (the R logo) in the subtitle using html formatting. In this case, we also need the `htmltools` package.


```{r thecode5, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)
library(htmltools)

# create and display the gt table 
data %>%
  gt() %>%
    tab_header(title = html("<span style='color:red;'>A <strong>red</strong> title</span>"),
               subtitle = tagList(
                 tags$div(style = css(`text-align` = "center"),
                          HTML(web_image("https://www.r-project.org/logo/Rlogo.png")
                     )
                   )
                 )
               )
```
</div>
</div>






# Conclusion
***

We now know how to customize your table titles with the `gt` package. There is much more you can do using this package, so feel free to visit the [gt table section](package/gt.html) of the gallery to learn more about it and check other examples.

<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```




