---
title: "Introduction to reactable"
descriptionMeta: "This post explains how to use the reactable package to build interactive tables in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build an [interactive table](table.html) with the `reactable` package. It provides several reproducible examples with explanation and R code."
sectionText: "Correlogram section"
sectionLink: "correlogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/correlogram.html"
url: "350-introduction-to-reactable.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Default table with reactable
***

Simply pass a table or dataframe as the first argument to the `reactable` function.

```{r thecode, echo=TRUE, out.width = "100%", fig.height=7}
library("reactable") # load the package
data(iris) # load the iris dataset

reactable(iris)
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/intro-reactable1.html" height="700" width="700" style="border:none;"></iframe></center>

</div>


# Column grouping and aggregating
***

Use the `groupBy` argument to group rows by a column. The column will be hidden and a toggle will appear above the table to show and hide the groups.


```{r thecode2, echo=TRUE, out.width = "100%", fig.height=7}
reactable(iris,
          groupBy = "Species", # Group by the Species column
          columns = list(
            Sepal.Length = colDef(aggregate = "max"), # max value
            Sepal.Width = colDef(aggregate = "mean"), # mean value
            Petal.Length = colDef(aggregate = "min"), # min value
            Petal.Width = colDef(aggregate = "sum") # sum value
            )
        )
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/intro-reactable2.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Custom theme
***

Use the `theme` argument to **customize the appearance** of the table. You can also use the `striped` or `highlight` arguments to **add stripes** or **highlight** the hovered row.

```{r thecode3, echo=TRUE, out.width = "100%", fig.height=7}
reactable(
  iris[1:30, ],
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  theme = reactableTheme(
    borderColor = "black",
    stripedColor = "gold",
    highlightColor = "lightblue",
    cellPadding = "8px 12px"
  )
)
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/intro-reactable3.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Conclusion
***

This post explains how to build an interactive [table](table.html) with `reactable`. This package can **offer much more** than what is presented here.

I encourage you to check the [reactable page](package/reactable.html) to learn more about it.

</div>






<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
