---
title: "Customize your footer in your gt tables"
descriptionMeta: "This post explains how to build customize the footer in a table with the gt package."
descriptionTop: "This post explains how to customize the footer in a table with the gt package. It provides several reproducible examples with explanation and R code."
sectionText: "Table section"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/waffle.html"
url: "351-custom-footer-and-references-in-gt-table.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---



```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# The `gt` package
***

The [gt](https://r-graph-gallery.com/package/gt.html) package is an excellent way to **create and customize nice table output** in R. You can read more about this package on its [github page](https://github.com/rstudio/gt). You can install it directly from the CRAN by running the following:

```{r eval=FALSE}
install.packages("gt")
```

<br/>


# Default gt table
***

<div class = "row">

Tables from `gt` are sober but highly customizable

<div class = "col-md-12  col-sm-12 align-self-center">

```{r thecode, echo=TRUE, fig.height=7, out.width="100%"}
library(gt)
library(dplyr)

# dataset
data = data.frame(
  Planet = c("Earth", "Mars", "Jupiter", "Venus"),
  Moons = c(1, 2, 79, 0),
  Distance_from_Sun = c(149.6, 227.9, 778.3, 108.2),
  Diameter = c(12742, 6779, 139822, 12104)
)

# create and display the gt table (equivalent to "gt(data)")
data %>%
  gt()
```
</div>
</div>






# Add a simple footer in markdown
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The most practical and intuitive way to create a gt table is to use it in **combination** of the `dplyr` package.

Thanks to the `tab_footnote()` function, we can super easily add a footer. And with the `md()` function we can write it in **markdown**.

```{r thecode2, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_footnote(footnote = md("This text is the footer of this **table**"))
```
</div>
</div>






# Footer with reference
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

If your footer is intended to **refer to a table element**, you can add the `location` argument to specify which part of the table your footnote refers to. 


```{r thecode3, echo=TRUE, fig.height=7, out.width="100%"}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_footnote(footnote = md("Measured in **millions** of Km"),
                 locations = cells_column_labels(columns = Distance_from_Sun))
```
</div>
</div>




# Footer with mutliple references
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Adding **multiple references** is just as easy as adding one: just add as much as you need the `tab_footnote()` function


```{r thecode4, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_footnote(footnote = md("Measured in **millions** of Km"),
                 locations = cells_column_labels(columns = Distance_from_Sun)) %>%
    tab_footnote(footnote = md("Measured in **Km**"),
                 locations = cells_column_labels(columns = Diameter))
```
</div>
</div>






# Change the type of element that indicates the reference
***


<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In this example, we now indicate the footer with letters (in upper case) thanks to the `opt_footnote_marks()` function.

```{r thecode5, echo=TRUE, out.width = "100%", fig.height=7}
library(gt)
library(dplyr)

# create and display the gt table 
data %>%
  gt() %>%
    tab_footnote(footnote = md("Measured in **millions** of Km"),
                 locations = cells_column_labels(columns = Distance_from_Sun)) %>%
    tab_footnote(footnote = md("Measured in **Km**"),
                 locations = cells_column_labels(columns = Diameter)) %>%
    tab_footnote(footnote = md("The original data are from *Some Organization*")) %>%
    opt_footnote_marks(marks = "LETTERS")
```
</div>
</div>






# Conclusion
***

We now know how to customize your table footer with the `gt` package. There is much more you can do using this package, so feel free to visit the [gt table section](package/gt.html) of the gallery to learn more about it and check other examples.

<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```




