---
title: "Column formatting with reactable"
descriptionMeta: "This post explains how to use the reactable package to build interactive tables with custom formatting in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build an [interactive table](table.html) with the `reactable` package. It provides several reproducible examples on how to **format the columns**"
sectionText: "Correlogram section"
sectionLink: "correlogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/correlogram.html"
url: "351-column-formatting-in-reactable.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Default table with reactable
***

Simply pass a table or dataframe as the first argument to the `reactable` function.

```{r thecode, echo=TRUE, out.width = "100%", fig.height=7}
library("reactable") # load the package

# Create the dataset 
df = data.frame(
  date = as.POSIXct(c("2019-01-02 3:22:15", "2019-03-15 09:15:55", "2019-09-22 14:20:00"), tz = "America/New_York"),
  currency = c(1000, 2000, 3000),
  temperature = c(10, 20, 30),
  percentage = c(0.12, 0.23, 0.34)
)

tab = reactable(df)

library(htmlwidgets)
saveWidget(tab, file=paste0(getwd(), "/HtmlWidget/column-formatting-reactable-1.html"))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/column-formatting-reactable-1.html" height="700" width="700" style="border:none;"></iframe></center>

</div>


# Date formatting
***

Use the `colFormat` function to **format the date**. You can also use the `format` argument to **customize the date format**. 

There is 3 date format: `datetime`, `date` and `time`. 


```{r thecode2, echo=TRUE, out.width = "100%", fig.height=7}
datetimes = as.POSIXct(c("2019-01-02 3:22:15", "2019-03-15 09:15:55", "2019-09-22 14:20:00"),
                        tz = "America/New_York")
data = data.frame(
  datetime = datetimes,
  date = datetimes,
  time = datetimes,
  time_24h = datetimes,
  datetime_pt_BR = datetimes
)

tab = reactable(data, columns = list(
                datetime = colDef(format = colFormat(datetime = TRUE)),
                date = colDef(format = colFormat(date = TRUE)),
                time = colDef(format = colFormat(time = TRUE)),
                time_24h = colDef(format = colFormat(time = TRUE, hour12 = FALSE)),
                datetime_pt_BR = colDef(format = colFormat(datetime = TRUE, locales = "pt-BR"))
))

# Save the widget
library(htmlwidgets)
saveWidget(tab, file=paste0(getwd(), "/HtmlWidget/column-formatting-reactable-2.html"))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/column-formatting-reactable-2.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Currency formatting
***

In the `colFormat` function, use the `currency` argument to **format the currency**. You mainly have to specify the `locales` argument to **customize the currency format**.

```{r thecode3, echo=TRUE, out.width = "100%", fig.height=7}
data = data.frame(
  USD = c(12.12, 2141.213, 0.42, 1.55, 34414),
  EUR = c(10.68, 1884.27, 0.37, 1.36, 30284.32),
  INR = c(861.07, 152122.48, 29.84, 110, 2444942.63),
  JPY = c(1280, 226144, 44.36, 164, 3634634.61),
  MAD = c(115.78, 20453.94, 4.01, 15, 328739.73)
)

tab = reactable(data, columns = list(
  USD = colDef(format = colFormat(currency = "USD", separators = TRUE, locales = "en-US")),
  EUR = colDef(format = colFormat(currency = "EUR", separators = TRUE, locales = "de-DE")),
  INR = colDef(format = colFormat(currency = "INR", separators = TRUE, locales = "hi-IN")),
  JPY = colDef(format = colFormat(currency = "JPY", separators = TRUE, locales = "ja-JP")),
  MAD = colDef(format = colFormat(currency = "MAD", separators = TRUE, locales = "ar-MA"))
))

# Save the widget
library(htmlwidgets)
saveWidget(tab, file=paste0(getwd(), "/HtmlWidget/column-formatting-reactable-3.html"))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/column-formatting-reactable-3.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Numeric value formatting
***

```{r thecode4, echo=TRUE, out.width = "100%", fig.height=7}
df = data.frame(
  temperature = c(10, 20, 30),
  percentage = c(0.12, 0.23, 0.34)
)

tab = reactable(df, columns = list(
  temperature = colDef(format = colFormat(suffix = " °C")),
  percentage = colDef(format = colFormat(percent = TRUE, prefix = "Percent: "))
))


library(htmlwidgets)
saveWidget(tab, file=paste0(getwd(), "/HtmlWidget/column-formatting-reactable-4.html"))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/column-formatting-reactable-4.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Conclusion
***

This post explains how to build an interactive [table](table.html) with `reactable` and how to format the columns.

I encourage you to check the [reactable page](package/reactable.html) to learn more about it.

</div>






<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```