---
title: "Highlight specific elements in a line chart"
descriptionMeta: "This post explains how to showcase elements in a line chart with ggplot2 and gghighlight. It considers a dataset that includes multiple time series to see how the package behave."
descriptionTop: "This post explains how to **highlight elements** in a line chart with [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [gghighlight](https://www.r-graph-gallery.com/package/gghighlight.html). Highlighting elements on a chart is a very powerful dataviz technique. It leads the reader to the important message you are trying to convey, making the figure more insightful."
sectionText: "Linechart section"
sectionLink: "line-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "www.data-to-viz.com/graph/line.html"
url: "354-highlight-specific-elements-in-line-chart.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Dataset
***

The dataset has three columns:

- `Date`: This column contains dates starting from January 1, 2020, and continuing for 100 days.

- `Value`: This column holds numbers. These numbers are created by adding random amounts to each other in a cumulative way. It's like keeping a running total of random values.

- `Type`: This column categorizes each row with one of three labels: `a`, `b`, or `c`.

```{r out.width = "100%"}
# dataset with 3 lines named "a", "b" and "c"
set.seed(1)
period = 100
df = data.frame(Date = seq(as.Date("2020-01-01"),
                           by = "day",
                           length.out = period), 
                Value = c(cumsum(rnorm(period)),
                          cumsum(rnorm(period)),
                          cumsum(rnorm(period))),
                Type = c(rep("a", period),
                         rep("b", period),
                         rep("c", period)))
```

<br/>


# Package
*** 

For this post, we need the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [gghighlight](https://www.r-graph-gallery.com/package/gghighlight.html) packages.

```{r}
# Install the packages if not already done
# install.packages("ggplot2")
# install.packages("gghighlight")

# Load the package
library(ggplot2)
library(gghighlight)
```











# Default line chart in ggplot2
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Here's what the default line chart output looks like with `ggplot2`:


</div>


<div class = "col-md-12  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
ggplot(df) +
  geom_line(aes(Date, Value, colour = Type))
```
</div>
</div>





# Highlight specific lines
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

With the `gghighlight()` function and some criterion, we can indicate **which line** will be displayed.

In this case, it's all lines where the maximum `Value` is greater than 10.

</div>


<div class = "col-md-12  col-sm-12">

```{r thecode3, echo=TRUE, fig.height=7, message=FALSE, warning=FALSE, comment=FALSE, out.width="100%"}
ggplot(df) +
  geom_line(aes(Date, Value, colour = Type)) +
  gghighlight(max(Value) > 10)
```
</div>
</div>



# Properties of faded out lines
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gghighlight()` function has an argument named **unhighlighted_params** that allows us to define what are the properties to use for the lines that does not meet our criterion.

In this case, we reduce the line width, change the color to blue and lower the opacity (`alpha`). We also change the line to `dashed`. 

</div>


<div class = "col-md-12  col-sm-12">
```{r}
ggplot(df) +
  geom_line(aes(Date, Value, colour = Type), linewidth=1) +
  gghighlight(max(Value) > 10,
              unhighlighted_params = list(linewidth = 0.3,
                                          colour = alpha("blue", 0.7),
                                          linetype = "dashed"))
```
</div>
</div>





# A more complex example
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gghighlight` package is also very useful when **combined with other packages** such as [hrbrthemes](https://r-graph-gallery.com/package/hrbrthemes.html) and [patchwork](https://r-graph-gallery.com/package/patchwork.html).

In this example we use the power of `gghighlight` with the simplicity of `patchwork` for stacking graphs and `hrbrthemes'` beautiful themes. 

</div>


<div class = "col-md-12  col-sm-12">
```{r thecode5, fig.height=6}
library(hrbrthemes)
library(patchwork)

plot1 = ggplot(df) +
  geom_line(aes(Date, Value, colour = Type), linewidth=0.4, color='#4393C3') +
  gghighlight(max(Value) > 10,
              unhighlighted_params = list(linewidth = 0.3,
                                          colour = alpha("darkred", 0.7),
                                          linetype = "dotted"),
              use_direct_label = FALSE) +
  theme_bw() + xlab("") + ylab("")

plot2 = ggplot(df) +
  geom_line(aes(Date, Value, colour = Type), linewidth=0.4, color='#4393C3') +
  gghighlight(min(Value) < -10,
              unhighlighted_params = list(linewidth = 0.3,
                                          colour = alpha("darkred", 0.7),
                                          linetype = "dotted"),
              use_direct_label = FALSE) +
  theme_bw() 

plot1 / plot2 + plot_annotation(title = 'This chart is built with gghighlight')
```
</div>
</div>





# Conclusion
***

In this post, we look at how to use the [gghighlight](https://www.r-graph-gallery.com/package/gghighlight.html) package to **highlight specific elements** of a multi-part line chart. To find out more about how to customize a line chart, see the [dedicated section](https://r-graph-gallery.com/line-plot.html). 
























<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```