---
title: "Area chart with gradient"
descriptionMeta: "This post explains how to create an area chart with a color gradient between the bottom line and the line chart."
descriptionTop: "This post explains how to create an `area chart` with a **color gradient** between the bottom line and the line chart.<br>We'll go through several examples with reproducible R code using [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [ggpattern](https://r-graph-gallery.com/package/ggpattern.html)."
sectionText: "Area Chart"
sectionLink: "area-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/area.html"
url: "357-area-chart-with-gradient"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following libraries:

```{r}
library(ggplot2)
library(ggpattern)
library(hrbrthemes)
```

<br/>


# Dataset
***

We create a simple dataset with 2 columns, randomly generated thanks to `runif()` and `rnorm()` functions.


```{r}
set.seed(0)
n = 10
x = 1:n
y = x + rnorm(n=n, mean=100, sd=20)
df = data.frame(xValue = x,
                yValue = y)
```




# Most basic area chart
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

To create an area chart with R, all we need to do is add the `geom_area()` function from [ggplot2](https://r-graph-gallery.com/ggplot2-package.html).

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
ggplot(df, aes(x=xValue, y=yValue)) +
  geom_area(colour='black', fill='lightblue')
```
</div>
</div>




# Add a simple gradient color
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

To add a color gradient in place of the light blue, we need to use the [ggpattern](https://www.r-graph-gallery.com/package/ggpattern.html) library and define which colors will be used for the graph.

The bottom color is `white` and the top color is `red`.

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
ggplot(df, aes(xValue, yValue)) + 
  geom_area_pattern(data = df,
                    pattern = "gradient",
                    pattern_fill = "white",
                    pattern_fill2 = "red")
```
</div>
</div>





# Gradient color with transparency
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

To add **transparency** to the graphic, simply use colors that are naturally transparent. For example, here we change **white to transparent grey**. 

The bottom color is a transpaent `light grey` (#00000010) and the top color is `blue`.

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
ggplot(df, aes(xValue, yValue)) + 
  geom_area_pattern(data = df,
                    pattern = "gradient",
                    fill = "#00000010",
                    pattern_fill  = "#00000010",
                    pattern_fill2 = "blue")
```
</div>
</div>






# Fully customized area chart wiht gradient
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

If we take what we've seen above and add other features like changing the `theme`, removing axis lines etc., you end up with an even more beautiful graph.

In this case, we:
- **add a black line** for the upper part of the area chart
- **change the theme** with the [hrbrthemes](https://r-graph-gallery.com/package/hrbrthemes.html) library
- **remove axis lines**

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
ggplot(df, aes(xValue, yValue)) + 
  geom_area_pattern(data = df,
                    pattern = "gradient", 
                    fill = "#00000000",
                    pattern_fill  = "#00000000",
                    pattern_fill2 = "magenta") + 
  geom_line(data = df, colour = "black", linewidth = 0.8) +
  geom_point(shape = 16, size = 4.5, colour = "purple") +
  geom_point(shape = 16, size = 2.5, colour = "white") +
  ggtitle("Area chart with a color gradient and line with data points") + 
  theme_bw() +
  theme(plot.title = element_text(size = 14),
        panel.border       = element_blank(),
        axis.line.x        = element_line(),
        text               = element_text(size = 12),
        axis.ticks         = element_blank(),
        axis.text.y        = element_text(margin = margin(0,15,0,0, unit = "pt"))) +
  scale_alpha_identity() + labs(x="",y="")
```
</div>
</div>


# Conclusion

This post explained how to add a color gradient to an area chart using the [ggpattern](https://r-graph-gallery.com/package/ggpattern.html) library. For more area chart examples, see the [dedicated section](https://r-graph-gallery.com/area-chart.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
