---
title: "Density plot with labels on lines"
descriptionMeta: "This post explains how to add a text or labels on the line of a density chart. It uses the ggplot2 and geomtextpath packages."
descriptionTop: "This post explains how to add a text or labels on the line of a density chart. It uses the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [geomtextpath](https://www.r-graph-gallery.com/package/geomtextpath.html)  packages for creating the chart and making it nice."
sectionText: "Density section"
sectionLink: "density-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "www.data-to-viz.com/graph/density.html"
url: "361-density-plot-with-labels-on-lines.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">





# Package
*** 

For this post, we need the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [geomtextpath](https://www.r-graph-gallery.com/package/geomtextpath.html) packages.

```{r}
# Install the packages if not already done
# install.packages("ggplot2")
# install.packages("geomtextpath")

# Load the package
library(ggplot2)
library(geomtextpath)
```











# Default density chart in ggplot2
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Here's what the default density chart output looks like with `ggplot2`:

</div>


<div class = "col-md-12  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
data(iris)

ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density() +
  theme(legend.position = "none")
```
</div>
</div>





# Add labels
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

With the `geom_textdensity()` function (inspired from the `geom_density()` func from [ggplot2](https://r-graph-gallery.com/ggplot2-package.html)), we can add the label of species directly on each density line.


</div>


<div class = "col-md-12  col-sm-12">

```{r thecode3, echo=TRUE, fig.height=7, message=FALSE, warning=FALSE, comment=FALSE, out.width="100%"}
library(hrbrthemes)
data(iris)

ggplot(iris, aes(x = Sepal.Length, colour = Species, label = Species)) +
  geom_textdensity() +
  theme_bw() + guides(color = 'none')
```
</div>
</div>



# Change texts properties
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `geom_textdensity()` can take multiple arguments in order to change the properties of the labels:

- `size`: size of the text

- `fontface`: style of the font

- `vjust`: vertical adjustment

- `hjust`: horizontal adjustment

The 2 last arguments can either be float (generally **between -1 and 1**) or string such as `xmid` (or `ymid`), `xmax` (or `ymax`) and `auto` (default).

</div>


<div class = "col-md-12  col-sm-12">
```{r}
library(hrbrthemes)
data(mtcars)

mtcars$labels = ifelse(mtcars$vs==0, "Type 0", "Type 1")
ggplot(mtcars, aes(x = qsec, colour = as.factor(labels), label = as.factor(labels))) +
  geom_textdensity(size = 6, fontface = 4, # size of 6, bold italic
                   vjust = -0.4, hjust = "ymid") +
  theme_bw() + guides(color = 'none')
```
</div>
</div>







# Conclusion
***

In this post, we look at how to use the [geomtexpath](https://www.r-graph-gallery.com/package/geomtextpath.html) package to **create density plot with text**. To find out more about how to customize a density plot, see the [dedicated section](https://r-graph-gallery.com/density-plot.html). 
























<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```