---
title: "Scatter plot with trend lines and labels on top"
descriptionMeta: "This post explains how to add a text or labels on the trend line in a scatter with one or multiple groups. It uses the ggplot2 and geomtextpath packages."
descriptionTop: "This post explains how to add a text or labels on the trend line in a scatter with one or multiple groups. It uses the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [geomtextpath](https://www.r-graph-gallery.com/package/geomtextpath.html)  packages for creating the chart and making it nice."
sectionText: "Scatter plot section"
sectionLink: "scatterplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "www.data-to-viz.com/graph/scatter.html"
url: "362-scatter-plot-with-trend-line-and-labels.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">





# Package
*** 

For this post, we need the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [geomtextpath](https://www.r-graph-gallery.com/package/geomtextpath.html) packages. 

```{r}
# Install the packages if not already done
# install.packages("ggplot2")
# install.packages("geomtextpath")

# Load the package
library(ggplot2)
library(geomtextpath)
```











# Default scatter plot in ggplot2
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Here's what the default scatter plot output looks like with `ggplot2`:

</div>


<div class = "col-md-12  col-sm-12">

```{r thecode1, echo=FALSE, out.width = "100%", fig.height=7}
data(iris)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  theme(legend.position = "none")
```
</div>
</div>





# Add trend line with label on top
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

With the `geom_labelsmooth()` function, we add a trend line with a label on it!

It has the following arguments:

- `fill`: the background color of the label

- `method`: type of trend wanted. In our case, `lm` means the ordinary least squared estimator (linear regression). Check the [function documentation](https://rdrr.io/cran/geomtextpath/man/geom_textsmooth.html) for available possibilities

- `size`, `linewidth` and `boxlinewidth`: define the properties of the text and its box

</div>


<div class = "col-md-12  col-sm-12">

```{r thecode2, echo=TRUE, fig.height=7, message=FALSE, warning=FALSE, comment=FALSE, out.width="100%"}
library(hrbrthemes)
data(iris)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  geom_labelsmooth(aes(label = 'My Label'), fill = "white",
                method = "lm", formula = y ~ x,
                size = 6, linewidth = 2, boxlinewidth = 0.6) +
  theme_bw() + guides(color = 'none') # remove legend
```
</div>
</div>



# Multiple groups trend lines with labels
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In the case of a multi-group scatter plot, the `geom_labelsmooth()` function works just as simply:

- we add `color=Species`

- we change `label = 'My Label'` to `label = Species`

And that's it!

</div>


<div class = "col-md-12  col-sm-12">

```{r thecode3, echo=TRUE, fig.height=7, message=FALSE, warning=FALSE, comment=FALSE, out.width="100%"}
library(hrbrthemes)
data(iris)

ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  geom_labelsmooth(aes(label = Species), fill = "white",
                method = "lm", formula = y ~ x,
                size = 3, linewidth = 1, boxlinewidth = 0.4) +
  theme_bw() + guides(color = 'none') # remove legend
```
</div>
</div>





# Conclusion
***

In this post, we look at how to use the [geomtexpath](https://www.r-graph-gallery.com/package/geomtextpath.html) package to **create scatter plots with trend lines and labels**. To find out more about how to customize a scatter plot, see the [dedicated section](https://r-graph-gallery.com/scatterplot.html). 
























<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```