---
title: "Customize colors in a kable table with kableExtra"
descriptionMeta: "This post explains how to customize colors in a kable output with the kableExtra package. We'll see how to add a color, several colors or a color gradient to a row or column."
descriptionTop: "This post explains how to customize colors in a kable output with kableExtra package. We'll go through several examples to see how to add a color, several colors or a color gradient to a row or column with [kableExtra](https://r-graph-gallery.com/package/kableExtra.html)."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/area.html"
url: "361-customize-colors-in-kableExtra"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
library(kableExtra)
```

<br/>


# Dataset
***

We create a simple dataset with 3 columns to showcase the **color features** of [kableExtra](https://r-graph-gallery.com/package/kableExtra.html)


```{r}
df = data.frame(Temp = c(1,2,3,4),
                Rain = c(12, 42, 17, 9),
                Hum = c(21, 24, 71, 90)
                )
```




# Default kable table output
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The [kableExtra](https://r-graph-gallery.com/package/kableExtra.html) relies on the `kable` package and allows the use of the `%>%` (pipe) symbole. The main function is named `kbl()` and is similar to `kable()`.

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
df %>%
  kbl() %>%
  kable_styling()
```
</div>
</div>




# Change color of a row or column
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `row_spec()` and `column_spec()` functions have a `color` and `background` arguments that will **change the colors**, either of the background or the cell content. 

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
df %>%
  kbl(align = "c") %>% # center the columns
  kable_styling(full_width = F) %>%
  column_spec(2, color = "red") %>%
  row_spec(3, color = "blue") %>%
  column_spec(1, background = "green") %>%
  row_spec(4, background = "yellow")
```
</div>
</div>



# Use a list of colors
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `color` and `background` argument can also takes a vector of colors:

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
df %>%
  kbl(align = "c") %>% # center the columns
  kable_styling(full_width = F) %>%
  column_spec(3, background = c("blue", "red", "black", "blue"))
```
</div>
</div>




# Gradient of color
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In order to have a gradient of colors, we need to use the `colorRampPalette()` function in order to generate a vector of colors.

We also sort the dataframe with the `Rain` column so that the **colors will be proportional** to the values in this column.

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
# gradient color
colfunc = colorRampPalette(c("darkred", "magenta"))
n_color = nrow(df)
colors = colfunc(n_color) # generate nrow(df) (number of rows in df) colors, from darkred to magenta

df = df[order(df$Rain, decreasing = TRUE),] # sort (decreasing) by Rain values

df %>%
  kbl(align = "c") %>% # center the columns
  kable_styling(full_width = F) %>%
  column_spec(3, background = colors) %>%
  column_spec(2, color = colors)
```
</div>
</div>




# Conclusion

This post explained how to custom colors in a table using the [kableExtra](https://r-graph-gallery.com/package/kableExtra.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/kableExtra.html) or the [table section](https://r-graph-gallery.com/table.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
