---
title: "Images and links in a kable table with kableExtra"
descriptionMeta: "This post explains how to add images and hyperlinks in a kable output with the kableExtra package. We'll go through several examples with reproducible R code."
descriptionTop: "This post explains how to customize colors in a kable output with the [kableExtra](https://r-graph-gallery.com/package/kableExtra.html) package. We'll go through several examples with reproducible R code the with [kableExtra](https://r-graph-gallery.com/package/kableExtra.html)."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "363-add-images-and-links-in-kableExtra"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
library(kableExtra)
```

<br/>


# Dataset
***

We create a simple dataset with Nobel Prizes (2020). We use 3 columns: name, field and image.

```{r}
df = data.frame(name = c("E. Charpentier","R. Penrose", "L. Glück", "M. Houghton"),
                field = c("Chemistry", "Physics", "Litterature", "Medicine"),
                image = "" # keep it empty
                )
```




# Add images
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The [kableExtra](https://r-graph-gallery.com/package/kableExtra.html) relies on the `kable` package and allows the use of the `%>%` (pipe) symbole. The main function is named `kbl()` and is similar to `kable()`.

In order to add images in column, we need to use the `column_spec()` function with the `image` argument. Also, to change dimensions of images, we use the `spec_image()` function into the `image` arg. 

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
# we use images from the internet, but it works exactly the same
# for images locally stored on your computer!
path_images = c("https://www.nobelprize.org/images/charpentier-111763-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/penrose-111758-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/gluck-111767-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/houghton-111770-landscape-mini-2x.jpg")

df %>%
  kbl(booktabs = T, align = "c") %>%
  kable_paper(full_width = T) %>%
  column_spec(3, image = spec_image(path_images, 280, 200) # dimensions of the images
              )
```
</div>
</div>




# Add links
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `row_spec()` and `column_spec()` functions have a `color` and `background` arguments that will **change the colors**, either of the background or the cell content. 

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
urls = c("https://en.wikipedia.org/wiki/Emmanuelle_Charpentier",
         "https://en.wikipedia.org/wiki/Roger_Penrose",
         "https://en.wikipedia.org/wiki/Louise_Glück",
         "https://en.wikipedia.org/wiki/Michael_Houghton")

df %>%
  kbl(booktabs = T, align = "c") %>%
  kable_paper(full_width = T) %>%
  column_spec(1, link = urls)
```
</div>
</div>



# Combine images and links
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `color` and `background` argument can also takes a vector of colors:

</div>


<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
path_images = c("https://www.nobelprize.org/images/charpentier-111763-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/penrose-111758-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/gluck-111767-landscape-mini-2x.jpg",
                "https://www.nobelprize.org/images/houghton-111770-landscape-mini-2x.jpg")

urls = c("https://en.wikipedia.org/wiki/Emmanuelle_Charpentier",
         "https://en.wikipedia.org/wiki/Roger_Penrose",
         "https://en.wikipedia.org/wiki/Louise_Glück",
         "https://en.wikipedia.org/wiki/Michael_Houghton")

df %>%
  kbl(booktabs = T, align = "c") %>%
  kable_material(c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(1, link = urls, bold=TRUE) %>%
  column_spec(1, link = urls, bold=TRUE) %>%
  column_spec(3, image = spec_image(path_images, 280, 200))
```
</div>
</div>







# Conclusion

This post explained how to add images and links in a table using the [kableExtra](https://r-graph-gallery.com/package/kableExtra.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/kableExtra.html) or the [table section](https://r-graph-gallery.com/table.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
