---
title: "Filtering, captioning and editing in tables with DT"
descriptionMeta: "This post explains how to add a filter parameter, a caption and edit cells in tables with the DT package. We'll go through several examples with reproducible R code."
descriptionTop: "This post explains how to add a **filter** parameter , a **caption** and **edit** cells in tables with the [DT](https://r-graph-gallery.com/package/dt.html) package. We'll go over several examples of how you can display interactive tables with customized features.<br>You can find out more about DT in the [dedicated section](https://r-graph-gallery.com/package/dt.html)."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "365-filtering-editing-captioning-in-dt-table"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
# install it if never done it before
# install.packages("DT")

library(DT)
```

<br/>



# Dataset
***

For this post, we'll use a dataset with different data types such as integer, numeric, factor, logical, and character.

```{r}
df <- data.frame(
  integer_col = as.integer(c(1, 2, 3)), # Integer column
  numeric_col = c(1.1, 2.2, 3.3), # Numeric column
  factor_col = factor(c("level1", "level2", "level3")), # Factor column
  logical_col = c(TRUE, FALSE, TRUE), # Logical column
  character_col = c("a", "b", "c") # Character column
)
```




# Filtering
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

By default, [DT](https://r-graph-gallery.com/package/dt.html) tables have no filters. However, the `datatable()` function has a `filter` argument with very useful properties, depending on the type of data.

Integer and numeric columns have **range sliders**, while other variables have a choice of all possible values. 

*Keep in mind*: make sure your variables have the **right type** so that the filters match. You may need to manually change the type of some of your columns. 

</div>

<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
table <- datatable(df,
  filter = "top"
) # put it at the top of the table

# save widget
library(htmltools)
saveWidget(table, file = "HtmlWidget/dt-filtering.html")
```

<center><iframe src="HtmlWidget/dt-filtering.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>




# Captioning
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Thanks to the `caption` argument, we can easily add a caption to our table. 

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
table <- datatable(
  df,
  caption = tags$caption(
    style = "caption-side: bottom; text-align: center;",
    "Table 1: ",
    em("This is a legend that you can customize")
  )
)

# save widget
library(htmltools)
saveWidget(table, file = "HtmlWidget/dt-captioning.html")
```

<center><iframe src="HtmlWidget/dt-captioning.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>



# Editing
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The cells can be edited directly into the table by **double clicking** on it, changing the value and then clicking outside. It's also possible to make **only specific** as editable.

In the case you want all cells to be **editable**, put `editable='cell'` in the `datatable()` function. Otherwise, put a list in argument of the `editable` argument like in the example below that allows editing in column 1, 3 and 5, for all rows:

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
table <- datatable(df,
  editable = list(
    target = "row",
    disable = list(columns = c(1, 3, 5))
  )
)

# save widget
library(htmltools)
saveWidget(table, file = "HtmlWidget/dt-editing.html")
```

<center><iframe src="HtmlWidget/dt-editing.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>





# All in once
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Here's what the output looks like when we put all together the features we've seen before:

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
df <- rbind(df, df, df) # add more rows
table <- datatable(df,
  editable = list(
    target = "row",
    disable = list(columns = c(1, 3, 5))
  ),
  caption = tags$caption(
    style = "caption-side: bottom; text-align: center;",
    "Table 1: ",
    em("This is a legend that you can customize")
  ),
  filter = "top"
)

# save widget
library(htmltools)
saveWidget(table, file = "HtmlWidget/dt-all-features.html")
```

<center><iframe src="HtmlWidget/dt-all-features.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>






# Conclusion

This post explained how to add a filter parameter, a caption and edit cells in [DT](https://r-graph-gallery.com/package/dt.html) table. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/dt.html) or the [table section](https://r-graph-gallery.com/table.html).






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```