---
title: "Display results of a regression model in a table"
descriptionMeta: "This post use the gtsummary package to show the results of a regression model in a table. We'll go through several examples with reproducible R code."
descriptionTop: "This post use the [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) package to show the results of a regression model in a table. We'll go through how to show p-values, confidence intervals, compare models and more with [gtsummary](https://r-graph-gallery.com/package/gtsummary.html)."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "366-regression-models-with-gtsummary"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
# install.packages("gtsummary")
library(gtsummary)
```

<br/>


# Default output for regression results
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) uses the `tbl_regression()` function to automatically generate results of the model in a table. 

In this case, we use the Titanic dataset, with the `Survived` variable as the target.

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
# create dataset
data("Titanic")
df = as.data.frame(Titanic)

# create the model
model = glm(Survived ~ Age + Class + Sex + Freq,
            family=binomial, data=df)

# generate table 
model %>%
  tbl_regression()
```
</div>
</div>




# Add more statistical details
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

If we take the code from above, we can **add other elements** to the table that will make the table much more informative and detailed:
- `add_glance_source_note()`: Adds a footnote to the table with information from the `glance()` function, like the number of observations, R-squared for regression models, etc
- `add_global_p()`: Appends a global p-value to the table which assesses the **overall significance** of categorical variables.
- `add_q()`: Incorporates **q-values** in the table, which are adjusted p-values used in multiple testing to control the false discovery rate.

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
# create dataset
data("Titanic")
df = as.data.frame(Titanic)

# create the model
model = glm(Survived ~ Age + Class + Sex + Freq,
            family=binomial, data=df)

# generate table 
model %>%
  tbl_regression(intercept=TRUE, conf.level=0.9) %>%
  add_glance_source_note() %>%
  add_global_p() %>%
  add_q() 
```
</div>
</div>



# Display results of different models
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

One easy way to show the results of 2 different models into a single table is to:
- **create a first table** with the first model (logistic regression)
- **create a second table** with the second model (Cox proportional hazards regression)
- **merge these tables** with `tbl_merge()`
- **add a spanner** for each model with the `tab_spanner` argument

In this case we use the `trial` dataset:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(survival)
library(gtsummary)

data(trial)

model_reglog = glm(response ~ trt + grade, data=trial, family = binomial) %>% tbl_regression()
model_cox = coxph(Surv(ttdeath, death) ~ trt + grade, data=trial) %>% tbl_regression()

tbl_merge(
  list(model_reglog, model_cox),
  tab_spanner = c("**Tumor Response**", "**Time to Death**")
)
```
</div>
</div>







# Conclusion

This post explained how to display the results of a regression model in a table using the [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/gtsummary.html) or the [table section](https://r-graph-gallery.com/table.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
