---
title: "Automate the creation of a summary table with gtsummary"
descriptionMeta: "This post use the gtsummary package to create a table that summarizes a dataset with descriptive, inferential statistics and more. We'll go through several examples with reproducible R code."
descriptionTop: "This post use the [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) package to create a table that summarizes a dataset with descriptive, inferential statistics and more. We'll go through several examples with reproducible code using the  [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) package."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "367-automate-data-summary-table-with-gtsummary"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
# install.packages("gtsummary")
library(gtsummary)
```

<br/>


# Default output for summary table
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) uses the `tbl_summary()` to generate the summary table and works well with the `%>%` symbol.

It automatically **detects data type** and use it to decides what type of statistics to compute. By default, it's:
- **median, 1st and 3rd quartile** for numeric columns
- **number of observations** and proportion for categorical columns

</div>


<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
library(gtsummary)

# create dataset
data("Titanic")
df = as.data.frame(Titanic)

# create the table
df %>%
  tbl_summary()
```
</div>
</div>




# Add p-values and statistical details
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

If you want to add p-values to the table, you **have to** add `by=variable_name` in the `tbl_summary()` function. This happens because p-values are used to compare things between them.

The variable in the `by` argument will be used to **split the dataset** into multiple sub-samples (2 if it's dichotomous, 3 if there are 3 distinct labels in the variable, etc). Those samples will be **compared** for each column in the dataset, and the test done depends on the type of data.

In this case, we add:
- `add_p()` to create a new column for p-values
- `add_overall()` to add a new column for descriptive statistics for the whole sample

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtsummary)

# create dataset
data("Titanic")
df = as.data.frame(Titanic)

# create the table
df %>%
  tbl_summary(by=Survived) %>%
  add_overall() %>%
  add_p() #%>%
  #add_stat_label()
```
</div>
</div>



# Add a column based on a custom function 
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

Thanks to the `add_stat()` function, we can create new column based on our own functions.

Below, we define an **anova** function that returns the p-values of an **ANOVA** and pass it to the `add_stat()` function.

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtsummary)

# create dataset
data("iris")
df = as.data.frame(iris)

my_anova = function(data, variable, by, ...) {
  result = aov(as.formula(paste(variable, "~", by)), data = data)
  summary(result)[[1]]$'Pr(>F)'[1] # Extracting the p-value for the group effect
}

# create the table
df %>%
  tbl_summary(by=Species) %>%
  add_overall() %>%
  add_p() %>%
  add_stat(fns = everything() ~ my_anova) %>%
  modify_header(
    list(
      add_stat_1 ~ "**p-value**",
      all_stat_cols() ~ "**{level}**"
    )
  ) %>%
  modify_footnote(
    add_stat_1 ~ "ANOVA")
```
</div>
</div>





# Conclusion

This post explained how to create summary table using the [gtsummary](https://r-graph-gallery.com/package/gtsummary.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/gtsummary.html) or the [table section](https://r-graph-gallery.com/table.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
