---
title: "Black and white barchart with textures, using ggpattern and ggplot2"
descriptionMeta: "This post explains how to create a black and white barchart with different background elements and textures using the ggpattern package. We'll go through several examples with reproducible R code."
descriptionTop: "This post explains how to create a **black and white** [barchart](https://r-graph-gallery.com/barplot.html) with different background elements and **textures** using the [ggpattern](https://r-graph-gallery.com/package/ggpattern.html) package.<br>We'll go through several examples with reproducible code."
sectionText: "Barplot"
sectionLink: "barplot.html"
DataToVizText: "Barplot"
DataToVizLink: "www.data-to-viz.com/graph/barplot.html"
url: "368-black-and-white-barchart"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE
)
```

<div class="container">


# Packages
***

For this post, we need to load [ggpattern](https://r-graph-gallery.com/package/ggpattern.html), [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and [hrbrthemes](https://r-graph-gallery.com/package/hrbrthemes.html) (for better looking chart).

```{r}
library(ggpattern)
library(ggplot2)
library(hrbrthemes)
```

<br/>


# Dataset
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In this post, we'll use a simple homemade dataset:

</div>

<div class = "col-md-12  col-sm-12">


```{r echo=TRUE}
df = data.frame(
  name=c("north","south","south-east","north-west","south-west"),
  val=sample(seq(7,15), 5)
)
```
</div>
</div>




# Barchart with ggplot2
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

With [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) and the `geom_bar()` function, we can easily create **simple barchart**:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
ggplot(df, aes(x=name, y=val, fill=name)) +
  geom_bar(stat="identity", alpha=.6, width=.4) +
  scale_fill_grey(start=0, end=0.8) +  # start and end define the range of grays
  theme_bw()
```
</div>
</div>



# Black and white barchart. 
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

In various cases, you can be obliged to have black and white charts. However, this does not imply that only can use **uniform colors**!

Actually, you can add textures to your chart elements and make them more visually apealling thanks to the [ggpattern](https://r-graph-gallery.com/package/ggpattern.html) library. 

This package has a `geom_col_pattern()` function for this specific use case!

The following arguments customize the appearance of the patterns:

- `fill = 'white'`: The base fill **color of the bars** (behind the pattern) is white.

- `colour = 'black'`: The **border color** of the bars is black.

- `pattern_density = 0.5`: The **density** of the pattern inside the bars.

- `pattern_fill = 'black'`: The **color of the pattern** itself.

- `pattern_colour = 'darkgrey'`: The **secondary color** used in the pattern.

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
ggplot(df, aes(x=name, y=val)) +
 geom_col_pattern(
    aes(pattern=name,
        pattern_angle=name,
        pattern_spacing=name
        ), 
    fill            = 'white',
    colour          = 'black', 
    pattern_density = 0.5, 
    pattern_fill    = 'black',
    pattern_colour  = 'darkgrey'
  ) +
  theme_bw()
```
</div>
</div>





# Conclusion

This post explained how to create a **textured black and white barchart** using the [ggpattern](https://r-graph-gallery.com/package/ggpattern.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/ggpattern.html) or the [barplot section](https://r-graph-gallery.com/barplot.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
