---
title: "Change the theme of gt tables with gtExtras"
descriptionMeta: "This post use the gtExtras package to change the theme from your gt tables"
descriptionTop: "This post use the [gtExtras](https://r-graph-gallery.com/package/gtExtras.html) package to change the theme from your [gt](https://r-graph-gallery.com/package/gt.html) tables. We'll go through several examples of which theme can be used with [gtExtras](https://r-graph-gallery.com/package/gtExtras.html)."
sectionText: "Table"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "369-custom-theme-in-gt-table-with-gtextras"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to load the following library:

```{r}
# install.packages("gtExtras")
library(gtExtras)
```

<br/>


# Default theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

For this post, we'll use the simple mtcars dataset that can be easily load using the code below.

The [gtExtras](https://r-graph-gallery.com/package/gtExtras.html) package relies on the [gt](https://r-graph-gallery.com/package/gt.html) package for the basic functions. Here's what the default theme looks like:


```{r}
data(mtcars)

head(mtcars) %>%
  gt()
```




# Excel theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_excel()` function offers a nice **Excel** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_excel()
```
</div>
</div>



# FiveThirtyEight theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_538()` function offers a nice **FiveThirtyEight** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_538()
```
</div>
</div>


# ESPN theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_espn()` function offers a nice **ESPN** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_espn()
```
</div>
</div>


# NY Times theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_nytimes()` function offers a nice **NY Times** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_nytimes()
```
</div>
</div>


# Dot matrix theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_dot_matrix()` function offers a nice **dot matrix** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_dot_matrix()
```
</div>
</div>


# Dark theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_dark()` function offers a nice **dark** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_dark()
```
</div>
</div>


# PFF theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_pff()` function offers a nice **PFF** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_pff()
```
</div>
</div>


# Guardian theme
***

<div class = "row">

<div class = "col-md-12  col-sm-12 align-self-center">

The `gt_theme_guardian()` function offers a nice **Guardian** theme:

</div>

<div class = "col-md-12  col-sm-12">


```{r, echo=TRUE}
library(gtExtras)

head(mtcars) %>%
  gt() %>%
  gt_theme_guardian()
```
</div>
</div>




# Conclusion

This post show all **table theme** available in a [gt](https://r-graph-gallery.com/package/gt.html) table using the [gtExtras](https://r-graph-gallery.com/package/gtExtras.html) library. For more of this package, see the [dedicated section](https://r-graph-gallery.com/package/gtExtras.html) or the [table section](https://r-graph-gallery.com/table.html). 






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
