---
title: "Basic beeswarm plot with R"
descriptionMeta: "A beeswarm plot or swarmplot is a type of data visualization that displays individual data points in a way that they don't overlap, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a simple beeswarm plot using the beeswarm package. We will provide reproducible code and examples."
descriptionTop: "A [beeswarm](beeswarm.html) plot or **swarmplot** is a type of data visualization that displays individual data points in a way that they **don't overlap**, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a simple [beeswarm](beeswarm.html) plot using the [beeswarm](https://github.com/aroneklund/beeswarm) package. We will provide reproducible code and examples."
sectionText: "Beeswarm"
sectionLink: "beeswarm.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "370-basic-beeswarm-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Packages
***

For this post, we need to install and load the [beeswarm](https://github.com/aroneklund/beeswarm) package. 

We can install it from CRAN using `install.packages("beeswarm")`. Then, we can load it:

```{r}
# install.packages("beeswarm")
library(beeswarm)
```

<br/>


# Dataset
***

Since [beeswarm](beeswarm.html) plots are made to visualize **individual data points**, we need a dataset that contains numerical values. Here, we'll use the `iris` dataset, which is a built-in dataset in R.

We can easily load it:

```{r}
data(iris)
print(head(iris)) # display first rows
```




# Most simple beeswarm
***

The package comes with a `beeswarm()` function that creates a beeswarm plot that can be used with only 1 argument: the **numerical variable** to visualize. Here, we'll use the `Sepal.Length` variable from the `iris` dataset.

```{r, echo=TRUE}
beeswarm(iris$Sepal.Length)
```




# Flip the plot
***

We can easily flip the plot by setting the `horizontal` argument to `TRUE` (or `vertical` to `FALSE`):

```{r, echo=TRUE}
beeswarm(iris$Sepal.Length, horizontal=TRUE)
```







# Going further
***
In this post, we learned how to create a simple [beeswarm](beeswarm.html) plot using the `beeswarm` package. We used the `beeswarm()` function to create a basic plot.

You might be interested in [how to custom the dots](371-custom-dots-beeswarm.html) or [how to display several groups on the beeswarm plot](372-grouped-beeswarm.html).






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
