---
title: "Customize dots in beeswarm plot"
descriptionMeta: "A beeswarm plot or swarmplot is a type of data visualization that displays individual data points in a way that they don't overlap, resulting in a *swarming* effect that resembles a swarm of bees.<br>In this post, we'll see how to custom the dots in a beeswarm plot. We will provide reproducible code and examples."
descriptionTop: "A [beeswarm](beeswarm.html) plot or **swarmplot** is a type of data visualization that displays individual data points in a way that they **don't overlap**, resulting in a *swarming* effect that resembles a swarm of bees.<br>In this post, we'll see how to **custom the dots** in a beeswarm plot using the [beeswarm](https://github.com/aroneklund/beeswarm) package. We will provide reproducible code and examples."
sectionText: "Beeswarm"
sectionLink: "beeswarm.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "371-custom-dots-beeswarm"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Packages
***

For this post, we need to install and load the [beeswarm](https://github.com/aroneklund/beeswarm) package. 

We can install it from CRAN using `install.packages("beeswarm")`. Then, we can load it:

```{r}
# install.packages("beeswarm")
library(beeswarm)
```

<br/>


# Dataset
***

Since [beeswarm](beeswarm.html) plots are made to visualize **individual data points**, we need a dataset that contains numerical values. Here, we'll use the `iris` dataset, which is a built-in dataset in R.

We can easily load it:

```{r}
data(iris)
```



# Change dots color
***

The package comes with a `beeswarm()` function, and thanks to its `col` argument, we can change the color of the dots. Here, we'll use the `blue` color.

```{r, echo=TRUE}
beeswarm(
  iris$Sepal.Length,
  col="blue"
)
```



# Change dots symbol
***

The `pch` argument allows to change the symbol of the dots. Here, we'll use the `16` symbol since it's the same as before but filled.

```{r, echo=TRUE}
beeswarm(
  iris$Sepal.Length,
  pch=16
)
```



# Change dots size
***

The `cex` argument allows to change the size of the dots. Here, we'll use a `2.5` size (default is 1).

```{r, echo=TRUE}
beeswarm(
  iris$Sepal.Length,
  cex=2.5
)
```


# Change dots position
***

The `method` argument allows to **change the position of the dots**. Here are the available **options**:

- `swarm` (default): the dots are placed randomly, but they don't overlap.
- `center`: the dots are symetrically placed around the center of the plot.
- `hex`: the dots are placed on a hexagonal grid.
- `square`: the dots are placed on a square grid.

In our case, we'll use the `center` method:

```{r, echo=TRUE}
beeswarm(
  iris$Sepal.Length,
  method="center",
  cex=2 # make the dots bigger
)
```




# Going further

This post explains how to **customize the dots** in a [beeswarm](beeswarm.html) plot.

You might also be interested in [how to flip a beeswarm plot](370-basic-beeswarm-plot.html) or [how to create a grouped beeswarm plot](372-grouped-beeswarm.html).






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
