---
title: "Grouped beeswarm plot"
descriptionMeta: "A beeswarm plot or swarmplot is a type of data visualization that displays individual data points in a way that they don't overlap, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a grouped beeswarm plot in R using the beeswarm package."
descriptionTop: "A [beeswarm](beeswarm.html) plot or **swarmplot** is a type of data visualization that displays individual data points in a way that they **don't overlap**, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a grouped [beeswarm](beeswarm.html) plot in R using the [beeswarm](https://github.com/aroneklund/beeswarm) package."
sectionText: "Beeswarm"
sectionLink: "beeswarm.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "372-grouped-beeswarm"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Packages
***

For this post, we need to install and load the [beeswarm](https://github.com/aroneklund/beeswarm) package. 

We can install it from CRAN using `install.packages("beeswarm")`. Then, we can load it:

```{r}
# install.packages("beeswarm")
library(beeswarm)
```

<br/>


# Dataset
***

Since [beeswarm](beeswarm.html) plots are made to visualize **individual data points**, we need a dataset that contains numerical values. Here, we'll use the `iris` dataset, which is a built-in dataset in R.

We can easily load it:

```{r}
data(iris)
```

```{r, include=FALSE}
iris <- do.call(rbind, replicate(7, iris, simplify = FALSE))
iris$Sepal.Length <- iris$Sepal.Length + rnorm(nrow(iris), 0, 0.2)
```

# Default grouped beeswarm plot
***

The package comes with a `beeswarm()` function, and thanks to the `~` **operator**, we can easily create a **grouped beeswarm plot**. Here, we'll use the `Species` column to group the data:

The following code basically means:

*plot `Sepal.Length` for each `Species` using the `iris` dataset*.

```{r, echo=TRUE}
beeswarm(
  Sepal.Length ~ Species,
  data=iris
)
```



# Use different colors
***

Even tough the default option does what we want, having the same color for all the `Species` makes the plot less readable. We can use the `col` argument to **change the color** of the dots:

```{r, echo=TRUE}
beeswarm(
  Sepal.Length ~ Species, 
  data=iris,
  col=c("orange", "lightblue", "magenta"),
  pch = 19 # fill the dots
)
```



# Custom position behavior 
***

When you have lots of data points, it can be useful to change the **position behavior** of the dots in order to avoid overlapping between groups. 

Fortunately, the `corral` argument allows you to change this. The available options are:

- `none` (default): no correction
- `gutter`: fix a **higher and lower limit** for each group
- `wrap`: similar to `gutter` but **add random noise** to the position of the dots
- `random`: **randomly position** the dots
- `omit`: **omit** the dots that would overlap

Here's an example with the `gutter` option:

```{r, echo=TRUE}
beeswarm(
  Sepal.Length ~ Species, 
  data=iris,
  col=c("orange", "lightblue", "magenta"),
  pch = 19, # fill the dots
  corral = "gutter"
)
```





# Going further
***

This post explains how to create a [grouped beeswarm](372-grouped-beeswarm.html) plot with R.

You might also be interested in [how to create a beeswarm plot with ggplot2](373-basic-beeswarm-with-ggbeeswarm.html).






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
