---
title: "Basic beeswarm plot with ggbeeswarm"
descriptionMeta: "A beeswarm plot or swarmplot is a type of data visualization that displays individual data points in a way that they don't overlap, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a basic beeswarm plot in R using the ggbeeswarm package."
descriptionTop: "A [beeswarm](beeswarm.html) plot or **swarmplot** is a type of data visualization that displays individual data points in a way that they **don't overlap**, resulting in a **swarming** effect that resembles a swarm of bees.<br>In this post, we'll see how to create a basic beeswarm plot in R using the [ggbeeswarm](package/ggbeeswarm.html) package."
sectionText: "Beeswarm"
sectionLink: "beeswarm.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "373-basic-beeswarm-with-ggbeeswarm"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Packages
***

For this post, we need to install and load the [ggbeeswarm](package/ggbeeswarm.html) package. 

We can install it from CRAN using `install.packages("ggbeeswarm")`. Then, we can load it:

```{r}
# install.packages("ggbeeswarm")
library(ggbeeswarm)
library(ggplot2)
```

<br/>


# Dataset
***

Since [beeswarm](beeswarm.html) plots are made to visualize **individual data points**, we need a dataset that contains numerical values. Here, we'll use the `iris` dataset, which is a built-in dataset in R.

We can easily load it:

```{r}
data(iris)
```

```{r, include=FALSE}
iris <- do.call(rbind, replicate(4, iris, simplify = FALSE))
iris$Sepal.Length <- iris$Sepal.Length + rnorm(nrow(iris), 0, 0.2)
```


# Beeswarm with ggplot2
***

The package comes with 2 main functions:

- `geom_beeswarm()`: to create a **beeswarm plot** with ggplot2
- `geom_quasirandom()`: to create a **quasirandom plot**, which is a **mix between** a beeswarm and a jitter plot.

In this post, we'll focus on the `geom_beeswarm()` function.

```{r, echo=TRUE}
ggplot(iris,aes(y=Sepal.Length,x='')) +
  geom_beeswarm()
```



# Flip the plot
***

We can super easily flip the plot by changing the `x` and `y` aesthetics:


```{r, echo=TRUE}
ggplot(iris,aes(x=Sepal.Length,y='')) +
  geom_beeswarm()
```


# Custom theme
***

We can change the color of the points and the overall theme of the plot using the `theme()` function:

```{r, echo=TRUE}
ggplot(iris,aes(y=Sepal.Length,x='')) +
  geom_beeswarm(color='blue') +
  theme_minimal()
```


# Change the position behavior
***

By default, the `geom_beeswarm()` function will use the `swarm` method to position the points. We can change this behavior using the `method` argument. Here are the **available options**:

- `swarm`: the default method
- `compactswarm`: similar to `swarm`, but the points are **more compact**
- `center`: the points are **centered** on the x-axis
- `hex`: the points are positioned in **hexagons**
- `square`: the points are positioned in **squares**

Here's an example with `center`:

```{r, echo=TRUE}
ggplot(iris,aes(y=Sepal.Length,x='')) +
  geom_beeswarm(method='center')
```


# Going further
***

This post explains how to create and customize a [beeswarm](beeswarm.html) plot with [ggbeeswarm](package/ggbeeswarm.html).

You might also be interested in [how to create a grouped beeswarm plot](374-grouped-beeswarm-with-ggbeeswarm.html) with ggplot2.






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
