---
title: "Grouped beeswarm plot with ggbeeswarm"
descriptionMeta: "A beeswarm plot or swarmplot is a type of data visualization that displays individual data points in a way that they don't overlap, resulting in a *swarming* effect that resembles a swarm of bees.<br>In this post, we'll see how to create a basic beeswarm plot in R using the ggbeeswarm package."
descriptionTop: "A [beeswarm](beeswarm.html) plot or **swarmplot** is a type of data visualization that displays individual data points in a way that they **don't overlap**, resulting in a *swarming* effect that resembles a swarm of bees.<br>In this post, we'll see how to create a basic beeswarm plot in R using the [ggbeeswarm](package/ggbeeswarm.html) package."
sectionText: "Beeswarm"
sectionLink: "beeswarm.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "374-grouped-beeswarm-with-ggbeeswarm"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Packages
***

For this post, we need to install and load the [ggbeeswarm](package/ggbeeswarm.html) package. 

We can install it from CRAN using `install.packages("ggbeeswarm")`. Then, we can load it:

```{r}
# install.packages("ggbeeswarm")
library(ggbeeswarm)
library(ggplot2)
```

<br/>


# Dataset
***

Since [beeswarm](beeswarm.html) plots are made to visualize **individual data points**, we need a dataset that contains numerical values. Here, we'll use the `iris` dataset, which is a built-in dataset in R.

We can easily load it:

```{r}
data(iris)
```



# Grouped beeswarm
***

We can easily create a **grouped beeswarm plot** by specifying a categorical variable in the `aes()` function.


```{r, echo=TRUE}
ggplot(iris,aes(x=Species, y=Sepal.Length)) +
  geom_beeswarm()
```



# Color by group
***

And in order to make the **plot more readable**, we can add some **color** to the points using the `colour` argument.

```{r, echo=TRUE}
ggplot(iris,aes(x=Species, y=Sepal.Length, colour=Species)) +
  geom_beeswarm()
```



# Customize colors
***

We can also **customize the colors** using the `scale_color_manual()` function. And thanks to the `theme_minimal()` function, we can make the plot a bit **more elegant**.

```{r, echo=TRUE}
ggplot(iris,aes(x=Species, y=Sepal.Length, colour=Species)) +
  geom_beeswarm() +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  theme_minimal()
```



# Going further
***
This post explains how to create a grouped beeswarm plot with R and the [ggbeeswarm](package/ggbeeswarm.html) package.

You might also be interested in [how to customize a beeswarm plot](373-basic-beeswarm-with-ggbeeswarm.html).






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
