---
title: "R Color Brewer's palettes"
descriptionMeta: "The RColorBrewer package offers several color palette for R. This post displays all of them to help you pick the right one."
descriptionTop: "The `RColorBrewer` package offers several color palette for R. This post displays all of them to help you pick the right one."
sectionText: "Color Section"
sectionLink: "ggplot2-color.html"
DataToVizText: "About line chart"
DataToVizLink: "data-to-viz.com/graph/line.html"
url: "38-rcolorbrewers-palettes"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `RColorBrewer` package is an unavoidable tool to manage colors with R. It offers several color palettes, as you can see in the attached figure. This page just allows to visualize the composition of each palette. [Other pages](ggplot2-color.html) explain how to use this tool.

There are 3 types of palettes :

- Sequential palettes are suited to ordered data that progress from low to high.  Lightness steps dominate the look of these schemes, with light colors for low data values to dark colors for high data values.

- Diverging palettes put equal emphasis on mid-range critical values and extremes at both ends of the data range. The critical class or break in the middle of the legend is emphasized with light colors and low and high extremes are emphasized with dark colors that have contrasting hues.

- Qualitative palettes do not imply magnitude differences between legend classes, and hues are used to create the primary visual differences between classes.  Qualitative schemes are best suited to representing nominal or categorical data.

</div>

<br><br>

<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=8}
library(RColorBrewer)
par(mar=c(3,4,2,2))
display.brewer.all()
```

</div>
</div>


```{r thecode, eval=FALSE}
```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedOther.html")
```
