---
title: "3d scatterplot with R"
descriptionMeta: "This post shows how to build a 3d scatterplot with R and the rgl package. It provides the reproducible code and makes it easy to apply it to your own dataset."
descriptionTop: "This post shows how to build a 3d scatterplot with R and the rgl package. It provides the reproducible code and makes it easy to apply it to your own dataset."
sectionText: "3d Section"
sectionLink: "3d.html"
DataToVizText: "3d: warning"
DataToVizLink: "data-to-viz.com/caveat/3d.html"
url: "3d_scatter_plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

Building a [3d scatterplot](3d.html) requires a dataset with 3 numeric variables, each being used on an axis. Here, the famous `iris` dataset is used.

The `rgl` package comes with the `plot3d()` function that is pretty close from the base R `plot()` function. Instead of providing just an `x` and a `y` argument, you also have to provide the `z` coordinate.

Note that the output is interactive by default. If you have X11 or a similar tool installed, you should be able to rotate the chart for a better user experience. A few command line are also provided in case you want to export the chart at `.html`, `.png` or `.Rmd` format.

</div>


<div class = "col-md-8  col-sm-12">

<center><iframe src="HtmlWidget/3dscatter.html" height="600" width="600" style="border:none;"></iframe></center>



</div>
</div>


```{r thecode, eval=FALSE}
# library
library(rgl)

# This is to output a rgl plot in a rmarkdown document. Note that you must add webgl=TRUE, results='hide' in the chunck header
#library(knitr)
#knit_hooks$set(webgl = hook_webgl)

# Data: the iris data is provided by R
data <- iris

# Add a new column with color
mycolors <- c('royalblue1', 'darkcyan', 'oldlace')
data$color <- mycolors[ as.numeric(data$Species) ]

# Plot
par(mar=c(0,0,0,0))
plot3d( 
  x=data$`Sepal.Length`, y=data$`Sepal.Width`, z=data$`Petal.Length`, 
  col = data$color, 
  type = 's', 
  radius = .1,
  xlab="Sepal Length", ylab="Sepal Width", zlab="Petal Length")

writeWebGL( filename="HtmlWidget/3dscatter.html" ,  width=600, height=600)

```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedOther.html")
```
