---
title: "Basic waffle chart in R"
descriptionMeta: "This post explains how to build a very basic waffle chart with R. It uses the waffle package, provides reproducible code and explain how input data must be formatted."
descriptionTop: "This post explains how to build a **very basic waffle chart** with R. It uses the [waffle](waffle.html) package, provides reproducible code and explain how input data must be formatted.<br><br>For a more <b>advanced usage</b> of waffle charts, check out [this post](406-geom-waffle-introduction.html)"
sectionText: "Waffle section"
sectionLink: "waffle.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "405-very-basic-waffle"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Libraries and dataset
***

The `waffle` package provides a `waffle()` function that allows to build waffle charts.

*Install the package with `install.packages("waffle")`.*

The input dataset is simple: we just have **3 groups**, and **each one has a value**.

```{r}
# library
#install.packages("waffle")
library(waffle)
 
# Create data
group <- c("group-1","group-2","group-3")
value <- c(13,5,22)
data <- data.frame(group,value)
```


# Most basic waffle
***

Thanks to the `waffle` function, we can easily build a waffle chart.

```{r thecode}
waffle(data)
```



# Change output
***

The default output will have 10 rows, but you can change this with the `rows` argument:

```{r thecode2}
waffle(data, rows = 5)
```



# Colors
***

You can change the color of the waffle chart with the `colors` argument. Here is an example with 3 colors:

```{r thecode3}
waffle(data, rows = 5, colors = c("#fb5607", "#ff006e", "#8338ec"))
```


# Legend position
***

You can change the position of the legend with the `legend_pos` argument. Here is an example with the legend on the bottom:

```{r thecode4}
waffle(data, rows = 5, colors = c("#fb5607", "#ff006e", "#8338ec"), legend_pos = "bottom")
```



# Going further
***

You might be interested in creating [waffle charts with ggplot](406-geom-waffle-introduction.html) and [waffle charts with subgroups](407-geom-waffle-multiple-groups.html).



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
