---
title: "Waffle chart with waffle and ggplot2"
descriptionMeta: "This post explains how to build a waffle chart with R using the geom_waffle() function. It uses the waffle package, provides reproducible code and explain how to customize the chart."
descriptionTop: "This post explains how to build a [waffle chart](waffle.html) with R using the `geom_waffle()` function. It uses the `waffle` package, provides reproducible code and explain how to **customize the chart**."
sectionText: "Waffle section"
sectionLink: "waffle.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "406-geom-waffle-introduction"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Libraries and dataset
***

*Install the package with `install.packages("waffle")`.*

The input dataset is simple: we just have **3 groups**, and **each one has a value**.

```{r}
# library
#install.packages("waffle")
library(waffle)
 
# Create data
group <- c("group-1","group-2","group-3")
value <- c(13,5,22)
data <- data.frame(group,value)
```



# Waffle with ggplot2
***

The `waffle` package provides a `geom_waffle()` function that allows to build **waffle charts with ggplot2**.

```{r thecode}
ggplot(data, aes(fill=group, values=value)) +
  geom_waffle() +
  theme_void()
```



# Colors
***

You can **customize the colors** of the waffle chart using the `scale_fill_manual()` function and the edge color with the `color` argument:

```{r thecode2}
ggplot(data, aes(fill=group, values=value)) +
  geom_waffle(color = "white") +
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9")) +
  theme_void()
```



# Labels
***

You can **change labels** in legend to the waffle chart:

```{r thecode3}
ggplot(data, aes(fill=group, values=value)) +
  geom_waffle() +
  scale_fill_manual(
    values = c("#999999", "#E69F00", "#56B4E9"),
    labels = c("First group", "Second group", "Third group")) +
  theme_void()
```




# Going further
***

You might be interested in how to create [waffle chart without ggplot](405-very-basic-waffle.html) and how to create [waffle charts with groups and subgroups](407-geom-waffle-multiple-groups.html).



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
