---
title: "Waffle chart with groups and subgroups"
descriptionMeta: "This post explains how to build a waffle chart when have both groups and subgroups using the geom_waffle() function. It uses the waffle package, provides reproducible code and explain how to customize the chart."
descriptionTop: "This post explains how to build a [waffle chart](waffle.html) when have both **groups and subgroups** using the `geom_waffle()` function. For an introduction to [waffle chart](waffle.html), check [this post](406-geom-waffle-introduction.html)"
sectionText: "Waffle section"
sectionLink: "waffle.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "407-geom-waffle-multiple-groups"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Libraries and dataset
***

First we need to load the `waffle` and `dplyr` libraries.

The input dataset contains **groups and subgroups**, with **values of their size**.

```{r}
# library
library(waffle)
library(dplyr)
 
# Create data
data <- data.frame(
  group = c("First group", "First group", "First group", "First group",
            "First group", "First group", "Second group", "Second group",
            "Second group", "Second group", "Third group", "Third group"),
  subgroup = c("A", "B", "C", "D", "E", "F", "A", "B", "C", "D", "A", "B"),
  value = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120)
)
```



# Grouped waffle chart
***

The `geom_waffle()` function can be used to build a waffle chart with groups and subgroups. The `fill` argument is used to color the chart by group.

```{r}
ggplot(data = data, aes(fill=subgroup, values=value)) +
  geom_waffle(color = "white", size = 1.125, n_rows = 6) +
  facet_wrap(~group, ncol=1) +
  theme_void()
```



# Custom colors
***

Thanks to the `scale_fill_manual()` function, it is possible to **customize the colors** of the waffle chart:

```{r}
ggplot(data = data, aes(fill=subgroup, values=value)) +
  geom_waffle(color = "white", size = 1.125, n_rows = 6) +
  facet_wrap(~group, ncol=1) +
  theme_void() +
  scale_fill_manual(values = c("#69b3a2", "#404080", "#FFA07A", "#FFD700", "#FF6347", "#4682B4"))
```




# Going further
***

You might be interested in this [introduction to geom_waffle()](406-geom-waffle-introduction.html) or more generally about [waffle charts](waffle.html).




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
