---
title: "Customized bump plot"
descriptionMeta: "This post explains how to build and customize bump plot with R. It uses the ggbump package, provides reproducible code and explain how input data must be formatted."
descriptionTop: "This post explains how to build and customize a **bump plot** with R. It uses the [ggbump](package/ggbump.html) package, provides reproducible code and explain how input data must be formatted.<br><br>For an introduction to bump plots, see [this introduction](408-basic-bump-plot.html)."
sectionText: "Bump plot"
sectionLink: "package/ggbump.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/parallel.html"
url: "409-customize-bump-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  fig.align='center',
  fig.width=8,
  fig.height=6
)
```

<div class="container">


# Libraries and dataset
***

The [ggbump](package/ggbump.html) package provides a `geom_bump()` function that allows to **build ggbump charts**.

Install the package with `install.packages("ggbump")`.

The input dataset is simple: we just have **3 groups**, with **one value** per group and per year. Here is how to build it:

```{r}
# Library
#install.packages("ggbump")
library(ggbump)
library(tidyverse)

# Create data
year <- rep(2019:2021, 3)
products_sold <- c(
  500, 600, 700,
  550, 650, 600,
  600, 400, 500
)
store <- c(
  "Store A", "Store A", "Store A",
  "Store B", "Store B", "Store B",
  "Store C", "Store C", "Store C"
)

# Create the new dataframe
df <- data.frame(
  year = year,
  products_sold = products_sold,
  store = store
)
```


# Simple bump plot
***

Thanks to the `geom_bump()` function, we can easily build a [ggbump](package/ggbump.html) chart.

```{r thecode}
ggplot(df, aes(x = year, y = products_sold, color = store)) +
  geom_bump(size = 2) +
  geom_point(size = 6)
```



# Change colors
***

It is possible to **add individual points** to the bump chart. This is done by adding a `geom_point()` layer.

```{r thecode2}
ggplot(df, aes(x = year, y = products_sold, color = store)) +
  geom_bump(size = 2) +
  geom_point(size = 6) +
  scale_color_brewer(palette = "Paired") +
  theme_minimal()
```


# Add labels and title
***

The `geom_text()` and `labs()` functions can be used to add labels and a title to the chart.

```{r thecode3}
ggplot(df, aes(x = year, y = products_sold, color = store)) +
  geom_bump(size = 2) +
  geom_point(size = 6) +
  geom_text(aes(label = store), nudge_y = 20, fontface = "bold", size=3) +
  scale_color_brewer(palette = "Paired") +
  theme_minimal() +
  labs(
    title = "Products sold per store",
    x = "Year",
    y = "Products sold"
  )
```


# Going further
***

You might be interested in

- learning how to use the [ggbump](package/ggbump.html) package
- the basics of [bump plots](408-basic-bump-plot.html)
- how to [highlight specific lines](web-bump-plot-with-highlights.html) in a bump plot



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
