---
title: "2d maps with rayshader"
descriptionMeta: "This post explains how to build and customize 2d maps with the rayshader package. It provides reproducible code and explain how input data must be formatted."
descriptionTop: "This post thoroughly details the process of building and customizing **2D maps** using the [rayshader](package/rayshader.html) package. It includes step-by-step, reproducible code examples and clearly explains the required formatting for input data.<br><br>By following this guide, you will learn how to create **visually appealing maps** and tailor them to meet your specific needs, ensuring your data is presented in the most effective way possible. Additionally, the post covers various customization options within **rayshader**, enabling you to enhance your maps with different visual **styles and effects**."
sectionText: "Map section"
sectionLink: "map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/map.html"
url: "410-map-2d-with-rayshader"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# Libraries
***

The [rayshader package](package/rayshader.html) makes it **simple** to create **shaded 2D relief maps**.

Since the package is on **CRAN**, you can **install it** with `install.packages("rayshader")`.

We also load the **`raster` package** to **access data** to work with.

```{r}
# install.packages("rayshader")
library(rayshader)
library(raster)
```


# Data format
***

The [rayshader](package/rayshader.html) package requires input data in an **elevation matrix format**. This is a specific matrix where each cell holds the **elevation value** for the corresponding map point.

Here's how to obtain the **elevation matrix** of the **Grand Canyon** from a raster file using the `raster` package.

```{r}
# Define a region for the SRTM data (example: Grand Canyon)
extent_gc <- extent(-113.0, -112.0, 36.0, 37.0)
srtm_gc <- getData("SRTM", lon = -112.5, lat = 36.5)
srtm_gc_cropped <- crop(srtm_gc, extent_gc)

# Convert the raster data to matrix
elevation_matrix <- raster_to_matrix(srtm_gc_cropped)
```


# Basic 2D map
***

Let's see how to create a basic 2D map with [rayshader](package/rayshader.html).

We use the `sphere_shade()` function to create the map. It takes the **elevation matrix** as input and returns a shaded map, and then plot it with the `plot_map()` function.

```{r}
elevation_matrix %>%
  sphere_shade(texture="desert", sunangle = 45) %>%
  plot_map()
```


# Change texture
***

[rayshader](package/rayshader.html) includes a **specialized function** to create textures: `create_texture()`. This function accepts **5 colors as input** and generates a texture suitable for use in the `sphere_shade()` function.

```{r}
custom_texture <- create_texture("#fff673","#55967a","#8fb28a","#55967a","#cfe0a9")

elevation_matrix %>%
  sphere_shade(texture=custom_texture, sunangle = 45) %>%
  plot_map()
```

# Change shade properties
***

The `add_shadow()` function can be used to add a shadow to the map. The `ray_shade()` function creates a shadow based on the sun angle, while the `ambient_shade()` function creates a shadow based on the ambient light.

```{r, cache=TRUE}
elevation_matrix %>%
  sphere_shade(texture="desert", sunangle = 45, zscale = 50) %>%
  add_shadow(ray_shade(elevation_matrix), 0.5) %>%  # Adds a ray-traced shadow
  add_shadow(ambient_shade(elevation_matrix), 0) %>%  # Adds an ambient shadow
  plot_map()
```



# Going further
***

You might be interested in

- creating [3d maps with rayshader](411-map-3d-with-rayshader.html)
- learning more about [rayshader with more examples](package/rayshader.html)



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
