---
title: "Customize CSS in ggiraph"
descriptionMeta: "This post explains how to use the ggiraph package to create interactive graph and how to customize it with additional CSS. It provides reproducible code and explanations."
descriptionTop: "This post explains how to use the [ggiraph](package/ggiraph.html) package to create **interactive graph** and how to customize it with additional CSS, with reproducible code and explanations.<br><br>For an introduction to `ggiraph`, check the [dedicated post](package/ggiraph.html)"
sectionText: "Interactive section"
sectionLink: "interactive-charts.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "412-customize-css-in-interactive-ggiraph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">


# Libraries and dataset
***

First we need to load the [ggiraph](package/ggiraph.html) and [ggplot2](ggplot2-package.html) libraries.

The input dataset is about **consumer confidence** in 9 different countries at differente dates:

```{r}
# library
library(ggplot2)
library(ggiraph)
library(tidyverse)

# Create data
data <- read.csv("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/dataConsumerConfidence.csv") %>%
  mutate(date = lubridate::my(Time)) %>%
  select(-Time) %>%
  pivot_longer(!date, names_to = "country", values_to = "value") %>%
  na.omit() %>%
  mutate(country = as.factor(country))
```



# Simple line chart with ggiraph
***

The `ggiraph` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **creating interactive charts**.

It offers a set of function to **add interactivity** such as `geom_point_interactive()` for an interactive version of `geom_point()`, `geom_sf_interactive()` for an interactive version of `geom_sf()` etc.

Let's start with a **simple interactive version** of this line chart. It uses the `geom_line_interactive()` function to replace the `geom_line()` one and make it interactive.

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = date,
    y = value,
    color = country,
    tooltip = country,
    data_id = country
  )) +
  geom_line_interactive(hover_nearest = TRUE) +
  theme_classic()

interactive_plot <- girafe(ggobj = plot)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-line-interactive-1.html")
```

<iframe src="HtmlWidget/geom-line-interactive-1.html" height="650" width="850" style="border:none;"></iframe>



# Change CSS
***

Although `CSS` is primarily used for styling **web pages**, it can also be relevant in customizing visuals in **R**, especially when dealing with web-based outputs like interactive charts or Shiny applications.

`CSS` code has the following structure: a property named (such as `font-size`), a colon (`:`), and the value of that property (such as `14px`). Each element is then separated by a `;`.

The easiest way to add `CSS` in [ggiraph](package/ggiraph.html) is to use the `girafe_options()` function.

Then we use `opts_hover(css = "fill:#ffe7a6;stroke:black;cursor:pointer;")` to **fill the lines** on hover with a given color and change the color of the lines to black.

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = date,
    y = value,
    color = country,
    tooltip = country,
    data_id = country
  )) +
  geom_line_interactive(hover_nearest = TRUE) +
  theme_classic()

interactive_plot <- girafe(ggobj = plot)

interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:#ffe7a6;stroke:black;cursor:pointer;"),
  opts_selection(type = "single", css = "fill:red;stroke:black;"),
  opts_toolbar(saveaspng = FALSE)
)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-line-interactive-2.html")
```

<iframe src="HtmlWidget/geom-line-interactive-2.html" height="650" width="850" style="border:none;"></iframe>


# Highlight a specific line
***

When displaying multiple lines in a single chart, it's easy to end up with a cluttered [spaghetti chart](https://www.data-to-viz.com/caveat/spaghetti.html). One effective way to enhance readability is by implementing a hover effect that elegantly **fades out** non-hovered lines while highlighting the selected line.

This can be achieved using CSS with the following properties:

1. `stroke`: Change the color of the hovered line (e.g., `stroke: #69B3A2;`)
2. `stroke-width`: Increase the width of the hovered line for emphasis
3. `transition`: Add a smooth transition effect (e.g., `transition: all 0.3s ease;`)
4. `opacity`: Reduce the opacity of non-hovered lines (e.g., `opacity: 0.5;`)
5. `filter`: Apply a grayscale effect to non-hovered lines (e.g., `filter: grayscale(90%);`)

```{r}
library(hrbrthemes) # for the `theme_ipsum()`

plot <- data %>%
  ggplot(mapping = aes(
    x = date,
    y = value,
    color = country,
    tooltip = country,
    data_id = country
  )) +
  geom_line_interactive(hover_nearest = TRUE) +
  theme_ipsum()

interactive_plot <- girafe(ggobj = plot)

interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "stroke:#69B3A2; stroke-width: 3px; transition: all 0.3s ease;"),
  opts_hover_inv("opacity:0.5;filter:saturate(10%);"),
  opts_toolbar(saveaspng = FALSE)
)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-line-interactive-3.html")
```

<iframe src="HtmlWidget/geom-line-interactive-3.html" height="650" width="850" style="border:none;"></iframe>


# More advanced usage
***

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = date,
    y = value,
    color = country,
    tooltip = country,
    data_id = country
  )) +
  geom_line_interactive(hover_nearest = TRUE) +
  theme_classic()

interactive_plot <- girafe(ggobj = plot)

hover_css <- "
  fill: #ffe7a6;
  fill-opacity: 0.5;
  stroke: black;
  stroke-width: 7px;
  stroke-dasharray: 5,5;
  transition: fill-opacity 0.5s, stroke-width 0.5s, stroke-dasharray 0.5s, filter 0.5s;
  filter: drop-shadow(0 0 5px rgba(0,0,0,0.5));
"

interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = hover_css),
  opts_toolbar(saveaspng = FALSE)
)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-line-interactive-4.html")
```

<iframe src="HtmlWidget/geom-line-interactive-4.html" height="650" width="850" style="border:none;"></iframe>


# Combine CSS with other ggiraph features
***

Since `CSS` provides an infinite number of possibilities, you can use it to create very original charts!

Hover CSS:

- `stroke: black;`: Sets the outline color of the hovered element to black.
- `stroke-width: 1px;`: Sets the width of the outline to 1 pixel.
- `r: 8px;`: Sets the radius of circular elements (like points) to 8 pixels, making them larger on hover.
- `transition: all 0.3s ease;`: Applies a smooth transition effect for all property changes, lasting 0.3 seconds.

Tooltip CSS:

- `background-color: #2C3E50;`: Sets a dark blue background for the tooltip.
- `color: #ECF0F1;`: Sets the text color to a light gray.
- `padding: 10px;`: Adds 10 pixels of space inside the tooltip around its content.
- `border-radius: 5px;`: Rounds the corners of the tooltip.
- `font-family: 'Arial', sans-serif;`: Sets the font to Arial or a sans-serif fallback.
- `font-size: 14px;`: Sets the text size to 14 pixels.
- `box-shadow: 0px 0px 10px rgba(0,0,0,0.5);`: Adds a subtle shadow effect to the tooltip.


```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = date,
    y = value,
    color = country,
    group = country,
    tooltip = paste("Country:", country, "<br>Date:", date, "<br>Value:", round(value, 2)),
    data_id = country
  )) +
  geom_line_interactive(size = 1.2, hover_nearest = TRUE) +
  geom_point_interactive(aes(size = value), alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size_continuous(range = c(1, 2)) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Interactive Country Data Visualization",
    subtitle = "Try to hover and click on the lines!",
    caption = "R-Graph-Gallery.com",
    x = "Date",
    y = "Consumer Confidence"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 16, face = "italic"),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "ivory", color = NA),
    plot.background = element_rect(fill = "ivory", color = NA)
  )

interactive_plot <- girafe(ggobj = plot)

hover_css <- "
  stroke: black;
  stroke-width: 1px;
  r: 8px;
  transition: all 0.3s ease;
"

tooltip_css <- "
  background-color: #2C3E50;
  color: #ECF0F1;
  padding: 10px;
  border-radius: 5px;
  font-family: 'Arial', sans-serif;
  font-size: 14px;
  box-shadow: 0px 0px 10px rgba(0,0,0,0.5);
"

interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = hover_css),
  opts_tooltip(css = tooltip_css, use_fill = TRUE),
  opts_selection(type = "multiple", only_shiny = FALSE),
  opts_zoom(min = 0.5, max = 2),
  opts_toolbar(saveaspng = TRUE, position = "topright", pngname = "country_data_plot"),
  opts_sizing(rescale = TRUE)
)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-line-interactive-5.html")
```

<iframe src="HtmlWidget/geom-line-interactive-5.html" height="750" width="890" style="border:none;"></iframe>


# Going further
***

You might be interested in:

- learn how [integrate JavaScript in ggiraph charts](413-use-javascript-in-ggiraph.html)
- discover all [ggiraph features](package/ggiraph.html)
- browse the [interactive chart section](interactive-charts.html)




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
