---
title: "Use JavaScript for ggiraph"
descriptionMeta: "This post explains how to use the ggiraph package to create interactive graph and how to use JavaScript to add additional effects for your ggplot interactive graphs."
descriptionTop: "This post explains how to use the [ggiraph](package/ggiraph.html) package to create **interactive graph** and how to use **JavaScript** to add additional effects for your ggplot interactive graphs, with reproducible code and explanations.<br><br>For an introduction to `ggiraph`, check the [dedicated post](package/ggiraph.html)"
sectionText: "Interactive section"
sectionLink: "interactive-charts.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "413-use-javascript-in-ggiraph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">


# Libraries and dataset
***

First we need to load the [ggiraph](package/ggiraph.html) and [ggplot2](ggplot2-package.html) libraries.

The input dataset is about **consumer confidence** in 9 different countries at differente dates:

```{r}
# library
library(ggplot2)
library(ggiraph)
library(tidyverse)
library(gapminder)
library(dplyr)
library(hrbrthemes)
library(viridis)

# The dataset is provided in the gapminder library
data <- gapminder %>%
  filter(year == "2007") %>%
  dplyr::select(-year) %>%
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country))
```



# Simple bubble chart
***

Let's start with a **simple interactive version** of this bubble chart. It uses the `geom_point_interactive()` function to replace the `geom_point()` one and make it interactive.

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = gdpPercap,
    y = lifeExp,
    size = pop,
    fill = continent,
    tooltip = paste("Country:", country, "<br>Life Expectancy:", lifeExp, "<br>GDP per capita:", gdpPercap),
    data_id = continent
  )) +
  geom_point_interactive(alpha = 0.5, shape = 21, color = "black") +
  scale_size(range = c(.1, 24), name = "Population (M)") +
  scale_fill_viridis(discrete = TRUE, guide = FALSE, option = "A") +
  theme_ipsum() +
  ylab("Life Expectancy") +
  xlab("Gdp per Capita") +
  theme(legend.position = "none")

interactive_plot <- girafe(ggobj = plot)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-point-interactive-1.html")
```

<iframe src="HtmlWidget/geom-point-interactive-1.html" height="650" width="850" style="border:none;"></iframe>



# JavaScript and the `onclick` argument
***

**JavaScript** is a powerful programming language primarily used to create **interactive and dynamic** content on web pages. When it comes to integrating **JavaScript** with R, particularly in the context of charts, one common use case involves the `onclick` **event handler**.

The `onclick` argument in **JavaScript** allows you to specify functionality that should occur when an element, like a part of a chart, is clicked.

Let's see a basic example where a simple **pop-up** appears when clicking on one of the bubble.

*Note: all quotes `"` inside JS code must be preceded by a `\` for R to understand them correctly.*

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = gdpPercap,
    y = lifeExp,
    size = pop,
    fill = continent,
    tooltip = paste("Country:", country, "<br>Life Expectancy:", lifeExp, "<br>GDP/capita:", gdpPercap),
    data_id = continent
  )) +
  geom_point_interactive(
    alpha = 0.5,
    shape = 21,
    color = "black",
    onclick = "alert(\"Hello from R-Graph-Gallery.com\");"
  ) +
  scale_size(range = c(.1, 24), name = "Population (M)") +
  scale_fill_viridis(discrete = TRUE, guide = FALSE, option = "A") +
  theme_ipsum() +
  ylab("Life Expectancy") +
  xlab("Gdp per Capita") +
  labs(title = "Try to click on a bubble!") +
  theme(legend.position = "none")

interactive_plot <- girafe(ggobj = plot)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-point-interactive-2.html")
```

<iframe src="HtmlWidget/geom-point-interactive-2.html" height="650" width="850" style="border:none;"></iframe>



# Change properties on click
***

The following code uses [D3.js](https://d3-graph-gallery.com/) to **change the properties** of the clicked point.

For example: `d3.select(this).style(\"fill\", \"red\");` tells to your computer that `this` (**the point** you just clicked on) change its `style` so that the `fill` parameter becomes `red`.

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = gdpPercap,
    y = lifeExp,
    size = pop,
    fill = continent,
    tooltip = paste("Country:", country, "<br>Life Expectancy:", lifeExp, "<br>GDP/capita:", gdpPercap),
    data_id = continent
  )) +
  geom_point_interactive(
    alpha = 0.5,
    shape = 21,
    color = "black",
    onclick = "
      // Change point color
      d3.select(this).style(\"fill\", \"red\");

      // Increase point size
      d3.select(this).attr(\"r\", function() {
        return parseFloat(d3.select(this).attr(\"r\")) + 1.5;
      });
    "
  ) +
  scale_size(range = c(.1, 24), name = "Population (M)") +
  scale_fill_viridis(discrete = TRUE, guide = FALSE, option = "A") +
  theme_ipsum() +
  ylab("Life Expectancy") +
  xlab("Gdp per Capita") +
  labs(title = "Try to click on a bubble!") +
  theme(legend.position = "none")

interactive_plot <- girafe(ggobj = plot)
htmltools::save_html(interactive_plot, "HtmlWidget/geom-point-interactive-3.html")
```

<iframe src="HtmlWidget/geom-point-interactive-3.html" height="650" width="850" style="border:none;"></iframe>



# A more advanced example
***

Since you can do anything you want with **JavaScript**, the only limitation is your imagination!

The following example shows how to add a **confetti effect** when clicking + some other simple features using the `onclick` argument and a bit of **JavaScript**

```{r}
plot <- data %>%
  ggplot(mapping = aes(
    x = gdpPercap,
    y = lifeExp,
    size = pop,
    fill = continent,
    tooltip = paste("Country:", country, "<br>Life Expectancy:", lifeExp, "<br>GDP/capita:", gdpPercap),
    data_id = continent
  )) +
  geom_point_interactive(
    alpha = 0.5,
    shape = 21,
    color = "black",
    onclick = paste0(
      "this.style.fill = this.style.fill === \"red\" ? this.getAttribute(\"original-fill\") : \"red\";",
      "this.classList.toggle(\"highlighted\");",
      "var tooltip = document.getElementById(\"custom-tooltip\");",
      "tooltip.innerHTML = \"A chart by: R-Graph-Gallery.com\";",
      "tooltip.style.display = \"block\";",
      "tooltip.style.left = (event.pageX + 10) + \"px\";",
      "tooltip.style.top = (event.pageY + 10) + \"px\";",
      "setTimeout(function() { tooltip.style.display = \"none\"; }, 1500);",
      "confetti({
        particleCount: 1000,
        spread: 70,
        origin: { y: 0.6 }
      });"
    )
  ) +
  scale_size(range = c(.1, 24), name = "Population (M)") +
  scale_fill_viridis(discrete = TRUE, guide = FALSE, option = "A") +
  theme_ipsum() +
  ylab("Life Expectancy") +
  xlab("Gdp per Capita") +
  labs(title = "Click on a bubble for amazing effects!") +
  theme(legend.position = "none")

interactive_plot <- girafe(ggobj = plot)

html_content <- htmltools::tags$html(
  htmltools::tags$head(
    htmltools::tags$script(src = "https://cdn.jsdelivr.net/npm/canvas-confetti@1.5.1/dist/confetti.browser.min.js"),
    htmltools::tags$style("
      .highlighted {
        stroke: black;
        stroke-width: 10px;
      }
      #custom-tooltip {
        position: absolute;
        background: white;
        border: 1px solid black;
        padding: 5px;
        display: none;
        pointer-events: none;
      }
    ")
  ),
  htmltools::tags$body(
    interactive_plot,
    htmltools::tags$div(id = "custom-tooltip")
  )
)
htmltools::save_html(html_content, "HtmlWidget/geom-point-interactive-4.html")
```

<iframe src="HtmlWidget/geom-point-interactive-4.html" height="650" width="850" style="border:none;"></iframe>


# Going further
***

You might be interested in:

- learn how [integrate CSS in ggiraph charts](412-customize-css-in-interactive-ggiraph.html)
- discover all [ggiraph features](package/ggiraph.html)
- browse the [interactive chart section](interactive-charts.html)




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
