---
title: "Combine charts in ggiraph"
descriptionMeta: "This post explains how to use the ggiraph package to create and link multiple interactive graphs, from very simple example to more complex ones."
descriptionTop: "This post explains how to use the [ggiraph](package/ggiraph.html) package to create and link **multiple interactive graphs**, from very simple example to more complex ones.<br><br>For an introduction to `ggiraph`, check the [dedicated post](package/ggiraph.html)"
sectionText: "Interactive section"
sectionLink: "interactive-charts.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "414-map-multiple-charts-in-ggiraph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Libraries
***

First of all, we need to load a bunch of libraries:


```{r}
library(ggiraph) # install.packages('ggiraph')
library(ggplot2) # install.packages('ggplot2')
library(dplyr) # install.packages('dplyr')
library(patchwork) # install.packages('patchwork')
library(tidyr) # install.packages('tidyr')
library(sf) # install.packages('sf')
set.seed(123)
```


# Most simple example
***

In order to **map the hover effect** between multiple charts, we **only** have to use the `data_id` argument. For example, in our case, we specify `data_id = car` since we want our mapping at the **car level**, using the `mtcars` dataset. 

Then we put the charts together using the [patchwork package](package/patchwork.html)

```{r}
library(ggiraph)
library(ggplot2)
library(dplyr)
library(patchwork)

data(mtcars)
mtcars$car <- rownames(mtcars)

# data_id in the aes mapping
p1 <- ggplot(mtcars, aes(wt, mpg, tooltip = car, data_id = car)) +
  geom_point_interactive(size = 4)

# data_id in the aes mapping
p2 <- ggplot(mtcars, aes(x = reorder(car, mpg), y = mpg, tooltip = car, data_id = car)) +
  geom_col_interactive() +
  coord_flip()

combined_plot <- p1 + p2 + plot_layout(ncol = 2)

interactive_plot <- girafe(ggobj = combined_plot)
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-1.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-1.html" height="650" width="850" style="border:none;"></iframe>


# Customize charts as usual
***

## Create dataset

Let's start by creating a simple dataset for our next charts:

```{r}
# Read the full world map
world_sf <- read_sf("https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/world.geojson")

# Create a sample dataset
happiness_data <- data.frame(
  Country = c(
    "France", "Germany", "United Kingdom",
    "Japan", "China", "Vietnam",
    "United States of America", "Canada", "Mexico"
  ),
  Continent = c(
    "Europe", "Europe", "Europe",
    "Asia", "Asia", "Asia",
    "North America", "North America", "North America"
  ),
  Happiness_Score = rnorm(mean = 30, sd = 20, n = 9),
  GDP_per_capita = rnorm(mean = 30, sd = 20, n = 9),
  Social_support = rnorm(mean = 30, sd = 20, n = 9),
  Healthy_life_expectancy = rnorm(mean = 30, sd = 20, n = 9)
)

# Join the happiness data with the full world map
world_sf <- world_sf %>%
  left_join(happiness_data, by = c("name" = "Country"))
```

## Create the charts

Since we only need to specify the column of our dataset in `data_id`, the rest of our charts is **customizable** as any other [ggplot2](ggplot2-package.html) chart!

```{r}
# Create the first chart (Scatter plot)
p1 <- ggplot(world_sf, aes(
  GDP_per_capita,
  Happiness_Score,
  tooltip = name,
  data_id = name,
  color = name
)) +
  geom_point_interactive(data = filter(world_sf, !is.na(Happiness_Score)), size = 4) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Create the second chart (Bar plot)
p2 <- ggplot(world_sf, aes(
  x = reorder(name, Happiness_Score),
  y = Happiness_Score,
  tooltip = name,
  data_id = name,
  fill = name
)) +
  geom_col_interactive(data = filter(world_sf, !is.na(Happiness_Score))) +
  coord_flip() +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Create the third chart (choropleth)
p3 <- ggplot() +
  geom_sf(data = world_sf, fill = "lightgrey", color = "lightgrey") +
  geom_sf_interactive(
    data = filter(world_sf, !is.na(Happiness_Score)),
    aes(fill = name, tooltip = name, data_id = name)
  ) +
  theme_void() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Combine the plots
combined_plot <- (p1 + p2) / p3 + plot_layout(heights = c(1, 2))

# Create the interactive plot
interactive_plot <- girafe(ggobj = combined_plot)
interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:black;stroke:black;")
)

# save as an html widget
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-2.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-2.html" height="800" width="850" style="border:none;"></iframe>

# Highlight all the items of a group
***

Previously, the interactivity and connection between charts were at the "country" level, meaning that when you **hovered over a country**, the same country was **highlighted** in other charts.

But what if we want to use the "continent" level instead? All we need to do is **change** `data_id = name` (country) to `data_id = Continent`. Additionally, we **adjust the color mapping** so that all countries within a given continent share the same color.

Give it a try by hovering over the charts below:

```{r}
# Create the first chart (Scatter plot)
p1 <- ggplot(world_sf, aes(
  GDP_per_capita,
  Happiness_Score,
  tooltip = name,
  data_id = Continent,
  color = Continent
)) +
  geom_point_interactive(data = filter(world_sf, !is.na(Happiness_Score)), size = 4) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Create the second chart (Bar plot)
p2 <- ggplot(world_sf, aes(
  x = reorder(name, Happiness_Score),
  y = Happiness_Score,
  tooltip = name,
  data_id = Continent,
  fill = Continent
)) +
  geom_col_interactive(data = filter(world_sf, !is.na(Happiness_Score))) +
  coord_flip() +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Create the third chart (choropleth)
p3 <- ggplot() +
  geom_sf(data = world_sf, fill = "lightgrey", color = "lightgrey") +
  geom_sf_interactive(
    data = filter(world_sf, !is.na(Happiness_Score)),
    aes(fill = Continent, tooltip = name, data_id = Continent)
  ) +
  theme_void() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )

# Combine the plots
combined_plot <- (p1 + p2) / p3 + plot_layout(heights = c(1, 2))

# Create the interactive plot
interactive_plot <- girafe(ggobj = combined_plot)
interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:black;stroke:black;")
)

# Save the interactive plot
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-3.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-3.html" height="800" width="850" style="border:none;"></iframe>



# Customise hover effect with CSS
***

Although `CSS` is primarily used for styling **web pages**, it can also be relevant in customizing visuals in **R**, especially when dealing with web-based outputs like **interactive charts** or **Shiny applications**.

`CSS` code has the following structure: a property named (such as `font-size`), a colon (`:`), and the value of that property (`14px` in the following example). Each element is then separated by a `;`.

_Note: If you haven't already, check the post on [how to use CSS in ggiraph](412-customize-css-in-interactive-ggiraph.html)._

In this example, we'll use the `girafe_options()` function and pass it several arguments:
- `opts_hover()` for passing the CSS when **hovering**
- `opts_tooltip()` for passing the CSS of the **tooltip**

```{r}
tooltip_css <- "
  border-radius: 12px;
  color: #333;
  background-color: white;
  padding: 10px;
  font-size: 14px;
"

hover_css <- "fill: #4e54c8;"

# Add interactivity
interactive_plot <- interactive_plot %>%
  girafe_options(
    opts_hover(css = hover_css),
    opts_tooltip(css = tooltip_css)
  )

# Save the interactive plot
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-4.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-4.html" height="800" width="850" style="border:none;"></iframe>


# More complex CSS
***

Since [CSS](412-customize-css-in-interactive-ggiraph.html) allows a super **high level of customization**, we can use it to create way more complex and advanced examples!

In the following case:

Tooltip CSS:

- `background`: Creates a gradient from light gray to slightly darker gray
- `border`: Removes any border
- `border-radius`: Rounds the corners of the tooltip
- `box-shadow`: Adds a subtle 3D effect with light and shadow
- `color`: Sets the text color to dark gray
- `font-family`: Specifies the font types to use
- `font-size`: Sets the text size
- `padding`: Adds space inside the tooltip
- `transition`: Enables smooth changes in properties over 0.5 seconds

Hover CSS:

- `fill`: Changes the main color of the hovered element to purple
- `stroke`: Adds a light blue outline
- `stroke-width`: Sets the thickness of the outline
- `cursor`: Changes the mouse cursor to a pointing hand
- `filter`: Increases brightness and adds a soft glow effect
- `transition`: Enables smooth changes in properties over 0.5 seconds

```{r}
tooltip_css <- "
  background: linear-gradient(145deg, #f0f0f0, #e6e6e6);
  border: none;
  border-radius: 12px;
  box-shadow: 3px 3px 10px #d1d1d1, -3px -3px 10px #ffffff;
  color: #333;
  font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
  font-size: 14px;
  padding: 12px;
  transition: all 0.5s ease-out;
"

hover_css <- "
  fill: #4e54c8;
  stroke: #8f94fb;
  cursor: pointer;
  filter: brightness(1.2) drop-shadow(0 0 5px rgba(78, 84, 200, 0.5));
  transition: all 0.5s ease-out;
"

# Add interactivity
interactive_plot <- interactive_plot %>%
  girafe_options(
    opts_hover(css = hover_css),
    opts_tooltip(css = tooltip_css)
  )

# Save the interactive plot
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-5.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-5.html" height="800" width="850" style="border:none;"></iframe>

# Hide the groups: more hover effects
***

One effective method to **emphasize** the group you're hovering over is to **reduce the transparency of other groups**, allowing the user's focus to naturally gravitate towards the desired elements!

In practice, we utilize the `opts_hover_inv()` (defines what happen to non-hovered groups) function along with some `CSS`:

- **`opacity: 0.5;`**: Adjusts the element's transparency to 50%, making it semi-transparent and less prominent.

- **`filter: saturate(20%);`**: Decreases the color saturation to 20% of its original level, giving the element a dull, nearly grayscale appearance.

```{r}
tooltip_css <- "
  border-radius: 12px;
  color: #333;
  background-color: white;
  padding: 10px;
  font-size: 14px;
  transition: all 0.5s ease-out;
"

hover_css <- "
  fill: #4e54c8;
  cursor: pointer;
  transition: all 0.5s ease-out;
"

# Add interactivity
interactive_plot <- interactive_plot %>%
  girafe_options(
    opts_hover(css = hover_css),
    opts_tooltip(css = tooltip_css),
    opts_hover_inv(css = "opacity:0.5;filter:saturate(20%);")
  )

# Save the interactive plot
htmltools::save_html(interactive_plot, "HtmlWidget/multiple-ggiraph-6.html")
```

<iframe src="HtmlWidget/multiple-ggiraph-6.html" height="800" width="850" style="border:none;"></iframe>

# Going further
***

You might be interested in:

- learn how [integrate CSS in ggiraph charts](412-customize-css-in-interactive-ggiraph.html)
- learn how [integrate JavaScript in ggiraph charts](413-use-javascript-in-ggiraph.html)
- discover all [ggiraph features](package/ggiraph.html)
- browse the [interactive chart section](interactive-charts.html)




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
