---
title: "Drawing Arrows in ggplot2"
descriptionMeta: "This post explains how to draw arrows in ggplot2 with R. It covers three different methods: drawing the simplest arrow possible, customizing arrow style, and creating curved arrows."
descriptionTop: "This post explains how to draw **arrows in ggplot2** with R. It covers three different methods: drawing the simplest arrow possible, customizing arrow style, and creating curved arrows. We'll provide reproducible code and explain how it works for beginners."
sectionText: "ggplot2"
sectionLink: "ggplot2-package.html#annotation"
DataToVizText: "Data To Viz"
DataToVizLink: "data-to-viz.com/caveat/annotation.html"
url: "415-arrows-in-ggplot-graph"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Libraries and dataset
***

We'll be using the `ggplot2` package to create our plots and draw arrows. Let's start by loading the necessary libraries and creating a simple dataset:

```{r}
# Load libraries
library(ggplot2)
library(dplyr)

# Create a simple dataset
set.seed(123)
df <- data.frame(
  x = 1:10,
  y = rnorm(10, mean = 5, sd = 2)
)
```

# Drawing the Simplest Arrow Possible
***

The simplest way to draw an arrow in ggplot2 is by using the `geom_segment()` function with the `arrow` parameter.

In this code:

1. We start with a basic scatter plot using `geom_point()`.
2. We add `geom_segment()` to draw the arrow.
3. Inside `geom_segment()`, we specify the start (`x`, `y`) and end (`xend`, `yend`) coordinates of the arrow.
4. The `arrow()` function is used to add an arrowhead to the end of the segment.

```{r simplest_arrow}
ggplot(df, aes(x, y)) +
  geom_point() +
  geom_segment(aes(x = 2, y = 2, xend = 8, yend = 8),
    arrow = arrow()
  ) +
  theme_minimal()
```



# Customizing Arrow Style
***

We can customize various aspects of the arrow, such as its **color**, **size**, **type**, and the properties of the **arrowhead**.

1. We set the `length` of the arrowhead to 0.5 cm using `unit(0.5, "cm")`.
2. We change the arrowhead `type` to "closed" for a filled arrowhead.
3. We set the `color` of the arrow to red.
4. We increase the `size` (thickness) of the arrow to 1.5.
5. We change the `linetype` to "dashed".

```{r custom_arrow}
ggplot(df, aes(x, y)) +
  geom_point() +
  geom_segment(aes(x = 2, y = 2, xend = 8, yend = 8),
    arrow = arrow(length = unit(0.5, "cm"), type = "closed"),
    color = "red",
    size = 1.5,
    linetype = "dashed"
  ) +
  theme_minimal()
```

# Drawing Curved Arrows
***

To draw curved arrows, we can use `geom_curve()` instead of `geom_segment()`.

1. We replace `geom_segment()` with `geom_curve()`.
2. We add a `curvature` parameter to control the **bend of the arrow**. Negative values curve the arrow clockwise, while positive values curve it counterclockwise.
3. We adjust the `length` of the arrowhead to 0.3 cm for better proportions with the curve.
4. We set the arrow `color` to blue and its `size` to 1.2.

```{r curved_arrow}
ggplot(df, aes(x, y)) +
  geom_point() +
  geom_curve(aes(x = 2, y = 2, xend = 8, yend = 8),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    color = "blue",
    size = 1.2,
    curvature = -0.3
  ) +
  theme_minimal()
```



# Going further
***

You might be interested in:

- How [annotations work in ggplot](ggplot2-package.html#annotation)
- Learning how to [annotate a specific element](line_chart_annotation.html) in a single plot
- Exploring more [ggplot2 geoms](ggplot2-package.html) for different types of visualizations

<!-- Close container -->
</div>

```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedOther.html")
```