---
title: "Introduction to Thematic Maps with tmap"
descriptionMeta: "This post introduces the tmap package in R, explaining how to create thematic maps. It provides simple, reproducible code examples and explains the main concepts behind tmap."
descriptionTop: "This post introduces the [tmap](package/tmap.html) package in R, explaining how to create [thematic maps](https://www.r-graph-gallery.com/choropleth-map.html). It provides simple, reproducible code examples and explains the **main concepts** behind tmap."
sectionText: "Map section"
sectionLink: "map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "416-introduction-to-tmap"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Introduction to tmap
***

The [tmap](package/tmap.html) package is a powerful tool for creating thematic maps in R. It provides an intuitive and flexible way to visualize spatial data, similar to how `ggplot2` works for general data visualization.

# Libraries and dataset
***

First, let's install and load the [tmap](package/tmap.html) package. We'll use the built-in `World` dataset for our examples.

```{r}
# Install tmap if you haven't already
# install.packages("tmap")

# Load the library
library(tmap)

# Load the World dataset
data("World")
```

# Basic Concept: Layers
***

The main concept in [tmap](package/tmap.html) is building maps using layers. You start with `tm_shape()` to define the data, then add layers with various `tm_*()` functions.

Let's create a simple world map:

```{r}
tm_shape(World) +
  tm_polygons()
```

This creates a basic world map where each country is a polygon.

# Adding Data to the Map
***

We can add data to our map by specifying a variable in `tm_polygons()`:

```{r}
tm_shape(World) +
  tm_polygons("pop_est", palette = "Blues", title = "Population")
```

Here, we're coloring countries based on their population estimates.

# Customizing the Map
***

[tmap](package/tmap.html) offers many ways to customize your map. Let's add borders and adjust the color palette:

```{r}
tm_shape(World) +
  tm_polygons("pop_est", palette = "viridis", title = "Population") +
  tm_borders(col = "white", lwd = 0.5) +
  tm_layout(legend.outside = TRUE)
```

# Adding Multiple Layers
***

We can add multiple layers to our map. Let's add country names:

```{r}
tm_shape(World) +
  tm_polygons("pop_est", palette = "viridis", title = "Population") +
  tm_borders(col = "white", lwd = 0.5) +
  tm_text("name", size = 0.4) +
  tm_layout(legend.outside = TRUE)
```

# Interactive Maps
***

[tmap](package/tmap.html) can create both static and interactive maps. To switch to interactive mode:

```{r}
tmap_mode("view")

my_tmap <- tm_shape(World) +
  tm_polygons("pop_est", palette = "viridis", title = "Population")

tmap_save(my_tmap, filename = "HtmlWidget/416-introduction-to-tmap-interactive.html")
```

<center>
  <iframe src="HtmlWidget/416-introduction-to-tmap-interactive.html" height="600" width="600" style="border:none;"></iframe>
</center>

Remember to switch back to plot mode for static maps:

```{r}
tmap_mode("plot")
```

# Going further
***

You might be interested in
- [learning more about the tmap package](package/tmap.html)
- discover [more features of tmap](417-advanced-tmap-usage.html)

<!-- Close container -->
</div>

```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```