---
title: "Advanced Techniques with tmap"
descriptionMeta: "This post explores advanced techniques for creating thematic maps using the tmap package in R. It covers complex usages with clear code explanations and reproducible examples."
descriptionTop: "This post explores **advanced techniques** for creating [thematic maps](https://www.r-graph-gallery.com/choropleth-map.html) using the [tmap](package/tmap.html) package in R. It covers complex usages with clear code explanations and reproducible examples.<br><br>For an introduction to [tmap](package/tmap.html), check [this post](416-introduction-to-tmap.html)."
sectionText: "Map section"
sectionLink: "map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "417-advanced-tmap-techniques"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Libraries and dataset
***

We'll use the [tmap](package/tmap.html) package along with `sf` for some spatial data manipulation. We'll continue using the `World` dataset, but we'll also create some point data.

```{r}
library(tmap)
library(sf)

data("World")

# Create some point data
set.seed(123)
cities <- st_as_sf(data.frame(
  name = c("New York", "London", "Tokyo", "Sydney", "Rio de Janeiro"),
  lon = c(-74.006, -0.1276, 139.6503, 151.2093, -43.1729),
  lat = c(40.7128, 51.5074, 35.6762, -33.8688, -22.9068),
  pop = round(runif(5, 5e6, 20e6))
), coords = c("lon", "lat"), crs = 4326)
```

# Faceting with [tmap](package/tmap.html)
***

Faceting allows us to create small multiples based on a categorical variable.

This creates a separate **map for each continent**, allowing easy comparison of GDP per capita across different regions.

```{r}
tm_shape(World) +
  tm_polygons("gdp_cap_est", style = "quantile", palette = "viridis") +
  tm_facets(by = "continent") +
  tm_layout(frame = FALSE)
```


# Combining Different Geometries
***

We can combine different types of geometries in a single map.

This map shows **countries colored by income group** and **cities** represented by bubbles sized according to population.

```{r}
tm_shape(World) +
  tm_polygons("income_grp", palette = "YlOrRd") +
  tm_shape(cities) +
  tm_bubbles(size = "pop", col = "blue", alpha = 0.5, scale = 2) +
  tm_text("name", size = 0.7) +
  tm_layout(legend.outside = TRUE, frame = FALSE)
```


# Custom Projections
***

[tmap](package/tmap.html) allows us to easily change map projections.

The `projection` argument in `tm_shape()` allows us to **specify any projection** supported by the PROJ library. Let's create a map using the **Robinson** projection:

```{r}
tm_shape(World, projection = "+proj=robin") +
  tm_polygons("pop_est", style = "quantile", palette = "-viridis") +
  tm_layout(
    legend.outside = TRUE,
    title = "World Population (Robinson Projection)",
    frame = FALSE
  )
```


# Advanced Layout Customization
***

We can extensively customize the layout of our maps.

This example demonstrates how to customize the **title**, **legend**, and add elements like a **compass** and **scale bar**.

```{r}
tm_shape(World) +
  tm_polygons("gdp_cap_est", palette = "RdYlBu", style = "pretty", title = "GDP per Capita") +
  tm_layout(
    title = "Global GDP per Capita",
    title.position = c("center", "top"),
    title.size = 1.5,
    legend.title.size = 1,
    legend.text.size = 0.6,
    legend.position = c("right", "bottom"),
    frame = FALSE
  ) +
  tm_compass(position = c("left", "top")) +
  tm_scale_bar(position = c("left", "bottom"))
```



# Going further
***

You might be interested in

- this [introduction to tmap](416-introduction-to-tmap.html)
- [learning more about the tmap package](package/tmap.html)

<!-- Close container -->
</div>

```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```