---
title: "Correlation between discrete variable"
descriptionMeta: "Studying the relationship between 2 discrete variables is complicated since an usual scatterplot suffers overplotting. Here is a workaround using base R."
descriptionTop: "Studying the relationship between 2 discrete variables is complicated since an usual [scatterplot](scatterplot.html) suffers overplotting. Here is a workaround using base R."
sectionText: "Scatter section"
sectionLink: "scatterplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/scatter.html"
url: "5-correlation-of-discrete-variables"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Scatterplot with variable size
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

An usual [scatterplot](scatterplot.html) would suffer overplotting when used for discrete variables: dots would be drawn on top of each other, making the chart unreadable.

The workaround suggested here makes dot size proportional to the number of datapoints behind it. On top of that, the exact number can be represented in the bubble thanks to the `text()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
#Let's create 2 discrete variables 
a <- c(1,1,3,4,5,5,1,1,2,3,4,1,3,2,1,1,5,1,4,3,2,3,1,0,2)
b <- c(1,2,3,5,5,5,2,1,1,3,4,3,3,4,1,1,4,1,4,2,2,3,0,0,1)
 
#I count the occurence of each couple of values. Eg : number of time a=1 and b=1, number of time a=1 and b=2 etc...
AA <- xyTable(a,b)
 
#Now I can plot this ! I represent the dots as big as the couple occurs often
coeff_bigger <- 2
plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col=rgb(0,0,1,0.5) , xlab= "value of a" , ylab="value of b" , xlim=c(0,6) , ylim=c(0,6) )
text(AA$x , AA$y , AA$number )
 
#Note : It's easy to make a function that will compute this kind of plot automaticaly :
represent_discrete_variable <- function(var1, var2 , coeff_bigger){
  AA=xyTable(var1,var2)
  plot(AA$x , AA$y , cex=AA$number*coeff_bigger  , pch=16 , col="chocolate1" , xlab= "value of a" , ylab="value of b" )
  text (AA$x , AA$y , AA$number )
}
```
</div>
</div>



```{r thecode, eval=FALSE}
```




# Other workarounds
***
Other workarounds could be considered in this situation:

- [Heatmap](heatmap.html)
- [2d density chart](2d-density-chart.html)
- Jittering
- [Boxplot](boxplot.html)









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
