---
title: "Basic R: X axis labels on several lines"
descriptionMeta: "How to display the X axis on several lines in base R."
descriptionTop: "It is a common practice to display the X axis label on several lines. Here is an example applied to a boxplot."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "70-boxplot-with-categories-on-multiple-lines"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It can be handy to display X axis labels on several lines. For instance, to add the number of values present in each box of a [boxplot](boxplot.html).

How it works:  

- Change the names of your categories using the `names()` function. 
- Use `\n` to start new line
- Increase the distance between the labels and the X axis with the `mgp` argument of the `par()` function. It avoids overlap with the axis.

<u>Note:</u> `mgp` is a numeric vector of length 3, which sets the axis label locations relative to the edge of the inner plot window. Default value : c(3,1,0). First value : location the labels (xlab and ylab in plot). Second value : location of the tick-mark labels (what we want to lower). Third Value : position of the tick marks
      
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Create 2 vectors
a <- sample(2:24, 20 , replace=T)
b <- sample(4:14, 8 , replace=T)
 
# Make a list of these 2 vectors
C <- list(a,b)
 
# Change the names of the elements of the list :
names(C) <- c(paste("Category 1\n n=" , length(a) , sep=""), paste("Category 2\n n=" , length(b) , sep=""))
 
# Change the mgp argument: avoid text overlaps axis
par(mgp=c(3,2,0))
 
# Final Boxplot
boxplot(C , col="#69b3a2" , ylab="value" )
```
</div>
</div>



```{r, eval=FALSE}
# Create 2 vectors
a <- sample(2:24, 20 , replace=T)
b <- sample(4:14, 8 , replace=T)
 
# Make a list of these 2 vectors
C <- list(a,b)
 
# Change the names of the elements of the list :
names(C) <- c(paste("Category 1\n n=" , length(a) , sep=""), paste("Category 2\n n=" , length(b) , sep=""))
 
# Change the mgp argument: avoid text overlaps axis
par(mgp=c(3,2,0))
 
# Final Boxplot
boxplot(C , col="#69b3a2" , ylab="value" )
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
