---
title: "Base R margins: a cheatsheet."
descriptionMeta: "This post is dedicated to margin management in base R. It describes how to use the par function to call the mar and oma parameters: the 2 types of margin."
descriptionTop: "This post is dedicated to margin management in `base R`. It describes how to use the `par` function to call the `mar` and `oma` parameters: the 2 types of margin."
sectionText: "Line chart Section"
sectionLink: "line-plot.html"
DataToVizText: "About line chart"
DataToVizLink: "data-to-viz.com/graph/line.html"
url: "74-margin-and-oma-cheatsheet"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

There are 2 margin areas in `base R` plots: `margin` and `oma`.  You can control their size calling the `par()` function before your plot and giving the corresponding arguments:

- `mar()` for margin.
- `oma()` for outer margin area.

For both arguments, you must give four values giving the desired space in the bottom, left, top and right part of the chart respectively. For instance, `par(mar=c(4,0,0,0))` draws a margin of size 4 only on the bottom of the chart.

You can use `mai()` and `omi()` if you want to set the areas in inches and not in lines.

<u>Note</u>: this post is inspired from [here](http://research.stowers-institute.org/efg/R/Graphics/Basics/mar-oma/), have a look for more details.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Margins area
par(oma=c(3,3,3,3)) # all sides have 3 lines of space 
par(mar=c(5,4,4,2) + 0.1)

# Plot 
plot(0:10, 0:10, type="n", xlab="X", ylab="Y") # type="n" hides the points 

# Place text in the plot and color everything plot-related red 
text(5,5, "Plot", col="red", cex=2) 
box(col="red") 

# Place text in the margins and label the margins, all in forestgreen  
mtext("Margins", side=3, line=2, cex=2, col="forestgreen")  
mtext("par(mar=c(b,l,t,r))", side=3, line=1, cex=1, col="forestgreen")  
mtext("Line 0", side=3, line=0, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 1", side=3, line=1, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 2", side=3, line=2, adj=1.0, cex=1, col="forestgreen")  
mtext("Line 3", side=3, line=3, adj=1.0, cex=1, col="forestgreen")  
box("figure", col="forestgreen")  
  
# Label the outer margin area and color it blue  
# Note the 'outer=TRUE' command moves us from the figure margins to the outer margins.  
mtext("Outer Margin Area", side=1, line=1, cex=2, col="blue", outer=TRUE)  
mtext("par(oma=c(b,l,t,r))", side=1, line=2, cex=1, col="blue", outer=TRUE)  
mtext("Line 0", side=1, line=0, adj=0.0, cex=1, col="blue", outer=TRUE)  
mtext("Line 1", side=1, line=1, adj=0.0, cex=1, col="blue", outer=TRUE)  
mtext("Line 2", side=1, line=2, adj=0.0, cex=1, col="blue", outer=TRUE)  
box("outer", col="blue")  
```

</div>
</div>


```{r thecode, eval=FALSE}
```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
