---
title: "Split base R plot window with layout()"
descriptionMeta: "The layout() function of R allows to split the plot window in areas with custom sizes."
descriptionTop: "The `layout()` function of `R` allows to split the plot window in areas with custom sizes. Here are a few examples illustrating how to use it with reproducible code and explanation."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "75-split-screen-with-layout/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# 2 rows
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`Layout` divides the device up into as many rows and columns as there are in matrix `mat`.

Here I create the matrix with `matrix(c(1,2), ncol=1)` -> 1 column, 2 rows. This is what I get in the chart!

<u>Note:</u> this could be done using `par(mfrow=c(1,2))` as well. But this option does not allow the customization we'll see further in this post.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,2), ncol=1) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
```
</div>
</div>

```{r, eval=FALSE}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,2), ncol=1) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
```










# 2 columns
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here I create the matrix with `matrix(c(1,2), ncol=2)` -> 2 columns, 1 row. This is what I get in the chart! 

<u>Note:</u> if you swap to `c(2,1)`, second chart will be on top, first at the bottom

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,2), ncol=2) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
```
</div>
</div>

```{r, eval=FALSE}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,2), ncol=2) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
```








# Subdivide second row
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`matrix(c(1,1,2,3), nrow=2)` creates a matrix of 2 rows and 2 columns. First 2 panels will be for the first chart, the third for chart2 and the last for chart 3.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,1,2,3), nrow=2,  byrow=TRUE) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
boxplot(b , xlab="b" , col=rgb(0.4,0.2,0.3,0.5) , las=2)
```
</div>
</div>

```{r, eval=FALSE}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Create the layout
nf <- layout( matrix(c(1,1,2,3), nrow=2, byrow=TRUE) )

# Fill with plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
boxplot(b , xlab="b" , col=rgb(0.4,0.2,0.3,0.5) , las=2)
```













# Custom proportions
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

You can custom columns and row proportions with `widths` and `heights`.

Here, `widths=c(3,1)` means first column takes three quarters of the plot window width, second takes one quarter.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Set the layout
nf <- layout(
  matrix(c(1,1,2,3), ncol=2, byrow=TRUE), 
  widths=c(3,1), 
  heights=c(2,2)
)

#Add the plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
boxplot(b , xlab="b" , col=rgb(0.4,0.2,0.3,0.5) , las=2)
```
</div>
</div>

```{r, eval=FALSE}
# Dummy data
a <- seq(129,1)+4*runif(129,0.4)
b <- seq(1,129)^2+runif(129,0.98)
 
# Set the layout
nf <- layout(
  matrix(c(1,1,2,3), ncol=2, byrow=TRUE), 
  widths=c(3,1), 
  heights=c(2,2)
)

#Add the plots
hist(a , breaks=30 , border=F , col=rgb(0.1,0.8,0.3,0.5) , xlab="distribution of a" , main="")
boxplot(a , xlab="a" , col=rgb(0.8,0.8,0.3,0.5) , las=2)
boxplot(b , xlab="b" , col=rgb(0.4,0.2,0.3,0.5) , las=2)
```







<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
