---
title: "ggplot2 heatmap"
descriptionMeta: "This document provides several examples of heatmaps built with R and ggplot2. It describes the main customization you can apply, with explanation and reproducible code."
descriptionTop: "This document provides several examples of [heatmaps](heatmap.html) built with `R` and [ggplot2](ggplot2-package.html). It describes the main customization you can apply, with explanation and reproducible code.<br><u>Note:</u> The native [heatmap() function](215-the-heatmap-function.html) provides more options for data normalization and clustering. Consider it as a valuable option."
sectionText: "Heatmap section"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/heatmap.html"
url: "79-levelplot-with-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic heatmap with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the most basic heatmap you can build with `R` and `ggplot2`, using the `geom_tile()` function.  

Input data must be a long format where each row provides an observation. At least 3 variables are needed per observation:

- `x`: position on the X axis
- `y`: position on the Y axis
- `fill`: the numeric value that will be translated in a color


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# Heatmap 
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile()
```
</div>
</div>



```{r thecode, eval=FALSE}
```





# Control color palette
***

```{r thecode2, echo=FALSE, out.width = c("33%","33%","33%"), fig.height=7, fig.show='hold'}
# Library
library(ggplot2)
library(hrbrthemes)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# Give extreme colors:
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="blue") +
  theme_ipsum()

# Color Brewer palette
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_distiller(palette = "RdPu") +
  theme_ipsum()

# Color Brewer palette
library(viridis)
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE) +
  theme_ipsum()
```


Color palette can be changed like in any [ggplot2](ggplot2-package.html) chart. Above are 3 examples using different methods:

- `scale_fill_gradient()` to provide extreme colors of the palette
- `scale_fill_distiller)` to provide a [ColorBrewer](package/rColorBrewer.html) palette
- `scale_fill_viridis()` to use Viridis. Do not forget `discrete=FALSE` for a continuous variable.



```{r thecode2, eval=FALSE}
```








# From wide input format
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is a common issue to have a wide matrix as input, as for the `volcano` dataset. In this case, you need to tidy it with the `gather()` function of the `tidyr` package to visualize it with [ggplot](ggplot2-package.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)
library(tidyr)
library(tibble)
library(hrbrthemes)
library(dplyr)

# Volcano dataset
#volcano

# Heatmap 
volcano %>%
  
  # Data wrangling
  as_tibble() %>%
  rowid_to_column(var="X") %>%
  gather(key="Y", value="Z", -1) %>%
  
  # Change Y to numeric
  mutate(Y=as.numeric(gsub("V","",Y))) %>%

  # Viz
  ggplot(aes(X, Y, fill= Z)) + 
    geom_tile() +
    theme_ipsum() +
    theme(legend.position="none")
```
</div>
</div>



```{r thecode3, eval=FALSE}
```









# Turn it interactive with `plotly`
***

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

One of the nice feature of [ggplot2](ggplot2-package.html) is that charts can be turned interactive in seconds thanks to `plotly`. You just need to wrap your chart in an object and call it in the `ggplotly()` function. 

Often, it is a good practice to custom the text available in the tooltip.

<u>Note: </u>try to hover cells to see the tooltip, select an area to zoom in.
</div>


<div class = "col-md-7  col-sm-12">
<center><iframe src="HtmlWidget/ggplotlyHeatmap.html" height="700" width="700" style="border:none;"></iframe></center>
</div>
</div>


```{r, echo=TRUE, eval=FALSE}
# Library
library(ggplot2)
library(hrbrthemes)
library(plotly)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# new column: text for tooltip:
data <- data %>%
  mutate(text = paste0("x: ", x, "\n", "y: ", y, "\n", "Value: ",round(Z,2), "\n", "What else?"))

# classic ggplot, with text in aes
p <- ggplot(data, aes(X, Y, fill= Z, text=text)) + 
  geom_tile() +
  theme_ipsum()

ggplotly(p, tooltip="text")

# save the widget
# library(htmlwidgets)
# saveWidget(pp, file=paste0( getwd(), "/HtmlWidget/ggplotlyHeatmap.html"))

```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
