---
title: "Histogram with colored tail"
descriptionMeta: "This post explains how to build a histogram with colored tails in Basic R."
descriptionTop: "A [histogram](https://www.data-to-viz.com/graph/histogram.html) displays the distribution of a numeric variable. This posts explains how to color both tails of the distribution in Basic R, without any package. This can be useful to highlight a part of the distribution."
sectionText: "Histogram Section"
sectionLink: "histogram.html"
DataToVizText: "About histogram"
DataToVizLink: "data-to-viz.com/graph/histogram.html"
url: "82-boxplot-on-top-of-histogram.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">




<div class = "row">
  
<div class = "col-md-6  col-sm-12 align-self-center">

This example demonstrates how to color parts of the [histogram](histogram.html). First of all, the `hist` function must be called without plotting the result using the `plot=F` option. It allows to store the position of each bin in an object (`my_hist` here).  

Those bin borders are now available in the `$breaks` slot of the object, what allows to build a color vector using `ifelse` statements. Finally, this color vector can be used in a `plot` call.

</div>

  
<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Create data
my_variable <- rnorm(2000, 0 , 10)
 
# Calculate histogram, but do not draw it
my_hist <- hist(my_variable , breaks=40  , plot=F)
 
# Color vector
my_color <- ifelse(my_hist$breaks < -10, rgb(0.2,0.8,0.5,0.5) , ifelse (my_hist$breaks >=10, "purple", rgb(0.2,0.2,0.2,0.2) ))
 
# Final plot
plot(my_hist, col=my_color , border=F , main="" , xlab="value of the variable", xlim=c(-40,40) )
```
</div>
</div>


```{r eval=FALSE}
# Create data
my_variable=rnorm(2000, 0 , 10)
 
# Calculate histogram, but do not draw it
my_hist=hist(my_variable , breaks=40  , plot=F)
 
# Color vector
my_color= ifelse(my_hist$breaks < -10, rgb(0.2,0.8,0.5,0.5) , ifelse (my_hist$breaks >=10, "purple", rgb(0.2,0.2,0.2,0.2) ))
 
# Final plot
plot(my_hist, col=my_color , border=F , main="" , xlab="value of the variable", xlim=c(-40,40) )
```


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```







