---
title: "Tukey Test and boxplot in R"
descriptionMeta: "Tukey test compares all possible pairs of means for a set of categories. This post explains how to perform it in R and represent its result on a boxplot."
descriptionTop: "A Tukey test compares all possible pair of means for a set of categories. This post explains how to perform it in R and host to represent the result on a [boxplot](boxplot.html)."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "84-tukey-test"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# The Tukey test
***
<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">


[Tukey test](https://en.wikipedia.org/wiki/Tukey's_range_test) is a single-step multiple comparison procedure and statistical test. It is a post-hoc analysis, what means that it is used in conjunction with an ANOVA. 

It allows to find means of a factor that are significantly different from each other, comparing all possible pairs of means with a t-test like method. ([Read more](http://vassarstats.net/textbook/ch14pt2.html) for the exact procedure)

In R, the `multcompView` allows to run the Tukey test thanks to the `TukeyHSD()` function. It also offers a chart that shows the mean difference for each pair of group.
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# library
library(multcompView)
 
# Create data
set.seed(1)
treatment <- rep(c("A", "B", "C", "D", "E"), each=20) 
value=c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T), sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) , sample(10:20, 20 , replace=T) )
data=data.frame(treatment,value)
 
# What is the effect of the treatment on the value ?
model=lm( data$value ~ data$treatment )
ANOVA=aov(model)
 
# Tukey test to study each pair of treatment :
TUKEY <- TukeyHSD(x=ANOVA, 'data$treatment', conf.level=0.95)
 
# Tuckey test representation :
plot(TUKEY , las=1 , col="brown")
```
</div>
</div>



```{r, eval=FALSE}
# library
library(multcompView)
 
# Create data
set.seed(1)
treatment <- rep(c("A", "B", "C", "D", "E"), each=20) 
value=c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T), sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) , sample(10:20, 20 , replace=T) )
data=data.frame(treatment,value)
 
# What is the effect of the treatment on the value ?
model=lm( data$value ~ data$treatment )
ANOVA=aov(model)
 
# Tukey test to study each pair of treatment :
TUKEY <- TukeyHSD(x=ANOVA, 'data$treatment', conf.level=0.95)
 
# Tuckey test representation :
plot(TUKEY , las=1 , col="brown")
```







# Tukey test result on top of boxplot
***
<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The previous chart showed no significant difference between groups A and C, and between D and B. 

It is possible to represent this information in a [boxplot](boxplot.html). Group A and C are represented using a similar way: same color, and same 'b' letter on top. And so on for B-D and for E.


</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# I need to group the treatments that are not different each other together.
generate_label_df <- function(TUKEY, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- TUKEY[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }
 
# Apply the function on my dataset
LABELS <- generate_label_df(TUKEY , "data$treatment")
 
 
# A panel of colors to draw each group with the same color :
my_colors <- c( 
  rgb(143,199,74,maxColorValue = 255),
  rgb(242,104,34,maxColorValue = 255), 
  rgb(111,145,202,maxColorValue = 255)
  )
 
# Draw the basic boxplot
a <- boxplot(data$value ~ data$treatment , ylim=c(min(data$value) , 1.1*max(data$value)) , col=my_colors[as.numeric(LABELS[,1])] , ylab="value" , main="")
 
# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max( a$stats[nrow(a$stats),] )
 
#Add the labels
text( c(1:nlevels(data$treatment)) , a$stats[nrow(a$stats),]+over , LABELS[,1]  , col=my_colors[as.numeric(LABELS[,1])] )
```
</div>
</div>



```{r, eval=FALSE}
# I need to group the treatments that are not different each other together.
generate_label_df <- function(TUKEY, variable){
 
     # Extract labels and factor levels from Tukey post-hoc 
     Tukey.levels <- TUKEY[[variable]][,4]
     Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
     
     #I need to put the labels in the same order as in the boxplot :
     Tukey.labels$treatment=rownames(Tukey.labels)
     Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
     return(Tukey.labels)
     }
 
# Apply the function on my dataset
LABELS <- generate_label_df(TUKEY , "data$treatment")
 
 
# A panel of colors to draw each group with the same color :
my_colors <- c( 
  rgb(143,199,74,maxColorValue = 255),
  rgb(242,104,34,maxColorValue = 255), 
  rgb(111,145,202,maxColorValue = 255)
  )
 
# Draw the basic boxplot
a <- boxplot(data$value ~ data$treatment , ylim=c(min(data$value) , 1.1*max(data$value)) , col=my_colors[as.numeric(LABELS[,1])] , ylab="value" , main="")
 
# I want to write the letter over each box. Over is how high I want to write it.
over <- 0.1*max( a$stats[nrow(a$stats),] )
 
#Add the labels
text( c(1:nlevels(data$treatment)) , a$stats[nrow(a$stats),]+over , LABELS[,1]  , col=my_colors[as.numeric(LABELS[,1])] )
```


<br><br>
<u>Note</u>: Tukey test is also called:  Tukey's range test / Tukey method / Tukey's honest significance test / Tukey's HSD (honest significant difference) test / Tukey-Kramer method




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
