---
title: "Ordering boxplots in base R"
descriptionMeta: "This post is dedicated to boxplot ordering in base R. It describes 3 common use cases of reordering issue with code and explanation"
descriptionTop: "This post is dedicated to [boxplot](boxplot.html) ordering in base R. It describes 3 common use cases of reordering issue with code and explanation."
sectionText: "Boxplot Section"
sectionLink: "boxplot.html"
DataToVizText: "Boxplot pitfalls"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "9-ordered-boxplot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">



# Reordering category by median
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The most common need is to reorder categories by increasing median. It allows to quickly spot what group has the highest value and how categories are ranked.

It is doable using the `reorder()` function in combination with the `with()` function as suggested below:
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
# Create data : 7 varieties / 20 samples per variety / a numeric value for each sample
variety <- rep( c("soldur", "silur", "lloyd", "pescadou", "X4582", "Dudur", "Classic"), each=20)
note <- c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T),
        sample(1:7, 30 , replace=T), sample(3:10, 70 , replace=T) )
data <- data.frame(variety, note)
 
# Create a vector named "new_order" containing the desired order, and ask to boxplot using this vector !
new_order <- with(data, reorder(variety , note, median , na.rm=T))
 
# And draw the boxplot
boxplot(data$note ~ new_order , ylab="sickness" , col="#69b3a2", boxwex=0.4 , main="")
```
</div>
</div>



```{r, eval=FALSE}
# Create data : 7 varieties / 20 samples per variety / a numeric value for each sample
variety <- rep( c("soldur", "silur", "lloyd", "pescadou", "X4582", "Dudur", "Classic"), each=20)
note <- c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T),
        sample(1:7, 30 , replace=T), sample(3:10, 70 , replace=T) )
data <- data.frame(variety, note)
 
# Create a vector named "new_order" containing the desired order
new_order <- with(data, reorder(variety , note, median , na.rm=T))
 
# Draw the boxplot using this new order
boxplot(data$note ~ new_order , ylab="sickness" , col="#69b3a2", boxwex=0.4 , main="")
```



# Give a specific order
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

[Boxplot](boxplot.html) categories are provided in a column of the input data frame. This column needs to be a `factor`, and has several `levels`. Categories are displayed on the chart following the order of this factor, often in alphabetical order.

Sometimes, we need to show groups in a specific order (A,D,C,B here). This can be done by reordering the levels, using the `factor()` function.

</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
#Creating data 
names <- c(rep("A", 20) , rep("B", 20) , rep("C", 20), rep("D", 20))
value <- c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T), 
       sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) )
data <- data.frame(names,value)
 
# Classic boxplot (A-B-C-D order)
# boxplot(data$value ~ data$names)
 
# I reorder the groups order : I change the order of the factor data$names
data$names <- factor(data$names , levels=c("A", "D", "C", "B"))
 
#The plot is now ordered !
boxplot(data$value ~ data$names , col=rgb(0.3,0.5,0.4,0.6) , ylab="value" , 
    xlab="names in desired order")
```
</div>
</div>



```{r, eval=FALSE}
#Creating data 
names <- c(rep("A", 20) , rep("B", 20) , rep("C", 20), rep("D", 20))
value <- c( sample(2:5, 20 , replace=T) , sample(6:10, 20 , replace=T), 
       sample(1:7, 20 , replace=T), sample(3:10, 20 , replace=T) )
data <- data.frame(names,value)
 
# Classic boxplot (A-B-C-D order)
# boxplot(data$value ~ data$names)
 
# I reorder the groups order : I change the order of the factor data$names
data$names <- factor(data$names , levels=c("A", "D", "C", "B"))
 
#The plot is now ordered !
boxplot(data$value ~ data$names , col=rgb(0.3,0.5,0.4,0.6) , ylab="value" , 
    xlab="names in desired order")
```










# Grouped and ordered boxplot {#grouped}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

In a grouped boxplot, categories are organized in groups and subgroups. For instance, let's take several varieties (group) that are grown in high or low temperature (subgroup).

Here both subgroups are represented one beside each other, and groups are ranked by increasing median:
</div>


<div class = "col-md-6  col-sm-12">

```{r, echo=FALSE, out.width = "100%", fig.height=7}
variety <- rep( c("soldur", "silur", "lloyd", "pescadou", "X4582", "Dudur", "Classic"), each=40)
treatment <- rep(c(rep("high" , 20) , rep("low" , 20)) , 7)
note <- c( rep(c(sample(0:4, 20 , replace=T) , sample(1:6, 20 , replace=T)),2), 
          rep(c(sample(5:7, 20 , replace=T), sample(5:9, 20 , replace=T)),2), 
          c(sample(0:4, 20 , replace=T) , sample(2:5, 20 , replace=T), 
          rep(c(sample(6:8, 20 , replace=T) , sample(7:10, 20 , replace=T)),2) ))
data=data.frame(variety, treatment ,  note)
 
# Reorder varieties (group) (mixing low and high treatments for the calculations)
new_order <- with(data, reorder(variety , note, mean , na.rm=T))
 
# Then I make the boxplot, asking to use the 2 factors : variety (in the good order) AND treatment :
par(mar=c(3,4,3,1))
myplot <- boxplot(note ~ treatment*new_order , data=data  , 
        boxwex=0.4 , ylab="sickness",
        main="sickness of several wheat lines" , 
        col=c("slateblue1" , "tomato") ,  
        xaxt="n")
 
# To add the label of x axis
my_names <- sapply(strsplit(myplot$names , '\\.') , function(x) x[[2]] )
my_names <- my_names[seq(1 , length(my_names) , 2)]
axis(1, 
     at = seq(1.5 , 14 , 2), 
     labels = my_names , 
     tick=FALSE , cex=0.3)

# Add the grey vertical lines
for(i in seq(0.5 , 20 , 2)){ 
  abline(v=i,lty=1, col="grey")
  }
 
# Add a legend
legend("bottomright", legend = c("High treatment", "Low treatment"), 
       col=c("slateblue1" , "tomato"),
       pch = 15, bty = "n", pt.cex = 3, cex = 1.2,  horiz = F, inset = c(0.1, 0.1))
```
</div>
</div>



```{r, eval=FALSE}
# Create dummy data
variety <- rep( c("soldur", "silur", "lloyd", "pescadou", "X4582", "Dudur", "Classic"), each=40)
treatment <- rep(c(rep("high" , 20) , rep("low" , 20)) , 7)
note <- c( rep(c(sample(0:4, 20 , replace=T) , sample(1:6, 20 , replace=T)),2), 
          rep(c(sample(5:7, 20 , replace=T), sample(5:9, 20 , replace=T)),2), 
          c(sample(0:4, 20 , replace=T) , sample(2:5, 20 , replace=T), 
          rep(c(sample(6:8, 20 , replace=T) , sample(7:10, 20 , replace=T)),2) ))
data=data.frame(variety, treatment ,  note)
 
# Reorder varieties (group) (mixing low and high treatments for the calculations)
new_order <- with(data, reorder(variety , note, mean , na.rm=T))
 
# Then I make the boxplot, asking to use the 2 factors : variety (in the good order) AND treatment :
par(mar=c(3,4,3,1))
myplot <- boxplot(note ~ treatment*new_order , data=data  , 
        boxwex=0.4 , ylab="sickness",
        main="sickness of several wheat lines" , 
        col=c("slateblue1" , "tomato") ,  
        xaxt="n")
 
# To add the label of x axis
my_names <- sapply(strsplit(myplot$names , '\\.') , function(x) x[[2]] )
my_names <- my_names[seq(1 , length(my_names) , 2)]
axis(1, 
     at = seq(1.5 , 14 , 2), 
     labels = my_names , 
     tick=FALSE , cex=0.3)

# Add the grey vertical lines
for(i in seq(0.5 , 20 , 2)){ 
  abline(v=i,lty=1, col="grey")
  }
 
# Add a legend
legend("bottomright", legend = c("High treatment", "Low treatment"), 
       col=c("slateblue1" , "tomato"),
       pch = 15, bty = "n", pt.cex = 3, cex = 1.2,  horiz = F, inset = c(0.1, 0.1))
```



<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
