---
title: "Parallel chart with the MASS library"
descriptionMeta: "This post explains how to build a parallel coordinate chart with R and the MASS library. Examples with reproducible code are provided."
descriptionTop: "This post explains how to build a [parallel coordinate chart](parallel-plot.html) with `R` and the `MASS` library. Note that using ggplot2 is probably a better option."
sectionText: "Parallel coord section"
sectionLink: "parallel-plot.html"
DataToVizText: "ggplot2 option"
DataToVizLink: "data-to-viz.com/graph/parallel.html"
url: "93-parrallel-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# The `parcoord()` function of the `MASS` library.
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `MASS` library provides the `parcoord()` function that automatically builds parallel coordinates chart.

The input dataset must be a data frame composed by numeric variables only. Each variable will be used to build one vertical axis of the chart.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# You need the MASS library
library(MASS)
 
# Vector color
my_colors <- colors()[as.numeric(iris$Species)*11]
 
# Make the graph !
parcoord(iris[,c(1:4)] , col= my_colors  )
```
</div>
</div>



```{r thecode, eval=FALSE}
```










# Reorder variables {#reorder}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is important to find the best variable order in your parallel coordinates chart. To change it, just change the order in the input dataset.

<u>Note</u>: the `RColorBrewer` package is used to generate a nice and reliable color palette.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# You need the MASS library
library(MASS)
 
# Vector color
library(RColorBrewer)
palette <- brewer.pal(3, "Set1") 
 my_colors <- palette[as.numeric(iris$Species)]

# Make the graph !
parcoord(iris[,c(1,3,4,2)] , col= my_colors  )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```







# Highlight a group {#highlight}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Data visualization aims to highlight a story in the data. If you are interested in a specific group, you can highlight it as follow:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# You need the MASS library
library(MASS)
 
# Let's use the Iris dataset as an example
data(iris)
 
# Vector color: red if Setosa, grey otherwise.
isSetosa <- ifelse(iris$Species=="setosa","red","grey")

# Make the graph !
parcoord(iris[,c(1,3,4,2)] , col=isSetosa  )
```
</div>
</div>



```{r thecode3, eval=FALSE}
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
