---
title: "Most basic violin plot with ggplot2"
descriptionMeta: "Learn how to build a basic violin plot with R and ggplot2. Reproducible R code is provided, different input formats are considered."
descriptionTop: "A [violin plot](violin.html) allows to compare the distribution of several groups by displaying their densities. See how to build it with `R` and `ggplot2` below. Learn more about violin chart theory in [data-to-viz](https://www.data-to-viz.com/graph/violin.html)."
sectionText: "Violin Section"
sectionLink: "violin.html"
DataToVizText: "Violin theory"
DataToVizLink: "data-to-viz.com/graph/violin.html"
url: "area_basic"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```


<div class="container">




# Basic violin plot
***
  
<div class = "row">
  
<div class = "col-md-6  col-sm-12 align-self-center">
Building a [violin plot](violin.html) with `ggplot2` is pretty straightforward thanks to the dedicated `geom_violin()` function.

```{r}
# Library
library(ggplot2)

# create a dataset
data <- data.frame(
  name=c( rep("A",500), rep("B",500), rep("B",500), rep("C",20), rep('D', 100)  ),
  value=c( rnorm(500, 10, 5), rnorm(500, 13, 1), rnorm(500, 18, 1), rnorm(20, 25, 4), rnorm(100, 12, 1) )
)

# Most basic violin chart
p <- ggplot(data, aes(x=name, y=value, fill=name)) + # fill=name allow to automatically dedicate a color for each group
  geom_violin()

#p
```

</div>
  
<div class = "col-md-6  col-sm-12">
```{r, warning=FALSE, echo=FALSE, out.width = "100%", fig.width=5, fig.height=5}
p
```
</div>
</div>



# Note on input format
***
Ggplot2 expects input data to be in a *long* format: each row is dedicated to *one* observation. Your input needs 2 column:    
- a *categorical* variable for the X axis: it needs to be have the class `factor`  
- a *numeric* variable for the Y axis: it needs to have the class `numeric`
<br>


<div class = "row">
  
<div class = "col-md-6  col-sm-12">
### &rarr; From long format
You already have the good format. It's going to be a breeze to plot it with `geom_violin()` as follow:
```{r, out.width = "100%"}
# Library
library(ggplot2)
library(dplyr)

# Create data
data <- data.frame(
  name=c( rep("A",500), rep("B",500), rep("B",500), rep("C",20), rep('D', 100)  ),
  value=c( rnorm(500, 10, 5), rnorm(500, 13, 1), rnorm(500, 18, 1), rnorm(20, 25, 4), rnorm(100, 12, 1) ) %>% round(2)
)
```


```{r echo=FALSE}
library(knitr)
library(kableExtra)
kable(head(data,4), align="c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```



```{r, out.width = "100%"}
# Basic violin
ggplot(data, aes(x=name, y=value, fill=name)) + 
  geom_violin()
```
</div>


<div class = "col-md-6  col-sm-12">
### &rarr; From wide format
In this case we need to reformat the input. This is possible thanks to the `gather()` function of the `tidyr` library that is part of the `tidyverse`.
```{r}
# Let's use the iris dataset as an example:
data_wide <- iris[ , 1:4]
```

```{r echo=FALSE}
kable(head(data_wide,4), align="c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "center")
```

```{r, out.width = "100%"}
library(tidyr)
data_wide %>% 
  gather(key="MesureType", value="Val") %>%
  ggplot( aes(x=MesureType, y=Val, fill=MesureType)) +
    geom_violin()
```

</div>

</div>












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```







