---
title: "Basic scatterplot matrix"
descriptionMeta: "This post explains how to build a scatterplot matrix with basic R, without any packages. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build a [scatterplot matrix](correlogram.html) with basic R, without any packages. It provides several reproducible examples with explanation and R code."
sectionText: "Correlogram section"
sectionLink: "correlogram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/correlogram.html"
url: "98-basic-scatterplot-matrix"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Scatterplot matrix with the native `plot()` function
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

For a set of data variables (dimensions) X1, X2, ??? , Xk, the scatter plot matrix shows all the pairwise [scatterplots](scatterplot.html) of the variables on a single view with multiple scatterplots in a matrix format.

The native `plot()` function does the job pretty well as long as you just need to display scatterplots. For more option, check the [correlogram section](correlogram.html)

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Data: numeric variables of the native mtcars dataset
data <- mtcars[ , c(1,3:6)]
 
# Plot
plot(data , pch=20 , cex=1.5 , col="#69b3a2")
```
</div>
</div>



```{r thecode, eval=FALSE}
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
