---
title: "The title that will appear in big at the top of the page"
descriptionMeta: "This post describes how to build a hexbin map, with input data available at geojson format. It provides a ste-by-step explanation and reproducible code."
descriptionTop: "The description that appears at the top of the page. You can include some `markdown` here!"
sectionText: "Hexbin map section"
sectionLink: "hexbin-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/hexbinmap.html"
url: "blog-post-template"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# About
***
The gallery always give credit to creators!! Please put a link to your homepage, to the original blogpost, to the package you've created or anything else you want people to be aware of about this project!


# Libraries
***

Please add a section with all the libraries that need to be loaded. And how to install them if it is tricky

```{r}
library(tidyverse)
```


# Dataset
***
Tell us a bit about what data we're gonna study

# Let's blog!
***
- Please keep the syntax above for titles!
- Try to start simple, and iterate on the examples until you get a nice chart!


# Layout: code and explanation side by side
***


<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Some explanation appearing on the left

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# The plot on the right. Code will appear below
plot(1,1)
```
</div>
</div>


```{r thecode, eval=FALSE}
```














# Layout: normal
***
You can also just write text followed by code followed by img

```{r}
plot(1,1)
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
# This add some images at the end of the blogpost with related chart types
# You can choose: Map, Correlation, Flow, Other, Ranking, Evolution, PartOfWhole, Distribution
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
