---
title: "Interactive bubble chart"
descriptionMeta: "This post explains how to build an interactive bubble chart with R, using ggplot2 and the ggplotly() function of the plotly packages."
descriptionTop: "This post explains how to build an interactive [bubble chart](bubble-chart.html) with R, using `ggplot2` and the `ggplotly()` function of the plotly package."
sectionText: "Bubble section"
sectionLink: "bubble-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/bubble.html"
url: "bubble_chart_interactive_ggplotly"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic bubble chart with `geom_point()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This post follows the previous [step by step description](320-the-basis-of-bubble-plot.html) of building [bubble chart](bubble-chart.html) with [ggplot2](ggplot2.html).

The idea is to turn the chart interactive:

- You can zoom by selecting an area of interest
- Hover a circle to get information about it
- Export to png
- Slide axis
- Double click to re-initialize.

This is done thanks to the `ggplotly()` function of the `plotly` package that turn any ggplot2 chart object interactive. Note the little trick to custom the tooltip content.

</div>


<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/ggplotlyBubblechart.html" height="700" width="700" style="border:none;"></iframe></center>

</div>
</div>




```{r, eval=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(plotly)
library(viridis)
library(hrbrthemes)

# The dataset is provided in the gapminder library
library(gapminder)
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)

# Interactive version
p <- data %>%
  mutate(gdpPercap=round(gdpPercap,0)) %>%
  mutate(pop=round(pop/1000000,2)) %>%
  mutate(lifeExp=round(lifeExp,1)) %>%
  
  # Reorder countries to having big bubbles on top
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country)) %>%
  
  # prepare text for tooltip
  mutate(text = paste("Country: ", country, "\nPopulation (M): ", pop, "\nLife Expectancy: ", lifeExp, "\nGdp per capita: ", gdpPercap, sep="")) %>%
  
  # Classic ggplot
  ggplot( aes(x=gdpPercap, y=lifeExp, size = pop, color = continent, text=text)) +
    geom_point(alpha=0.7) +
    scale_size(range = c(1.4, 19), name="Population (M)") +
    scale_color_viridis(discrete=TRUE, guide=FALSE) +
    theme_ipsum() +
    theme(legend.position="none")

# turn ggplot interactive with plotly
pp <- ggplotly(p, tooltip="text")
pp

# save the widget
# library(htmlwidgets)
# saveWidget(pp, file=paste0( getwd(), "/HtmlWidget/ggplotlyBubblechart.html"))
```






```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
