---
title: "Choropleth map with R"
descriptionMeta: "This post explains how to build a choropleth map with R. It shows how to load geospatial data, bind it to a numeric value and draw the map. Reproducible code is provided."
descriptionTop: "This post explains how to build a [choropleth map](choropleth-map.html) with R. It shows how to load geospatial data, bind it to a numeric value and draw the map."
sectionText: "Choropleth section"
sectionLink: "choropleth-map.html"
DataToVizText: "About Maps"
DataToVizLink: "data-to-viz.com/graph/choropleth.html"
url: "choropleth-map-in-r"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">


Two inputs are needed to build a choropleth map:

- A geospatial object providing region boundaries
- A numeric variable that will be used to cotrol region color









# Find and download a shape file
***
<u>Note</u>: this step is described in detail [here](168-load-a-shape-file-into-r.html). Read it if you are not familiar with geospatial data management in R.

The region boundaries required to make maps are usually stored in geospatial objects. Those objects can come from [shapefiles](168-load-a-shape-file-into-r.html), [geojson files](325-background-map-from-geojson-format-in-r.html) or provided in a [R package](278-the-maps-library.html). See the [map](maps.html) section for possibilities.

Let's get a geospatial object from a shape file available [here](http://thematicmapping.org/downloads/world_borders.php).

```{r, eval=FALSE}
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="DATA/world_shape_file.zip")
# You now have it in your current working directory, have a look!

# Unzip this file. You can do it with R (as below), or clicking on the object you downloaded.
system("unzip DATA/world_shape_file.zip")
#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
```

<br>

And let's load it in R

```{r}
# Read this shape file with the rgdal library. 
library(rgdal)
my_spdf <- readOGR( 
  dsn= paste0(getwd(),"/DATA/world_shape_file/") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# Select Africa only
africa <- my_spdf[my_spdf@data$REGION==2 , ]

# -- > Now you have a Spdf object (spatial polygon data frame). You can start doing maps!
```







# Plot the map
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The hardest part is to build the color palette that will be used on the map. The `RColorBrewer` package makes it a breeze to get a nice set of colors thanks to its `brewer.pal()` function.

Once the palette is created, `POP2005` is cut in several bins using the `cut()` function, and the approriate color is attributed to each bin.

Finally, build the map with `plot()` and give the color vector to the `col` argument to get the choropleth.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode9, echo=FALSE, out.width = "100%", fig.height=7}
# Plot
plot(africa , xlim=c(-20,60) , ylim=c(-40,40))
```

</div>
</div>

```{r thecode9, eval=FALSE}
```













# Visualize the numeric variable
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

In a choropleth map, each region has a color that represents the value of a numeric variable.

In this example, this variable is already available in the `data` slot of the geospatial object. You can have a look to it typing `africa@data` in the terminal.

This object is a data frame. It is thus possible to visualize the distribution of `POP2005` with a classic histogram made in [ggplot2](ggplot2-package.html)

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=4}
# library
library(dplyr)
library(ggplot2)

# Make sure the variable you are studying is numeric
africa@data$POP2005 <- as.numeric( africa@data$POP2005 )

# Distribution of the population per country?
africa@data %>% 
  ggplot( aes(x=as.numeric(POP2005))) + 
    geom_histogram(bins=20, fill='#69b3a2', color='white')
```

</div>
</div>

```{r thecode, eval=FALSE}
```










# Build the choropleth
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The hardest part is to build the color palette that will be used on the map. The `RColorBrewer` package makes it a breeze to get a nice set of colors thanks to its `brewer.pal()` function.

Once the palette is created, `POP2005` is cut in several bins using the `cut()` function, and the approriate color is attributed to each bin.

Finally, build the map with `plot()` and give the color vector to the `col` argument to get the choropleth.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Palette of 30 colors
library(RColorBrewer)
my_colors <- brewer.pal(9, "Reds") 
my_colors <- colorRampPalette(my_colors)(30)
 
# Attribute the appropriate color to each country
class_of_country <- cut(africa@data$POP2005, 30)
my_colors <- my_colors[as.numeric(class_of_country)]
 
# Make the plot
plot(africa , xlim=c(-20,60) , ylim=c(-40,40), col=my_colors ,  bg = "#A6CAE0")
```

</div>
</div>

```{r thecode2, eval=FALSE}
```












<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
