---
title: "Connected scatterplot with R and ggplot2"
descriptionMeta: "This post explains how to build a basic connected scatterplot with R and ggplot2. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to build a basic [connected scatterplot](connected-scatterplot.html) with R and [ggplot2](ggplot2-package.html). It provides several reproducible examples with explanation and R code."
sectionText: "Connected scatter section"
sectionLink: "connected-scatterplot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/connectedscatter.html"
url: "connected_scatterplot_ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic connected scatterplot: `geom_point()` and `geom_line()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A [connected scatterplot](connected-scatterplot.html) is basically a hybrid between a [scatterplot](scatterplot.html) and a [line plot](line-plot.html). Thus, you just have to add a `geom_point()` on top of the `geom_line()` to build it.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# Plot
data %>%
  tail(10) %>%
  ggplot( aes(x=date, y=value)) +
    geom_line() +
    geom_point()
```


</div>
</div>



```{r thecode, eval=FALSE}
```










# Customize the connected scatterplot
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Custom the general theme with the `theme_ipsum()` function of the `hrbrthemes` package. Add a title with `ggtitle()`. Custom circle and line with arguments like `shape`, `size`, `color` and more.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# Plot
data %>%
  tail(10) %>%
  ggplot( aes(x=date, y=value)) +
    geom_line( color="grey") +
    geom_point(shape=21, color="black", fill="#69b3a2", size=6) +
    theme_ipsum() +
    ggtitle("Evolution of bitcoin price")
```


</div>
</div>



```{r thecode2, eval=FALSE}
```









# Connected scatterplot to show an evolution
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The [connected scatterplot](connected-scatter.html) can also be a powerfull technique to tell a story about the evolution of 2 variables. Let???s consider a dataset composed of 3 columns:

- Year
- Number of baby born called Amanda this year
- Number of baby born called Ashley

The scatterplot beside allows to understand the evolution of these 2 names. Note that the code is pretty different in this case. `geom_segment()` is used of `geom_line()`. This is because `geom_line()` automatically sort data points depending on their X position to link them.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(babynames)
library(ggrepel)
library(tidyr)

# data
data <- babynames %>% 
  filter(name %in% c("Ashley", "Amanda")) %>%
  filter(sex=="F") %>%
  filter(year>1970) %>%
  select(year, name, n) %>%
  spread(key = name, value=n, -1)

# plot
data %>% 
  ggplot(aes(x=Amanda, y=Ashley, label=year)) +
     geom_point() +
     geom_segment(aes(
                    xend=c(tail(Amanda, n=-1), NA), 
                    yend=c(tail(Ashley, n=-1), NA)
                  )
      ) 
```


</div>
</div>



```{r thecode3, eval=FALSE}
```









<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It makes sense to add arrows and labels to guide the reader in the chart:
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(babynames)
library(ggrepel)
library(tidyr)

# data
data <- babynames %>% 
  filter(name %in% c("Ashley", "Amanda")) %>%
  filter(sex=="F") %>%
  filter(year>1970) %>%
  select(year, name, n) %>%
  spread(key = name, value=n, -1)

# Select a few date to label the chart
tmp_date <- data %>% sample_frac(0.3)

# plot 
data %>% 
  ggplot(aes(x=Amanda, y=Ashley, label=year)) +
     geom_point(color="#69b3a2") +
     geom_text_repel(data=tmp_date) +
     geom_segment(color="#69b3a2", 
                  aes(
                    xend=c(tail(Amanda, n=-1), NA), 
                    yend=c(tail(Ashley, n=-1), NA)
                  ),
                  arrow=arrow(length=unit(0.3,"cm"))
      ) +
      theme_ipsum()
```


</div>
</div>



```{r thecode4, eval=FALSE}
```











<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
