---
title: "Customize plotly tooltip"
descriptionMeta: "This post explains how to customize the tooltip in a plotly chart in R. It provides reproducible code and explanation how to improve the default tooltip."
descriptionTop: "This post explains how to **customize the tooltip** in a [plotly](https://r-graph-gallery.com/plotly.html) chart in R. It provides reproducible code and explanation how to **improve the default tooltip**."
sectionText: "Plotly section"
sectionLink: "plotly.html"
DataToVizText: "Interactive charts"
DataToVizLink: "r-graph-gallery.com/interactive-charts.html"
url: "customize-plotly-tooltip"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px;">




# Packages

First, we need to load a few packages:

- [plotly](https://r-graph-gallery.com/plotly.html) for the interactive chart
- [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) for creating the initial plot
- [hrbrthemes](https://r-graph-gallery.com/hrbrthemes.html) for the theme
- `viridis` for the color palette
- `gapminder` for the dataset
- `dplyr` for data manipulation

```{r}
# Visualisation packages
library(ggplot2)
library(plotly)
library(viridis)
library(hrbrthemes)

# Data packages
library(gapminder)
library(dplyr) 
```


# Dataset

The `gapminder` dataset is used in this example. It contains data about life expectancy, population and GDP per capita for many countries over time.

Here, we:

- filter the data to keep only the year 2007
- remove the year column
- reorder the countries to have big bubbles on top

```{r}
data <- gapminder %>% filter(year=="2007") %>% dplyr::select(-year)
data = data %>%  
  # Reorder countries to having big bubbles on top
  arrange(desc(pop)) %>%
  mutate(country = factor(country, country))
```




# Default plot
***

```{r}
p = data %>%
  ggplot( aes(x=gdpPercap, y=lifeExp, size = pop, color = continent)) +
    geom_point(alpha=0.7) +
    scale_size(range = c(1.4, 19), name="Population (M)") +
    scale_color_viridis(discrete=TRUE, guide=FALSE) +
    theme_ipsum() +
    theme(legend.position="none")

# turn ggplot interactive with plotly
pp = ggplotly(p)

# save the widget
#library(htmlwidgets)
#saveWidget(pp, file="HtmlWidget/plotlyDefaultTooltip.html")
```

<center><iframe src="HtmlWidget/plotlyDefaultTooltip.html" height="700" width="700" style="border:none;"></iframe></center>


# Customize tooltip
***

As you can see, it's a bit **unfortunate** that the tooltip does not tell us **what is the country**. Let's customize it to add this information.

For this, we:

- create a **new column** `text` in the dataset that contains the **information we want to display** in the tooltip using the `mutate()` function
- add `text` as an aesthetic in the `ggplot()` function

And that's it!

```{r}
data = data %>% 
  mutate(text = paste(
    "Country: ", country,
    "\nPopulation (M): ", pop,
    "\nLife Expectancy: ", lifeExp,
    "\nGdp per capita: ", gdpPercap, sep="")
  )

p = data %>%
  ggplot( aes(x=gdpPercap, y=lifeExp, size=pop, color=continent, text=text)) +
    geom_point(alpha=0.7) +
    scale_size(range = c(1.4, 19), name="Population (M)") +
    scale_color_viridis(discrete=TRUE, guide=FALSE) +
    theme_ipsum() +
    theme(legend.position="none")

# turn ggplot interactive with plotly
pp = ggplotly(p, tooltip="text")

# save the widget
#library(htmlwidgets)
#saveWidget(pp, file="HtmlWidget/plotlyDefaultTooltip2.html")
```

<center><iframe src="HtmlWidget/plotlyDefaultTooltip2.html" height="700" width="700" style="border:none;"></iframe></center>



# More advanced customization
***

The previous example is a good start, but we can go further. Here, we:

- use **bold** text to make the tooltip more readable
- add a **dollar sign** before the GDP per capita
- use the `round()` function to make the numbers more readable
- add a **title** to the tooltip
- change the **background color** of the tooltip to white
- change the **font color** of the tooltip to black

```{r}
data <- data %>%
  mutate(text = paste(
    "<span style='font-size:16px;'><b>", country, "</b></span><br>",
    "<br><b>Population (M):</b> ", pop,
    "<br><b>Life Expectancy (years):</b> ", round(lifeExp, 2),
    "<br><b>Gdp per capita:</b> $", round(gdpPercap, 2),
    sep=""))

p <- data %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, size=pop, color=continent, text=text)) +
    geom_point(alpha=0.7) +
    scale_size(range = c(1.4, 19), name="Population (M)") +
    scale_color_viridis(discrete=TRUE, guide=FALSE) +
    theme_ipsum() +
    theme(legend.position="none")

pp <- ggplotly(p, tooltip="text")
pp <- pp %>%
  layout(
    hoverlabel = list(bgcolor = "white",
                      font = list(size = 12, color = "black"))
  )

# save the widget
#library(htmlwidgets)
#saveWidget(pp, file="HtmlWidget/plotlyDefaultTooltip3.html")
```

<center><iframe src="HtmlWidget/plotlyDefaultTooltip3.html" height="700" width="700" style="border:none;"></iframe></center>


# Conclusion
***

This post explains how to customize the tooltip in a [plotly](https://r-graph-gallery.com/plotly.html) chart in R.

You can learn more about [interactive charts](r-graph-gallery.com/interactive-charts.html) on the R graph gallery.


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
