---
title: "Histogram with several groups - ggplot2"
descriptionMeta: "How to build histograms showing the distribution of several groups with R and ggplot2."
descriptionTop: "A [histogram](https://www.data-to-viz.com/graph/histogram.html) displays the distribution of a numeric variable. A common task is to compare this distribution through several groups. This document explains how to do so using R and [ggplot2](ggplot2.html)."
sectionText: "Histogram Section"
sectionLink: "histogram.html"
DataToVizText: "About histogram"
DataToVizLink: "data-to-viz.com/graph/histogram.html"
url: "histogram_several_group.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```


<div class="container">


# Several histograms on the same axis
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If the number of group or variable you have is relatively low, you can display all of them on the same axis, using a bit of transparency to make sure you do not hide any data.

<u>Note</u>: with 2 groups, you can also build a [mirror histogram](density_mirror_ggplot2.html)
```{r}
# library
library(ggplot2)
library(dplyr)
library(hrbrthemes)

# Build dataset with different distributions
data <- data.frame(
  type = c( rep("variable 1", 1000), rep("variable 2", 1000) ),
  value = c( rnorm(1000), rnorm(1000, mean=4) )
)

# Represent it
p <- data %>%
  ggplot( aes(x=value, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_ipsum() +
    labs(fill="")
```

</div>

<div class = "col-md-6  col-sm-12 align-self-center">
```{r, out.width=c('100%'), echo=FALSE}
p
```

</div>
</div>










# Using small multiple
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If the number of group you need to represent is high, drawing them on the same axis often results in a <u>cluttered</u> and unreadable figure.  

A good workaroung is to use <u>small multiple</u> where each group is represented in a fraction of the plot window, making the figure easy to read. This is pretty easy to build thanks to the `facet_wrap()` function of ggplot2.

<u>Note:</u> read more about the dataset used in this example [here](https://www.data-to-viz.com/story/OneNumOneCatSeveralObs.html).
</div>

<div class = "col-md-6  col-sm-12 align-self-center">
```{r}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(forcats)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>%
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0))

# plot
p <- data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_histogram(alpha=0.6, binwidth = 5) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)") +
    facet_wrap(~text)
```
</div>
</div>


```{r out.width='80%', echo=FALSE, fig.align='center', fig.height=7}
p
```








<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
