---
title: "Multi groups line chart with ggplot2"
descriptionMeta: "This post explains how to build a line chart that represents several groups with ggplot2. It provides several examples with explanation and reproducible code."
descriptionTop: "This post explains how to build a [line chart](line-plot.html) that represents several groups with [ggplot2](ggplot2.html). It provides several examples with explanation and reproducible code."
sectionText: "Line chart Section"
sectionLink: "line-plot.html"
DataToVizText: "About line chart"
DataToVizLink: "data-to-viz.com/graph/line.html"
url: "line-chart-several-groups-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Basic version
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

If you're not familiar with the `geom_line()` function, you should probably have a look to the [most basic line chart](line-chart-ggplot2.html) first.

Here, the input data frame is composed by 3 columns:

- An *ordered* numeric variable for the X axis
- Another numeric variable for the Y axis
- A categorical variable that specify the group of the observation

The idea is to draw one line per group. This is doable by specifying a different color to each group with the `color` argument of `ggplot2`.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=6}
# Libraries
library(ggplot2)
library(babynames) # provide the dataset: a dataframe called babynames
library(dplyr)

# Keep only 3 names
don <- babynames %>% 
  filter(name %in% c("Ashley", "Patricia", "Helen")) %>%
  filter(sex=="F")
  
# Plot
don %>%
  ggplot( aes(x=year, y=n, group=name, color=name)) +
    geom_line()
```


</div>
</div>



```{r thecode, eval=FALSE}
```










# Customize the grouped line chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Several options are available to customize the line chart appearance:

- Add a title with `ggtitle()`.
- Change line style with arguments like `shape`, `size`, `color` and more.
- Use the `viridis` package to get a nice color palette.
- Custom the general theme with the `theme_ipsum()` function of the `hrbrthemes` package.

More generally, visit the [ggplot2 section] for more ggplot2 related stuff.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=6}
# Libraries
library(ggplot2)
library(babynames) # provide the dataset: a dataframe called babynames
library(dplyr)
library(hrbrthemes)
library(viridis)

# Keep only 3 names
don <- babynames %>% 
  filter(name %in% c("Ashley", "Patricia", "Helen")) %>%
  filter(sex=="F")
  
# Plot
don %>%
  ggplot( aes(x=year, y=n, group=name, color=name)) +
    geom_line() +
    scale_color_viridis(discrete = TRUE) +
    ggtitle("Popularity of American names in the previous 30 years") +
    theme_ipsum() +
    ylab("Number of babies born")
```


</div>
</div>



```{r thecode2, eval=FALSE}
```







# Notes
***

- Read more about line charts theory [here](https://www.data-to-viz.com/graph/line.html)
- Be careful: a line chart with too many groups results in a [spaghetti chart](https://www.data-to-viz.com/caveat/spaghetti.html), which is a [bad practice](https://www.data-to-viz.com/caveats.html).
- Visit [data-to-viz](https://www.data-to-viz.com) for line chart alternatives.
 


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
