---
title: "Dual Y axis with R and ggplot2"
descriptionMeta: "This post describes how to build a dual Y axis chart using R and ggplot2. It uses the sec.axis attribute to add the second Y axis."
descriptionTop: "This post describes how to build a dual Y axis chart using `R` and `ggplot2`. It uses the `sec.axis` attribute to add the second Y axis. Note that this kind of chart has major [drawbacks](https://www.data-to-viz.com/caveats.html). Use it with care."
sectionText: "Line chart Section"
sectionLink: "line-plot.html"
DataToVizText: "About line chart"
DataToVizLink: "data-to-viz.com/graph/line.html"
url: "line-chart-dual-Y-axis-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Visualizing 2 series with `R` and `ggplot2`
***

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

Let's consider a dataset with 3 columns:

- `date`
- first serie to display: fake `temperature`. Range from 0 to 10.
- second serie: fake `price`. Rangee from 0 to 100.

One could easily build 2 line charts to study the evolution of those 2 series using the code below. 

But even if strongly [unadvised](https://www.data-to-viz.com/caveats.html), one sometimes wants to display both series on the same chart, thus needing a second Y axis.

</div>


<div class = "col-md-8  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=5}
# Libraries
library(ggplot2)
library(dplyr)
library(patchwork) # To display 2 charts together
library(hrbrthemes)

# Build dummy data
data <- data.frame(
  day = as.Date("2019-01-01") + 0:99,
  temperature = runif(100) + seq(1,100)^2.5 / 10000,
  price = runif(100) + seq(100,1)^1.5 / 10
)

# Most basic line chart
p1 <- ggplot(data, aes(x=day, y=temperature)) +
  geom_line(color="#69b3a2", size=2) +
  ggtitle("Temperature: range 1-10") +
  theme_ipsum()
  
p2 <- ggplot(data, aes(x=day, y=price)) +
  geom_line(color="grey",size=2) +
  ggtitle("Price: range 1-100") +
  theme_ipsum()

# Display both charts side by side thanks to the patchwork package
p1 + p2
```


</div>
</div>



```{r thecode, eval=FALSE}
```





# Adding a second Y axis with `sec.axis()`: the idea
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`sec.axis()` does not allow to build an entirely new Y axis. It just builds a second Y axis based on the first one, applying a mathematical transformation.

In the example below, the second Y axis simply represents the first one multiplied by 10, thanks to the `trans` argument that provides the `~.*10` mathematical statement.

Note that because of that you can't easily control the second axis lower and upper boundaries. We'll see a trick below in the tweaking section.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode1, echo=FALSE, out.width = "100%", fig.height=5}
# Start with a usual ggplot2 call:
ggplot(data, aes(x=day, y=temperature)) +
  
  # Custom the Y scales:
  scale_y_continuous(
    
    # Features of the first axis
    name = "First Axis",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~.*10, name="Second Axis")
  ) +
  
  theme_ipsum()
```


</div>
</div>



```{r thecode1, eval=FALSE}
```










# Show 2 series on the same line chart thanks to `sec.axis()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

We can use this `sec.axis` mathematical transformation to display 2 series that have a different range.

Since the price has a maximum value that is 10 times biggeer than the maximum temperature:

- the second Y axis is like the first multiplied by 10 (`trans=~.*10`).
- the value be display in the second variable `geom_line()` call must be divided by 10 to mimic the range of the first variable.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Value used to transform the data
coeff <- 10

ggplot(data, aes(x=day)) +
  
  geom_line( aes(y=temperature)) + 
  geom_line( aes(y=price / coeff)) + # Divide by 10 to get the same range than the temperature
  
  scale_y_continuous(
    
    # Features of the first axis
    name = "First Axis",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Second Axis")
  )
```


</div>
</div>



```{r thecode2, eval=FALSE}
```















# Dual Y axis customization with `ggplot2`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

A feew usual tricks to make the chart looks better:

- `ipsum` theme to remove the black background and improve the general style
- add a title
- customize the Y axes to pair them with their related line.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Value used to transform the data
coeff <- 10

# A few constants
temperatureColor <- "#69b3a2"
priceColor <- rgb(0.2, 0.6, 0.9, 1)

ggplot(data, aes(x=day)) +
  
  geom_line( aes(y=temperature), size=2, color=temperatureColor) + 
  geom_line( aes(y=price / coeff), size=2, color=priceColor) +
  
  scale_y_continuous(
    
    # Features of the first axis
    name = "Temperature (Celsius °)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Price ($)")
  ) + 
  
  theme_ipsum() +

  theme(
    axis.title.y = element_text(color = temperatureColor, size=13),
    axis.title.y.right = element_text(color = priceColor, size=13)
  ) +

  ggtitle("Temperature down, price up")
  
```


</div>
</div>



```{r thecode3, eval=FALSE}
```











# Barplot with overlapping line chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is totally possible to usee the same tricks with other `geoms`. 

Here is an example displaying a [line chart](line-plot.html) on top of a [barplot](barplot.html).
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=7}
# Value used to transform the data
coeff <- 10

# A few constants
temperatureColor <- "#69b3a2"
priceColor <- rgb(0.2, 0.6, 0.9, 1)

ggplot(head(data, 80), aes(x=day)) +
  
  geom_bar( aes(y=temperature), stat="identity", size=.1, fill=temperatureColor, color="black", alpha=.4) + 
  geom_line( aes(y=price / coeff), size=2, color=priceColor) +
  
  scale_y_continuous(
    
    # Features of the first axis
    name = "Temperature (Celsius °)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Price ($)")
  ) + 
  
  theme_ipsum() +

  theme(
    axis.title.y = element_text(color = temperatureColor, size=13),
    axis.title.y.right = element_text(color = priceColor, size=13)
  ) +

  ggtitle("Temperature down, price up")
  
```


</div>
</div>



```{r thecode4, eval=FALSE}
```




<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
