---
title: "Line chart annotation with ggplot2"
descriptionMeta: "Annotation is a crucial part of a time sery visual. This post shows how to highlight main parts of a line chart with text, circles, lines and more."
descriptionTop: "Annotation is a crucial part of a [time sery](time-series.html) visual. This post shows how to highlight main parts of a [line chart](line-plot.html) with text, circles, lines and more."
sectionText: "Time series section"
sectionLink: "time-series.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/connectedscatter.html"
url: "line_chart_annotation"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 80px">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The `ggplot2` package recognizes the `date` format and automatically uses a specific type of X axis. If the time variable isn't at the `date` format, this won't work. Always check with `str(data)` how variables are understood by R. If not read as a date, use [lubridate](time-series.html) to convert it. Read more about this [here](time-series.html).

On the chart beside, dates are displayed using a neat format: month + year.

<u>Note</u>: the gallery offers a section dedicated to [line charts](line-plot.html).

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Libraries
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# plot
data %>% 
  ggplot( aes(x=date, y=value)) +
    geom_line(color="#69b3a2") +
    ylim(0,22000) +
    annotate(geom="text", x=as.Date("2017-01-01"), y=20089, 
             label="Bitcoin price reached 20k $\nat the end of 2017") +
    annotate(geom="point", x=as.Date("2017-12-17"), y=20089, size=10, shape=21, fill="transparent") +
    geom_hline(yintercept=5000, color="orange", size=.5) +
    theme_ipsum()
```
</div>
</div>



```{r thecode, eval=FALSE}
```


















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
