---
title: "Line plot with log scale"
descriptionMeta: "How to build a line chart that has a log scale for its Y axis: a post with explanation and reproducible code."
descriptionTop: "This post explaines how to build a line chart with a log scale for its Y axis, using the `scale_y_log10` function."
sectionText: "Lollipop section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "line-plot-log-scale"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">





<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

It is sometimes useful to use a log scale for a numeric variable. Indeed, it allows to "magnify" the lower part of the curve.

This is possible thanks to the `scale_y_log10()` function. Control the horizontal grid lines with `breaks`, and the axis limits with `limits`. (Note that 0 is not allowed, since log(0) is not defined.)

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)

# Create dummy data
data <- data.frame(
  x=seq(10,100),
  y=seq(10,100)/2+rnorm(90)
)

# Make the plot
ggplot(data, aes(x=x, y=y)) +
  geom_line() +
  scale_y_log10( breaks=c(1,5,10,15,20,50,100), limits=c(1,100) )
```

</div>
</div>



```{r thecode, eval=FALSE}
```





# Without log transform
***


<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the same chart without the log transform:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Library
library(ggplot2)

# Create dummy data
data <- data.frame(
  x=seq(10,100),
  y=seq(10,100)/2+rnorm(90)
)

# Make the plot
ggplot(data, aes(x=x, y=y)) +
  geom_line() 
```

</div>
</div>



```{r thecode2, eval=FALSE}
```





<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedEvolution.html")
```
