---
title: "Interactive network chart with R"
descriptionMeta: "The networkD3 package allows to build interactive network diagram directly from R. It allows to zoom, hover nodes, reorganize the layout and more. This post explains how to use the library with reproducible examples."
descriptionTop: "The `networkD3` package allows to build interactive [network diagram](network.html) directly from `R`. It allows to zoom, hover nodes, reorganize the layout and more. This post explains how to use the library with reproducible examples."
sectionText: "Network section"
sectionLink: "network.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/network.html"
url: "network-interactive"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic network diagram
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Here is a very basic interactive network diagram built with the `networkD3` package. Its creation follows those steps:

- Input has to be at edge list format: each row specify a link between 2 nodes
- The `simpleNetwork()` function allows to plot the interactive chart directly
- The `saveWidget()` function allows to save the chart in a standalone `.html` file if needed

</div>


<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/networkInteractive1.html" height="300" width="300" style="border:none;"></iframe></center>

</div>
</div>




```{r, eval=FALSE}
# Libraries
library(igraph)
library(networkD3)

# create a dataset:
data <- data_frame(
  from=c("A", "A", "B", "D", "C", "D", "E", "B", "C", "D", "K", "A", "M"),
  to=c("B", "E", "F", "A", "C", "A", "B", "Z", "A", "C", "A", "B", "K")
)

# Plot
p <- simpleNetwork(data, height="100px", width="100px")

# p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/networkInteractive1.html"))
```













# Interactive network customization
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Many option are available to customize the interactive diagram.

- some options allow to customize the node, links and label feature, like `nodeColour` or `fontSize`. They are easy to understand and you can get a list of them with `help(simpleNetwork)`.
- other options allow to control the layout, i.e. how nodes are distributed on the window. `networkD3` uses a force algorithm: each nodes are gonna be attracted / repulsed depending on their connection degree?


</div>


<div class = "col-md-6  col-sm-12">

<center><iframe src="HtmlWidget/networkInteractive2.html" height="500" width="400" style="border:none;"></iframe></center>

</div>
</div>




```{r, eval=FALSE}
# Libraries
library(igraph)
library(networkD3)

# create a dataset:
data <- data_frame(
  from=c("A", "A", "B", "D", "C", "D", "E", "B", "C", "D", "K", "A", "M"),
  to=c("B", "E", "F", "A", "C", "A", "B", "Z", "A", "C", "A", "B", "K")
)

# Plot
p <- simpleNetwork(data, height="100px", width="100px",        
        Source = 1,                 # column number of source
        Target = 2,                 # column number of target
        linkDistance = 10,          # distance between node. Increase this value to have more space between nodes
        charge = -900,                # numeric value indicating either the strength of the node repulsion (negative value) or attraction (positive value)
        fontSize = 14,               # size of the node names
        fontFamily = "serif",       # font og node names
        linkColour = "#666",        # colour of edges, MUST be a common colour for the whole graph
        nodeColour = "#69b3a2",     # colour of nodes, MUST be a common colour for the whole graph
        opacity = 0.9,              # opacity of nodes. 0=transparent. 1=no transparency
        zoom = T                    # Can you zoom on the figure?
        )

#p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/networkInteractive2.html"))
```










</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```
