---
title: "Combine multiple ggplots with **patchwork**"
descriptionMeta: "This post explains how to combine multiple ggplot objects into a single plot using the patchwork package. It provides several reproducible examples with explanations and R code."
descriptionTop: "The `patchwork` package in R is a powerful tool that allows for the easy combination of multiple [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) plots into a single, cohesive visualization."
sectionText: "Patchwork section"
sectionLink: "patchwork-chart.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/patchwork.html"
url: "package-patchwork"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# In a nutshell:
***

**goal**: easily combine multiple ggplot objects into a cohesive visualization

✍️ **author**: Thomas Lin Pedersen

📘 **documentation**: [github](https://patchwork.data-imaginist.com/)

⭐️ *highly acclaimed and widely used in the ggplot2 community*

# Installation
***
Getting started with `patchwork` is simple. 

You can install `patchwork` directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("patchwork")
```

# Key features
***

The `patchwork` package in R is designed to seamlessly combine multiple [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) plots. 

It provides an intuitive syntax and a set of operators to **arrange, annotate, and align** plots, making it an indispensable tool for creating complex visualizations from individual ggplot objects.

# Basic usage 
***

Combining plots with `patchwork` is as simple as using mathematical operators:

```{r}
library(ggplot2)
library(patchwork)

p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))

p1 + p2
```

### Gallery of patchwork examples
***

<section class="bg" id="portfolio" style="padding-top: 10px"> 
<!-- Add relevant images and descriptions for patchwork examples here -->
</section>

<!-- Close container -->
</div>
```

