---
title: "Parallel coordinates chart with ggally"
descriptionMeta: "ggally is a ggplot2 extension. It allows to build parallel coordinates chart thanks to the ggparcoord() function. Check several reproducible examples in this post."
descriptionTop: "`ggally` is a `ggplot2` extension. It allows to build [parallel coordinates charts](parallel-plot.html) thanks to the `ggparcoord()` function. Check several reproducible examples in this post."
sectionText: "Parallel coord section"
sectionLink: "parallel-plot.html"
DataToVizText: "About parallel coord. charts"
DataToVizLink: "data-to-viz.com/graph/parallel.html"
url: "parallel-plot-ggally"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is the most basic parallel coordinates chart you can build with `R`, the `ggally` packages and its `ggparcoord()` function.

The input dataset must be a data frame with several numeric variables, each being used as a vertical axis on the chart. Columns number of these variables are specified in the `columns` argument of the function.

<u>Note</u>: here, a categoric variable is used to color lines, as specified in the `groupColumn` variable.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=6}
# Libraries
library(GGally)

# Data set is provided by R natively
data <- iris

# Plot
ggparcoord(data,
    columns = 1:4, groupColumn = 5
    ) 
```
</div>
</div>



```{r thecode, eval=FALSE}
```














# Custom color, theme, general appearance {#custom}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This is pretty much the same chart as te previous one, except for the following customizations:

- color palette is improved thanks to the `viridis` package
- title is added with `title`, and customized in `theme`
- dots are added with `showPoints`
- a bit of transparency is applied to lines with `alphaLines`
- `theme_ipsum()` is used for the general appearance


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=6}
# Libraries
library(hrbrthemes)
library(GGally)
library(viridis)

# Data set is provided by R natively
data <- iris

# Plot
ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for the Iris Data",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    plot.title = element_text(size=10)
  )
```
</div>
</div>



```{r thecode2, eval=FALSE}
```

















# Scaling {#scale}
***

Scaling transforms the raw data to a new scale that is common with other variables. It is a crucial step to compare variables that do not have the same unit, but can also help otherwise as shown in the example below.

The `ggally` package offers a `scale` argument. Four possible options are applied on the same dataset below:

- `globalminmax` &rarr; No scaling
- `uniminmax` &rarr; Standardize to Min = 0 and Max = 1
- `std` &rarr; Normalize univariately (substract mean & divide by sd)
- `center` &rarr; Standardize and center variables



<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode31, echo=FALSE, out.width = c("100%"), fig.height=5}
ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="globalminmax",
    showPoints = TRUE, 
    title = "No scaling",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")
```

</div>

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode32, echo=FALSE, out.width = c("100%"), fig.height=5}
ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="uniminmax",
    showPoints = TRUE, 
    title = "Standardize to Min = 0 and Max = 1",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")
```

</div>

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode33, echo=FALSE, out.width = c("100%"), fig.height=5}
ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="std",
    showPoints = TRUE, 
    title = "Normalize univariately (substract mean & divide by sd)",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")
```

</div>

<div class = "col-md-6  col-sm-12 align-self-center">

```{r thecode34, echo=FALSE, out.width = c("100%"), fig.height=5}
ggparcoord(data,
    columns = 1:4, groupColumn = 5, order = "anyClass",
    scale="center",
    showPoints = TRUE, 
    title = "Standardize and center variables",
    alphaLines = 0.3
    ) + 
  scale_color_viridis(discrete=TRUE) +
  theme_ipsum()+
  theme(
    legend.position="none",
    plot.title = element_text(size=13)
  ) +
  xlab("")
```

</div>
</div>




```{r thecode31, eval=FALSE}
```

```{r thecode32, eval=FALSE}
```

```{r thecode33, eval=FALSE}
```

```{r thecode34, eval=FALSE}
```










# Highlight a group {#highlight}
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Data visualization aims to highlight a story in the data. If you are interested in a specific group, you can highlight it as follow:

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode4, echo=FALSE, out.width = "100%", fig.height=6}
# Libraries
library(GGally)
library(dplyr)

# Data set is provided by R natively
data <- iris

# Plot
data %>%
  arrange(desc(Species)) %>%
  ggparcoord(
    columns = 1:4, groupColumn = 5, order = "anyClass",
    showPoints = TRUE, 
    title = "Original",
    alphaLines = 1
    ) + 
  scale_color_manual(values=c( "#69b3a2", "#E8E8E8", "#E8E8E8") ) +
  theme_ipsum()+
  theme(
    legend.position="Default",
    plot.title = element_text(size=10)
  ) +
  xlab("")
```
</div>
</div>



```{r thecode4, eval=FALSE}
```










<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
