---
title: "ggplot2 Piechart"
descriptionMeta: "How to build a pie chart with ggplot2 to visualize the proportion of a set of groups. Several examples with reproducible code provided."
descriptionTop: "A [pie chart](pie-plot.html) is a circle divided into sectors that each represent a proportion of the whole. This page explains how to build one with the `ggplot2` package."
sectionText: "Pie Chart section"
sectionLink: "pie-plot.html"
DataToVizText: "Why you should not do it"
DataToVizLink: "data-to-viz.com/caveat/pie.html"
url: "piechart-ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container" style="padding-top: 100px">




# Most basic pie chart
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

`ggplot2` does not offer any specific geom to build piecharts. The trick is the following:

- input data frame has 2 columns: the group names (`group` here) and its value (`value` here)
- build a stacked barchart with one bar only using the `geom_bar()` function.
- Make it circular with `coord_polar()`

The result is far from optimal yet, keep reading for improvements.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create Data
data <- data.frame(
  group=LETTERS[1:5],
  value=c(13,7,9,21,2)
)

# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)
```
</div>
</div>



```{r thecode, eval=FALSE}
```










# Improve appearance
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Previous version looks pretty bad. We need to:

- remove useless numeric labels
- remove grid and grey background

It's better now, just need to add labels directly on chart.
</div>


<div class = "col-md-6  col-sm-12">

```{r thecode2, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create Data
data <- data.frame(
  group=LETTERS[1:5],
  value=c(13,7,9,21,2)
)

# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() # remove background, grid, numeric labels
  
```

</div>
</div>



```{r thecode2, eval=FALSE}
```












# Adding labels with `geom_text()`
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

Since ggplot2 v2.2.0, the position of labels can easily be set with `position_stack()`.


</div>


<div class = "col-md-6  col-sm-12">

```{r thecode3, echo=FALSE, out.width = "100%", fig.height=7}
# Load ggplot2
library(ggplot2)

# Create Data
data <- data.frame(
  group=LETTERS[1:5],
  value=c(13,7,9,21,2)
)

# Basic piechart
ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() + # remove background, grid, numeric labels

  geom_text(aes(label = group), position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette="Set1")
  
```

</div>
</div>



```{r thecode3, eval=FALSE}
```









<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
