---
title: "Interactive stacked area chart with R and plotly"
descriptionMeta: "The plotly package allows to build interactive charts directly from R. Here is a application to stacked area chart, with explanation and reproducible code."
descriptionTop: "The `plotly` package allows to build interactive charts directly from `R`. Here is a application to [stacked area chart](stacked-area-graph.html), using both the `plot_ly()` and `ggplotly()` functions."
sectionText: "Stacked area section"
sectionLink: "stacked-area-graph.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/stackedarea.html"
url: "stacked-area-chart-plotly"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Using the `ggplotly()` function
***

<div class = "row">

<div class = "col-md-4  col-sm-12 align-self-center">

This post follows the previous [basic stacked area chart](136-stacked-area-chart.html) built with [ggplot2](ggplot2.html).

The idea is to turn the chart interactive:

- You can zoom by selecting an area of interest
- Hover the line to get exact time and value
- Export to png
- Slide axis
- Double click to re-initialize.

This is done thanks to the `ggplotly()` function of the `plotly` package that turn any `ggplot2` chart object interactive.

</div>


<div class = "col-md-8  col-sm-12">

<center><iframe src="HtmlWidget/ggplotlyStackedareachart.html" height="600" width="700" style="border:none;"></iframe></center>

</div>
</div>




```{r, eval=FALSE}
# Libraries
library(ggplot2)
library(dplyr)
library(babynames)
library(viridis)
library(hrbrthemes)
library(plotly)

# Load dataset from github
data <- babynames %>% 
  filter(name %in% c("Ashley", "Amanda", "Jessica",    "Patricia", "Linda", "Deborah",   "Dorothy", "Betty", "Helen")) %>%
  filter(sex=="F")

# Plot
p <- data %>% 
  ggplot( aes(x=year, y=n, fill=name, text=name)) +
    geom_area( ) +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Popularity of American names in the previous 30 years") +
    theme_ipsum() +
    theme(legend.position="none")

# Turn it interactive
p <- ggplotly(p, tooltip="text")
p

# save the widget
# library(htmlwidgets)
# saveWidget(p, file=paste0( getwd(), "/HtmlWidget/ggplotlyStackedareachart.html"))
```















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
