---
title: "Upset plot"
descriptionMeta: "An Upset plot is an alternative to the Venn Diagram used to deal with more than 3 sets. It is built in R thanks to the upSetR package."
descriptionTop: "An Upset plot is an alternative to the [Venn Diagram](venn-diagramm.html) used to deal with more than 3 sets. It is built in `R` thanks to the `upSetR` package."
sectionText: "Venn diagram section"
sectionLink: "venn-diagram.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/venn.html"
url: "upset-plot"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">




# Most basic
***





```{r theCode, echo=FALSE, fig.out="100%", fig.align="center"}
# Specific library
library(UpSetR)

# Dataset
input <- c(
  M.acuminata = 759,
  P.dactylifera = 769,
  A.thaliana = 1187,
  O.sativa = 1246,
  S.bicolor = 827,
  B.distachyon = 387,
  "P.dactylifera&M.acuminata" = 467,
  "O.sativa&M.acuminata" = 29,
  "A.thaliana&O.sativa" = 6,
  "S.bicolor&A.thaliana" = 9,
  "O.sativa&P.dactylifera" = 32,
  "S.bicolor&P.dactylifera" = 49,
  "S.bicolor&M.acuminata" = 49,
  "B.distachyon&O.sativa" = 547,
  "S.bicolor&O.sativa" = 1151,
  "B.distachyon&A.thaliana" = 10,
  "B.distachyon&M.acuminata" = 9,
  "B.distachyon&S.bicolor" = 402,
  "M.acuminata&A.thaliana" = 155,
  "A.thaliana&P.dactylifera" = 105,
  "B.distachyon&P.dactylifera" = 25,
  "S.bicolor&O.sativa&P.dactylifera" = 42,
  "B.distachyon&O.sativa&P.dactylifera" = 12,
  "S.bicolor&O.sativa&B.distachyon" = 2809,
  "B.distachyon&O.sativa&A.thaliana" = 18,
  "S.bicolor&O.sativa&A.thaliana" = 40,
  "S.bicolor&B.distachyon&A.thaliana" = 14,
  "O.sativa&B.distachyon&M.acuminata" = 28,
  "S.bicolor&B.distachyon&M.acuminata" = 13,
  "O.sativa&M.acuminata&P.dactylifera" = 35,
  "M.acuminata&S.bicolor&A.thaliana" = 21,
  "B.distachyon&M.acuminata&A.thaliana" = 7,
  "O.sativa&M.acuminata&A.thaliana" = 13,
  "M.acuminata&P.dactylifera&A.thaliana" = 206,
  "P.dactylifera&A.thaliana&S.bicolor" = 4,
  "O.sativa&A.thaliana&P.dactylifera" = 6,
  "S.bicolor&O.sativa&M.acuminata" = 64,
  "S.bicolor&M.acuminata&P.dactylifera" = 19,
  "B.distachyon&A.thaliana&P.dactylifera" = 3,
  "B.distachyon&M.acuminata&P.dactylifera" = 12,
  "B.distachyon&S.bicolor&P.dactylifera" = 23,
  "M.acuminata&B.distachyon&S.bicolor&A.thaliana" = 54,
  "P.dactylifera&S.bicolor&O.sativa&M.acuminata" = 62,
  "B.distachyon&O.sativa&M.acuminata&P.dactylifera" = 18,
  "S.bicolor&B.distachyon&O.sativa&A.thaliana" = 206,
  "B.distachyon&M.acuminata&O.sativa&A.thaliana" = 29,
  "O.sativa&M.acuminata&A.thaliana&S.bicolor" = 71,
  "M.acuminata&O.sativa&P.dactylifera&A.thaliana" = 28,
  "B.distachyon&M.acuminata&O.sativa&A.thaliana" = 7,
  "B.distachyon&S.bicolor&P.dactylifera&A.thaliana" = 11,
  "B.distachyon&O.sativa&P.dactylifera&A.thaliana" = 5,
  "A.thaliana&P.dactylifera&S.bicolor&O.sativa" = 21,
  "M.acuminata&S.bicolor&P.dactylifera&A.thaliana" = 23,
  "M.acuminata&B.distachyon&S.bicolor&P.dactylifera" = 24,
  "M.acuminata&O.sativa&S.bicolor&B.distachyon" = 368,
  "P.dactylifera&B.distachyon&S.bicolor&O.sativa" = 190,
  "P.dactylifera&B.distachyon&S.bicolor&O.sativa&A.thaliana" = 258,
  "P.dactylifera&M.acuminata&S.bicolor&B.distachyon&O.sativa" = 685,
  "M.acuminata&S.bicolor&B.distachyon&O.sativa&A.thaliana" = 1458,
  "S.bicolor&M.acuminata&P.dactylifera&O.sativa&A.thaliana" = 149,
  "B.distachyon&M.acuminata&P.dactylifera&O.sativa&A.thaliana" = 80,
  "M.acuminata&S.bicolor&B.distachyon&P.dactylifera&A.thaliana" = 113,
  "M.acuminata&S.bicolor&B.distachyon&P.dactylifera&O.sativa&A.thaliana" = 7674
)

# Plot
upset(fromExpression(input), 
      nintersects = 40, 
      nsets = 6, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )
```

To visualize the intersection between more than 3 sets, the best option is to use an UpSet plot.

Here is an example provided by the `UpsetR` package that displays information concerning the Banana genome.

The total size of each set is represented on the left barplot. Every possible intersection is represented by the bottom plot, and their occurence is shown on the top barplot.

<u>Source</u>: the [upSetR documentation](https://github.com/hms-dbmi/UpSetR)



```{r theCode, eval=FALSE}
```
















<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
