---
title: "Vaccination Impact Heatmap with R and ggplot2"
descriptionMeta: "This tutorial explains how to create an advanced temporal heatmap with R and ggplot2. We'll use **measles data from US states between 1928 and 2011** as an example, highlighting the effect of vaccine introduction in 1963."
descriptionTop: "This tutorial explains how to create an advanced temporal [heatmap](heatmap.html) built with `R` and [ggplot2](ggplot2-package.html). We'll use measles data from US states between 1928 and 2011 as an example, highlighting the effect of vaccine introduction in 1963."
sectionText: "Heatmap section"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/heatmap.html"
url: "vaccination-heatmap"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# About
***

This chart visualizes **measles cases across US states** between 1928-2011 using a **temporal heatmap**. It highlights the impact of the 1963 measles vaccine introduction, demonstrating through historical public health data how vaccination successfully reduced disease incidence across the United States.

The visualization is based on **historical public health records** and clearly demonstrates the effectiveness of vaccination programs, providing insights into one of the most successful public health interventions in US history. 

This chart was adapted from [Ben Moore's post](https://benjaminlmoore.wordpress.com/2015/04/09/recreating-the-vaccination-heatmaps-in-r/). Thanks to him for sharing this insightful chart!


# Load required libraries 
***
This visualization only requires two R packages: `dplyr` for data preparation, and `ggplot2` for creating the heatmap.
```{r}
# Libraries
library(dplyr)
library(ggplot2)
```

# Read and prepare the data 
***

This code reads a CSV file containing measles cases per million inhabitants per year and US state. First, it imports the data using `readr::read_csv()`. See how the data is structured from the `print()` output. Then, it creates a new dataset "mdf" where **states are converted to factors** and **arranged in reverse alphabetical order** - this ordering will determine how states are displayed from top to bottom in our final heatmap.

```{r}
# Read the data 
measles = readr::read_csv("measles_data_1928-2011.csv")

# Show the first few rows of the data
print(head(measles))

# Prepare the data: rename column and sort states alphabetically
mdf <- measles %>% 
  mutate(State = factor(State, levels = rev(sort(unique(State)))))
```

# Create color palette
***

This code creates a **custom color gradient** for the heatmap that will represent different levels of measles cases. It combines two color palettes: a spectrum from white through blue, green, yellow, and orange (for lower case counts up to 500), followed by an orange-to-red gradient (for higher case counts from 500-4000). The `colorRampPalette()` function smoothly interpolates between these colors to create 100 total color steps.

```{r}
cols <- c(
  colorRampPalette(c("white", "blue", "green", "yellow", "orange"))(20),  # Full spectrum until 500
  colorRampPalette(c("orange", "red"))(80)  # Orange to red gradient for 500-4000
)
```

# Create advanced temporal heatmap 
***

This code generates a comprehensive **temporal heatmap visualization** in R using `ggplot2` to display measles cases across US states from 1928 to 2011. It also highlights the **year of vaccine introduction**. 

### Core Plot Structure:

- Creates a heatmap using `ggplot2` with states on y-axis, years on x-axis, and case numbers determining color
- Uses `geom_tile()` to create the rectangular cells with white borders

### Color and Scale Settings:

- Applies the custom color gradient defined earlier
- Sets color scale limits from 0 to 4,000 cases per million
- Formats legend labels with 'k' suffix (0k to 4k)

### Timeline Features:

- X-axis spans 1928-2012 with decade intervals
- Adds black vertical line at 1963 (vaccine introduction)
- Includes annotation **"Vaccine introduced"** near the line

### Visual Styling:

- Uses `theme_minimal()` with customized elements
- Removes grid lines and unnecessary aesthetic elements
- Sets horizontal legend at bottom of plot
- Uses Helvetica font family throughout
- Customizes text sizes (smaller for states, larger for years)

### Title and Labels:

- Adds main title "Measles cases per million by US state, 1928-2011"
- Removes x and y-axis labels
- Positions legend title "Cases per million" at top of color bar


```{r, out.width = "100%"}
ggplot(mdf, aes(x = Year, y = State, fill = value)) + 
geom_tile(colour = "white", linewidth = 0.5, 
            width = .9, height = .9) + 
theme_minimal() +
scale_fill_gradientn(colours = cols,
                    limits = c(0, 4000),
                    breaks = seq(0, 4000, by = 1000),
                    labels = c("0k", "1k", "2k", "3k", "4k"),
                    na.value = rgb(246/255, 246/255, 246/255),
                    guide = guide_colourbar(ticks = TRUE, 
                                          nbin = 50,
                                          barheight = .5, 
                                          label = TRUE, 
                                          barwidth = 10,
                                          title = "Cases per million",
                                          title.position = "top",
                                          title.hjust = 0.5)) +
scale_x_continuous(expand = c(0,0), 
                  breaks = seq(1930, 2010, by = 10),
                  limits = c(1928, 2012)) +
geom_vline(xintercept = 1963, color = "black", size = 0.5) +
  theme(legend.position = c(.5, -.13),
      legend.direction = "horizontal",
      legend.text = element_text(colour = "grey20"),
      plot.margin = grid::unit(c(.5,.5,1.5,.5), "cm"),
      axis.text.y = element_text(size = 6, family = "Helvetica", 
                                hjust = 1),
      axis.text.x = element_text(size = 8),
      axis.ticks.y = element_blank(),
      panel.grid = element_blank(),
      title = element_text(hjust = -.07, face = "bold", vjust = 1, 
                          family = "Helvetica"),
      text = element_text(family = "Helvetica")) +
ggtitle("Measles cases per million by US state, 1928-2011") +
annotate("text", label = "Vaccine introduced", x = 1963, y = 53, 
         vjust = 0.9, hjust = -0.1, size = I(3), family = "Helvetica") +
  xlab("") + ylab("")
```


<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
