---
title: "Violin plot with included boxplot and sample size in ggplot2"
descriptionMeta: "An implementation of a violin plot with included boxplot and sample size provided on the X axis. R reproducible code provided, use of the ggplot2 library."
descriptionTop: "It can be handy to include a [boxplot](boxplot.html) in the [violin plot](violin plot) to see both the distribution of the data and its summary statistics. Moreover, adding sample size of each group on the X axis is often a necessary step. Here is how to do it with <code>R</code> and [ggplot2](ggplot2-package.html)."
sectionText: "Violin Section"
sectionLink: "violin.html"
DataToVizText: "Boxplot caveat"
DataToVizLink: "data-to-viz.com/caveat/boxplot.html"
url: "violin_and_boxplot_ggplot2"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"

---


```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```


<div class="container" style="padding-top: 100px">


Building a [violin plot](violin.html) with `ggplot2` is pretty straightforward thanks to the dedicated `geom_violin()` function. It is possible to use `geom_boxplot()` with a small `width` in addition to display a boxplot that provides summary statistics.

Moreover, note a small trick that allows to provide sample size of each group on the X axis: a new column called `myaxis` is created and is then used for the X axis.

```{r, warning=FALSE, out.width = "90%", fig.width=8, fig.height=5, fig.align="center"}
# Libraries
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

# create a dataset
data <- data.frame(
  name=c( rep("A",500), rep("B",500), rep("B",500), rep("C",20), rep('D', 100)  ),
  value=c( rnorm(500, 10, 5), rnorm(500, 13, 1), rnorm(500, 18, 1), rnorm(20, 25, 4), rnorm(100, 12, 1) )
)

# sample size
sample_size = data %>% group_by(name) %>% summarize(num=n())

# Plot
data %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(name, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=value, fill=name)) +
    geom_violin(width=1.4) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("A Violin wrapping a boxplot") +
    xlab("")
```













<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```
