---
title: "Customized Dorling cartogram with R"
descriptionMeta: "This post explains how to build a Dorling Cartogram with R and ggplot2. Explanation and reproducible code provided."
descriptionTop: "This blogpost explains how to build a **Dorling Cartogram** with `R`. A dorling cartogram is a variation of a [cartogram](https://r-graph-gallery.com/cartogram.html), close to a [bubble map](https://r-graph-gallery.com/bubble-map.html).<br/>The chart is implemented using the `ggplot2`, `cartogram`, `sf` and a few other packages. This article includes a step-by-step walkthrough of the map construction, together with its final reproducible code."
sectionText: "Cartogram section"
sectionLink: "bubble-map.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/bubble-map.html"
url: "web-dorling-cartogram-with-R.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
editor_options: 
  chunk_output_type: console
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE, 
  message=FALSE
)
```

<div class="container" style="padding-top: 100px">


A Dorling cartogram replaces polygons with circles proportional to a given variable, respecting the spatial position of the polygons as much as possible. So normally **only one variable can be represented with a Dorling cartogram**. 

This tutorial will show how these cartograms can be adapted to represent the value of sub-variables. For this tutorial, the main variable that we will plot as a Dorling cartogram is **the total agricultural land per country**. The sub-variables we are going to add are **arable land area** (~crop land) and **permanent grass area**.



# 1. Load data
***

To create our customized Dorling cartogram, we will need these packages:

```{r, warning=FALSE, message=FALSE}
library(rnaturalearth)
library(sf)
library(tidyverse)
library(cartogram)
library(ggforce)
```

We will download the world basemap with {rnaturalearth}. Data about agricultural land use have been downloaded from [FAOStat](https://www.fao.org/faostat/en/#home).  Data may be downloaded at [this link](https://raw.githubusercontent.com/BjnNowak/dorling_map/main/data/FAOSTAT_land_use_2020.csv) or loaded directy in R as shown below:

```{r, warning=FALSE, message=FALSE}
world <- ne_countries(scale = 110, type = "countries", returnclass = "sf")%>%
  # Convert WGS84 to projected crs (here Robinson)
  sf::st_transform(world_ne, crs="ESRI:54030")

data <- read_csv('https://raw.githubusercontent.com/BjnNowak/dorling_map/main/data/FAOSTAT_land_use_2020.csv')
```

For data cleaning, we will create one column per variable of interest (total agricultural land, crop land and grass land), then join the data to the map.

```{r, warning=FALSE, message=FALSE}
clean <- data%>%
  select(iso_a3='Area Code (ISO3)',item=Item,surface=Value)%>%
  # Simplify variable names
  mutate(item=case_when(
    item=="Agricultural land"~"total",
    item=="Arable land"~"crop",
    item=="Permanent meadows and pastures"~"grass"
  ))%>%
  # Pivot from long to wide
  pivot_wider(names_from=item,values_from=surface)

# Join data to world map
map <- world%>%
  left_join(clean)%>%
  drop_na(total)

```

Before going further, we will make a simple choropleth showing total agricultural land per country: 

```{r, warning=FALSE, message=FALSE}
ggplot(map, aes(fill=total))+
  geom_sf()
```

It is this map that will be converted as a Dorling cartogram.

# 2. Make Dorling cartogram
***

It is really straightforward to make a Dorling cartogram with the {cartogram} library!

```{r, warning=FALSE, message=FALSE}
# Making Dorling cartogram based on total agricultural land
dorl<-cartogram::cartogram_dorling(
  map, weight='total', k = 5,
  m_weight = 1, itermax = 1000
)
```

As the circles layout may be confusing, I like to use a "traditional" map as background when plotting a Dorling cartogram:

```{r, warning=FALSE, message=FALSE}
# Set colors
col_world <- "#9CB4BF"
col_back <- "#1D201F"

# Set theme
theme_custom <- theme_void()+
  theme(plot.background = element_rect(fill=col_back,color=NA))

ggplot()+
  # World basemap
  geom_sf(
    world,mapping=aes(geometry=geometry),
    fill=col_world,color=alpha("dimgrey",0.25)
  )+
  # Dorling cartogram
  geom_sf(
    dorl,mapping=aes(geometry=geometry),
    fill=alpha("dimgrey",0.75),color=alpha("white",0.2)
  )+
  theme_custom
```

Up to this point, we have a classic Dorling cartogram, with circles proportional to total agricultural land, which is easy to plot with `geom_sf()`. But, in order to add supplementary variables to our map, we will extract some features from this map: **the radius and the coordinates of the centroids for each circus**. 

```{r, warning=FALSE, message=FALSE}
# Compute area and radius for each circus of the cartogram
dorl<-dorl%>%
  mutate(
    # Compute area
    ar=as.numeric(st_area(dorl)),
    # Compute radius based on area
    rad=as.numeric(sqrt(ar/pi))
  )

# Extract centroids for each circle
centr <- dorl%>%
  st_centroid()%>%
  st_coordinates()

# Combine data
dorl2 <- tibble(dorl,X=centr[,1],Y=centr[,2])%>%
  arrange(-total)
```

With these features, we may now make the same map as before, but using the `geom_circle()` function from the {ggforce} package.

```{r, warning=FALSE, message=FALSE}
ggplot()+
  # World basemap
  geom_sf(
    world,mapping=aes(geometry=geometry),
    fill=col_world,color=alpha("dimgrey",0.25)
  )+
  # Draw Dorling cartogram with geom_circle()
  ggforce::geom_circle(
    data = dorl2, aes(x0 = X, y0 = Y, r = rad),
    fill=alpha("dimgrey",0.75),color=alpha("white",0.2)
  )+
  theme_custom
```

# 3. Add supplementary variables to cartogram
***

The interest of calculating such features is that we can now use this workflow to compute the related radius for the surface under crop or grass (centroids will be the same).

```{r, warning=FALSE, message=FALSE}
dorl2 <- dorl2 %>%
  mutate(
    ratio_crop = crop/total,
    ratio_grass = grass/total
  )%>%
  mutate(
    rad_crop=sqrt(rad*rad*ratio_crop),
    rad_grass=sqrt(rad*rad*ratio_grass)
  )
```

This way we can add the crop (or grass) area for each country to the map:

```{r, warning=FALSE, message=FALSE}
col_crop <- "#f2e901"
col_grass <- "#51c26f"

ggplot()+
  # World basemap
  geom_sf(
    world,mapping=aes(geometry=geometry),
    fill=col_world,color=alpha("dimgrey",0.25)
  )+
  # Draw Dorling cartogram with geom_circle()
  ggforce::geom_circle(
    data = dorl2, aes(x0 = X, y0 = Y, r = rad),
    fill=alpha("dimgrey",0.75),color=alpha("white",0.2)
  )+
  # Draw circle for crops (or grass)
  ggforce::geom_circle(
    data = dorl2, aes(x0 = X, y0 = Y, r = rad_crop),
    fill=col_crop,color=NA
  )+
  theme_custom
```

But with this method it is not possible to plot at the same time the total area, the crop area and the grass area. To overcome this limit, we will plot two half circles (one for crop, one for grass) on the map.

The function below, adapted from [this post](https://stackoverflow.com/questions/28185743/draw-a-half-circle-with-ggplot2) on StackOverflow allows to draw circle portions: 

```{r, warning=FALSE, message=FALSE}
circleFun <- function(
    center=c(0,0),   # center of the circle 
    diameter=1,      # diameter 
    npoints=100,     # number of points to draw the circle
    start=0, end=2   # start point/end point
  ){
      tt <- seq(start*pi, end*pi, length.out=npoints)
      tb <- tibble(
        x = center[1] + diameter / 2 * cos(tt), 
        y = center[2] + diameter / 2 * sin(tt)
      )
    return(tb)
}
```

As an example, we will use this function to draw two half circles (one for crop, one for grass) for the first country of the map, China:

```{r, warning=FALSE, message=FALSE}
# Half circle for crops
half_crop <- bind_cols(
  iso_a3 = rep(dorl2$iso_a3[1],100),
  circleFun(
    c(dorl2$X[1],dorl2$Y[1]),dorl2$rad_crop[1]*2, start=1.5, end=2.5
  ))

# Half circle for grass
half_grass <- bind_cols(
  iso_a3 = rep(dorl2$iso_a3[1],100),
  circleFun(
    c(dorl2$X[1],dorl2$Y[1]),dorl2$rad_grass[1]*2, start=0.5, end=1.5
  ))
```

These half-circles will be added to the map using the `geom_polygon()` function, which allows to join points.

```{r, warning=FALSE, message=FALSE}
ggplot()+
  # World basemap
  geom_sf(
    world,mapping=aes(geometry=geometry),
    fill=col_world,color=alpha("dimgrey",0.25)
  )+
  # Draw Dorling cartogram with geom_circle()
  ggforce::geom_circle(
    data = dorl2, aes(x0 = X, y0 = Y, r = rad),
    fill=alpha("dimgrey",0.75),color=alpha("white",0.2)
  )+
  # Draw half circle for crop with geom_polygon
  geom_polygon(
    half_crop,
    mapping=aes(x,y,group=iso_a3),
    fill=col_crop,color=NA
  )+ 
  # Draw half circle for grass with geom_polygon
  geom_polygon(
    half_grass,
    mapping=aes(x,y,group=iso_a3),
    fill=col_grass,color=NA
  )+ 
  theme_custom
```

Now that we detailed the workflow for one country, we just have to extend it to the rest of the world, **using a for loop**.

```{r, warning=FALSE, message=FALSE}
# Make loop for all countries
for (i in 2:dim(dorl2)[1]){

  # Draw for crops
  temp_crop <- bind_cols(
    iso_a3 = rep(dorl2$iso_a3[i],100),
    circleFun(
      c(dorl2$X[i],dorl2$Y[i]),dorl2$rad_crop[i]*2, start=1.5, end=2.5
    ))
  # Draw for grass
  temp_grass <- bind_cols(
    iso_a3 = rep(dorl2$iso_a3[i],100),
    circleFun(
      c(dorl2$X[i],dorl2$Y[i]),dorl2$rad_grass[i]*2, start=0.5, end=1.5
    ))
  
  half_crop<-half_crop%>%
    bind_rows(temp_crop)
  
  half_grass<-half_grass%>%
    bind_rows(temp_grass)
}

# Make map
ggplot()+
  # World basemap
  geom_sf(
    world,mapping=aes(geometry=geometry),
    fill=col_world,color=alpha("dimgrey",0.25)
  )+
  # Draw Dorling cartogram with geom_circle()
  ggforce::geom_circle(
    data = dorl2, aes(x0 = X, y0 = Y, r = rad),
    fill=alpha("dimgrey",0.75),color=alpha("white",0.2)
  )+
  # Draw half circle for crop with geom_polygon
  geom_polygon(
    half_crop,
    mapping=aes(x,y,group=iso_a3),
    fill=col_crop,color=NA
  )+ 
  # Draw half circle for grass with geom_polygon
  geom_polygon(
    half_grass,
    mapping=aes(x,y,group=iso_a3),
    fill=col_grass,color=NA
  )+ 
  theme_custom
```

Now feel free to adapt this tutorial to your project! For example, it is also possible to represent more than two sub-variables per circle, by modifying the start and end arguments of the `circleFun()` function.


**Author** [Benjamin Nowak](https://twitter.com/BjnNowak)
<br>
**Data** [FAOStat](https://www.fao.org/faostat/en/#home)

<!-- Close container -->
</div>

```{r, echo=FALSE}
# Correlation | Ranking | Evolution.. 
htmltools::includeHTML("htmlChunkRelatedMap.html")
```
