---
title: "miRNA dysregulation dice plot with ggdiceplot"
descriptionMeta: "A dice plot showing miRNA dysregulation across compounds and organs, using categorical fill to encode regulation direction (up, down, unchanged). Built with R and the ggdiceplot package."
descriptionTop: "A dice plot showing **miRNA dysregulation** across compounds and organs, using categorical fill to encode regulation direction. Built with R and the [ggdiceplot](package/ggdiceplot.html) package."
sectionText: "Heatmap"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "web-mirna-dice-plot-with-ggdiceplot"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container" style="padding-top: 100px">


# About
***

This page showcases a **categorical fill** dice plot using the [ggdiceplot](package/ggdiceplot.html) package. Instead of mapping fill to a continuous gradient, we map it to a discrete variable (regulation direction), making it easy to see patterns of up- and down-regulation at a glance.

You can find the package source on [GitHub](https://github.com/maflot/ggdiceplot).


# Packages
***

```{r}
# install.packages("ggdiceplot")
library(ggdiceplot)
library(ggplot2)
```

<br/>


# Dataset
***

We use the bundled `sample_dice_miRNA` dataset, which simulates miRNA dysregulation across **compounds** and **organs**, with a categorical `direction` variable (Up, Down, Unchanged).

```{r}
data("sample_dice_miRNA", package = "ggdiceplot")
df_dice <- sample_dice_miRNA
head(df_dice)
```

Each row represents one miRNA–compound–organ combination. The `Organ` variable will be mapped to pip positions (the `dots` aesthetic), and `direction` to fill color.

<br/>


# Basic miRNA dice plot
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

This plot places **miRNAs** on the x-axis, **compounds** on the y-axis. Each tile has up to 4 pips (one per organ: Lung, Liver, Brain, Kidney). Pip color indicates regulation direction.

We use `scale_fill_manual()` with a diverging blue–red scheme to distinguish down- from up-regulation.

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=5}
direction_colors <- c(Down = "#2166ac", Unchanged = "grey80", Up = "#b2182b")

ggplot(df_dice, aes(x = miRNA, y = Compound)) +
  geom_dice(
    aes(dots = Organ, fill = direction, width = 0.8, height = 0.8),
    show.legend = TRUE,
    pip_scale   = 1.0,
    ndots       = length(levels(df_dice$Organ)),
    x_length    = length(levels(df_dice$miRNA)),
    y_length    = length(levels(df_dice$Compound))
  ) +
  scale_fill_manual(values = direction_colors, name = "Regulation") +
  theme_dice() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5),
    axis.text.y = element_text(hjust = 1),
    panel.grid  = element_blank()
  ) +
  labs(
    title = "DicePlot: log2FC direction per miRNA, compound and organ",
    x = "miRNA", y = "Compound"
  )
```
</div>
</div>


```{r thecode, eval=FALSE}
```

<br/>


# Customizing the theme
***

`theme_dice()` provides a clean minimal look optimized for dice plots. You can layer additional `theme()` calls on top for fine-tuning. Let's compare `theme_dice()` with `theme_minimal()`:

<div class = "row">

<div class = "col-md-6  col-sm-12">

```{r, out.width = "100%", fig.height=5}
direction_colors <- c(Down = "#2166ac", Unchanged = "grey80", Up = "#b2182b")

ggplot(df_dice, aes(x = miRNA, y = Compound)) +
  geom_dice(
    aes(dots = Organ, fill = direction, width = 0.8, height = 0.8),
    show.legend = TRUE,
    pip_scale   = 1.0,
    ndots       = length(levels(df_dice$Organ)),
    x_length    = length(levels(df_dice$miRNA)),
    y_length    = length(levels(df_dice$Compound))
  ) +
  scale_fill_manual(values = direction_colors, name = "Regulation") +
  theme_dice() +
  labs(title = "theme_dice()", x = "miRNA", y = "Compound")
```

</div>

<div class = "col-md-6  col-sm-12">

```{r, out.width = "100%", fig.height=5}
ggplot(df_dice, aes(x = miRNA, y = Compound)) +
  geom_dice(
    aes(dots = Organ, fill = direction, width = 0.8, height = 0.8),
    show.legend = TRUE,
    pip_scale   = 1.0,
    ndots       = length(levels(df_dice$Organ)),
    x_length    = length(levels(df_dice$miRNA)),
    y_length    = length(levels(df_dice$Compound))
  ) +
  scale_fill_manual(values = direction_colors, name = "Regulation") +
  theme_minimal() +
  labs(title = "theme_minimal()", x = "miRNA", y = "Compound")
```

</div>
</div>

<br/>


# Adjusting pip scale
***

The `pip_scale` parameter controls how large pips are relative to the tile. Let's compare different values side by side:

```{r, fig.width=12, fig.height=4}
library(patchwork)

make_plot <- function(ps) {
  ggplot(df_dice, aes(x = miRNA, y = Compound)) +
    geom_dice(
      aes(dots = Organ, fill = direction, width = 0.8, height = 0.8),
      pip_scale = ps,
      ndots     = length(levels(df_dice$Organ)),
      x_length  = length(levels(df_dice$miRNA)),
      y_length  = length(levels(df_dice$Compound))
    ) +
    scale_fill_manual(values = direction_colors) +
    theme_dice() +
    labs(title = paste0("pip_scale = ", ps)) +
    theme(legend.position = "none")
}

make_plot(0.5) + make_plot(0.75) + make_plot(1.0)
```


<!-- Close container -->
</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
