---
title: "Interactive Mekko Chart with ggplot2 and ggiraph"
descriptionMeta: "This post explains how to create an interactive Mekko chart in R to visualize gender distribution in sports using ggplot2 and ggiraph."
descriptionTop: "This post explains how to create an interactive Mekko chart in R to visualize gender distribution in sports using [ggplot2](package/ggplot2.html) and [ggiraph](package/ggiraph.html). The chart **highlights gender differences** in federated sports licenses in Spain."
sectionText: "Mekko chart"
sectionLink: "package/ggiraph.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/mekko.html"
url: "web-mekko-chart-with-interactivity"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
   warning = FALSE,
   message = FALSE,
   fig.align = "center"
)
```

<div class="container">

# About
***

This post demonstrates how to visualize the **gender distribution in sports** in Spain using an interactive Mekko chart. The chart uses data from the Consejo Superior de Deportes (2022) and illustrates how sports licenses are distributed among men and women across different sports categories.

It has been created by [Laura Navarro](https://laura-navarro-s.github.io/). Thanks to her for sharing this beautiful chart!

# Libraries
***

To create this chart, we use several R packages:

- `tidyverse` for data manipulation
- `showtext` for custom fonts
- [ggiraph](package/ggiraph.html) for adding interactivity to the ggplot-based Mekko chart

```{r}
library(tidyverse)
library(showtext)
library(ggiraph)
```

# Setting Up Fonts
***

To enhance the visual appeal, we use Google Fonts. The `showtext` package allows us to easily include and use these fonts in our plots.

```{r}
# Adding Google Fonts
font_add_google("DM Serif Display", "abril")
font_add_google("Tajawal", "tawa")
showtext_auto()
```

# Loading the Data
***

The dataset is loaded directly from a GitHub repository, and it contains information about the number of federated sports licenses held by men and women in different sports.

```{r}
# Load the data
fedes <- read.csv("https://raw.githubusercontent.com/lau-cloud/30DayChartChallenge2024/main/5_diverging/5_diverging.csv", encoding = "UTF-8")
```

# Data Preparation
***

The data is prepared by calculating the percentage difference between the number of men and women holding federated licenses in each sport. The data is then reshaped into a long format for easier plotting.

```{r}
# Data wrangling
fedes_clean <- fedes %>%
   mutate(diferencia = ((mujeres / total) * 100) - ((hombres / total) * 100)) %>%
   pivot_longer(cols = c("hombres", "mujeres"), names_to = "sexo", values_to = "federados") %>%
   mutate(proporcion = federados / total)

# Convert sport names to title case
fedes_clean$deporte <- str_to_sentence(fedes_clean$deporte, locale = "es")
```

# Creating Mekko Chart Rectangles
***

The Mekko chart is created by defining the dimensions of each rectangle to represent different sports and the distribution of licenses between men and women.

```{r}
# Create rectangles for Mekko chart
fedes_mosaic <- fedes_clean %>%
   group_by(deporte) %>%
   mutate(
      prop_federados = federados / sum(federados),
      tot_group = sum(federados)
   ) %>%
   ungroup() %>%
   arrange(diferencia)

fedes_mosaic2 <- fedes_mosaic %>%
   group_by(sexo) %>%
   arrange(desc(diferencia)) %>%
   mutate(
      ymax = cumsum(tot_group) / sum(tot_group),
      ymin = (ymax - (tot_group / sum(tot_group)))
   ) %>%
   ungroup() %>%
   group_by(deporte) %>%
   arrange(desc(sexo)) %>%
   mutate(
      xmax = cumsum(prop_federados),
      xmin = xmax - prop_federados,
      data_id = paste0(deporte, sexo), # For interactivity
      tooltip = paste0(
         "<b>", as.character(deporte), "</b>", "<br>",
         sexo, ": ", "<b>", round(proporcion * 100, 1), "</b>", "%<br>",
         "Total federados/as: ", federados, "<br>"
      )
   )
```

# Plotting the Mekko Chart
***

The `ggplot2` package is used to create the Mekko chart, and [ggiraph](package/ggiraph.html) is used to add interactive tooltips.

```{r, fig.width=9, fig.height=9}
# Plotting the chart
p1 <- ggplot(fedes_mosaic2) +
   geom_rect_interactive(aes(
      ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = sexo,
      data_id = data_id, tooltip = tooltip
   ), colour = "white", size = 0.1) +
   labs(
      fill = NULL, colour = NULL,
      title = "Deporte federado en España",
      subtitle = "Licencias federadas según sexo. El ancho de la barra\n indica el peso de cada deporte sobre el total",
      caption = "Fuente: Consejo Superior de Deportes, año 2022. / Laura Navarro"
   ) +
   theme_minimal() +
   scale_fill_manual(values = c("#D6DAC8", "#824D74")) +
   scale_x_continuous(labels = c("0.00" = "0%", "0.25" = "25%", "0.50" = "50%", "0.75" = "75%", "1.00" = "100%")) +
   theme(
      text = element_text(),
      legend.position = "top",
      plot.title = element_text(hjust = 0.5, size = 24),
      plot.subtitle = element_text(hjust = 0.5, size = 12, color = "darkgrey"),
      plot.caption = element_text(color = "grey", hjust = 0.5, size = 12, margin = margin(20, 0, 0, 0)),
      plot.margin = margin(1, 1, 1.5, 1.2, "cm"),
      axis.text.y = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.x = element_text(size = 12),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
   )

p1
```

# Adding Interactivity
***

[ggiraph](package/ggiraph.html) makes the plot interactive by enabling hover effects and tooltips that display additional information.

```{r}
css <- "
   background:white;
   font-family: Helvetica;
   font-size:12px;
   padding:3pt;
   border-radius:5px
"

# Adding interactivity
plot <- girafe(
   ggobj = p1,
   options = list(
      opts_hover(css = "stroke-width:1.5;stroke:black;"),
      opts_selection(only_shiny = FALSE, type = "single", css = "stroke:black;"),
      opts_tooltip(offx = 20, offy = -10, use_fill = FALSE, use_stroke = TRUE, delay_mouseout = 1000, css = css),
      opts_sizing(rescale = FALSE)
   ),
   width_svg = 7.5,
   height_svg = 9
)

# Save as HTML widget
htmltools::save_html(plot, "HtmlWidget/interactive-mekko-chart.html")
```

<iframe src="HtmlWidget/interactive-mekko-chart.html" width="900" height="900" style="border:none;"></iframe>

# Going Further
***

Explore more:

- How to make [interactive charts with ggiraph](package/ggiraph.html)
- Customize your chart with [themes](package/ggthemes.html)

<!-- Close container -->
</div>

```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedFlow.html")
```