---
title: "Ridgeline plot with inside plot and annotations"
descriptionMeta: "How to create a ridgeline plot with an inside plot for the legend and annotations in R with ggplot2. Examples of data visualization with the ridgeline plot."
descriptionTop: "A [ridgeline plot](https://data-to-viz.com/graph/ridgeline.html) is a type of chart that displays the **distribution of a numeric variable** for several groups. This type of chart can gives us **lot of information** in a single graph!<br><br>This post explains how to create a ridgeline plot with an **inside plot** for the legend and **annotations** in R with [ggplot2](https://r-graph-gallery.com/ggplot2-package.html)."
sectionText: "Ridgeline section"
sectionLink: "ridgeline-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/ridgeline.html"
url: "web-ridgeline-plot-with-inside-plot-and-annotations"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
editor_options: 
  chunk_output_type: console
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.align = "center"
)
```

<div class="container" style="padding-top: 100px">

# About
***

This page showcases the work of [Ansgar Wolsing](https://twitter.com/_ansgar) for the TidyTuesday challenge. Thanks to him for accepting sharing his work here!

The charts is representing a [ridgeline plot](https://data-to-viz.com/graph/ridgeline.html) with
an **inside plot for the legend** and annotations. The data are about the **rent prices in San Francisco**
and the keywords they are related to. 



# Load packages
***

First, we need to load a few packages that will be used to create the chart.

*Remember that you can install those packages by running `install.packages("package_name")` in your R console.*

```{r}
library(tidyverse)
library(ggtext)
library(ggdist)
library(glue)
library(patchwork)
```

<br>

# Load and prepare the dataset
***

For this post, we need 3 datasets: `rent`, `rent_title_words` and `df_plot`.

The `df_plot` is actually the **main one**, and others are used for **annotations purposes**.

```{r}
rent = read_csv("DATA/rent.csv")
rent_title_words = read_csv("DATA/rent_title_words.csv")
df_plot = read_csv("DATA/df_plot.csv")

# sort dataframe by mean_price
df_plot <- df_plot %>% arrange(desc(mean_price))
df_plot$word <- factor(df_plot$word, levels = unique(df_plot$word))
```



<br>

## Create the chart
***

- **Main Plot Construction (named `p`)**
  - `stat_halfeye()` is used for density plots and `stat_summary()` for showing medians
  - `annotate()` adds static text annotations
  - `scale_()` functions customize scales and colors, including a manual color scale using `MetBrewer::met.brewer()`
  - `coord_flip()` flips the axes to change the plot orientation

- **Legend Construction (`p_legend`)**
  - We use a subset of data (`rent_title_words`) filtered for the word *beautiful*
  - And `geom_curve` to draw arrows pointing to specific elements

- **Inserting the Legend into the Main Plot**
  - The `inset_element` function combines the main plot (`p`) and the legend (`p_legend`) by embedding the legend within the main plot's space

```{r fig.height=9, fig.width=7}
mean_price <- mean(rent$price, na.rm = TRUE)
median_price <- median(rent$price, na.rm = TRUE)
n_rental_posts <- nrow(subset(rent, !is.na(title)))

bg_color <- "grey97"
font_family <- "Fira Sans"

plot_subtitle = glue("Adjectives used to describe houses and apartments
in the San Francisco Bay Area in the titles of rental posts on Craigslist and 
how they are related to rental prices. Titles from 
{scales::number(n_rental_posts, big.mark = ',')} rental posts on Craigslist
between 2000 and 2018. 
The 15 most frequent adjectives are shown.
")

p <- df_plot %>% 
  ggplot(aes(word, price)) +
  stat_halfeye(fill_type = "segments", alpha = 0.3) +
  stat_interval() +
  stat_summary(geom = "point", fun = median) +
  annotate("text", x = 16, y = 0, label = "(\U00F8 bedrooms)",
           family = "Fira Sans", size = 3, hjust = 0.5) +
  stat_summary(
    aes(y = beds),
    geom = "text",
    fun.data = function(x) {
      data.frame(
        y = 0,
        label = sprintf("(%s)", scales::number(mean(ifelse(x > 0, x, NA), na.rm = TRUE), accuracy = 0.1)))},
    family = font_family, size = 2.5
    ) +
  geom_hline(yintercept = median_price, col = "grey30", lty = "dashed") +
  annotate("text", x = 16, y = median_price + 50, label = "Median rent",
           family = "Fira Sans", size = 3, hjust = 0) +
  scale_x_discrete(labels = toupper) +
  scale_y_continuous(breaks = seq(2500, 20000, 2500)) +
  scale_color_manual(values = MetBrewer::met.brewer("VanGogh3")) +
  coord_flip(ylim = c(0, 10000), clip = "off") +
  guides(col = "none") +
  labs(
    title = toupper("Nice and Clean - Relatively Low Rent?"),
    subtitle = plot_subtitle,
    caption = "Axis capped at 10,000 USD.<br>
    Data: Pennington, Kate (2018). 
    Bay Area Craigslist Rental Housing Posts, 2000-2018.<br>
    Retrieved from github.com/katepennington/historic_bay_area_craigslist_housing_posts/blob/master/clean_2000_2018.csv.zip.
    <br>
    Visualization: Ansgar Wolsing",
    x = NULL,
    y = "Rent in USD"
  ) +
  theme_minimal(base_family = font_family) +
  theme(
    plot.background = element_rect(color = NA, fill = bg_color),
    panel.grid = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.1, color = "grey75"),
    plot.title = element_text(family = "Fira Sans SemiBold"),
    plot.title.position = "plot",
    plot.subtitle = element_textbox_simple(
      margin = margin(t = 4, b = 16), size = 10),
    plot.caption = element_textbox_simple(
      margin = margin(t = 12), size = 7
    ),
    plot.caption.position = "plot",
    axis.text.y = element_text(hjust = 0, margin = margin(r = -10), family = "Fira Sans SemiBold"),
    plot.margin = margin(4, 4, 4, 4)
  )

# create the dataframe for the legend (inside plot)
df_for_legend <- rent_title_words %>% 
  filter(word == "beautiful")

p_legend <- df_for_legend %>% 
  ggplot(aes(word, price)) +
  stat_halfeye(fill_type = "segments", alpha = 0.3) +
  stat_interval() +
  stat_summary(geom = "point", fun = median) +
  annotate(
    "richtext",
    x = c(0.8, 0.8, 0.8, 1.4, 1.8),
    y = c(1000, 5000, 3000, 2400, 4000),
    label = c("50 % of prices<br>fall within this range", "95 % of prices", 
              "80 % of prices", "Median", "Distribution<br>of prices"),
    fill = NA, label.size = 0, family = font_family, size = 2, vjust = 1,
  ) +
  geom_curve(
    data = data.frame(
      x = c(0.7, 0.80, 0.80, 1.225, 1.8),
      xend = c(0.95, 0.95, 0.95, 1.075, 1.8), 
      y = c(1800, 5000, 3000, 2300, 3800),
      yend = c(1800, 5000, 3000, 2100, 2500)),
    aes(x = x, xend = xend, y = y, yend = yend),
    stat = "unique", curvature = 0.2, size = 0.2, color = "grey12",
    arrow = arrow(angle = 20, length = unit(1, "mm"))
  ) +
  scale_color_manual(values = MetBrewer::met.brewer("VanGogh3")) +
  coord_flip(xlim = c(0.75, 1.3), ylim = c(0, 6000), expand = TRUE) +
  guides(color = "none") +
  labs(title = "Legend") +
  theme_void(base_family = font_family) +
  theme(plot.title = element_text(family = "Fira Sans SemiBold", size = 9,
                                  hjust = 0.075),
        plot.background = element_rect(color = "grey30", size = 0.2, fill = bg_color))

# Insert the custom legend into the plot
p + inset_element(p_legend, l = 0.6, r = 1.0,  t = 0.99, b = 0.7, clip = FALSE)
```

<br/><br/>

## Going further
***

This post explains how to create a [ridgeline plot](https://r-graph-gallery.com/ridgeline-plot.html) with an **inside plot for the legend** and **annotations** in R with [ggplot2](https://r-graph-gallery.com/ggplot2-package.html).

You might be interested by the [ggridges](https://r-graph-gallery.com/ridgeline-plot.html) package, that is not used here but that can be **very useful** for this kind of chart, and the [ggstatsplot](https://r-graph-gallery.com/web-violinplot-with-ggstatsplot.html) package that can help you to **annotate** your chart while plotting distributions.


<!-- Close container -->
</div>

```{r, echo=FALSE}
# Correlation | Ranking | Evolution.. 
htmltools::includeHTML("htmlChunkRelatedDistribution.html")
```