---
title: "Basic dice plot with ggdiceplot"
descriptionMeta: "A dice plot is a visualization where each tile acts as a die face, with dot positions representing different categorical variables. This post shows how to create a basic dice plot in R using the ggdiceplot package."
descriptionTop: "A dice plot is a visualization where each tile acts as a die face, with dot positions representing different **categorical variables**. This post shows how to create a basic dice plot in R using the [ggdiceplot](package/ggdiceplot.html) package."
sectionText: "Heatmap"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "web-basic-dice-plot-with-ggdiceplot"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container">


# About
***

This page showcases the [ggdiceplot](package/ggdiceplot.html) package, a [ggplot2](ggplot2-package.html) extension for creating dice-based categorical visualizations. You can find the source code on [GitHub](https://github.com/maflot/ggdiceplot).

If you use this package, please cite: Flotho, Flotho & Keller, "DicePlot: a package for high-dimensional categorical data visualization," *Bioinformatics*, 2026.


# Packages
***

For this post, we need to install and load the [ggdiceplot](package/ggdiceplot.html) package.

We can install it from CRAN using `install.packages("ggdiceplot")`. Then, we can load it:

```{r}
# install.packages("ggdiceplot")
library(ggdiceplot)
library(ggplot2)
```

<br/>


# Dataset
***

We use the bundled `sample_dice_data2` dataset, which contains simulated log2 fold-change and q-values for **oral microbial taxa** across disease conditions and specimen sites.

```{r}
data("sample_dice_data2", package = "ggdiceplot")
toy_data <- sample_dice_data2
head(toy_data)
```

The data is in **long format**: one row per combination of taxon, disease, and specimen. The `dots` aesthetic will map to `disease`, meaning each pip position on a tile represents a specific disease condition.

<br/>


# Basic dice plot
***

<div class = "row">

<div class = "col-md-6  col-sm-12 align-self-center">

The simplest dice plot shows categorical **presence/absence**. Each tile is a die face; each pip position corresponds to one level of the `dots` variable (here: disease).

Key parameters:

- `ndots`: number of pip positions (matches the number of disease levels)
- `x_length` / `y_length`: grid dimensions for the 1:1 aspect ratio
- `pip_scale`: pip diameter as fraction of available space

</div>


<div class = "col-md-6  col-sm-12">

```{r thecode, echo=FALSE, out.width = "100%", fig.height=7}
ggplot(toy_data, aes(x = specimen, y = taxon)) +
  geom_dice(
    aes(dots = disease, width = 0.9, height = 0.9),
    na.rm       = TRUE,
    show.legend = TRUE,
    pip_scale   = 0.9,
    ndots       = length(unique(toy_data$disease)),
    x_length    = length(unique(toy_data$specimen)),
    y_length    = length(unique(toy_data$taxon))
  ) +
  theme_dice() +
  labs(x = "Specimen", y = "Taxon", title = "Disease presence across taxa and specimens")
```
</div>
</div>


```{r thecode, eval=FALSE}
```


<br/>

# Adding color: fill mapped to log2 fold-change
***

Map `fill` to a continuous variable like `lfc` (log2 fold-change) to encode **effect size** via a diverging color scale. Missing values appear as white pips.

```{r, fig.height=7}
lo  <- floor(min(toy_data$lfc, na.rm = TRUE))
up  <- ceiling(max(toy_data$lfc, na.rm = TRUE))
mid <- (lo + up) / 2

ggplot(toy_data, aes(x = specimen, y = taxon)) +
  geom_dice(
    aes(dots = disease, fill = lfc, width = 0.9, height = 0.9),
    na.rm       = TRUE,
    show.legend = TRUE,
    pip_scale   = 0.9,
    ndots       = length(unique(toy_data$disease)),
    x_length    = length(unique(toy_data$specimen)),
    y_length    = length(unique(toy_data$taxon))
  ) +
  scale_fill_gradient2(
    low = "#40004B", high = "#00441B", mid = "white",
    na.value = "white", limit = c(lo, up), midpoint = mid,
    name = "Log2FC"
  ) +
  theme_dice() +
  labs(x = "Specimen", y = "Taxon")
```

# Adding size: significance encoding
***

Map `size` to `-log10(q)` to make more significant pips larger. When combined with fill, this creates a **domino plot** that encodes both effect size and significance.

```{r, fig.height=7}
minsize <- floor(min(-log10(toy_data$q), na.rm = TRUE))
maxsize <- ceiling(max(-log10(toy_data$q), na.rm = TRUE))
midsize <- ceiling(quantile(-log10(toy_data$q), 0.5, na.rm = TRUE))

ggplot(toy_data, aes(x = specimen, y = taxon)) +
  geom_dice(
    aes(dots = disease, fill = lfc, size = -log10(q), width = 0.9, height = 0.9),
    na.rm       = TRUE,
    show.legend = TRUE,
    pip_scale   = 0.9,
    ndots       = length(unique(toy_data$disease)),
    x_length    = length(unique(toy_data$specimen)),
    y_length    = length(unique(toy_data$taxon))
  ) +
  scale_fill_gradient2(
    low = "#40004B", high = "#00441B", mid = "white",
    na.value = "white", limit = c(lo, up), midpoint = mid,
    name = "Log2FC"
  ) +
  scale_size_continuous(
    range  = c(2, 8),
    limits = c(minsize, maxsize),
    breaks = c(minsize, midsize, maxsize),
    labels = c(10^minsize, 10^-midsize, 10^-maxsize),
    name   = "q-value"
  ) +
  theme_minimal() +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y     = element_text(size = 12),
    legend.key       = element_blank(),
    legend.key.size  = unit(0.8, "cm")
  ) +
  labs(x = "Specimen", y = "Taxon")
```


<!-- Close container -->
</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
