---
title: "Interactive table with images and charts"
descriptionMeta: "Interactive table made in R with reactable that has images and charts in it."
descriptionTop: "Interactive table made in `R` with [reactable](https://r-graph-gallery.com/package/reactable.html) that has **images** and **charts** in it.<br> This blogpost guides you through the construction of this table made by [Tanya Shapiro](https://github.com/tashapiro)."
sectionText: "Table section"
sectionLink: "table.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com.html"
url: "web-interactive-table-with-images-charts-and-more"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
editor_options: 
  chunk_output_type: console
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE, 
  message=FALSE
)
```

<div class="container" style="padding-top: 100px">

## About
***

This page showcases the work of [Tanya Shapiro](https://github.com/tashapiro). Thanks to her for agreeing to share her work here!

The table uses the [reactable](https://r-graph-gallery.com/package/reactable.html) package to create an **interactive table** on the best female tennis player and their titles. It uses Wikipedia data and has the following features on each row:

- clickable columns

- barplot

- photo of the player

- some statistics

- footer and header


## Load packages
***

Let's start by loading the packages needed to build the table:

```{r}
library(tidyverse)
library(reactablefmtr)
library(reactable)
library(htmltools)
library(sysfonts)
library(showtext)
library(webshot2)
```

<br>

## Load and prepare the dataset
***

Today's data can be easily accessed on the github of the gallery by using the following code:

```{r data}
# REMOVE WHEN MERGING
# url = 'https://raw.githubusercontent.com/holtzy/R-graph-gallery/master/DATA/tennis-best-players.csv'
# df = readRDS(url)
df = read.csv2("DATA/tennis-best-players.csv", sep=',')
```

<br>

## Create the table
***

Now let's create the table:

```{r }
# create custom color palette for scale fill
pal_scale <- c("#F4FFFD", "#E9DAEC", "#A270E5", "#43009A")

table <- reactable(df %>% select(rank, player, region, australian_open, french_open, us_open, wimbledon, titles),
  theme = reactableTheme(
    style = list(fontFamily = "sans-serif"),
    borderColor = "#DADADA"
  ),
  defaultPageSize = 11,
  defaultColDef = colDef(
    vAlign = "center",
    align = "center",
    headerVAlign = "center",
    style = color_scales(df, span = 4:7, colors = pal_scale),
    headerStyle = list(fontFamily = "sans-serif"),
    width = 90
  ),
  columnGroups = list(
    colGroup(name = "", columns = c("player", "region", "titles"), headerStyle = list(fontFamily = "sans-serif"), align = "left"),
    colGroup(name = "Event", columns = c("australian_open", "us_open", "french_open", "wimbledon"), headerStyle = list(fontFamily = "Roboto"))
  ),
  columns = list(
    rank = colDef(show = FALSE),
    player = colDef(
      name = "Player (First Title - Last Title)",
      align = "left", width = 250,
      cell = function(value) {
        # image <- img(src = paste0("https://raw.githubusercontent.com/holtzy/R-graph-gallery/main/img/fromTheWeb/",str_replace_all(tolower(value)," ","_"),".png"), style = "height: 33px;", alt = value)
        image <- img(src = paste0("https://raw.githubusercontent.com/tashapiro/tanya-data-viz/main/tennis/images/", str_replace_all(tolower(value), " ", "_"), ".png"), style = "height: 33px;", alt = value)
        tagList(
          div(style = "display: inline-block;vertical-align:middle;width:50px", image),
          div(
            style = "display: inline-block;vertical-align:middle;",
            div(style = "vertical-align:middle;", value),
            div(style = "vertical-align:middle;font-size:8pt;color:#8C8C8C;", paste0("(", df[df$player == value, ]$years), ")")
          )
        )
      }
    ),
    region = colDef(
      name = "Region",
      align = "left",
      cell = function(value, index) {
        image <- img(src = value, style = "width:60px;height:20px;", alt = value)
        player <- df$player[index]
        if (player %in% c("Monica Seles", "Molla Bjurstedt Mallory")) {
          tagList(div(style = "display:inline-block;vertical-align:middle;width:80px", image, "*"))
        } else {
          tagList(div(style = "display:inline-block;vertical-align:middle;width:50px", image))
        }
      },
      width = 120
    ),
    australian_open = colDef(name = "AU Open"),
    french_open = colDef(name = "FR Open"),
    us_open = colDef(name = "US Open"),
    wimbledon = colDef(name = "Wmbl"),
    titles = colDef(
      name = "Total Titles",
      width = 180,
      class = "border-left",
      align = "left",
      cell = data_bars(df,
        fill_color = "#7814ff",
        text_position = "outside-end",
        bar_height = 10,
        text_size = 12,
        min_value = 5,
        max_value = 32,
        background = "transparent"
      )
    )
  )
)



# add title, subtitle, footnote and source
# note, I downloaded fonts locally - sans-serif & Font Awesome Branded Icons
table_final <- table %>%
  # title & subtitle
  htmlwidgets::prependContent(
    tagList( # get tennis logo for the headr
      tags$img(src = "https://pngimg.com/uploads/tennis/tennis_PNG10416.png", style = "width:50px;height:34px;display:inline-block;vertical-align:middle;"),
      # tags$h1("trophy  ",style="font-family:'Font Awesome 6 Free';margin-bottom:0;display:inline-block;vertical-align:middle;padding-right:10px;"),
      tags$div("Grand Slam Legends", style = "font-size:32px;font-weight:bold;font-family:sans-serif;margin-bottom:0;display:inline-block;vertical-align:middle;"),
      tags$h3("Top Women's Tennis Players by Singles Championship Titles", style = "font-family:sans-serif;margin-bottom:0;margin-top:0;font-weight:400;color:#8C8C8C;padding-left:10px;")
    )
  ) %>%
  # footnote and source
  htmlwidgets::appendContent(
    tags$div("* Player represented more than one country during career. Most recent country shown.", style = "font-family:Roboto;color:black;font-size:9pt;border-bottom-style:solid;border-top-style:solid;width:910px;padding-bottom:8px;padding-top:8px;border-color:#DADADA;"),
    tags$div(
      tags$div("Data: Wikipedia as of November 2022 | Graphic: ", style = "display:inline-block;vertical-align:middle;"),
      tags$div("twitter", style = "font-family:'Font Awesome 6 Brands';display:inline-block;vertical-align:middle;"),
      tags$div("tanya_shapiro", style = "display:inline-block;vertical-align:middle;"),
      tags$div("github", style = "font-family:'Font Awesome 6 Brands';display:inline-block;vertical-align:middle;"),
      tags$div("tashapiro", style = "display:inline-block;vertical-align:middle;"),
      style = "font-family:sans-serif;color:#8C8C8C;font-size:10pt;width:910px;padding-top:8px;display:inline-block;vertical-align:middle;"
    )
  )
```


```{r include=FALSE}
library(htmlwidgets)
html = "HtmlWidget/web-interactive-table-with-images-and-charts.html"
saveWidget(table_final, html)
```

Here's what the final result looks like. Try clicking on a column title to change the row order automatically!

<center><iframe src="HtmlWidget/web-interactive-table-with-images-and-charts.html" height="700" width="1000" style="border:none;"></iframe></center>


# Conclusion

This post explains how to reproduce an interactive table built with [reactable](https://r-graph-gallery.com/package/reactable.html) by [Tanya Shapiro](https://github.com/tashapiro). The table has very nice features such as clickable columns, gradient colors and inline charts.

If you want to learn more about `reactable`, check out the [dedicated section](https://r-graph-gallery.com/package/reactable.html). Check also the [table section](https://r-graph-gallery.com/table.html)

<br>

<!-- Close container -->
</div>

```{r, echo=FALSE}
# Correlation | Ranking | Evolution.. 
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```