---
title: "R Consortium ISC Grants: Voronoi Treemap Visualization"
descriptionMeta: "This post demonstrates how to create a Voronoi treemap in R to visualize R Consortium ISC grants using ggplot2, VoronoiPlus, and ggtext."
descriptionTop: "This post demonstrates how to create a Voronoi treemap in R to visualize the distribution of R Consortium ISC grants by project year using [ggplot2](ggplot2-package.html), [VoronoiPlus](package/VoronoiPlus.html), and [ggtext](package/ggtext.html). The chart **highlights the size of funding** for each project."
sectionText: "Voronoi Treemap"
sectionLink: "package/VoronoiPlus.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/treemap.html"
url: "web-voronoi-treemap-r-consortium"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
   warning = FALSE,
   message = FALSE,
   fig.align = "center",
   fig.width = 12,
   fig.height = 7
)
```

<div class="container">

# About
***

This post visualizes the **R Consortium ISC Grants** by creating a Voronoi [treemap](treemap.html). Each polygon in the [treemap](treemap.html) represents a different grant project, with the area of the polygon proportional to the funding amount. The [treemap](treemap.html) highlights the distribution of funds over different years and provides an insightful overview of the projects that received grants from the R Consortium Infrastructure Steering Committee (ISC).

It has been created by [Nicola Rennie](https://nrennie.rbind.io/). Thanks to her for sharing this beautiful chart!

# Libraries
***

To create this visualization, we will use several R packages: `tidyverse` for data manipulation, `showtext` for custom fonts, `ggtext` for enhanced text styling, and `VoronoiPlus` for generating the Voronoi [treemap](treemap.html).

Install VoronoiPlus with `devtools::install_github("AllanCameron/VoronoiPlus")`

```{r}
library(tidyverse)
library(showtext)
library(ggtext)
# library(nrBrand)
library(glue)
library(sf)
library(VoronoiPlus)
```

# Dataset
***

The dataset is part of the TidyTuesday initiative and includes information about grants awarded by the R Consortium ISC. The dataset contains columns for the year the grant was awarded, whether the project was funded, and the project title.

```{r}
tuesdata <- tidytuesdayR::tt_load("2024-02-20")
isc_grants <- tuesdata$isc_grants
```

The data is processed by selecting relevant columns, computing Voronoi polygons based on the grant data, and preparing labels for each year.

```{r}
# Data wrangling
grant_data <- isc_grants |>
   select(year, funded, title)

vor_data <- voronoi_treemap(funded ~ year + title, data = grant_data)
set.seed(1234)
groups <- filter(vor_data, level == 1)
subgroups <- filter(vor_data, level == 2) |>
   group_by(group) |>
   mutate(alpha = runif(1, 0, 0.6)) |>
   ungroup()

year_labels <- groups |>
   select(group, x, y) |>
   st_as_sf(coords = c("x", "y")) |>
   group_by(group) |>
   summarise(geometry = st_combine(geometry)) |>
   st_cast("POLYGON") |>
   st_centroid() %>%
   mutate(
      x = st_coordinates(.)[, 1],
      y = st_coordinates(.)[, 2]
   ) |>
   st_drop_geometry()
```

# Fonts and Colors
***

Custom fonts and colors are loaded to enhance the aesthetics of the plot. Google fonts `Roboto` and `Roboto Slab` are used for text, while a color palette from `rcartocolor` is applied to represent different years.

```{r}
# Load fonts
font_add_google("Roboto", "roboto")
font_add_google("Roboto Slab", "robotoslab")
showtext_auto()

# Define colors and fonts
bg_col <- "#fafafa"
text_col <- "black"
highlight_col <- "#1c68bc"

cols_vec <- rcartocolor::carto_pal(length(unique(grant_data$year)) + 1, "Prism")[1:length(unique(grant_data$year))]
names(cols_vec) <- unique(grant_data$year)

body_font <- "roboto"
title_font <- "robotoslab"
```

# Plotting
***

The final visualization is a Voronoi treemap that uses custom polygons to represent different grant projects, with labels indicating the year.

```{r}
# Define text
social <- nrBrand::social_caption(
   bg_colour = bg_col,
   icon_colour = highlight_col,
   font_colour = text_col,
   font_family = body_font
)
title <- glue("<br><span style='font-size: 48px; font-weight: bold; font-family:{title_font};'>R Consortium ISC Grants</span><br><br>")
st <- glue("<span style='font-family:{body_font};'>The R Consortium Infrastructure Steering Committee (ISC) Grant Program has been awarding grants since 2016. They will accept proposals again between March 1 and April 1, 2024 (and then again in the fall). Each polygon represents a different project, with the size of the area representing the funding amount.</span><br><br>")
cap <- paste0(title, st, "**Data**: R Consortium ISC<br>**Graphic**:", social)

# Plot
ggplot() +
   geom_polygon(
      data = groups,
      mapping = aes(x = x, y = y, group = group, fill = group),
      colour = text_col,
      linewidth = 0.9
   ) +
   geom_polygon(
      data = subgroups,
      mapping = aes(x = x, y = y, group = group, alpha = alpha),
      fill = bg_col,
      colour = text_col,
      linewidth = 0.3
   ) +
   geom_polygon(
      data = groups,
      mapping = aes(x = x, y = y, group = group),
      colour = "transparent",
      fill = bg_col,
      alpha = 0.3,
      linewidth = 0.9
   ) +
   geom_text(
      data = year_labels,
      mapping = aes(x = x, y = y, label = group),
      colour = text_col,
      family = title_font,
      fontface = "bold",
      size = 12
   ) +
   scale_alpha_identity() +
   scale_fill_manual(values = cols_vec) +
   labs(
      caption = cap
   ) +
   coord_equal() +
   theme_void(base_size = 24, base_family = body_font) +
   theme(
      legend.position = "none",
      plot.margin = margin(-16, 0, 0, 0),
      plot.background = element_rect(fill = bg_col, colour = bg_col),
      panel.background = element_rect(fill = text_col, colour = text_col),
      plot.caption = element_textbox_simple(
         colour = text_col,
         hjust = 0,
         halign = 0,
         margin = margin(l = 10, t = 8),
         lineheight = 0.5,
         family = body_font
      )
   )
```

# Going further
***

You might be interested in:

- Learning more about [Voronoi diagrams](voronoi-diagram.html)
- How to create [an interactive Voronoi treemap](interactive-voronoi-treemap.html)
- Explore [other visualization techniques](visualization-techniques.html)

<!-- Close container -->
</div>

```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedPartOfWhole.html")
```
