---
title: "Grouped line chart with small multiple"
descriptionMeta: "This post explains how to build a grouped line chart with small multiples. Each item is the same grouped line chart, with one specific group higlighted each time"
descriptionTop: "This post explains how to build a grouped line chart with small multiples. Each item is the same grouped line chart, with one specific group higlighted each time"
sectionText: "Line chart section"
sectionLink: "lollipop-plot.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com/graph/lollipop.html"
url: "web-grouped-line-chart-small-multiple"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">




# Introduction
***


This plot has been initially published by Abdul Madjid, as a contribution to the [TidyTuesday challenge](https://github.com/AbdoulMa/TidyTuesday/tree/main/2021_w19). 



# Packages
***

The code works pretty much the same way, but it is important to note that the X axis can represent a categorical variable as well. In this case, the lollipop chart is a good replacement of the [barplot](barplot.html).

The [ggtext](https://www.r-graph-gallery.com/custom-fonts-in-R-and-ggplot2.html) package allows to customize fonts on the chart. It notably allows to use specific fonts, like `Lato` in this example. You can read more about its usage [here](https://www.r-graph-gallery.com/custom-fonts-in-R-and-ggplot2.html).

```{r thecode2, echo=TRUE, out.width = "100%", fig.height=7}
# Libraries that we need
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(geojsonio)
library(sf)
library(broom)
library(cowplot)
library(hexbin)
library(showtext)

# Loading fonts
font_add_google("DM Serif Display", "abril")
font_add_google("Tajawal", "tawa")
showtext_auto()
```


# Dataset
***

Description of the dataset

```{r}
data <- read.csv("https://raw.githubusercontent.com/lau-cloud/30DayMapChallenge/main/parking_areas/TRAMS.csv", sep = ",", header = T)

# Loading Barcelona's district boundaries
districts <- st_read("https://raw.githubusercontent.com/lau-cloud/30DayMapChallenge/main/parking_areas/0301040100_Districtes_UNITATS_ADM.json",
  stringsAsFactors = FALSE,
  as_tibble = TRUE
)

# Filtering types
trams_verda <- c("VR", "VM")

tram_AZL <- data |>
  filter(TIPUS_TRAM == "AZL")

tram_VR <- data |>
  filter(TIPUS_TRAM %in% trams_verda)

tram_DUM <- data |>
  filter(TIPUS_TRAM == "DUM")
```

# Ridgeline with female names
***

```{r}
mapa_AZL <- ggplot() +
  geom_hex(
    data = tram_AZL, aes(x = LONGITUD_I, y = LATITUD_I),
    color = "white", alpha = 0.8, bins = 35
  ) +
  geom_sf(
    data = districts, fill = "transparent",
    color = "black", linewidth = .25
  ) +
  labs(
    title = "Blue Zone",
    caption = ""
  ) +
  theme_void() +
  scale_fill_gradient(
    low = "#D1F0E5",
    high = "#306D75",
    breaks = c(1, 20, 40),
    name = "Blue spaces",
    guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth = unit(10, units = "mm"), label.position = "bottom", title.position = "top", nrow = 1)
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(color = "black", size = 8),
    text = element_text(color = "black"),
    plot.subtitle = element_text(hjust = 0.5, size = 8, color = "black"),
    plot.title = element_text(hjust = 0.5, size = 16, family = "tawa"),
    plot.caption = element_text(hjust = 1, size = 7, color = "black")
  )

mapa_AZL
```

# Ridgeline with male names
***

```{r}
mapa_VR <- ggplot() +
  geom_hex(
    data = tram_VR, aes(x = LONGITUD_I, y = LATITUD_I),
    color = "white", alpha = 0.8, bins = 35
  ) +
  geom_sf(
    data = districts, fill = "transparent",
    color = "black", linewidth = .25
  ) +
  labs(
    title = "Green Zone",
    caption = ""
  ) +
  theme_void() +
  scale_fill_gradient(
    low = "#E1FCC1",
    high = "#5DAA01",
    breaks = c(1, 60, 120),
    name = "Green spaces",
    guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth = unit(10, units = "mm"), label.position = "bottom", title.position = "top", nrow = 1)
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(color = "black", size = 8),
    text = element_text(color = "black"),
    plot.subtitle = element_text(hjust = 0.5, size = 8, color = "black"),
    plot.title = element_text(hjust = 0.5, size = 16, family = "tawa"),
    plot.caption = element_text(hjust = 0.5, size = 7, color = "black")
  )


mapa_VR
```


# thrid
***

```{r}
mapa_DUM <- ggplot() +
  geom_hex(
    data = tram_DUM, aes(x = LONGITUD_I, y = LATITUD_I),
    color = "white", alpha = 0.8, bins = 35
  ) +
  geom_sf(
    data = districts, fill = "transparent",
    color = "black", linewidth = .25
  ) +
  labs(
    title = "Urban Distribution of Goods Zone"
  ) +
  theme_void() +
  scale_fill_gradient(
    low = "#FFD2D2",
    high = "#CD0000",
    breaks = c(1, 12, 25),
    name = "DUM spaces",
    guide = guide_legend(keyheight = unit(2.5, units = "mm"), keywidth = unit(10, units = "mm"), label.position = "bottom", title.position = "top", nrow = 1)
  ) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(color = "black", size = 8),
    text = element_text(color = "black"),
    plot.subtitle = element_text(hjust = 0.5, size = 8, color = "black"),
    plot.title = element_text(hjust = 0.5, size = 16, family = "tawa"),
    plot.caption = element_text(hjust = 1, size = 7, color = "black")
  )

mapa_DUM
```

# Combine plots
***

```{r}
# Title
ggdraw() +
  draw_text("Parking Areas in Barcelona",
    size = 24, family = "abril", fontface = "bold"
  ) -> header_plot

# Caption
ggdraw() +
  draw_text("Data: OpenData BCN | Visualization: Laura Navarro",
    size = 16, family = "tawa", color = "grey", hjust = 0.5
  ) -> caption_plot

# Grid plots
grid_plots <- plot_grid(
  mapa_AZL,
  mapa_VR,
  mapa_DUM,
  nrow = 1,
  ncol = 3
)

# Final plot
plot_grid(

  header_plot,
  grid_plots,
  caption_plot,

  ## plot settings
  nrow = 3,
  ncol = 1,
  rel_heights = c(1, 5, 1)
)
```

<!-- Close container -->
</div>




```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedRanking.html")
```
