---
title: "Domino plot for gene expression with ggdiceplot"
descriptionMeta: "A domino plot combines fill color and dot size to encode both effect size and statistical significance in a single dice-based visualization. Built with R and the ggdiceplot package."
descriptionTop: "A **domino plot** combines fill color and dot size to encode both effect size and statistical significance in a single dice-based visualization. Built with R and the [ggdiceplot](package/ggdiceplot.html) package."
sectionText: "Heatmap"
sectionLink: "heatmap.html"
DataToVizText: "Data to Viz"
DataToVizLink: "data-to-viz.com"
url: "web-domino-plot-with-ggdiceplot"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_rgg.html
      css: style.css
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')
```

<div class="container" style="padding-top: 100px">


# About
***

This page demonstrates a **domino plot** — a dice plot where both **fill** and **size** aesthetics are mapped to data variables. This encodes two quantitative dimensions (e.g., effect size and significance) on top of the categorical structure of the dice.

The [ggdiceplot](package/ggdiceplot.html) package is available on [CRAN](https://CRAN.R-project.org/package=ggdiceplot) and [GitHub](https://github.com/maflot/ggdiceplot).

If you use this package, please cite: Flotho, Flotho & Keller, "DicePlot: a package for high-dimensional categorical data visualization," *Bioinformatics*, 2026.

As a teaser, here is the plot we're going to build:

<br>

```{r teaser, echo=FALSE, out.width = "80%", fig.height=9}
library(ggplot2)
library(ggdiceplot)
library(dplyr)
library(tidyr)

# Create a gene-expression-style dataset
set.seed(42)
genes      <- c("SPP1", "APOE", "SERPINA1", "PINK1", "ANGPT1", "ANGPT2", "APP", "CLU", "ABCA7")
cell_types <- c("Astrocyte", "Microglia", "Neuron", "Oligodendrocyte", "OPC", "Endothelial")
contrasts  <- c("AD-CT", "MS-CT", "FTD-CT", "ASD-CT", "HD-CT")

zebra_sim <- expand.grid(
  gene      = genes,
  cell_type = cell_types,
  contrast  = contrasts,
  stringsAsFactors = FALSE
) %>%
  mutate(
    logFC = rnorm(n(), mean = 0, sd = 1.5),
    FDR   = runif(n(), min = 1e-6, max = 0.2)
  ) %>%
  # Introduce some sparsity — not all genes significant in all combinations
  mutate(
    logFC = ifelse(runif(n()) < 0.3, NA_real_, logFC),
    FDR   = ifelse(is.na(logFC), NA_real_, FDR)
  ) %>%
  mutate(
    cell_type = factor(cell_type, levels = sort(cell_types)),
    contrast  = factor(contrast, levels = contrasts),
    gene      = factor(gene, levels = genes)
  )

lo  <- floor(min(zebra_sim$logFC, na.rm = TRUE))
up  <- ceiling(max(zebra_sim$logFC, na.rm = TRUE))
mid <- (lo + up) / 2
minsize <- floor(min(-log10(zebra_sim$FDR), na.rm = TRUE))
maxsize <- ceiling(max(-log10(zebra_sim$FDR), na.rm = TRUE))
midsize <- ceiling(quantile(-log10(zebra_sim$FDR), 0.5, na.rm = TRUE))

ggplot(zebra_sim, aes(x = gene, y = cell_type)) +
  geom_dice(
    aes(dots = contrast, fill = logFC, size = -log10(FDR)),
    na.rm       = TRUE,
    show.legend = TRUE,
    ndots       = length(contrasts),
    x_length    = length(genes),
    y_length    = length(cell_types)
  ) +
  scale_fill_gradient2(
    low = "#40004B", high = "#00441B", mid = "white",
    na.value = "white", limit = c(lo, up), midpoint = mid,
    name = "Log2FC"
  ) +
  scale_size_continuous(
    limits = c(minsize, maxsize),
    breaks = c(minsize, midsize, maxsize),
    labels = c(10^minsize, 10^-midsize, 10^-maxsize),
    name   = "FDR"
  ) +
  theme_minimal() +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y     = element_text(size = 12),
    legend.text     = element_text(size = 12),
    legend.title    = element_text(size = 12),
    legend.key       = element_blank(),
    legend.key.size  = unit(0.8, "cm")
  ) +
  labs(x = "Gene", y = "Cell Type", title = "ZEBRA-style Domino Plot: DEGs across diseases and cell types")
```

<br>


# Packages
***

```{r}
# install.packages("ggdiceplot")
library(ggdiceplot)
library(ggplot2)
library(dplyr)
library(tidyr)
```

<br/>


# Dataset
***

A domino plot works best with data that has **three categorical axes** (x, y, and dots) plus one or two **quantitative variables** (fill and/or size).

Here we simulate a gene expression dataset inspired by the [ZEBRA](https://doi.org/10.1093/nar/gkac010) cross-disease analysis, with genes on the x-axis, cell types on the y-axis, and disease contrasts as pip positions:

```{r}
set.seed(42)
genes      <- c("SPP1", "APOE", "SERPINA1", "PINK1", "ANGPT1", "ANGPT2", "APP", "CLU", "ABCA7")
cell_types <- c("Astrocyte", "Microglia", "Neuron", "Oligodendrocyte", "OPC", "Endothelial")
contrasts  <- c("AD-CT", "MS-CT", "FTD-CT", "ASD-CT", "HD-CT")

zebra_sim <- expand.grid(
  gene      = genes,
  cell_type = cell_types,
  contrast  = contrasts,
  stringsAsFactors = FALSE
) %>%
  mutate(
    logFC = rnorm(n(), mean = 0, sd = 1.5),
    FDR   = runif(n(), min = 1e-6, max = 0.2)
  ) %>%
  # Introduce sparsity: not all genes are significant in all conditions
  mutate(
    logFC = ifelse(runif(n()) < 0.3, NA_real_, logFC),
    FDR   = ifelse(is.na(logFC), NA_real_, FDR)
  ) %>%
  mutate(
    cell_type = factor(cell_type, levels = sort(cell_types)),
    contrast  = factor(contrast, levels = contrasts),
    gene      = factor(gene, levels = genes)
  )

head(zebra_sim)
```

<br/>

# Building the domino plot step by step
***

## Step 1: Basic structure with fill only

Start by mapping `fill` to `logFC` with a diverging purple–green gradient. Pips with `NA` values appear white (absent from the die face).

```{r, fig.height=9}
lo  <- floor(min(zebra_sim$logFC, na.rm = TRUE))
up  <- ceiling(max(zebra_sim$logFC, na.rm = TRUE))
mid <- (lo + up) / 2

p <- ggplot(zebra_sim, aes(x = gene, y = cell_type)) +
  geom_dice(
    aes(dots = contrast, fill = logFC),
    na.rm       = TRUE,
    show.legend = TRUE,
    ndots       = length(contrasts),
    x_length    = length(genes),
    y_length    = length(cell_types)
  ) +
  scale_fill_gradient2(
    low = "#40004B", high = "#00441B", mid = "white",
    na.value = "white", limit = c(lo, up), midpoint = mid,
    name = "Log2FC"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12)) +
  labs(x = "Gene", y = "Cell Type")

p
```

<br>

## Step 2: Add size for significance

Now map `size` to `-log10(FDR)`. Larger pips indicate stronger statistical significance. The combination of color (effect direction/magnitude) and size (confidence) creates the full domino plot.

```{r, fig.height=9}
minsize <- floor(min(-log10(zebra_sim$FDR), na.rm = TRUE))
maxsize <- ceiling(max(-log10(zebra_sim$FDR), na.rm = TRUE))
midsize <- ceiling(quantile(-log10(zebra_sim$FDR), 0.5, na.rm = TRUE))

p <- ggplot(zebra_sim, aes(x = gene, y = cell_type)) +
  geom_dice(
    aes(dots = contrast, fill = logFC, size = -log10(FDR)),
    na.rm       = TRUE,
    show.legend = TRUE,
    ndots       = length(contrasts),
    x_length    = length(genes),
    y_length    = length(cell_types)
  ) +
  scale_fill_gradient2(
    low = "#40004B", high = "#00441B", mid = "white",
    na.value = "white", limit = c(lo, up), midpoint = mid,
    name = "Log2FC"
  ) +
  scale_size_continuous(
    limits = c(minsize, maxsize),
    breaks = c(minsize, midsize, maxsize),
    labels = c(10^minsize, 10^-midsize, 10^-maxsize),
    name   = "FDR"
  )

p
```

<br>

## Step 3: Polish the theme

Fine-tune text sizes, legend appearance, and add a descriptive title.

```{r, fig.height=9}
p +
  theme_minimal() +
  theme(
    axis.text.x     = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y     = element_text(size = 12),
    legend.text     = element_text(size = 12),
    legend.title    = element_text(size = 12),
    legend.key       = element_blank(),
    legend.key.size  = unit(0.8, "cm")
  ) +
  labs(
    x     = "Gene",
    y     = "Cell Type",
    title = "ZEBRA-style Domino Plot: DEGs across diseases and cell types"
  )
```


# Preparing your own data
***

To create a domino plot with your own data, ensure the input is in **long format** with one row per dot:

```{r, eval=FALSE}
# Example: reshape wide gene expression results into long format
library(dplyr)
library(tidyr)

my_data <- my_results %>%
  # One row per gene x cell_type x contrast
  filter(PValue < 0.05) %>%
  group_by(gene, cell_type, contrast) %>%
  summarise(logFC = mean(logFC), FDR = min(FDR), .groups = "drop") %>%
  # Fill in missing combinations with NA
  complete(gene, cell_type, contrast,
           fill = list(logFC = NA_real_, FDR = NA_real_))
```

The key columns are:

- **x** and **y**: grid position variables (e.g., gene, cell_type)
- **dots**: categorical variable for pip positions (e.g., contrast/disease)
- **fill**: quantitative or categorical variable for pip color (e.g., logFC)
- **size** (optional): quantitative variable for pip diameter (e.g., -log10(FDR))


<!-- Close container -->
</div>



```{r, echo=FALSE}
htmltools::includeHTML("htmlChunkRelatedCorrelation.html")
```
