---
title: "Create Interactive Chord Diagrams with **chorddiag**"
logo: "chorddiag"
descriptionMeta: "This post explains how to create interactive chord diagrams using the chorddiag package in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `chorddiag` package in R is designed to create **interactive chord diagrams** using the D3 JavaScript library.<br/>This post showcases the **key features** of `chorddiag` and provides a set of **graph examples** using the package."
documentationLink: "github.com/mattflor/chorddiag"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `chorddiag` package in R is designed to create **interactive chord diagrams** using the D3 JavaScript library.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of **functions** that make it easy to create and customize chord diagrams for visualizing relationships between entities.

✍️ **author** &rarr; Matthias Flor

📘 **documentation** &rarr; [github](https://github.com/mattflor/chorddiag)

⭐️ *more than 100 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(chorddiag)
m <- matrix(
   c(
      11975, 5871, 8916, 2868,
      1951, 10048, 2060, 6171,
      8010, 16145, 8090, 8045,
      1013, 990, 940, 6907
   ),
   byrow = TRUE,
   nrow = 4, ncol = 4
)
groupNames <- c("black", "blonde", "brown", "red")
groupColors <- c("#000000", "#FFDD89", "#957244", "#F26223")
chorddiag(m, groupNames = groupNames, groupColors = groupColors)
```
</div>
</div>

# Installation
***
To get started with `chorddiag`, you can install it from GitHub using the `devtools` package:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("mattflor/chorddiag")
```

# Basic usage 
***

The `chorddiag` package allows you to create chord diagrams using the `chorddiag()` function.

Here's a basic example:

```{r}
library(chorddiag)

m <- matrix(
   c(
      11975, 5871, 8916, 2868,
      1951, 10048, 2060, 6171,
      8010, 16145, 8090, 8045,
      1013, 990, 940, 6907
   ),
   byrow = TRUE,
   nrow = 4, ncol = 4
)
groupNames <- c("black", "blonde", "brown", "red")
groupColors <- c("#000000", "#FFDD89", "#957244", "#F26223")
interactive_plot <- chorddiag(m, groupNames = groupNames, groupColors = groupColors)

# save the plot as an HTML file
htmlwidgets::saveWidget(interactive_plot, "../HtmlWidget/chord_diagram-1.html")
```

<center>
   <iframe src="../HtmlWidget/chord_diagram-1.html" alt="chord diagram" height="640" width="100%" style="border:none;"/></iframe>
</center>

# Key features
***

## &rarr; Customizing Colors
You can customize the colors of your chord diagram using the `groupColors` parameter.

Example:

```{r}
interactive_plot <- chorddiag(m, groupNames = groupNames, groupColors = c("blue", "yellow", "green", "red"))

# save the plot as an HTML file
htmlwidgets::saveWidget(interactive_plot, "../HtmlWidget/chord_diagram-2.html")
```

<center>
   <iframe src="../HtmlWidget/chord_diagram-2.html" height="640" width="100%"  alt="chord diagram"  style="border:none;"/></iframe>
</center>

<br>

## &rarr; Adjusting Margins
Use the `margin` parameter to adjust the margins of your chord diagram.

Example:

```{r}
interactive_plot <- chorddiag(m, groupNames = groupNames, margin = 100)

# save the plot as an HTML file
htmlwidgets::saveWidget(interactive_plot, "../HtmlWidget/chord_diagram-3.html")
```

<center>
   <iframe src="../HtmlWidget/chord_diagram-3.html" height="640" width="100%"  alt="chord diagram"  style="border:none;"/></iframe>
</center>

<br>

## &rarr; Customizing Labels
You can customize the labels of your chord diagram using the `groupnamePadding` and `groupnameFontsize` parameters.

Example:

```{r}
interactive_plot <- chorddiag(m, groupNames = groupNames, groupnamePadding = 20, groupnameFontsize = 12)

# save the plot as an HTML file
htmlwidgets::saveWidget(interactive_plot, "../HtmlWidget/chord_diagram-4.html")
```

<center>
   <iframe src="../HtmlWidget/chord_diagram-4.html" height="640" width="100%"  alt="chord diagram"  style="border:none;"/></iframe>
</center>

<br>

## &rarr; Adding Tooltips
`chorddiag` allows you to add tooltips to your chord diagram using the `tooltipGroupConnector` parameter.

Example:

```{r}
interactive_plot <- chorddiag(m, groupNames = groupNames, tooltipGroupConnector = " to ")

# save the plot as an HTML file
htmlwidgets::saveWidget(interactive_plot, "../HtmlWidget/chord_diagram-5.html")
```

<center>
   <iframe src="../HtmlWidget/chord_diagram-5.html" height="640" width="100%"  alt="chord diagram"  style="border:none;"/></iframe>
</center>

<br>

# Gallery of chorddiag examples
***


<!-- Close container -->
</div>