---
title: "Combine Multiple Plots with **cowplot**"
logo: "cowplot"
descriptionMeta: "This post explains how to combine multiple ggplot2 plots into a single layout using the cowplot package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `cowplot` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **combining multiple plots** into a single layout.<br/>This post showcases the **key features** of `cowplot` and provides a set of **graph examples** using the package."
documentationLink: "wilkelab.org/cowplot/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

<br>

<div class = "alert alert-warning" style="padding-top: 30px;">
   **Warning!** The `cowplot` package is very similar to the [**patchwork**](patchwork.html) package. [**Patchwork**](patchwork.html) is easier to use (intuitive syntax) and provides the same main features!
</div>

<br>

# Quick start
***
The `cowplot` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **combining multiple plots** into a single layout.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of **functions** that make it easy to arrange plots in a grid, add labels, and create publication-ready figures.

✍️ **author** &rarr; Claus O. Wilke

📘 **documentation** &rarr; [wilkelab.org/cowplot](https://wilkelab.org/cowplot/)

⭐️ *more than 700 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(cowplot)
p1 <- ggplot(mtcars) +
   geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) +
   geom_boxplot(aes(gear, disp, group = gear))
plot_grid(p1, p2)
```
</div>
</div>

# Installation
***
To get started with `cowplot`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("cowplot")
```

# Basic usage 
***

The `cowplot` package allows you to combine multiple plots using the `plot_grid()` function.

Here's a basic example:

```{r}
library(ggplot2)
library(cowplot)

p1 <- ggplot(mtcars) +
   geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) +
   geom_boxplot(aes(gear, disp, group = gear))
plot_grid(p1, p2)
```

# Key features
***

<br>

## &rarr; Combining Plots
You can combine multiple plots using the `plot_grid()` function.

Example:

```{r}
plot_grid(p1, p2, p1, ncol = 3)
```

<br>

## &rarr; Arranging Plots
Use the `nrow` and `ncol` arguments to control the layout.

Example:

```{r}
plot_grid(p1, p2, nrow = 2)
```

<br>

## &rarr; Adding Labels
You can add labels to your plots using the `labels` argument.

Example:

```{r}
plot_grid(p1, p2, labels = c("A", "B"))
```

<br>

## &rarr; Adjusting Relative Sizes
Control the relative sizes of plots with the `rel_widths` and `rel_heights` arguments.

Example:

```{r}
plot_grid(p1, p2, rel_widths = c(2, 1))
```

<br>

# Gallery of cowplot examples
***

The gallery is filled with examples using the `cowplot` package! Browse them below:

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../web-double-ridgeline-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Double ridgeline plot <hr> <p class=explanation_portfolio>A double ridgeline plot with two groups. </div> </div><img alt="Double ridgeline plot" class="img-fluid imgOfPortfolio" src=../img/graph/web-double-ridgeline-plot.png> </a></div> <div class="portfolio-item col-sm-6 col-md-8"><a class=portfolio-link href=../web-lollipop-plot-with-R-the-office.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Lollipop plot <hr> <p class=explanation_portfolio>A lollipop plot with a dual Y axis. </div> </div><img alt="Lollipop plot with dual Y axis" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-lollipop-plot-with-R-the-office.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-scatterplot-corruption-and-human-development.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Scatterplot <hr> <p class=explanation_portfolio>A scatterplot with individual annotations. </div> </div><img alt="Scatterplot with individual annotations" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-scatterplot-corruption-and-human-development-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"> <div class=portfolio-item><a class=portfolio-link href=../web-streamchart-with-ggstream.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Streamchart <hr> <p class=explanation_portfolio>A streamchart with box annotations. </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-streamchart-with-ggstream-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-10"> <div class=portfolio-item><a class=portfolio-link href=../web-triple-map-into-a-single-chart.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Triple map <hr> <p class=explanation_portfolio>A triple map into a single chart. </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/web-triple-map-into-a-single-chart.png> </a></div> </div> </div> </div> </section>

<!-- Close container -->
</div>