---
title: "Create interactive table with **DT**"
logo: "DT"
descriptionMeta: "This post explains how display R data object (such as dataframe) as HTML pages with filtering, editing and many other features with the `DT` package. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how display R data object (such as dataframe) as `HTML` pages with **filtering**, **editing** and even use **callback javascript functions** with the `DT` package. It provides several reproducible examples with explanation and R code.<br>We showcase the **key features** of `DT` and provides a set of **table examples** using the package."
documentationLink: "github.com/rstudio/DT"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `DT` package is a interface for the `DataTable` JavaScript library in R. It can be used for rendering interactive and dynamic tables in **R Markdown** documents and **Shiny** web applications.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It (mainly) offers a `datatable()` function that creates an **HTML widget** that display a table. This table is interactive and dynamic thanks to the use of `JavaScript`. 

✍️ **author** &rarr; Yihui Xie

📘 **documentation** &rarr; [github](https://github.com/rstudio/DT)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.wiDTh = "85%", fig.align='center'}
library(DT)
data(iris)

datatable(iris, class = 'cell-border stripe',
          caption = 'Table 1: The mtcars dataset is a famous dataset about cars and their properties.',
          filter = 'top', options = list(pageLength = 5, autoWidth = TRUE))

#library(htmlwidgets)
#saveWidget(table, file="../HtmlWidget/dt-table-quick.html")
```

<center><iframe src="../HtmlWidget/dt-table-quick.html" height="500" width="600" style="border:none;"></iframe></center>

</div>
</div>

# Installation
***
To get started with `DT`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("DT")
```

# Basic usage 
***

The `DT` package allows you to combine multiple plots using simple **arithmetic operators** like `+`, `/`, and `|`.

Here's a basic example:

```{r}
library(DT)
data(mtcars)

datatable(mtcars)

# save the widget
# library(htmlwidgets)
# saveWidget(table, file="../HtmlWidget/dt-table-basic.html")
```

<center><iframe src="../HtmlWidget/dt-table-basic.html" height="700" width="700" style="border:none;"></iframe></center>

# Key features
***

<br>

## &rarr; Use CSS classes
In the `datatable()` function, there is a `cell` argument that can take CSS classes. The available ones are: `display`, `cell-border`, `compact`	`hover`, `nowrap`, `order-column`, `row-border`, `stripe`. 

Example using `cell-border`, `stripe`, `hover` and `compact`:

```{r}
library(DT)

datatable(
  mtcars, 
  class = 'cell-border stripe hover compact'
)

# save the widget
# library(htmlwidgets)
# saveWidget(table, file="../HtmlWidget/dt-table-1.html")
```

<center><iframe src="../HtmlWidget/dt-table-1.html" height="700" width="700" style="border:none;"></iframe></center>

<br>

## &rarr; Add caption
Thanks to the `caption` argument and the `htmltools` library, you can add customized caption to the table.

Example:

```{r}
library(DT)
library(htmltools)
data(mtcars)

datatable(
  mtcars,
  caption = tags$caption(
    style = 'caption-side: bottom; text-align: center;',
    'Table 1: ', em('The mtcars dataset is a dataset about cars properties')
  )
)

# save widget
# library(htmltools)
# saveWidget(table, file="../HtmlWidget/dt-table-2.html")
```

<center><iframe src="../HtmlWidget/dt-table-2.html" height="700" width="700" style="border:none;"></iframe></center>

<br>

## &rarr; Filtering
The `datatable()` function has a filtering option. Depending on the type of variable, the filtering might differ:
- **quantitative** columns: range slider
- **qualitative** columns: select from all possible categories

Example:

```{r}
library(DT)
data(mtcars)

datatable(mtcars,
          filter = 'top', # put filters on top of the table
          )

# save widget
# library(htmltools)
# saveWidget(table, file="../HtmlWidget/dt-table-3.html")
```

<center><iframe src="../HtmlWidget/dt-table-3.html" height="700" width="700" style="border:none;"></iframe></center>

<br>

## &rarr; Callback JavaScript functions
You can apply JavaScript functions to your tables with the `callback` argument. The following example starts the table on the second page.

Example:

```{r}
library(DT)
data(mtcars)

datatable(mtcars,
          callback = JS('table.page("next").draw(false);'))

# save widget
# library(htmltools)
# saveWidget(table, file="../HtmlWidget/dt-table-4.html")
```

<center><iframe src="../HtmlWidget/dt-table-4.html" height="700" width="700" style="border:none;"></iframe></center>

<br>

## &rarr; Custom style with colors, urls...
For this example, we'll create a specific dataset for better understanding.

Example:

```{r}
library(DT)

# create dataset
data = matrix(c(
  '<b>Bold</b>', '<em>Emphasize</em>', '<a href="https://r-graph-gallery.com/package/dt.html">Click here</a>',
  '<a href="#" onclick="alert(\'This message is displayed thanks to DT table!\');">Click there</a>'),
  2)
#change columns name
colnames(data) = c('<span style="color:red">Red column</span>', '<em>Italic column</em>')

datatable(data,
          escape = FALSE # this argument allows the content of the cells to be interpreted as HTML
          )

# save widget
# library(htmltools)
# saveWidget(table, file="../HtmlWidget/dt-table-5.html")
```

<center><iframe src="../HtmlWidget/dt-table-5.html" height="700" width="700" style="border:none;"></iframe></center>

# Gallery of DT examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../365-filtering-editing-captioning-in-dt-table.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Filtering, Captioning and Editing<p class=explanation_portfolio>With the DT package you can have editable with filtering feature and a caption</div></div><img alt="Filtering, Captioning and Editing"class="img-fluid imgOfPortfolio"src=../img/graph/365-filtering-editing-captioning-in-dt-table.png></a></div><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../364-add-html-content-in-dt-table.html><div class=portfolio-hover><div class=portfolio-hover-content><p>HTML in table cells<p class=explanation_portfolio>How to add HTML content in the cells of a DT table for better looking tables with links, images and more</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/364-add-html-content-in-dt-table.png></a></div></div></div></div></section>

<!-- Close container -->
</div>
