---
title: "Create beeswarm plot with ggbeeswarm"
logo: "ggbeeswarm"
descriptionMeta: "The ggbeeswarm package in R is an extension of the ggplot2 package, designed to simplify the process of creating beeswarm plots. This post showcases the key features of ggbeeswarm and provides a set of graph examples using the package."
descriptionTop: "The `ggbeeswarm` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of creating **beeswarm plots**. This post showcases the **key features** of `ggbeeswarm` and provides a set of **graph examples** using the package."
documentationLink: "github.com/eclarke/ggbeeswarm"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `ggbeeswarm` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of creating **beeswarm plots**.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers 2 main functions to create beeswarm plots: `geom_quasirandom()` and `geom_beeswarm()`.

✍️ **author** &rarr; Erik Clarke

📘 **documentation** &rarr; [github](https://github.com/eclarke/ggbeeswarm)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(ggbeeswarm)

# Create a simple beeswarm plot
ggplot(iris, aes(x = Species, y = Sepal.Length, colour=Species)) +
  geom_beeswarm() +
  theme_minimal() +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))
```
</div>
</div>

# Installation
***
To get started with `ggbeeswarm`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggbeeswarm")
```

# Basic usage 
***

The `ggbeeswarm` package makes it easy to create beeswarm plots using the `geom_beeswarm()` function.

Here's a basic example:

```{r}
library(ggplot2)
library(ggbeeswarm)

ggplot(iris, aes(x = Species, y = Sepal.Length, colour=Species)) +
  geom_beeswarm()
```

# Key features
***

This section is currently a work in progress! Check back soon for more information.

## &rarr; Customizing the plot

```{r}
# 
```

## &rarr; Adding color

Example:

```{r}
#
```


# Gallery of ggbeeswarm examples
***


<!-- Close container -->
</div>
