---
title: "Create bumbplot with **ggbump**"
logo: "ggbump"
descriptionMeta: "This post explains how to combine multiple ggplot2 plots into a single layout using the ggbump package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggbump` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **combining multiple plots** into a single layout.<br/>This post showcases the **key features** of `ggbump` and provides a set of **graph examples** using the package."
documentationLink: "github.com/davidsjoberg/ggbump"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  fig.align='center',
  fig.width=6,
  fig.height=4
)
```

<div class="container">

# Quick start
***
The `ggbump` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to create **bump plots**, which is a variant of the [parallel coordinates plot](../parallel-plot.html)

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a `geom_bump()` function that creates bump plots with [ggplot2](../ggplot2-package.html). The package is particularly useful for visualizing **ranked data** and **ranking changes** over time.

✍️ **author** &rarr; David Sjoberg

📘 **documentation** &rarr; [github](https://github.com/davidsjoberg/ggbump/)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(tidyverse)
library(ggbump)

year <- rep(2019:2021, 4)
revenue <- c(
  100, 200, 300,
  150, 250, 100,
  200, 300, 400,
  300, 200, 100
)

company <- c(
  "Company A", "Company A", "Company A",
  "Company B", "Company B", "Company B",
  "Company C", "Company C", "Company C",
  "Company D", "Company D", "Company D"
)

df <- data.frame(
  year = year,
  revenue = revenue,
  company = company
)

ggplot(df, aes(x = year, y = revenue, color = company)) +
  geom_bump() +
  theme_minimal() +
  labs(
    title = "Revenue by Company",
    subtitle = "2019-2021"
  )
```
</div>
</div>

# Installation
***
To get started with `ggbump`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggbump")
```

# Basic usage 
***

We start by creating a simple dataset with three columns: `year`, `revenue`, and `company`.

```{r}
year <- rep(2019:2021, 3)
revenue <- c(
  100, 200, 300,
  150, 250, 100,
  200, 300, 400
)
company <- c(
  "Company A", "Company A", "Company A",
  "Company B", "Company B", "Company B",
  "Company C", "Company C", "Company C"
)
df <- data.frame(
  year = year,
  revenue = revenue,
  company = company
)
```

Next, we create a `ggplot` object using the `df` dataset and the `geom_bump` function from `ggbump`.

```{r}
ggplot(df, aes(x = year, y = revenue, color = company)) +
  geom_bump()
```

# Key features
***

## &rarr; Individual points

You can easily combine `geom_bump()` with `geom_point()` to display individual points on the plot.

Example:

```{r}
ggplot(df, aes(x = year, y = revenue, color = company)) +
  geom_bump(size=2) +
  geom_point(size=5)
```

<br>

## &rarr; Custom style

`ggbump` allows you to **customize the appearance** of your bump plot using various arguments and ggplot2 themes.

Example:

```{r}
ggplot(df, aes(x = year, y = revenue, color = company)) +
  geom_bump(size=1.5) +
  geom_point(size=4) +
  scale_color_brewer(palette = "Accent") +
  theme_minimal()
```



# Gallery of ggbump examples
***

The examples from the gallery will help you understand the capabilities of the `ggbump` package and **how to use it** in concrete cases.

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../408-basic-bump-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Most basic bump plot <hr> <p class=explanation_portfolio>How to build a simple bump plot with ggplot2. </div> </div><img alt="Dumbell chart with gap column on the right" class="img-fluid imgOfPortfolio" src=../img/graph/408-basic-bump-plot-1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../409-customize-bump-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Customize bump plot <hr> <p class=explanation_portfolio>Learn how to change colors and add individual points to a bump plot. </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/409-customize-bump-plot-1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../409-customize-bump-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Labels and annotations <hr> <p class=explanation_portfolio>Discover how to add labels and annotations to a bump plot. </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/409-customize-bump-plot-2.png> </a></div> </div> </div> </div> </section>

<!-- Close container -->
</div>
