---
title: "Highlight specific elements in your charts with **gghighlight**"
logo: "gghighlight"
descriptionMeta: "This post explains how to showcase specific lines, bars or points in you ggplot2 plots using the gghighlight package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `gghighlight` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **highlighting** elements.<br/>This post showcases the **key features** of `gghighlight` and provides a set of **graph examples** using the package."
documentationLink: "yutannihilation.github.io/gghighlight/articles/gghighlight.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `gghighlight` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **highlighting** elements.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It mainly offers a `gghighlight()` function that highlights geoms.

✍️ **author** &rarr; Hiroaki Yutani

📘 **documentation** &rarr; [github](https://github.com/yutannihilation/gghighlight)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(gghighlight)

set.seed(123)
period <- 365
df <- data.frame(
  Date = seq(as.Date("2020-01-01"),
    by = "day",
    length.out = period
  ),
  Value = c(
    cumsum(rnorm(period)),
    cumsum(rnorm(period)),
    cumsum(rnorm(period))
  ),
  Type = c(
    rep("a", period),
    rep("b", period),
    rep("c", period)
  )
)

ggplot(df) +
  geom_line(aes(Date, Value, colour = Type)) +
  gghighlight(max(Value) > 15)
```
</div>
</div>

# Installation
***
To get started with `gghighlight`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("gghighlight")
```

# Basic usage 
***

The `gghighlight` package allows you to showcase specific elements in your ggplot2 charts.

Here's a basic example on a line charts with multiple lines. At the end of the usual `ggplot2` call, the `gghighlight()` function is called with a single argument. 

This argument is the condition that a group should respect to be highlighted.

```{r}
library(ggplot2)
library(gghighlight)

# dataset with 3 lines named "a", "b" and "c"
set.seed(1)
period <- 365
df <- data.frame(
  Date = seq(as.Date("2020-01-01"),
    by = "day",
    length.out = period
  ),
  Value = c(
    cumsum(rnorm(period)),
    cumsum(rnorm(period)),
    cumsum(rnorm(period))
  ),
  Type = c(
    rep("a", period),
    rep("b", period),
    rep("c", period)
  )
)

ggplot(df) +
  geom_line(aes(Date, Value, colour = Type)) +
  gghighlight(min(Value) < -15) # highlight the group if its min value is below -15
```

# Key features
***

## &rarr; Highlight a histogram with multiple groups
Use a facetted version with `facet_wrap(~ Species)` and the gghighlight feature allow us to showcase distribution of each group while **maintaining the background distributions** for other variables.

Example:

```{r}
library(ggplot2)
library(gghighlight)

# load the iris dataset
data(iris)

# show distribution of Sepal.Length of each species
ggplot(iris, aes(Sepal.Length, fill = Species)) +
  geom_histogram() +
  gghighlight() +
  facet_wrap(~Species)
```

## &rarr; Highlight points in a scatter plot
You can highlight specific points by indicating the properties of the points in question.

Example:

```{r}
library(ggplot2)
library(gghighlight)

# create dataset
set.seed(1)
df_scatter <- data.frame(
  xvalue = rnorm(n = 100),
  yvalue = rnorm(n = 100)
)

# show distribution of Sepal.Length of each species
ggplot(df_scatter, aes(xvalue, yvalue)) +
  geom_point() +
  gghighlight(yvalue > 1.5, xvalue < 1)
```

## &rarr; Specify properties of unhighlighted
You can define the style of elements outside your filters using the `unhighlighted_params` argument.

Example:

```{r}
library(ggplot2)
library(gghighlight)

ggplot(df) +
  geom_line(aes(Date, Value, colour = Type)) +
  gghighlight(min(Value) < -15,
    unhighlighted_params = list(
      linewidth = 0.3,
      colour = alpha("lightblue", 0.8)
    )
  )
```




# Gallery of gghighlight examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=web-line-chart-small-multiple-all-group-greyed-out.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Line chart with small multiple<hr><p class=explanation_portfolio>A line chart with small multiple where all groups are repeated                 (but greyed-out) on each pannel.</div></div><img alt='Line chart with small multiple'class='img-fluid imgOfPortfolio'src=../img/fromTheWeb/we-line-chart-consumer-confidence.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=354-highlight-specific-elements-in-line-chart.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Highlight group on line chart<hr><p class=explanation_portfolio>Easily highlight interesting groups on a line chart using                     the gghighlight package</div></div><img alt='Highlight group on line chart'class='img-fluid imgOfPortfolio'src=../img/graph/354-highlight-specific-elements-in-line-chart-1.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=354-highlight-specific-elements-in-line-chart.html><div class=portfolio-hover><div class=portfolio-hover-content><p>highlight groups<hr><p class=explanation_portfolio>Combination of gghighlight, patchwork and hrbrthemes to get                     2 combined highlighted line charts.</div></div><img alt='highlight groups'class='img-fluid imgOfPortfolio'src=../img/graph/354-highlight-specific-elements-in-line-chart-2.png></a></div></div></div></section>







<!-- Close container -->
</div>
