---
title: "Make ggplot2 dynamic and interactive with **ggiraph**"
logo: "ggiraph"
descriptionMeta: "This post explains how to create interactive ggplot2 graphs with a high level of customization using the **ggiraph** extension. It provides several reproducible examples and detailed explanations."
descriptionTop: "The `ggiraph` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of creating **interactive** and **dynamic** graphs. It's an htmlwidget, which means it's highly compatible with RMarkdown/Quarto document and Shiny application.<br/><br/>This post showcases the **key features** of `ggiraph` and provides a set of **graph examples** using the package."
documentationLink: "ardata.fr/ggiraph-book/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***

The `ggiraph` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **creating interactive charts**.

Try to **hover** over the chart below:

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of function to **add interactivity** such as `geom_point_interactive()` for an interactive version of `geom_point()`, `geom_sf_interactive()` for an interactive version of `geom_sf()` etc.

✍️ **author** &rarr; David Gohel

📘 **documentation** &rarr; [github](https://github.com/davidgohel/ggiraph)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "65%", fig.align='center'}
library(ggplot2)
library(ggiraph)

myplot <- ggplot(
  data = diamonds,
  mapping = aes(x = color, fill = cut, data_id = cut)
) +
  theme_minimal() +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 3
  )

interactive_plot <- girafe(ggobj = myplot)
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-1.html")
```

<iframe src="../HtmlWidget/ggiraph-1.html" height="400" width="400" style="border:none;"></iframe>

</div>
</div>

# Installation
***
To get started with `ggiraph`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggiraph")
```

# Basic usage 
***

Here's an example using the `geom_point_interactive()` function that "replaces" the original `geom_point()` from [ggplot2](../ggplot2-package.html). It will also draw circles on the chart, but those circle will be interactive.

This geom has some new arguments like `hover_nearest = TRUE` that makes sure we always consider the **nearest point** to be hovered.



```{r}
library(ggiraph)
library(tidyverse)

mtcars_db <- rownames_to_column(mtcars, var = "carname")

myplot <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = disp, y = qsec,
    # here we add interactive aesthetics
    tooltip = carname, data_id = carname
  )
) +
  geom_point_interactive(
    size = 3, hover_nearest = TRUE
  )

interactive_plot <- girafe(ggobj = myplot)
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-2.html")
```

<iframe src="../HtmlWidget/ggiraph-2.html" height="500" width="700" style="border:none;"></iframe>

<u>Note</u>: the `tooltip` property of the `aes` function is used to control the information to be displayed in the tooltip.

# Key features
***

<br/>

## &rarr; Combining Plots

`ggiraph` becomes very powerful when **multiple plots** are displayed in the same charts. Due to its efficient mapping, we can easily create **"connected" plots** with cool hover effects!

_Note: check the [dedicated post on how to combine plots in ggiraph](../414-map-multiple-charts-in-ggiraph.html) to learn more about it._

### Combine 2 charts

```{r}
library(ggiraph)
library(tidyverse)
library(patchwork)

mtcars_db <- rownames_to_column(mtcars, var = "carname")

# First plot: Scatter plot
scatter <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = disp, y = qsec,
    tooltip = carname, data_id = carname
  )
) +
  geom_point_interactive(
    size = 3, hover_nearest = TRUE
  ) +
  labs(
    title = "Displacement vs Quarter Mile",
    x = "Displacement", y = "Quarter Mile"
  ) +
  theme_bw()

# Second plot: Bar plot
bar <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = reorder(carname, mpg), y = mpg,
    tooltip = paste("Car:", carname, "<br>MPG:", mpg),
    data_id = carname
  )
) +
  geom_col_interactive(fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Miles per Gallon by Car",
    x = "Car", y = "Miles per Gallon"
  ) +
  theme_bw()

# Combine the plots using patchwork
combined_plot <- scatter + bar +
  plot_layout(ncol = 2)

# Create a single interactive plot with both subplots
interactive_plot <- girafe(ggobj = combined_plot)

# Set options for the interactive plot
interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:cyan;stroke:black;cursor:pointer;"),
  opts_selection(type = "single", css = "fill:red;stroke:black;")
)

htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-3.html")
```

<iframe src="../HtmlWidget/ggiraph-3.html" height="600" width="820" style="border:none;"></iframe>

### Combine 3 charts

It is even possible to create more complex examples by adding, for example, a map:

<iframe src="../HtmlWidget/multiple-ggiraph-6.html" height="600" width="820" style="border:none;"></iframe>

<a class="btn btn-secondary text-uppercase"
    href = "../414-map-multiple-charts-in-ggiraph.html"
    >Code</a>





<br/><br/>

## &rarr; Use your own CSS for better styling

Although `CSS` is primarily used for styling **web pages**, it can also be relevant in customizing visuals in **R**, especially when dealing with web-based outputs like **interactive charts** or **Shiny applications**.

`CSS` code has the following structure: a property named (such as `font-size`), a colon (`:`), and the value of that property (`14px`). Each element is then separated by a `;`.

_Note: If you haven't already, check the posts on [how to use CSS in ggiraph](../412-customize-css-in-interactive-ggiraph.html) and how to [combine multiple plots and use CSS in ggiraph](../414-map-multiple-charts-in-ggiraph.html)._

In this example, we'll use the `girafe_options()` function and pass it several arguments:

- `opts_hover()` for passing the CSS when **hovering**
- `opts_tooltip()` for passing the CSS of the **tooltip**

```{r}
library(ggiraph)
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(viridis)
library(scales)

# Prepare data
mtcars_db <- rownames_to_column(mtcars, var = "carname") %>%
  mutate(efficiency = ifelse(mpg > mean(mpg), "Above Average", "Below Average"))

# Create the plot
barplot <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = reorder(carname, mpg),
    y = mpg,
    tooltip = paste(
      "<strong>Car:</strong>", carname,
      "<br><strong>MPG:</strong>", round(mpg, 1),
      "<br><strong>Horsepower:</strong>", hp,
      "<br><strong>Weight:</strong>", wt, "tons"
    ),
    data_id = carname,
    fill = mpg,
    color = efficiency
  )
) +
  geom_col_interactive(width = 0.7) +
  coord_flip() +
  scale_fill_viridis(option = "C") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Miles per Gallon by Car") +
  theme_ipsum(base_size = 14, plot_title_size = 22) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10),
    panel.background = element_rect(fill = "white")
  )

# Create interactive plot
interactive_plot <- girafe(
  ggobj = barplot,
  width_svg = 10,
  height_svg = 8,
  options = list(
    opts_hover(css = "fill:#6f1d1b;"),
    opts_hover_inv(css = "opacity:0.3;"),
    opts_selection(type = "multiple", css = "fill:#FF851B;stroke:black;"),
    opts_toolbar(saveaspng = FALSE, position = "topright", delay_mouseout = 10000),
    opts_tooltip(
      css = "background-color:black;color:white;padding:30px;border-radius:10px;box-shadow:10px 10px 10px rgba(0,0,0,0.3);font-family:Arial;font-size:20px;",
      opacity = 0.9,
      use_fill = TRUE
    ),
    opts_sizing(rescale = TRUE),
    opts_zoom(max = 2)
  )
)

# Save the plot
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-4.html")
```

<iframe src="../HtmlWidget/ggiraph-4.html" height="650" width="750" style="border:none;"></iframe>







<br/><br/>

## &rarr; `onclick` effects with JavaScript

Using the `onclick` argument, you can define **actions** to be performed when the user **clicks** on certain parameters using JavaScript (JS).

_Note: If you haven't already, check the post on [how to use JavaScript in ggiraph](../413-use-javascript-in-ggiraph.html)._


In the following example, we use JS to make the rectangles **change colour** and **display confetti** when the user clicks.

_Note: check the post on [how to use Javascript in ggiraph](../413-use-javascript-in-ggiraph.html)._


```{r}
library(ggplot2)
library(ggiraph)
library(htmltools)

my_text <- "the R-Graph-Gallery"
p1 <- ggplot(
  data = diamonds,
  mapping = aes(x = color, fill = cut, data_id = cut)
) +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 10,
    `data-blah` = my_text,
    extra_interactive_params = c("data-blah"),
    onclick = paste0(
      "this.style.fill = this.style.fill === \"red\" ? this.getAttribute(\"original-fill\") : \"red\";",
      "this.classList.toggle(\"highlighted\");",
      "var tooltip = document.getElementById(\"custom-tooltip\");",
      "tooltip.innerHTML = \"A chart by: \" + this.getAttribute(\"data-blah\");",
      "tooltip.style.display = \"block\";",
      "tooltip.style.left = (event.pageX + 10) + \"px\";",
      "tooltip.style.top = (event.pageY + 10) + \"px\";",
      "setTimeout(function() { tooltip.style.display = \"none\"; }, 1500);",
      "confetti({
        particleCount: 1000,
        spread: 70,
        origin: { y: 0.6 }
      });"
    )
  ) +
  theme_classic()

interactive_plot <- girafe(ggobj = p1)

html_content <- htmltools::tags$html(
  htmltools::tags$head(
    htmltools::tags$script(src = "https://cdn.jsdelivr.net/npm/canvas-confetti@1.5.1/dist/confetti.browser.min.js"),
    htmltools::tags$style("
      .highlighted {
        stroke: black;
        stroke-width: 2px;
      }
      #custom-tooltip {
        position: absolute;
        background: white;
        border: 1px solid black;
        padding: 5px;
        display: none;
        pointer-events: none;
      }
    ")
  ),
  htmltools::tags$body(
    interactive_plot,
    htmltools::tags$div(id = "custom-tooltip")
  )
)

htmltools::save_html(html_content, "../HtmlWidget/ggiraph-5.html")
```

<iframe src="../HtmlWidget/ggiraph-5.html" height="650" width="750" style="border:none;"></iframe>









<br/><br/>

## &rarr; Interactive maps

Interactivity is a very interesting feature, **especially** for [choropleth maps](../choropleth-map.html), as it is very intuitive to **fly over a specific region** and want to know more.

In the following example, we see how to use `ggiraph` with [sf](sf.html) to create an interactive choropleth map with additional information:

```{r}
library(ggplot2)
library(ggthemes)
library(sf)
library(ggiraph)
library(dplyr)

atlas <- readr::read_rds(
  "https://github.com/viniciusoike/restateinsight/raw/main/static/data/atlas_sp_hdi.rds"
)

pop_hdi <- atlas |>
  st_drop_geometry() |>
  mutate(
    group_hdi = findInterval(HDI, seq(0.65, 0.95, 0.05), left.open = FALSE),
    group_hdi = factor(group_hdi)
  ) |>
  group_by(group_hdi) |>
  summarise(score = sum(pop, na.rm = TRUE)) |>
  ungroup() |>
  mutate(share = score / sum(score) * 100) |>
  na.omit() |>
  mutate(
    y_text = if_else(group_hdi %in% c(0, 7), share + 3, share - 3),
    label = paste0(round(share, 1), "%"),
    data_id = as.character(group_hdi) # Add data_id to pop_hdi
  )

atlas <- atlas |>
  mutate(group_hdi = findInterval(HDI, seq(0.65, 0.95, 0.05), left.open = FALSE))

pmap <- ggplot(atlas) +
  geom_sf_interactive(aes(fill = HDI, data_id = group_hdi, tooltip = paste("HDI:", HDI)), lwd = 0.05, color = "white") +
  scale_fill_fermenter(
    name = "",
    breaks = seq(0.65, 0.95, 0.05),
    direction = 1,
    palette = "YlGnBu"
  ) +
  labs(
    title = "HDI in Sao Paulo, BR (2010)",
    subtitle = "Microregion HDI in Sao Paulo",
    caption = "Source: Atlas Brasil"
  ) +
  theme_map() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

x_labels <- c(
  "0.650 or less", "0.650 to 0.699", "0.700 to 0.749", "0.750 to 0.799",
  "0.800 to 0.849", "0.850 to 0.899", "0.900 to 0.949", "0.950 or more"
)

pcol <- ggplot(pop_hdi, aes(group_hdi, share, fill = group_hdi)) +
  geom_col_interactive(aes(data_id = data_id, tooltip = paste("Share:", label))) +
  geom_hline(yintercept = 0) +
  geom_text_interactive(
    aes(y = y_text, label = label, color = group_hdi, data_id = data_id),
    size = 2
  ) +
  coord_flip() +
  scale_x_discrete(labels = x_labels) +
  scale_fill_brewer(palette = "YlGnBu") +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 8), # Reduced title size
    axis.text.y = element_text(size = 5), # Reduced y-axis text size
    axis.text.x = element_blank(),
    aspect.ratio = 1.5
  )

p_hdi_atlas <- pmap + pcol + plot_layout(widths = c(3, 1))
p_hdi_atlas <- pmap + inset_element(pcol, left = 0.5, bottom = 0, right = 1, top = 0.5)

interactive_plot <- girafe(
  ggobj = p_hdi_atlas,
  options = list(
    opts_hover(css = "fill:orange;"),
    opts_hover_inv(css = "opacity:0.5;"),
    opts_selection(type = "single", only_shiny = FALSE)
  )
)

htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-6.html")
```

</div>

<center><iframe src="../HtmlWidget/ggiraph-6.html" height="1200" width="100%" style="max-width:1200px; border:none;"></iframe></center>

<div class="container">

<br/><br/>

# Gallery of ggiraph examples
***

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../412-customize-css-in-interactive-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Custom the fill properties using CSS <hr> <p class=explanation_portfolio>Learn how to add your own custom CSS to your interactive ggiraph charts</p> </div> </div><img alt="ggiraph with custom css" class="img-fluid imgOfPortfolio" src=../img/graph/412-customize-css-in-interactive-ggiraph-1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../412-customize-css-in-interactive-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Change properties with hover <hr> <p class=explanation_portfolio>Discover how to add CSS properties for an infinite customization in ggiraph</p> </div> </div><img alt="ggiraph with custom css" class="img-fluid imgOfPortfolio" src=../img/graph/412-customize-css-in-interactive-ggiraph-2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../412-customize-css-in-interactive-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Advanced interactive use case <hr> <p class=explanation_portfolio>Learn how to combine CSS and JavaScript to create amazing interactive charts</p> </div> </div><img alt="ggiraph with custom css" class="img-fluid imgOfPortfolio" src=../img/graph/412-customize-css-in-interactive-ggiraph-3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../413-use-javascript-in-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Change plot properties on user click</p> <hr> <p class=explanation_portfolio>Discover how to use the <code>onclick</code> argument to add javascript features in your ggiraph charts</p> </div> </div><img alt="ggiraph with javascript and onclick argument" class="img-fluid imgOfPortfolio" src=../img/graph/413-use-javascript-in-ggiraph-1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../413-use-javascript-in-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Confetti effect</p> <hr> <p class=explanation_portfolio>Add a confetti effect each time you click on your chart</p> </div> </div><img alt="ggiraph with javascript and onclick argument" class="img-fluid imgOfPortfolio" src=../img/graph/413-use-javascript-in-ggiraph-2.png> </a></div><div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../414-map-multiple-charts-in-ggiraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Combine plots</p> <hr> <p class=explanation_portfolio>Learn how to combine any charts together for interconnected dashboards</p> </div> </div><img alt="ggiraph with javascript and onclick argument" class="img-fluid imgOfPortfolio" src=../img/graph/414-map-multiple-charts-in-ggiraph.png> </a></div>  </div> </div> </section>

<!-- Close container -->
</div>
