---
title: "Make ggplot2 dynamic and interactive with **ggiraph**"
logo: "ggiraph"
descriptionMeta: "This post explains how to create interactive ggplot2 graphs with a high level of customization using the **ggiraph** extension. It provides several reproducible examples and detailed explanations."
descriptionTop: "The `ggiraph` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of creating **interactive** and **dynamic** graphs. It's an htmlwidget, which means it's highly compatible with RMarkdown/Quarto document and Shiny application.<br/><br/>This post showcases the **key features** of `ggiraph` and provides a set of **graph examples** using the package."
documentationLink: "ardata.fr/ggiraph-book/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***

The `ggiraph` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **creating interactive charts**.

Try to **hover** the chart below:

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of function to **add interactivity** such as `geom_point_interactive()` for an interactive version of `geom_point()`, `geom_sf_interactive()` for an interactive version of `geom_sf()` etc.

✍️ **author** &rarr; David Gohel

📘 **documentation** &rarr; [github](https://github.com/davidgohel/ggiraph)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "65%", fig.align='center'}
library(ggplot2)
library(ggiraph)

myplot <- ggplot(
  data = diamonds,
  mapping = aes(x = color, fill = cut, data_id = cut)
) +
  theme_minimal() +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 3
  )

interactive_plot <- girafe(ggobj = myplot)
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-1.html")
```

<iframe src="../HtmlWidget/ggiraph-1.html" height="400" width="400" style="border:none;"></iframe>

</div>
</div>

# Installation
***
To get started with `ggiraph`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggiraph")
```

# Basic usage 
***

Here's an example of the `geom_point_interactive()` function that "replaces" the original `geom_point()` from [ggplot2](../ggplot2-package.html).

Note that what makes the chart **interactive** in the `tooltip` (column of dataset that contains tooltips to be displayed when mouse is over elements) and `data_id` (column of dataset that contains id to be associated with elements) arguments.

```{r}
library(ggiraph)
library(tidyverse)

mtcars_db <- rownames_to_column(mtcars, var = "carname")

myplot <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = disp, y = qsec,
    # here we add iteractive aesthetics
    tooltip = carname, data_id = carname
  )
) +
  geom_point_interactive(
    size = 3, hover_nearest = TRUE
  )

interactive_plot <- girafe(ggobj = myplot)
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-2.html")
```

<iframe src="../HtmlWidget/ggiraph-2.html" height="700" width="700" style="border:none;"></iframe>

# Key features
***

## &rarr; Combining Plots

`ggiraph` becomes very powerful when **multiple plots** are displayed in the same charts. Due to its efficient mapping, we can easily create **"connected" plots** with cool hover effects!

Example:

```{r}
library(ggiraph)
library(tidyverse)
library(patchwork)

mtcars_db <- rownames_to_column(mtcars, var = "carname")

# First plot: Scatter plot
scatter <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = disp, y = qsec,
    tooltip = carname, data_id = carname
  )
) +
  geom_point_interactive(
    size = 3, hover_nearest = TRUE
  ) +
  labs(
    title = "Displacement vs Quarter Mile",
    x = "Displacement", y = "Quarter Mile"
  ) +
  theme_bw()

# Second plot: Bar plot
bar <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = reorder(carname, mpg), y = mpg,
    tooltip = paste("Car:", carname, "<br>MPG:", mpg),
    data_id = carname
  )
) +
  geom_col_interactive(fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Miles per Gallon by Car",
    x = "Car", y = "Miles per Gallon"
  ) +
  theme_bw()

# Combine the plots using patchwork
combined_plot <- scatter + bar +
  plot_layout(ncol = 2)

# Create a single interactive plot with both subplots
interactive_plot <- girafe(ggobj = combined_plot)

# Set options for the interactive plot
interactive_plot <- girafe_options(
  interactive_plot,
  opts_hover(css = "fill:cyan;stroke:black;cursor:pointer;"),
  opts_selection(type = "single", css = "fill:red;stroke:black;"),
  opts_toolbar(saveaspng = FALSE)
)

htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-3.html")
```

<iframe src="../HtmlWidget/ggiraph-3.html" height="600" width="820" style="border:none;"></iframe>

## &rarr; Use your own CSS for better styling

Thanks to the `options` argument in the `girafe()` function, we can provide lots of customization feature with **CSS**:

- colors
- hover effects
- shapes
- durations

Example:

```{r}
library(ggiraph)
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(viridis)
library(scales)

# Prepare data
mtcars_db <- rownames_to_column(mtcars, var = "carname") %>%
  mutate(efficiency = ifelse(mpg > mean(mpg), "Above Average", "Below Average"))

# Create the plot
barplot <- ggplot(
  data = mtcars_db,
  mapping = aes(
    x = reorder(carname, mpg),
    y = mpg,
    tooltip = paste(
      "<strong>Car:</strong>", carname,
      "<br><strong>MPG:</strong>", round(mpg, 1),
      "<br><strong>Horsepower:</strong>", hp,
      "<br><strong>Weight:</strong>", wt, "tons"
    ),
    data_id = carname,
    fill = mpg,
    color = efficiency
  )
) +
  geom_col_interactive(width = 0.7) +
  coord_flip() +
  scale_fill_viridis(option = "C") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Miles per Gallon by Car") +
  theme_ipsum(base_size = 14, plot_title_size = 22) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 10),
    panel.background = element_rect(fill = "white")
  )

# Create interactive plot
interactive_plot <- girafe(
  ggobj = barplot,
  width_svg = 10,
  height_svg = 8,
  options = list(
    opts_hover(css = "fill:#6f1d1b;"),
    opts_hover_inv(css = "opacity:0.3;"),
    opts_selection(type = "multiple", css = "fill:#FF851B;stroke:black;"),
    opts_toolbar(saveaspng = FALSE, position = "topright", delay_mouseout = 10000),
    opts_tooltip(
      css = "background-color:black;color:white;padding:30px;border-radius:10px;box-shadow:10px 10px 10px rgba(0,0,0,0.3);font-family:Arial;font-size:20px;",
      opacity = 0.9,
      use_fill = TRUE
    ),
    opts_sizing(rescale = TRUE),
    opts_zoom(max = 2)
  )
)

# Save the plot
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-4.html")
```

<iframe src="../HtmlWidget/ggiraph-4.html" height="650" width="750" style="border:none;"></iframe>

## &rarr; `onclick` effects with JavaScript

Using the `onclick` argument, you can define **actions** to be performed when the user **clicks** on certain parameters using JavaScript (JS).

In the following example, we use JS to make the rectangles **change colour** and **display confetti** when the user clicks.

```{r}
library(ggplot2)
library(ggiraph)
library(htmltools)

my_text <- "the R-Graph-Gallery"
p1 <- ggplot(
  data = diamonds,
  mapping = aes(x = color, fill = cut, data_id = cut)
) +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 10,
    `data-blah` = my_text,
    extra_interactive_params = c("data-blah"),
    onclick = paste0(
      "this.style.fill = this.style.fill === \"red\" ? this.getAttribute(\"original-fill\") : \"red\";",
      "this.classList.toggle(\"highlighted\");",
      "var tooltip = document.getElementById(\"custom-tooltip\");",
      "tooltip.innerHTML = \"A chart by: \" + this.getAttribute(\"data-blah\");",
      "tooltip.style.display = \"block\";",
      "tooltip.style.left = (event.pageX + 10) + \"px\";",
      "tooltip.style.top = (event.pageY + 10) + \"px\";",
      "setTimeout(function() { tooltip.style.display = \"none\"; }, 1500);",
      "confetti({
        particleCount: 1000,
        spread: 70,
        origin: { y: 0.6 }
      });"
    )
  ) +
  theme_classic()

interactive_plot <- girafe(ggobj = p1)

html_content <- htmltools::tags$html(
  htmltools::tags$head(
    htmltools::tags$script(src = "https://cdn.jsdelivr.net/npm/canvas-confetti@1.5.1/dist/confetti.browser.min.js"),
    htmltools::tags$style("
      .highlighted {
        stroke: black;
        stroke-width: 2px;
      }
      #custom-tooltip {
        position: absolute;
        background: white;
        border: 1px solid black;
        padding: 5px;
        display: none;
        pointer-events: none;
      }
    ")
  ),
  htmltools::tags$body(
    interactive_plot,
    htmltools::tags$div(id = "custom-tooltip")
  )
)

htmltools::save_html(html_content, "../HtmlWidget/ggiraph-5.html")
```

<iframe src="../HtmlWidget/ggiraph-5.html" height="650" width="750" style="border:none;"></iframe>

## &rarr; Interactive maps

Interactivity is a very interesting feature, **especially** for [choropleth maps](../choropleth-map.html), as it is very intuitive to **fly over a specific region** and want to know more.

In the following example, we see how to use `ggiraph` with [sf](sf.html) to create an interactive choropleth map with additional information:

```{r}
library(ggthemes)
library(sf)

atlas <- readr::read_rds(
  "https://github.com/viniciusoike/restateinsight/raw/main/static/data/atlas_sp_hdi.rds"
)
choropleth <- ggplot(atlas) +
  geom_sf_interactive(aes(
    fill = HDI,
    tooltip = paste(
      "<strong>HDI:</strong>", HDI,
      "<br><strong>Population:</strong>", round(pop, 1)
    ),
    data_id = code_udh
  ), lwd = 0.05, color = "white") +
  scale_fill_fermenter(
    name = "",
    breaks = seq(0.65, 0.95, 0.05),
    direction = 1,
    palette = "YlGnBu"
  ) +
  labs(
    title = "HDI in Sao Paulo, BR (2010)",
    subtitle = "Microregion HDI in Sao Paulo",
    caption = "Source: Atlas Brasil"
  ) +
  theme_map() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 18, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )


interactive_plot <- girafe(ggobj = choropleth)
htmltools::save_html(interactive_plot, "../HtmlWidget/ggiraph-6.html")
```

<iframe src="../HtmlWidget/ggiraph-6.html" height="1000" width="750" style="border:none;"></iframe>

# Gallery of ggiraph examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-dumbbell-chart-with-a-gap-column.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dumbbell plot with gap section<hr><p class=explanation_portfolio>An extended dumbbell plot originally published by the Pew research center. Clean design with a gap section on the right hand side.</div></div><img alt="Dumbell chart with gap column on the right"class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-dumbbell-chart-with-a-gap-column-2.png></a></div><div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../line-chart-dual-Y-axis-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dual Y axis<hr><p class=explanation_portfolio>A few examples showing how to use the sec.axis argument to add a second Y axis.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/line-chart-dual-Y-axis-ggplot2_2.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><a class=portfolio-link href=../web-circular-lollipop-plot-with-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Circular lollipop chart<hr><p class=explanation_portfolio>A circular lollipop chart with customized layout, in circle legend and groups.<br><br>A work by <b>Cédric Scherer</b> for the <b>TidyTuesday</b> challenge.</div></div><img alt="Circular lollipop chart with R and ggplot2"class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-circular-lollipop-plot-with-ggplot2-square.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><div class=portfolio-item><a class=portfolio-link href=../web-lineplots-and-area-chart-the-economist.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Line and Area charts by The Economist<hr><p class=explanation_portfolio>Learn how to mimick the style of The Economist with libraries like ggplot2, ggiraph, grid, ggtext and more.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-lineplots-and-area-chart-the-economist.png></a></div></div></div></div></section>

<!-- Close container -->
</div>
