---
title: "Enhance Your ggplot2 Visualizations with **ggpattern**"
logo: "ggpattern"
descriptionMeta: "This post explains how to add patterns to ggplot2 visualizations with R and the ggpattern package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggpattern` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to add **pattern aesthetics** to your ggplot2 visualizations.<br/>This post showcases the **key features** of `ggpattern` and provides a set of **graph examples** using the package."
documentationLink: "github.com/coolbutuseless/ggpattern"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Quick start
***
The `ggpattern` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to add **pattern aesthetics** to your ggplot2 visualizations.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It allows you to **add patterns** like stripes, crosses, and more to your ggplot2 plots, making your visualizations more informative and engaging.

✍️ **author** &rarr; Mike FC

📘 **documentation** &rarr; [github](https://coolbutuseless.github.io/package-ggpattern/)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggpattern)
# Example ggpattern code here
```
</div>
</div>

```{r, eval=FALSE}
# Example ggpattern code here
```

# Installation
***
Getting started with `ggpattern` is straightforward. 

First, ensure you have `ggplot2` installed. Then, you can install `ggpattern` directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggpattern", repos = "https://cinc.rud.is")
```

# Basic usage 
***
The `ggpattern` package comes with a variety of **pattern aesthetics** that can be easily integrated into your ggplot2 visualizations.

Here is an example using the famous `mtcars` dataset, adding stripes to bars in a bar chart:

```{r}
# Example ggpattern code here
```

# Key features
***
Here is an overview of the main features offered by ggpattern:

## &rarr; Pattern Aesthetics
`ggpattern` allows you to add various types of patterns to your ggplot2 visualizations, such as stripes, crosses, and more.

Example:

```{r}
# Example ggpattern code here
```

<!-- Close container -->
</div>
