---
title: "Filled areas with geometric and image-based patterns"
logo: "ggpattern"
descriptionMeta: "This post explains how to add patterns to ggplot2 visualizations with R and the ggpattern package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggpattern` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to add **pattern aesthetics** to your ggplot2 visualizations.<br/>This post showcases the **key features** of `ggpattern` and provides a set of **graph examples** using the package."
documentationLink: "github.com/coolbutuseless/ggpattern"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">


# Quick start
***
The `ggpattern` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to add **pattern aesthetics** to your ggplot2 visualizations.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It allows you to add patterns like **stripes**, **crosses**, **custom images** and more to your ggplot2 plots, making your visualizations more informative and engaging. 


✍️ **author** &rarr; Mike FC

📘 **documentation** &rarr; [github](https://coolbutuseless.github.io/package-ggpattern/)

⭐️ *more than 500 stars on github*

</div>


<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
# libs
library(ggplot2)
library(ggpattern)

# Create data
data <- data.frame(
  name=c("A","B","C","D","E") ,  
  value=c(3, 12, 5, 18, 45)
  )

# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_col_pattern(
    aes(pattern_fill = name), 
    pattern = 'stripe',
    fill    = 'white',
    colour  = 'black'
  )
```
</div>
</div>

This package can be especially useful if you need to publish some **black and white** content.


```{r, eval=FALSE}
# libs
library(ggplot2)
library(ggpattern)

# Create data
data <- data.frame(
  name=c("A","B","C","D","E") ,  
  value=c(3, 12, 5, 18, 45)
  )

# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_col_pattern(
    aes(pattern_fill = name), 
    pattern = 'stripe',
    fill    = 'white',
    colour  = 'black'
  )
```


# Installation
***
Getting started with `ggpattern` is straightforward. 

First, ensure you have `ggplot2` installed. Then, you can install `ggpattern` directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggpattern", repos = "https://cinc.rud.is")
```


# Basic usage 
***
The `ggpattern` package comes with a variety of **pattern aesthetics** that can be easily integrated into your `ggplot2` visualizations.

Let's take an example with the `geom_col_pattern()` aesthetics. It does pretty much the same task as the `geom_col()` aesthetics of `ggplot2` that is used for [bar charts](https://r-graph-gallery.com/218-basic-barplots-with-ggplot2.html). 

But it adds options to use some **patterns** into rectangles.

The aesthetics comes with a number of options described in the next section. For each, you have to decide:

- if you want it to depend on a value of the dataset, put it inside the `aes()` call. Below, the color of the pattern is mapped to the `name` columnn since `pattern_fill` is **in** the `aes()` function.

- if you want it to be the same for all shapes, put it outside the `aes()` call. For instance, all bars will be using `crosshatch` below.

```{r}
# Barplot
ggplot(data, aes(x=name, y=value)) + 
  geom_col_pattern(
    aes(pattern_fill = name), 
    pattern = 'crosshatch',
    fill    = 'white',
    colour  = 'black'
  )
````



# Customization
***

It is possible to customize pretty much everything on the pattern you are building.

- `colour` and `fill` control the shape outline and the shape content colors.

- `pattern` controls the pattern type. It can be: 'stripe' (default), 'crosshatch', 'point', 'circle' or 'none'

- `pattern_alpha` controls the pattern transparency

- `pattern_angle` controls the angle 🤷‍♀️

- `pattern_colour` and `pattern_fill` control the outline and fill colors

Type `help(geom_col_pattern)` to get some more details on all the available options!


```{r}
# Bar chart using ggpattern with comments
ggplot(data, aes(x=name, y=value, fill= name)) + # Usual ggplot2 call, "name" for the x axis, "value" for the y axis
  geom_col_pattern(
    pattern = "circle",
    pattern_alpha = 0.8,
    pattern_fill = "white",
    pattern_colour  = '#69b3a2',
    pattern_size = 1.3
  )
```


# Mapping pattern to data
***
You can map the pattern features to the data in use. 

For instance, the following example use a very different look for each group of the dataset. To do so, the arguments described above are placed inside the `aes()` call:

```{r}
# Bar chart using ggpattern with comments
ggplot(data, aes(x=name, y=value)) + # Usual ggplot2 call, "name" for the x axis, "value" for the y axis
  geom_col_pattern(
    aes(
      pattern_fill = name,
      fill= name,
      pattern = name,
    ),  
    colour  = 'black'
  )
```


# Other aesthetics
***

Several other aesthetics are available, allowing to fill any kind of chart types with pattern. Names are always matching the original `ggplot2` names.

To put it in a nutshell:

- `geom_boxplot_pattern` can be used for [boxplot](https://r-graph-gallery.com/boxplot.html)
- `geom_density_pattern` for [density charts](https://r-graph-gallery.com/density-plot.html)
- `geom_map_pattern` for maps
- `geom_violin_pattern` for violin charts

And many more are on the list. Here is an overview of other examples in the gallery using the `ggpattern` library:

# Gallery of gtsummary examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items>

<div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../368-black-and-white-barchart.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Textured black and white barplot<p class=explanation_portfolio>How to add textures into the bars of a barplot</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/368-black-and-white-barchart-1.png></a></div>

<div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../357-area-chart-with-gradient.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Area chart with opaque gradient color<p class=explanation_portfolio>Color gradient in the area of an aread chart</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/357-area-chart-with-gradient-1.png></a></div>

<div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../357-area-chart-with-gradient.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Area chart with gradient color<p class=explanation_portfolio>Color gradient in the area of an aread chart with less opacity</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/357-area-chart-with-gradient-2.png></a></div>

<div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../357-area-chart-with-gradient.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Area chart with gradient color and data points<p class=explanation_portfolio>Color gradient in the area of an aread chart with less opacity and data points</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/357-area-chart-with-gradient-3.png></a></div>

</div></div></section>

<!-- Close container -->
</div>
