---
title: "Visualizing Networks with **ggraph**"
logo: "ggraph"
descriptionMeta: "This post explains how to create complex network graphs using the ggraph package in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggraph` package in R builds upon the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, enabling advanced **graph and network visualizations**.<br/>This post explores the **key features** of `ggraph` through a series of **graph visualization examples**."
documentationLink: "ggraph.data-imaginist.com/index.html"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `ggraph` package in R extends the capabilities of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package for creating sophisticated **graph visualizations**.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a rich set of **layouts** and **aesthetic options** that make it easy to represent complex network structures visually.

✍️ **author** &rarr; Thomas Lin Pedersen

📘 **documentation** &rarr; [github](https://ggraph.data-imaginist.com/index.html)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(ggraph)
library(tidygraph)
graph = as_tbl_graph(highschool)
ggraph(graph) +
  geom_edge_link() +
  geom_node_point() +
  theme_graph(background = 'white')
```
</div>
</div>

# Installation
***
To get started with `ggraph`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggraph")
```

# Basic usage 
***

The `ggraph` package allows you to visualize graphs and networks using **layouts** tailored to graph data.

Here's a basic example:

```{r}
library(ggraph)
graph = as_tbl_graph(highschool)

ggraph(graph) +
  geom_edge_link() +
  geom_node_point() +
  theme_graph(background = 'white')
```

# Key features
***

## &rarr; Map node color
You can map **node color** to your graph elements to encode additional data.

Example:

```{r}
library(ggraph)
graph <- as_tbl_graph(highschool) |> 
    mutate(Popularity = centrality_degree(mode = 'in'))

ggraph(graph, layout = 'fr') +
  geom_edge_link() +
  geom_node_point(aes(color = Popularity, size=10)) +
  theme_graph(background = 'white')
```

## &rarr; Map edge color
You can also map **edge color** to your graph elements to highlight relationships between nodes.

Example:

```{r}
library(ggraph)
graph <- as_tbl_graph(highschool)

ggraph(graph) +
  geom_edge_link(aes(color = factor(year))) +
  geom_node_point() +
  theme_graph(background = 'white')
```

## &rarr; Circular layout
The `ggraph` package supports a **circular layout** that can be used to represent network data in a circular form.

Example:

```{r}
library(ggraph)

ggraph(graph, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(aes(colour = factor(year))) + 
  coord_fixed() +
  theme_graph(background = 'white')
```

## &rarr; Treemap layout
The `ggraph` package also supports a **treemap layout** that can be used to represent hierarchical data in a tree structure.

Example:

```{r}
library(ggraph)
graph <- tbl_graph(flare$vertices, flare$edges)

ggraph(graph, 'treemap', weight = size) + 
  geom_node_tile(aes(fill = depth), size = 0.2) +
  theme_graph(background = 'white')
```


# Gallery of ggraph examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../311-add-labels-to-hierarchical-edge-bundling.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Hierarchical edge bundling</p><hr><p class=explanation_portfolio>A visually appealing way to represent interconnected data.</p></div></div><img alt="Force-directed graph" class="img-fluid imgOfPortfolio"src=../img/graph/311-add-labels-to-hierarchical-edge-bundling.png></a></div><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../314-custom-circle-packing-with-several-levels.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Circle packing<hr><p class=explanation_portfolio>Explore how to create a custom circle packing with several levels.</div></div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/314-custom-circle-packing-with-several-levels1.png></a></div><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../315-hide-first-level-in-circle-packing.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Text in circle packing<hr><p class=explanation_portfolio>Learn how to hide the first level of a circle packing and displaytext in the circles.</div></div><img alt="Circular graph layout with R and ggplot2" class="img-fluid imgOfPortfolio"src=../img/graph/315-hide-first-level-in-circle-packing2.png></a></div><div class="portfolio-item col-sm-6 col-md-6"><div class=portfolio-item><a class=portfolio-link href=../335-custom-ggraph-dendrogram.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Custom dendrogram<hr><p class=explanation_portfolio>Discover how to create and customize a dendrogram with ggraph</div></div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/335-custom-ggraph-dendrogram.png></a></div></div><div class="portfolio-item col-sm-6 col-md-6"><div class=portfolio-item><a class=portfolio-link href=../339-circular-dendrogram-with-ggraph.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Custom hierarchical edge bundling<hr><p class=explanation_portfolio>Explore how to customize color, shape and size</div></div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/339-circular-dendrogram-with-ggraph.png></a></div></div></div></div></section>

<!-- Close container -->
</div>