---
title: "Create Flow Diagrams with **ggsankey**"
logo: "ggsankey"
descriptionMeta: "This post explains how to create flow diagrams using the ggsankey package, an extension of ggplot2. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggsankey` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to **create flow visualizations**.<br/>This post showcases the **key features** of `ggsankey` and provides a set of **diagram examples** using the package."
documentationLink: "github.com/davidsjoberg/ggsankey"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `ggsankey` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to **create flow visualizations**.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of **functions** that make it easy to specify flow diagrams in a declarative manner.

✍️ **author** &rarr; David Sjoberg

📘 **documentation** &rarr; [github](https://github.com/davidsjoberg/ggsankey)

⭐️ *more than 250 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(ggsankey)

df <- mtcars %>%
  make_long(cyl, vs, am, gear, carb)

ggplot(df, aes(x = x, next_x = next_x, node = node, next_node = next_node, fill = factor(node), label = node)) +
  geom_sankey(flow.alpha = .6,
              node.color = "gray30") +
  geom_sankey_label(size = 3, color = "white", fill = "gray40") +
  scale_fill_viridis_d(drop = FALSE) +
  theme_sankey(base_size = 18) +
  labs(x = NULL) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = .5)) +
  ggtitle("Car features")
```
</div>
</div>

# Installation
***
To get started with `ggsankey`, you can install its developpment version directly from Github using the `install_github()` function:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("davidsjoberg/ggsankey")
```

# Basic usage 
***

The `ggsankey` package extends the grammar of graphics to include the description of flow diagrams, specifically, sankey, alluvial and sankey bump diagrams. You start with a dataframe in wide format, transform it using the `make_long()` function, and then add **flow-specific layers** to your ggplot.

Here's a basic example where we show how dimensions are linked using a **sankey diagram**: 

```{r}
library(tidyverse)
library(ggsankey)

# Create a simple dataset about education and career paths
df <- data.frame(
  education = c(rep("High School", 40), rep("Bachelor", 35), rep("Master", 25)),
  field = c(rep("Science", 20), rep("Arts", 20), rep("Business", 30), rep("Engineering", 30)),
  job = c(rep("Research", 25), rep("Teaching", 25), rep("Industry", 30), rep("Consulting", 20))
)

# Convert to long format for Sankey diagram
df_long <- df %>%
  make_long(education, field, job)

# Create the diagram
ggplot(df_long, 
       aes(x = x, 
           next_x = next_x, 
           node = node, 
           next_node = next_node,
           fill = factor(node))) +
  geom_sankey() +
  scale_fill_discrete(drop=FALSE)
```

# Key features
***

<br>

## &rarr; Embellishment

Labels with `geom_sankey_label()`nicely places labels in the center of nodes if given the same aesthetics.

`ggsankey` also comes with custom minimalistic themes that can be used. Here we use `theme_sankey()`:

```{r}
ggplot(df_long, 
       aes(x = x, 
           next_x = next_x, 
           node = node, 
           next_node = next_node,
           fill = factor(node))) +
  geom_sankey(alpha = 0.8) +
  scale_fill_viridis_d(option = "plasma", name = "Category") +
  theme_sankey(base_size = 14) +
  labs(title = "Education to Career Path Flow",
       x = NULL) +
  theme(legend.position = "bottom",
        legend.title = element_text(hjust = 0.5),
        legend.box.just = "center",
        plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) +
  scale_x_discrete(labels = c("Education", "Field", "Job"))
```

<br>

## &rarr; Continuous stacking versus centered gaps

**Alluvial diagrams** are very similiar to sankey diagrams but have no spaces between nodes and start at y = 0, instead of being centered around the x-axis.

This diagram shows how individuals progress from their education level through their field of study to their eventual job role. The `geom_alluvial()` function creates smooth flowing bands between the nodes, with the **width of each band** representing the number of individuals following each path.
```{r}
ggplot(df_long, 
       aes(x = x, 
           next_x = next_x, 
           node = node, 
           next_node = next_node,
           fill = factor(node),
           label = node)) +
  geom_alluvial(flow.alpha = .6) +
  geom_alluvial_text(size = 3, color = "black", space = 0.5) +  # Adjusted space and changed text color
  scale_fill_viridis_d(option = "plasma", drop = FALSE) +
  theme_alluvial(base_size = 18) +
  labs(x = NULL) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = .5)) +
  ggtitle("Education to Career Pathways") +
  scale_x_discrete(labels = c("Education", "Field", "Job"))
```

<br>

## &rarr; Intersecting flows show changing ranks

**Sankey bump plots** combine characteristics of bump charts and Sankey diagrams, visualizing how rankings and volumes **change over time**. When a group's value surpasses another, it shifts upward in the visualization, creating a distinctive **"bumping"** effect.

This visualization shows how different renewable energy sources have contributed to global electricity generation from 2000 to 2023. The `geom_sankey_bump()` function creates flowing streams that expand or contract to show changes in generation capacity, with technologies shifting positions when their relative contribution changes over time.
```{r}
library(tidyverse)
library(ggsankey)

# Read and prepare the data
renewable_data <- read.csv("https://ourworldindata.org/grapher/modern-renewable-prod.csv?v=1&csvType=filtered&useColumnShortNames=true") %>%
  # Select last 10 years for better visualization
  filter(Year >= 2000) %>%
  pivot_longer(
    cols = c(wind_generation__twh, hydro_generation__twh, 
             solar_generation__twh, other_renewables_including_bioenergy_generation__twh),
    names_to = "technology",
    values_to = "generation"
  ) %>%
  mutate(
    technology = case_when(
      technology == "wind_generation__twh" ~ "Wind",
      technology == "hydro_generation__twh" ~ "Hydro",
      technology == "solar_generation__twh" ~ "Solar",
      technology == "other_renewables_including_bioenergy_generation__twh" ~ "Other Renewables"
    )
  )

# Create the plot
ggplot(renewable_data, 
       aes(x = Year,
           node = technology,
           fill = technology,
           value = generation)) +
  geom_sankey_bump(space = 0, 
                   type = "alluvial", 
                   color = "transparent", 
                   smooth = 6) +
  scale_fill_viridis_d(option = "inferno", alpha = .8)  +
  scale_x_continuous(breaks = scales::pretty_breaks(), expand = c(0,0)) + 
  scale_y_continuous(labels = scales::comma, expand = c(0,0)) +
  theme_sankey_bump(base_size = 16) +
  labs(x = NULL,
       y = "Electricity Generation (TWh)",
       fill = "Technology",
       title = "Global Renewable Energy Generation by Source",
       subtitle = "Period: 2000-2023",
       caption = "Data sources: Ember (2024), Energy Institute - Statistical Review of World Energy (2024)\nProcessed by Our World in Data") +
  theme(legend.position = "bottom",
        legend.title = element_text(hjust = 0.5),
        legend.title.position = "top",
        plot.caption = element_text(hjust = 0, size = 10, face = "italic"))
```

<br>

<!-- Close container -->
</div>