---
title: "Enhancing Text in ggplot2 with **ggtext**"
logo: "ggtext"
descriptionMeta: "This post explains how to enhance text elements in ggplot2 plots using the ggtext package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `ggtext` package in R is an extension of [ggplot2](https://r-graph-gallery.com/ggplot2-package.html), designed to **improve text rendering** in plots.<br/>This post showcases the **key features** of `ggtext` and provides a set of **graph examples** using the package."
documentationLink: "github.com/wilkelab/ggtext"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Quick start
***
The `ggtext` package in R extends the text rendering capabilities of [ggplot2](https://r-graph-gallery.com/ggplot2-package.html), allowing for **rich text formatting** in plot elements.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It provides functions to render **markdown** and **HTML** text in ggplot2 plots, enabling more flexible and visually appealing text elements.

✍️ **author** &rarr; Claus O. Wilke

📘 **documentation** &rarr; [github](https://github.com/wilkelab/ggtext)

⭐️ *more than 600 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(ggtext)

ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    title = "Car <span style='color:blue;'>Weight</span> vs. <span style='color:red;'>MPG</span>",
    subtitle = "Relationship between <span style='color:blue;'>**weight**</span> and <span style='color:red;'>**miles per gallon**</span>"
  ) +
  theme(
    plot.title = element_markdown(),
    plot.subtitle = element_markdown()
  )
```

</div>
</div>

# Installation
***
To get started with `ggtext`, you can install it from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("ggtext")
```

# Basic usage 
***

The `ggtext` package allows you to use markdown and HTML formatting in text elements of ggplot2 plots. Here's a basic example:

In this example, we use HTML tags (`span`) in the subtitle to color specific words. The `element_markdown()` function in the `theme()` call enables the rendering of this formatted text.

```{r}
library(ggplot2)
library(ggtext)

ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    title = "Car Weight vs. MPG",
    subtitle = "Relationship between <span style='color:blue;'>weight</span> and <span style='color:red;'>miles per gallon</span>"
  ) +
  theme(plot.subtitle = element_markdown())
```


# Key features
***

## &rarr; Rich Text Formatting
You can use markdown and HTML to format text, including bold, italic, and colored text.

Example:

```{r}
library(ggplot2)
library(ggtext)

ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    title = "Car Weight vs. MPG",
    subtitle = "**Bold text**, *italic text*, and <span style='color:#72874EFF;'>**colored text**</span>"
  ) +
  theme(plot.subtitle = element_markdown())
```


<br><br>


## &rarr; `ggtext` as a legend

Using `ggtext`, we can change the axis labels so that they are colored to match the bar colors:

- **Inline Coloring**:
  - `color` defines bar and text colors.
  - `name` combines `item_name` and `group` with HTML tags for styling: `<i style='color:{color}'>**{item_name}**</i> ({group})`.

- **Theming**:
  - `axis.text.y = element_markdown()` applies HTML/CSS formatting to y-axis labels.

This has direct benefits:

- **No Legend Needed**: Labels are color-matched to bars, making the plot intuitive.
- **Enhanced Visualization**: Improved readability and aesthetics through direct color association.

```{r}
library(tidyverse)
library(ggtext)
library(glue)

data <- tibble(
  item_name = c("Item A", "Item B", "Item C", "Item D"),
  group = c("Group 1", "Group 2", "Group 3", "Group 4"),
  value = c(-1.5, 1.0, 2.5, 4.0)
)

data %>%
  mutate(
    color = c("#FF5733", "#33FF57", "#3357FF", "#FF33FF"),
    name = glue("<i style='color:{color}'>**{item_name}**</i> ({group})"),
    name = fct_reorder(name, value)
  ) %>%
  ggplot(aes(value, name, fill = color)) +
  geom_col(alpha = 0.5) +
  scale_fill_identity() +
  labs(caption = "Adapted example from the ***{ggtext}*** documentation<br>(using made-up data)") +
  theme(
    axis.text.y = element_markdown(),
    plot.caption = element_markdown(lineheight = 1.2)
  )
```


<br><br>

## &rarr; Box around text

`ggtext` can be used to create boxes around your text in different styles:

- `geom_richtext()`: Replaces the usual `geom_text()` to render the text in `label` with special formatting.
- The `label` entries use:
  - `**bold**` for bold text.
  - `<br>` for **line breaks**.
  - `<span style='color:green'>` for colored text.
  - `<sup>` and `<sub>` for **superscript** and **subscript**.
  - `<span style='font-size:16pt'>` for larger text.

This setup allows the plot to display text with **mixed styles** (e.g., bold, colored, italicized) and structured formats (e.g., line breaks, superscripts) within a single text label.


```{r}
library(ggplot2)
library(tibble)
library(ggtext)

df <- tibble(
  label = c(
    "Another text **in bold.**",
    "Different linebreaks<br>Linebreaks<br><br>Double Linebreaks",
    "*y*<sup>3</sup> + 7*y* + *D*<sub>*j*</sub>",
    "More <span style='color:green'>green text **in bold.**</span><br>And *italicized text.*<br>
    And some <span style='font-size:16pt; color:purple'>large</span> text."
  ),
  x = c(.3, .2, .6, .8),
  y = c(.7, .3, .2, .4),
  hjust = c(0.5, 0, 1, 0.5),
  vjust = c(0.5, 1, 0.5, 0.5),
  angle = c(0, 10, 30, -30),
  color = c("black", "green", "black", "purple"),
  fill = c("ivory", "white", "lightgreen", "white")
)

ggplot(df) +
  aes(
    x, y,
    label = label, angle = angle, color = color, fill = fill,
    hjust = hjust, vjust = vjust
  ) +
  geom_richtext() +
  geom_point(color = "black", size = 1) +
  scale_color_identity() +
  scale_fill_identity() +
  xlim(0, 1) +
  ylim(0, 1)
```


<br><br>

## Valid HTML/CSS properties

Unfortunatley, `ggtext` only supports a **limited amount of features**. Only the following will work:

- **bold** text when surrounded by `**` &rarr; \*\*some text that will renders bold\*\*
- **italic** text when surrounded by `*` &rarr; \*some text that will renders italic\*
- **color** text: `<span>` with a `color` value in `style` &rarr; \<span style='color:green'>green text\</span>
- **font size**: `<span>` with a `font-size` value in `style` &rarr; \<span style='font-size:30px'>large text\</span>
- **sub or superscript**: `<sub>` or `<sup>` &rarr; \<sub>subscript\</sub> and \<sup>subscript\</sup>

If you want access to more text formatting functions, we **strongly** recommend that you check out and use the brand new [marquee](https://marquee.r-lib.org/) package.

# Gallery of ggtext examples
***

The gallery is full of examples that use `ggtext` to improve their style. If you want to delve into some **real-life examples**, take a look at them!

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-circular-lollipop-plot-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular lollipop <hr> <p class=explanation_portfolio>Use of circular lollipop with nice style </div> </div><img alt="circular lollipop" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-circular-lollipop-plot-with-ggplot2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-line-chart-small-multiple-all-group-greyed-out.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Small multiple line chart <hr> <p class=explanation_portfolio>Discover how to create a small multiple line charts with a nice color schema </div> </div><img alt="Small multiple line chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/we-line-chart-consumer-confidence.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-extended-dumbbell-plot-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Dumbell chart <hr> <p class=explanation_portfolio>Learn how to use a dumbell chart in combination with a nice legend </div> </div><img alt="Dumbell chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-extended-dumbbell-plot-ggplot2-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-dumbell-chart.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Dumbell chart <hr> <p class=explanation_portfolio>Discover how to create a dumbell chart with a nice color schema </div> </div><img alt="dumbell chart" class="img-fluid imgOfPortfolio" src=../img/graph/web-dumbell-chart.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-line-chart-with-labels-at-end-of-line.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Line chart <hr> <p class=explanation_portfolio>Learn how to create a line chart with inline labels with nice colors </div> </div><img alt="Line chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-line-chart-with-labels-at-end-of-line.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-lineplots-and-area-chart-the-economist.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Combine line and area chart <hr> <p class=explanation_portfolio>Discover how to combine a line chart and an area chart to highlight trends </div> </div><img alt="Combine line and area chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-lineplots-and-area-chart-the-economist.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-lollipop-plot-with-r-mario-kart-64-world-records.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Dumbell chart <hr> <p class=explanation_portfolio>Learn how to create a dumbell chart with a nice legend and annotations </div> </div><img alt="Dumbell chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/lollipop-plot-with-r-mario-kart-64-world-records-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-ridgeline-plot-with-inside-plot-and-annotations.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Ridgeline plot <hr> <p class=explanation_portfolio>Discover how to create a ridgeline plot with a dedicated small chart as a legend </div> </div><img alt="Ridgeline plot" class="img-fluid imgOfPortfolio" src=../img/graph/web-ridgeline-plot-with-inside-plot-and-annotations-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-stacked-area-chart-inline-labels.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked area chart <hr> <p class=explanation_portfolio>Learn how to create a smoothed stacked area chart with inline labels </div> </div><img alt="Stacked area chart" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-stacked-area-chart-inline-labels.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-streamchart-with-ggstream.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Streamgraph <hr> <p class=explanation_portfolio>Discover how to create a streamgraph with a nice color schema and great annotations </div> </div><img alt="Streamgraph" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-streamchart-with-ggstream-square.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-treemap-with-annotations-labels-and-colors.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Treemap <hr> <p class=explanation_portfolio>Learn how to create a treemap that with a nice subtitle for the legend </div> </div><img alt="Treemap" class="img-fluid imgOfPortfolio" src=../img/graph/web-treemap-with-annotations-labels-and-colors.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-valued-dots-map-bertin.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Map of the France <hr> <p class=explanation_portfolio>Discover how to create a map of the France that uses the style of Jacques Bertin </div> </div><img alt="Map of the France" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-valued-dots-map-bertin.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-waffle-chart-share.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Waffle chart <hr> <p class=explanation_portfolio>Learn how to use waffle charts to represents share's of different categories with nice annotations </div> </div><img alt="Waffle chart" class="img-fluid imgOfPortfolio" src=../img/graph/web-waffle-chart-share.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-waffle-for-time-evolution.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Waffle chart <hr> <p class=explanation_portfolio>Discover how to use waffle charts to represent evolution and trends, in combination with great annotations </div> </div><img alt="Waffle chart" class="img-fluid imgOfPortfolio" src=../img/graph/web-waffle-for-time-evolution-square.png> </a></div> </div> </div> </section>


<!-- Close container -->
</div>