---
title: "Create beautiful tables with **gt**"
logo: "gt"
descriptionMeta: "This post explains how to make really nice looking tables with the gt package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `gt` package in R is a powerful tool for creating **elegant and customizable tables** for data visualization and reporting. It offers a user-friendly way to **design and style tables** in RMarkdown documents and Shiny applications."
documentationLink: "www.gt.rstudio.com"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `gt` package in R is a powerful tool for creating **elegant and customizable tables** for data visualization and reporting.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers options for formatting, styling, and theming tables, as well as support for handling complex data structures and creating publication-ready tables with ease.

✍️ **author** &rarr; Richard Iannone

📘 **documentation** &rarr; [github](https://gt.rstudio.com)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(gt)
library(dplyr)

# Create a simple data frame
data <- data.frame(
  Country = c("USA", "China", "India", "Brazil"),
  Population = c(331, 1441, 1393, 212),
  GDP = c(21.43, 14.34, 2.87, 1.49)
)


data %>%
  gt() %>%
  fmt_number(columns = c("Population", "GDP"), suffix = " M")

```
</div>
</div>

# Installation
***
To get started with `gt`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("gt")
```

# Basic usage 
***

We generally use the `dplyr` package in combination of `gt` for better code readability.

Here's a basic example with **default rendering**:

```{r}
library(gt)
library(dplyr)

# Create a simple data frame
data = data.frame(
  Country = c("USA", "China", "India", "Brazil"),
  Capitals = c("Washington D.C.", "Beijing", "New Delhi", "Brasília"),
  Population = c(331, 1441, 1393, 212),
  GDP = c(21.43, 14.34, 2.87, 1.49)
)

# Alternatively you can do (same output):
#gt(data)

# Use the gt function
data %>%
  gt()
```

# Key features
***

## &rarr; Title and Subtitle
You can add and customize **title and subtitle** with the `tab_header()` function. Moreover, by using the `md()` function we can insert some **markdown styling** inside the texts.

Example:

```{r}
data %>%
  gt() %>%
    tab_header(title = md("What a **nice title**"),
               subtitle = md("Pretty *cool subtitle* too, `isn't it?`"))
```

## &rarr; Footer and source
You can use the `tab_footnote()` function with the `footnote` and `locations` arguments to add references to your tables. 

Example:

```{r}
data %>%
  gt() %>%
  tab_footnote(
    footnote = "Source: James & al., 2020",
    locations = cells_body(columns = Country, rows = 3)
  )
```

## &rarr; Sub-header
The `tab_spanner()` function lets you **group columns** into categories.

Example:

```{r}
data %>%
  gt() %>%
   tab_spanner(
    label = "Number",
    columns = c(GDP, Population)) %>%
  tab_spanner(
    label = "Label",
    columns = c(Country, Capitals)
  )
```


# Gallery of gt examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../350-custom-title-in-gt-table.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Customize title<hr><p class=explanation_portfolio>You can apply some very nice customizations to your gt table</div></div><img alt='Customize title'class='img-fluid imgOfPortfolio'src=../img/graph/350-custom-title-in-gt-table.png></a></div><div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../351-custom-footer-and-references-with-gt-table.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Customize footer <hr><p class=explanation_portfolio>The footer of your gt tables can easily be customized</div></div><img alt='Customize footer 'class='img-fluid imgOfPortfolio'src=../img/graph/351-custom-footer-and-references-with-gt-table.png></a></div></div></div></section>

<!-- Close container -->
</div>
