---
title: "Enhance gt tables with **gtExtras**"
logo: "gtExtras"
descriptionMeta: "The `gtExtras` package in R is an extension of the gt package who provides lots of useful and easy-to-use functions that make gt tables even better!<br/>This post showcases the **key features** of `gtExtras` and provides a set of **table examples** using the package."
descriptionTop: "The `gtExtras` package in R is an extension of the [gt](https://r-graph-gallery.com/package/gt.html) package who provides lots of useful and easy-to-use functions that make gt tables even better!<br/>This post showcases the **key features** of `gtExtras` and provides a set of **table examples** using the package."
documentationLink: "jthomasmock.github.io/gtExtras/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `gtExtras` package in R is an extension of the [gt](https://r-graph-gallery.com/package/gt.html) package, designed to extend what can be done with [gt](https://r-graph-gallery.com/package/gt.html) with more than 60 new functions and features.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers several **table themes**, tools for **plotting** inside a table and other features such as **color gradients** for columns and rows. In total, it's more than **60 functions** that `gtExtras` has to offer!

✍️ **author** &rarr; Thomas Mock

📘 **documentation** &rarr; [github](https://jthomasmock.github.io/gtExtras/)

⭐️ *more than 100 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
#library(gt)
library(gtExtras)
```
</div>
</div>

# Installation
***
To get started with `gtExtras`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("gtExtras")
```

# Basic usage 
***

The `gtExtras` package relies on the [gt](https://r-graph-gallery.com/package/gt.html) package, which means that we need to use the `gt()` function:

```{r}
library(gtExtras)

head(iris) %>% 
  gt() %>% 
  gt_theme_excel()
```

# Key features
***

## &rarr; Theming
`gtExtras` offers a wide variety of **themes** that require just one line of code to use:

Example:

```{r}
library(gtExtras)

head(iris) %>% 
  gt() %>% 
  gt_theme_dot_matrix()
```

<br>

## &rarr; Color rows
You can use a **color palette** for your cells with the `gt_color_rows()` function, and even choose palette from other package like in the following example.

In the `palette` argument, you can also specify a vector of colors and it will **automatically generate the palette**. For example `palette=c("blue", "red")` will generate a palette starting from blue to red.

*Warning: if you select a non-numerical column in `gt_color_rows()`, an error will occur, as it is impossible to define which color level to choose for a categorical value.*

Example:

```{r}
library(gtExtras)

head(iris) %>%
  gt() %>%
  gt_color_rows(
    1:4, # column 1,2,3 and 4
    palette = "RColorBrewer::Spectral"
    )
```

<br>

## &rarr; Highlight specific row
The `gt_highlight_rows()` function allows to **highlight specific rows** with a background color and by putting text in bold.

Example:

```{r}
library(gtExtras)

head(iris, 6) %>%
  gt() %>%
  gt_highlight_rows(
    rows = c(1, 4), # rows to highlight
    target_col = 3, # which column to focus on
    bold_target_only = TRUE, # highlight target column 
    fill='darkred', # background color
    font_color = "#000051", # text color,
    alpha=0.5, # controls color opacity
  )
```

<br>

## &rarr; Plotting
`gtExtras` has multiple functions for plotting inside cells: `gt_plt_sparkline()` for a line chart, `gt_plt_dist()` for distribution charts (density, boxplot, histogram) and **lots of other**!

Example with `gt_plt_dist()`:

```{r}
library(gtExtras)
library(dplyr)

iris %>%
  group_by(Species) %>%
  summarize(SL_data = list(Sepal.Length)) %>%
  gt() %>%
  gt_plt_dist(SL_data)
```






# Gallery of gtExtras examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../368-plotting-in-cells-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Line chart<p class=explanation_portfolio>Line chart in each row of the table.</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/368-plotting-in-cells-with-gtextras-1.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../368-plotting-in-cells-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Distribution chart<p class=explanation_portfolio>Plot histogram, density charts or boxplot in table cells</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/368-plotting-in-cells-with-gtextras-2.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../368-plotting-in-cells-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Summary chart table<p class=explanation_portfolio>Create a table summary with charts with just one line of code!</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/368-plotting-in-cells-with-gtextras-4.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../369-custom-theme-in-gt-table-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Excel theme<p class=explanation_portfolio>Change the default theme to Excel table</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/369-custom-theme-in-gt-table-with-gtextras-1.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../369-custom-theme-in-gt-table-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dot matrix theme<p class=explanation_portfolio>Change the default theme to dot matrix table</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/369-custom-theme-in-gt-table-with-gtextras-2.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=../369-custom-theme-in-gt-table-with-gtextras.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dark theme<p class=explanation_portfolio>Change the default theme to Dark table</div></div><img alt=''class='img-fluid imgOfPortfolio'src=../img/graph/369-custom-theme-in-gt-table-with-gtextras-3.png></a></div></div></div></section>







<!-- Close container -->
</div>
