---
title: "Enhance gt tables with **gtExtras**"
logo: "gtExtras"
descriptionMeta: "The `gtExtras` package in R is an extension of the gt package who provides lots of useful and easy-to-use functions that make gt tables even better!<br/>This post showcases the **key features** of `gtExtras` and provides a set of **table examples** using the package."
descriptionTop: "The `gtExtras` package in R is an extension of the [gt](https://r-graph-gallery.com/package/gt.html) package who provides lots of useful and easy-to-use functions that make gt tables even better!<br/>This post showcases the **key features** of `gtExtras` and provides a set of **table examples** using the package."
documentationLink: "jthomasmock.github.io/gtExtras/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div class="container">

# Quick start
***
The `gtExtras` package in R is an extension of the [gt](https://r-graph-gallery.com/package/gt.html) package, designed to extend what can be done with [gt](https://r-graph-gallery.com/package/gt.html) with more than 60 new functions and features.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers several **table themes**, tools for **plotting** inside a table and other features such as **color gradients** for columns and rows. In total, it's more than **60 functions** that `gtExtras` has to offer!

✍️ **author** &rarr; Thomas Mock

📘 **documentation** &rarr; [github](https://jthomasmock.github.io/gtExtras/)

⭐️ *more than 100 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
#library(gt)
library(gtExtras)
```
</div>
</div>

# Installation
***
To get started with `gtExtras`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("gtExtras")
```

# Basic usage 
***

The `gtExtras` package relies on the [gt](https://r-graph-gallery.com/package/gt.html) package, which means that we need to use the `gt()` function:

```{r}
library(gtExtras)

head(iris) %>% 
  gt() %>% 
  gt_theme_excel()
```

# Key features
***

## &rarr; Theming
`gtExtras` offers a wide variety of **themes** that require just one line of code to use:

Example:

```{r}
library(gtExtras)

head(iris) %>% 
  gt() %>% 
  gt_theme_dot_matrix()
```

## &rarr; Color rows
You can use a **color palette** for your cells with the `gt_color_rows()` function, and even choose palette from other package like in the following example.

In the `palette` argument, you can also specify a vector of colors and it will automatically generate the palette. For example `c("blue", "red")` will generates a palette starting from blue to red.

*Warning: if you select a non-numerical column in `gt_color_rows()`, an error will occur, as it is impossible to define which color level to choose for a categorical value.*

Example:

```{r}
library(gtExtras)

head(iris) %>%
  gt() %>%
  gt_color_rows(
    1:4, # column 1,2,3 and 4
    palette = "RColorBrewer::Spectral"
    )
```

## &rarr; Specify properties of unhighlighted
You can define the style of elements outside your filters using the `unhighlighted_params` argument.

Example:

```{r}
library(gtExtras)

head(iris) %>% 
  gt() %>% 
  gt_theme_excel()
```




# Gallery of gtExtras examples
***

<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=web-line-chart-small-multiple-all-group-greyed-out.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Line chart with small multiple<hr><p class=explanation_portfolio>A line chart with small multiple where all groups are repeated                 (but greyed-out) on each pannel.</div></div><img alt='Line chart with small multiple'class='img-fluid imgOfPortfolio'src=../img/fromTheWeb/we-line-chart-consumer-confidence.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=354-highlight-specific-elements-in-line-chart.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Highlight group on line chart<hr><p class=explanation_portfolio>Easily highlight interesting groups on a line chart using                     the gtExtras package</div></div><img alt='Highlight group on line chart'class='img-fluid imgOfPortfolio'src=../img/graph/354-highlight-specific-elements-in-line-chart-1.png></a></div><div class='portfolio-item col-sm-6 col-md-4'><a class=portfolio-link href=354-highlight-specific-elements-in-line-chart.html><div class=portfolio-hover><div class=portfolio-hover-content><p>highlight groups<hr><p class=explanation_portfolio>Combination of gtExtras, patchwork and hrbrthemes to get                     2 combined highlighted line charts.</div></div><img alt='highlight groups'class='img-fluid imgOfPortfolio'src=../img/graph/354-highlight-specific-elements-in-line-chart-2.png></a></div></div></div></section>







<!-- Close container -->
</div>
