---
title: "Elevate Your ggplot2 Visuals with **hrbrthemes**"
logo: "hrbrthemes"
descriptionMeta: "This post provides an in-depth look at how to enhance your ggplot2 plots using the hrbrthemes package. It offers several reproducible examples with detailed explanations and R code."
descriptionTop: "The `hrbrthemes` package in R is a powerful extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package. It aims to provide **opinionated themes and typography** to make your ggplot2 plots more visually appealing and publication-ready.<br/>This post will delve into the **key features** of `hrbrthemes`, offering a comprehensive set of **graph examples** and explanations."
documentationLink: "github.com/hrbrmstr/hrbrthemes"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Introduction
***
The `hrbrthemes` package in R serves as a valuable extension to the renowned [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package. Created by Bob Rudis, `hrbrthemes` offers a collection of **opinionated themes and typography settings** that aim to elevate the visual appeal of your ggplot2 plots. Whether you are a data scientist, a researcher, or anyone interested in data visualization, this package helps you create **publication-quality** plots with minimal effort.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

The package is particularly useful for those who want to create **publication-quality** plots without spending too much time on customization. It provides a set of themes that are not only visually appealing but also adhere to good visualization practices.

✍️ **author** &rarr; Bob Rudis

📘 **documentation** &rarr; [github](https://github.com/hrbrmstr/hrbrthemes)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(hrbrthemes)
ggplot(mtcars, aes(x=mpg, y=disp)) + geom_point() + theme_ipsum()
```
</div>
</div>

# Installation and Setup
***
Before diving into the features, let's discuss how to get started with `hrbrthemes`. The package can be easily installed from CRAN using the `install.packages` function. Once installed, you can load it into your R session along with ggplot2 to start enhancing your plots.

```{r eval=FALSE}
install.packages("hrbrthemes")
library(hrbrthemes)
library(ggplot2)
```

# In-Depth Look at Features 
***

## Opinionated Themes
One of the standout features of `hrbrthemes` is its collection of opinionated themes. These themes, such as `theme_ipsum()`, `theme_modern_rc()`, and `theme_ft_rc()`, are designed to be visually appealing and publication-ready. They come with a set of pre-configured settings like font sizes, colors, and grid lines that adhere to good visualization practices.

Here's how you can apply the `theme_modern_rc()` to a scatter plot:

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) + geom_point() + theme_modern_rc()
```


# Conclusion
***
The `hrbrthemes` package is a must-have for anyone looking to create visually appealing, publication-quality plots in R. Its opinionated themes and enhanced typography settings make it easier than ever to produce stunning visualizations with ggplot2. Whether you are a seasoned data visualization expert or a beginner, `hrbrthemes` offers something for everyone.

<!-- Close container -->
</div>
