---
title: "Elevate Your ggplot2 Visuals with **hrbrthemes**"
logo: "hrbrthemes"
descriptionMeta: "This post provides an in-depth look at how to enhance your ggplot2 plots using the hrbrthemes package. It offers several reproducible examples with detailed explanations and R code."
descriptionTop: "The `hrbrthemes` package in R is a powerful extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package. It aims to provide **opinionated themes and typography** to make your ggplot2 plots more visually appealing and publication-ready.<br/>This post will delve into the **key features** of `hrbrthemes`, offering a comprehensive set of **graph examples** and explanations."
documentationLink: "github.com/hrbrmstr/hrbrthemes"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Introduction
***
The `hrbrthemes` package in R serves as a valuable extension to the renowned [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package. Created by Bob Rudis, `hrbrthemes` offers a collection of **opinionated themes and typography settings** that aim to elevate the visual appeal of your ggplot2 plots. Whether you are a data scientist, a researcher, or anyone interested in data visualization, this package helps you create **publication-quality** plots with minimal effort.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

The package is particularly useful for those who want to create **publication-quality** plots without spending too much time on customization. It provides a set of themes that are not only visually appealing but also adhere to good visualization practices.

✍️ **author** &rarr; Bob Rudis

📘 **documentation** &rarr; [github](https://github.com/hrbrmstr/hrbrthemes)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(hrbrthemes)
ggplot(mtcars, aes(x=mpg, y=disp)) + geom_point() + theme_ipsum()
```
</div>
</div>

# Installation and Setup
***
Before diving into the features, let's discuss how to get started with `hrbrthemes`. The package can be easily installed from CRAN using the `install.packages` function. Once installed, you can load it into your R session along with ggplot2 to start enhancing your plots.

```{r eval=FALSE}
install.packages("hrbrthemes")
library(hrbrthemes)
library(ggplot2)
```

# Basic usage

The `hrbrthemes` package offers a collection of opinionated **themes** that can be applied to your ggplot2 plots. These themes are designed to be visually appealing and publication-ready. They come with a set of **pre-configured settings** like font sizes, colors, and grid lines that adhere to good visualization practices.

Here's how you can apply the `theme_ipsum()` to a scatter plot:

```{r} 
ggplot(mtcars, aes(x=mpg, y=disp)) + geom_point() + theme_ipsum()
```

# In-Depth Look at Features 
***

## Opinionated Themes
One of the standout features of `hrbrthemes` is its collection of opinionated themes. These themes, such as `theme_ipsum()`, `theme_modern_rc()`, and `theme_ft_rc()`, are designed to be visually appealing and publication-ready. They come with a set of pre-configured settings like font sizes, colors, and grid lines that adhere to good visualization practices.

Here's how you can apply the `theme_modern_rc()` to a scatter plot:

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) + geom_point() + theme_modern_rc()
```


## Typography Settings
Another key feature of `hrbrthemes` is its typography settings. These settings allow you to customize the fonts and font sizes of your plots. For example, you can change the font family of your plot to `serif` and increase the font size of the axis labels to 14 points.

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) +
  geom_point() +
  theme_ipsum_rc(base_family = "serif", base_size = 14)
```

## Grid Lines
`hrbrthemes` also offers a set of grid line settings that allow you to **customize the grid lines** of your plots. For example, you can change the color of the grid lines to `lightblue` and increase the thickness of the grid lines to 1.2 points. We have to use the `theme()` function to apply these settings, to **both the major and minor** grid lines.

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) +
  geom_point() +
  theme_ipsum() +
  theme(panel.grid.major = element_line(size = 1.2, color = "lightblue")) +
  theme(panel.grid.minor = element_line(size = 1.2, color = "lightblue"))
```

## Custom Colors
The `hrbrthemes` package also offers a set of **custom colors** that can be used in your plots. You can use functions such as `scale_color_ipsum()` and `scale_fill_ipsum()` to apply these colors to your plots. For example, you can use the `scale_color_ipsum()` function to apply the `ipsum_blue()` color to the points in a scatter plot.

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) +
  geom_point(aes(color = factor(cyl))) +
  scale_color_ipsum() +
  theme_ipsum()
```

## Percent and comma scales
`hrbrthemes` also offers a set of **percent and comma scales** that can be used in your plots. These scales allow you to format the axis labels of your plots to include a percent sign or a comma. For example, you can use the `scale_x_percent()` function to format the x-axis labels of a scatter plot to include a percent sign.

```{r}
ggplot(mtcars, aes(x=mpg, y=disp)) +
  geom_point() +
  scale_x_percent() +
  theme_ipsum()
```


# Gallery of hrbrthemes examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../14-venn-diagramm.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Simple Venn Diagramm<hr><p class=explanation_portfolio>All possible logical relationships between several sets of data thanks to the VennDiagram package.</div></div><img alt="Venn Diagramm"class="img-fluid imgOfPortfolio"src=../img/graph/14-venn-diagramm2.png></a></div><div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../135-stacked-density-graph.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Density chart<hr><p class=explanation_portfolio>Density graph with multiple groups.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/135-stacked-density-graph1.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><a class=portfolio-link href=../48-grouped-barplot-with-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Grouped barplot<hr><p class=explanation_portfolio>Barplot with both groups and subgroups and a nice hrbrthemes theme.</div></div><img alt="Grouped barplot"class="img-fluid imgOfPortfolio"src=../img/graph/48-grouped-barplot-with-ggplot24.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><div class=portfolio-item><a class=portfolio-link href=../connected_scatterplot_ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Connected Scatter plot<hr><p class=explanation_portfolio>A connected scatter plot can be more informative than an ordinary scatter plot.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/connected_scatterplot_ggplot22.png></a></div><div class="portfolio-item col-sm-6 col-md-8"><div class=portfolio-item><a class=portfolio-link href=../341-stacked-barplot-with-negative-values.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Stacked Barplot<hr><p class=explanation_portfolio>Negative values in a stacked barplot for better visualization.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/341-stacked-barplot-with-negative-values.png></a></div></div></div></div></section>



# Conclusion
***
The `hrbrthemes` package is a must-have for anyone looking to create visually appealing, publication-quality plots in R. Its opinionated themes and enhanced typography settings make it easier than ever to produce stunning visualizations with ggplot2. Whether you are a seasoned data visualization expert or a beginner, `hrbrthemes` offers something for everyone.


<!-- Close container -->
</div>
