---
title: "Network Analysis with **igraph**"
logo: "igraph"
descriptionMeta: "This post explains how to perform network analysis and visualization using the igraph package in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `igraph` package in R is a powerful tool for **network analysis and visualization**. It provides a wide range of functions for creating, manipulating, and analyzing graphs and networks.<br/>This post showcases the **key features** of `igraph` and provides a set of **graph examples** using the package."
documentationLink: "r.igraph.org/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Quick start
***
The `igraph` package in R is a comprehensive toolkit for **network analysis and visualization**. It provides functionality for creating, manipulating, and analyzing various types of graphs and networks.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a wide range of **algorithms** and **visualization techniques** for working with network data.

✍️ **author** &rarr; Gábor Csárdi, Kirill Müller

📘 **documentation** &rarr; [igraph.org](https://r.igraph.org/)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(igraph)
# Create a simple graph
g <- make_ring(10)
# Plot the graph
plot(g, vertex.size = 30, vertex.label = 1:10, vertex.color = "lightblue", edge.arrow.size = 0.5)
```
</div>
</div>

# Installation
***
To get started with `igraph`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("igraph")
```

# Vocabulary
***

Key components of a graph in `igraph`:

- **Vertex (Node)**: Basic unit of a graph (e.g., a person in a social network)
- **Edge (Link)**: Connection between two vertices (e.g., a friendship)
- **Directed Edge**: One-way connection (e.g., following on Twitter/X)
- **Undirected Edge**: Two-way connection (e.g., Facebook friendship)
- **Weighted Edge**: Edge with a numerical value (e.g., distance between cities)
- **Degree**: Number of edges connected to a vertex
- **Path**: Sequence of edges connecting two vertices
- **Subgraph**: Smaller graph within a larger graph
- **Connected Component**: Subgraph where all vertices are connected


# Basic usage 
***

The `igraph` package allows you to **create**, **manipulate**, and **analyze** graphs and networks. 

Here's a basic example where:

- `make_ring(5)` creates a circular graph with 5 vertices
- `print(g)` displays basic information about the graph
- `plot(g)` visualizes the graph

```{r}
library(igraph)

# Create a simple graph
g <- make_ring(5)

# Print basic information about the graph
print(g)

# Plot the graph
plot(g)
```


# Key features
***

<br>

## &rarr; Creating Graphs

You can create graphs using various functions like `make_graph()`, `sample_gnm()`, or `graph_from_edgelist()`.

Example:

- `sample_gnm(10, 20)` creates a **random graph** with 10 vertices and 20 edges
- The resulting graph is visualized using `plot(g)`


```{r}
# Create a random graph
g <- sample_gnm(10, 20)
plot(g)
```

<br>

## &rarr; Graph Analysis

`igraph` provides numerous functions for analyzing graph properties.

Example:

- `degree(g)` calculates the **degree centrality** for each vertex in the graph (number of connections each vertex has)

```{r}
# Calculate degree centrality
deg <- degree(g)
print(deg)
```

<br>

## &rarr; Community Detection

You can detect communities in networks using various algorithms.

Example:

- `cluster_louvain(g)` applies the **Louvain community detection** algorithm to the graph
- `plot(comm, g)` **visualizes the graph** with nodes colored by their detected communities


```{r}
# Detect communities using the Louvain method
comm <- cluster_louvain(g)
plot(comm, g)
```

<br>

## &rarr; Graph Layouts

`igraph` offers various layout algorithms for visualizing graphs.

Example:

- `layout_with_fr(g)` computes a **force-directed layout** for the graph
- The layout is then used in `plot(g, layout = layout)` to visualize the graph

```{r}
# Use a force-directed layout
layout <- layout_with_fr(g)
plot(g, layout = layout)
```

<br>

## Customizing Graph Appearance

- We create a **scale-free** network using `sample_pa()`
- Vertex sizes are set **proportional** to their degrees
- Vertex colors are assigned **based on degree** using a rainbow palette
- Edge widths are set based on **edge betweenness centrality**
- The graph is plotted with **curved edges** and **without vertex labels**


```{r}
# Create a scale-free network
g <- sample_pa(100, power = 1, m = 1, directed = FALSE)

# Customize the graph appearance
V(g)$size <- degree(g) * 2 # Vertex size based on degree
V(g)$color <- rainbow(100)[cut(degree(g), breaks = 100)] # Vertex color based on degree
E(g)$width <- edge_betweenness(g) * 0.03 # Edge width based on betweenness

# Plot the customized graph
plot(g,
  vertex.label = NA, # Remove vertex labels
  edge.curved = 0.2, # Add curve to edges
  layout = layout_with_fr(g)
)
```


<br>

## Visualizing Weighted Graphs

Explanation:


- We create a **weighted graph** from a data frame of edges
- Edge widths in the plot are **proportional** to their weights
- Vertices are arranged in a **circular layout**

```{r}
# Create a weighted graph
edges <- data.frame(
  from = sample(1:10, 20, replace = TRUE),
  to = sample(1:10, 20, replace = TRUE),
  weight = runif(20)
)
g <- graph_from_data_frame(edges, directed = FALSE)

# Plot the weighted graph
plot(g,
  edge.width = E(g)$weight * 8, # Edge width based on weight
  vertex.size = 20,
  vertex.color = "lightblue",
  vertex.label.color = "black",
  edge.color = "gray50",
  layout = layout_in_circle(g)
)
```


<br>

## Interactive Graph Visualization

For interactive graph visualization, you can use the `visNetwork` package, which integrates well with `igraph`:

- We create a graph using `sample_pa()`
- Node and edge data are prepared for `visNetwork`
- The interactive visualization allows zooming, panning, and node selection

```{r}
library(visNetwork)

# Create a graph
g <- sample_pa(50, power = 1, m = 1, directed = FALSE)

# Prepare node and edge data
nodes <- data.frame(
  id = 1:vcount(g),
  label = 1:vcount(g),
  size = degree(g) * 3
)
edges <- as_data_frame(g, what = "edges")

# Create an interactive visualization
interative_graph <- visNetwork(nodes, edges) %>%
  visIgraphLayout() %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)

htmltools::save_html(interative_graph, "../HtmlWidget/interactive-network.html")
```

<center><iframe src="../HtmlWidget/interactive-network.html" height="500" width="100%" style="border:none;"></iframe></center>

<br>

# Gallery of igraph examples
***

The Gallery is filled with **chart examples** that uses `igraph`! Browse them below:

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../247-network-chart-layouts.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Network layout</p> <hr> <p class=explanation_portfolio>Discover the different layout possibles</p> </div> </div><img alt="igraph different layout" class="img-fluid imgOfPortfolio" src=../img/graph/247-network-chart-layouts.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../248-igraph-plotting-parameters.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Customize style</p> <hr> <p class=explanation_portfolio>Learn how to customize your igraph plots</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/248-igraph-plotting-parameters.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../249-igraph-network-map-a-color.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Colors</p> <hr> <p class=explanation_portfolio>Discover how to change the colors of your igraph plots</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/249-igraph-network-map-a-color.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../251-network-with-node-size-based-on-edges-number.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Node size</p> <hr> <p class=explanation_portfolio>Custom node size in your igraph plot</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/251-network-with-node-size-based-on-edges-number.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../257-input-formats-for-network-charts.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Input format</p> <hr> <p class=explanation_portfolio>Learn what are the right input format for an igraph chart</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/257-input-formats-for-network-charts.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../309-intro-to-hierarchical-edge-bundling.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hierarchical Edge Bundling</p> <hr> <p class=explanation_portfolio>Use igraph to create hierarchical edge bundling</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/309-intro-to-hierarchical-edge-bundling.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../309-intro-to-hierarchical-edge-bundling.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hierarchical Edge Bundling</p> <hr> <p class=explanation_portfolio>Use igraph to create hierarchical edge bundling</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/309-intro-to-hierarchical-edge-bundling2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../309-intro-to-hierarchical-edge-bundling.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hierarchical Edge Bundling</p> <hr> <p class=explanation_portfolio>Use igraph to create hierarchical edge bundling</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/309-intro-to-hierarchical-edge-bundling3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../310-custom-hierarchical-edge-bundling.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Custom Hierarchical Edge Bundling</p> <hr> <p class=explanation_portfolio>Use igraph to customize hierarchical edge bundling</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/310-custom-hierarchical-edge-bundling3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../313-basic-circle-packing-with-several-levels.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circle packing</p> <hr> <p class=explanation_portfolio>Learn the basics of circle packing</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/313-basic-circle-packing-with-several-levels.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../314-custom-circle-packing-with-several-levels.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Custom circle packing</p> <hr> <p class=explanation_portfolio>Customize circle packing charts</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/314-custom-circle-packing-with-several-levels2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../335-custom-ggraph-dendrogram.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Dendrogram with igraph</p> <hr> <p class=explanation_portfolio>Learn how to create a dendrogram with igraph</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/335-custom-ggraph-dendrogram.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../339-circular-dendrogram-with-ggraph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular dendrograph</p> <hr> <p class=explanation_portfolio>Discover how to create a circular dendrogram</p> </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/339-circular-dendrogram-with-ggraph.png> </a></div> </div> </div> </section>


<!-- Close container -->
</div>