---
title: "Beaufitul HTML tables with **kableExtra**"
logo: "kableExtra"
descriptionMeta: "This post explains how to create and style tables using the kableExtra package. It provides several reproducible examples with explanation and R code."
descriptionTop: "This post explains how to create and style tables using the `kableExtra`.<br/>This post showcases the **key features** of `kableExtra` and provides a set of **table examples** using the package."
documentationLink: "github.com/haozhu233/kableExtra"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `kableExtra` package offers style features from **html and css** with syntax based on the pipe symbol `%>%`.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of features that relies on `kable()` output to create fully customized tables.

✍️ **author** &rarr; Hao Zhu

📘 **documentation** &rarr; [github](https://github.com/haozhu233/kableExtra)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(kableExtra)
data(iris)

iris[1:8, ] %>%
  kbl(align = "c") %>%
  kable_styling(bootstrap_options = "striped", font_size = 16) %>%
  kable_paper(full_width = F) %>%
  column_spec(2, color = spec_color(iris$Sepal.Width[1:8]),
              link = "https://r-graph-gallery.com/table.html") %>%
  column_spec(4, color = "white",
              background = spec_color(iris$Sepal.Length[1:8], end = 0.7),
              popover = paste("am:", iris$Petal.Width[1:8]))
```
</div>
</div>

# Installation
***
To get started with `kableExtra`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("kableExtra")
```

# Basic usage 
***

The `kableExtra` package has multiple themes: `kable_paper`, `kable_classic`, `kable_classic_2`, `kable_minimal`, `kable_material` and `kable_material_dark`.

Here's a basic example with the dark one:

<div class = "row">
<div class = "col-md-12  col-sm-12">
```{r}
library(kableExtra)
data(iris)

head(iris) %>%
  kbl() %>%
  kable_material_dark(c("striped", "hover"))
```
</div>
</div>

# Key features
***


## &rarr; Colors
You can personalized every color aspect of the table. In this example we show **5 different ways** to apply a color, a gradient of colors or a list of colors to kable columns. 

Example:

```{r}
library(kableExtra)
data(iris)
df = iris[order(iris$Petal.Length, decreasing = T),] # sort by petal length value

# gradient color
colfunc = colorRampPalette(c("purple", "orange"))
n_color = 8
colors = colfunc(n_color) # generate 8 colors, from purple to orange

# generate the table
head(df, 8) %>%
  kbl(align = "c") %>% # center the columns
  kable_paper(c("striped", "hover")) %>% # add strips and hover
  column_spec(2, color='red') %>%
  column_spec(3, background=colors) %>%
  column_spec(4, color=colors) %>%
  column_spec(5, background='lightblue') %>%
  column_spec(6, color=c("pink","pink","pink","yellow","pink","green","pink","pink"))
```


## &rarr; Images and Link
`kableExtra` allows you to add images and url links directly into columns.

Example:

```{r}
library(kableExtra)

df_img = data.frame(name = c("Grace Wahba","Ragnar Frisch"),
                    field = c("Machine Learning", "Econometrics"),
                    image = ""
                    )
df_img %>%
  kbl(booktabs = T, align = "c") %>%
  kable_paper(full_width = T) %>%
  column_spec(1, link = c("https://en.wikipedia.org/wiki/Grace_Wahba",
                          "https://en.wikipedia.org/wiki/Ragnar_Frisch")) %>%
  column_spec(2, bold=TRUE) %>%
  column_spec(3, image = spec_image(c("https://upload.wikimedia.org/wikipedia/commons/thumb/3/35/Grace_Wahba_1986.jpg/440px-Grace_Wahba_1986.jpg",
                                      "https://upload.wikimedia.org/wikipedia/commons/a/a1/Uio_frisch_2006_0025.jpg"),
                                    200, 200) # dimensions of the images
              )
```


## &rarr; Chart in cells
The `kableExtra` package proposes 3 plotting tools: `spec_hist`, `spec_boxplot`, and `spec_plot`. We split the iris dataset according to the `Species` column. 

Example:

```{r}
library(kableExtra)
data(iris)
df = data.frame(Species = c('Setosa', 'Versicolor', 'Virginica'),
                boxplots = "", histograms = "", plots1 = "", plots2 = "")
sepal_length_list = split(iris$Sepal.Length, iris$Species)

df %>%
  kbl(booktabs = TRUE) %>%
  kable_paper(full_width = TRUE) %>%
  column_spec(2, image = spec_boxplot(sepal_length_list)) %>%
  column_spec(3, image = spec_hist(sepal_length_list)) %>%
  column_spec(4, image = spec_plot(sepal_length_list, same_lim = TRUE)) %>%
  column_spec(5, image = spec_plot(sepal_length_list, polymin = 5))
```




# Gallery of kableExtra examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=web-dumbbell-chart-with-a-gap-column.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Custom colors with kableExtra<hr><p class=explanation_portfolio>This post explains how to custom colors, with unique color or gradient of colors, using the kableExtra package</div></div><img alt="Custom colors with kableExtra"class="img-fluid imgOfPortfolio"src=../img/graph/362-customize-colors-in-kableExtra-3.png></a></div><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=line-chart-dual-Y-axis-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Add image and link<p class=explanation_portfolio>A few examples showing how to add images and links into your kableExtra tables</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/363-add-images-and-links-in-kableExtra-1.png></a></div></div></div></div></section>

<!-- Close container -->
</div>
