---
title: "Combine Multiple Plots with **patchwork**"
logo: "patchwork"
descriptionMeta: "This post explains how to combine multiple ggplot2 plots into a single layout using the patchwork package. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `patchwork` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **combining multiple plots** into a single layout.<br/>This post showcases the **key features** of `patchwork` and provides a set of **graph examples** using the package."
documentationLink: "github.com/thomasp85/patchwork"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `patchwork` package in R is an extension of the [ggplot2](https://r-graph-gallery.com/ggplot2-package.html) package, designed to simplify the process of **combining multiple plots** into a single layout.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of **operators** that make it easy to specify complex layouts in a human-readable way.

✍️ **author** &rarr; Thomas Lin Pedersen

📘 **documentation** &rarr; [github](https://thomasp85.github.io/patchwork/)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(ggplot2)
library(patchwork)
p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p1 + p2
```
</div>
</div>

# Installation
***
To get started with `patchwork`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("patchwork")
```

# Basic usage 
***

The `patchwork` package allows you to combine multiple plots using simple **arithmetic operators** like `+`, `/`, and `|`.

Here's a basic example:

```{r}
p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p1 + p2
```

# Key features
***

## &rarr; Combining Plots
You can combine multiple plots using the `+` operator.

Example:

```{r}
p1 + p2 + p1
```

## &rarr; Arranging Plots
Use the `/` operator to stack plots on top of each other.

Example:

```{r}
p1 / p2
```

## &rarr; Nesting Plots
You can nest plots using parentheses.

Example:

```{r}
(p1 + p2) / p1
```

## &rarr; Adding Titles and Annotations
`patchwork` allows you to add titles, subtitles, and captions to your combined plots.

Example:

```{r}
p1 + p2 + plot_annotation(title = 'My Combined Plot')
```

# Gallery of patchwork examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class="portfolio-item col-sm-6 col-md-6"><a class=portfolio-link href=../web-dumbbell-chart-with-a-gap-column.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dumbbell plot with gap section<hr><p class=explanation_portfolio>An extended dumbbell plot originally published by the Pew research center. Clean design with a gap section on the right hand side.</div></div><img alt="Dumbell chart with gap column on the right"class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-dumbbell-chart-with-a-gap-column-2.png></a></div><div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../line-chart-dual-Y-axis-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Dual Y axis<hr><p class=explanation_portfolio>A few examples showing how to use the sec.axis argument to add a second Y axis.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/graph/line-chart-dual-Y-axis-ggplot2_2.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><a class=portfolio-link href=../web-circular-lollipop-plot-with-ggplot2.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Circular lollipop chart<hr><p class=explanation_portfolio>A circular lollipop chart with customized layout, in circle legend and groups.<br><br>A work by <b>Cédric Scherer</b> for the <b>TidyTuesday</b> challenge.</div></div><img alt="Circular lollipop chart with R and ggplot2"class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-circular-lollipop-plot-with-ggplot2-square.png></a></div><div class="portfolio-item col-sm-6 col-md-5"><div class=portfolio-item><a class=portfolio-link href=../web-lineplots-and-area-chart-the-economist.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Line and Area charts by The Economist<hr><p class=explanation_portfolio>Learn how to mimick the style of The Economist with libraries like ggplot2, patchwork, grid, ggtext and more.</div></div><img alt=""class="img-fluid imgOfPortfolio"src=../img/fromTheWeb/web-lineplots-and-area-chart-the-economist.png></a></div></div></div></div></section>

<!-- Close container -->
</div>
