---
title: "Create interactive table with **reactable**"
logo: "reactable"
descriptionMeta: "This post explains how to create interactive tables with the reactable package in R. It provides a set of reproducible examples and explanations for each feature."
descriptionTop: "The `reactable` package in R is designed to simplify the process of **creating interactive tables**.<br/>This post showcases the **key features** of `reactable` and provides a set of **table examples** using the package."
documentationLink: "github.com/glin/reactable"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set( warning=FALSE, message=FALSE)
```

<div class="container">

# Quick start
***
The `reactable` package in R is a powerful and flexible tool for creating **interactive and customizable** tables in web applications and R Shiny dashboards.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It mainly provides one function: `reactable()` &rarr; to create a table

✍️ **author** &rarr; Greg Lin

📘 **documentation** &rarr; [github](https://glin.github.io/reactable/index.html)

⭐️ *more than 500 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library("reactable")
reactable(mtcars)
```
</div>
</div>

# Installation
***
To get started with `reactable`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("reactable")
```

# Basic usage 
***

The `reactable` package provides one main function: `reactable()`. It takes a data frame as input and returns an interactive table.

Here's a basic example:

```{r warning=FALSE, message=FALSE}
reactable(mtcars,
          defaultColDef = colDef(minWidth = 100,
                                 width = 150,
                                 style = list(padding = "0 10px"), html=TRUE
))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="../HtmlWidget/reactable1.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Key features
***

## &rarr; Searching
You can enable searching by setting the `searchable` argument to `TRUE`.

Example:

```{r}
reactable(mtcars, searchable = TRUE)
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="../HtmlWidget/reactable2.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

## &rarr; Filtering
You can enable filtering by setting the `filterable` argument to `TRUE`.

Example:

```{r}
reactable(mtcars, filterable = TRUE, minRows = 10)
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="../HtmlWidget/reactable3.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

## &rarr; Grouping and Aggregating
You can enable **grouping and aggregating** by setting the `groupBy` argument to the name of a column and by adding the `aggregate` argument in each column definition.

Example:

```{r}
reactable(iris,
          groupBy = "Species", # Group by the Species column
          columns = list(
            Sepal.Length = colDef(aggregate = "max"), # max value
            Sepal.Width = colDef(aggregate = "mean", format = colFormat(digits = 1)), # mean value
            Petal.Length = colDef(aggregate = "unique"), # unique values
            Petal.Width = colDef(aggregate = "min") # min value
          )
        )
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="../HtmlWidget/reactable4.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

## &rarr; Theming
You can customize the appearance of the table by setting the `theme` argument to a `reactableTheme` object. You can also use the `striped` or `highlight` arguments to **add stripes** or **highlight** the hovered row.

Example:

```{r}
reactable(
  iris[1:30, ],
  striped = TRUE,
  highlight = TRUE,
  theme = reactableTheme(
    borderColor = "#dfe2e5",
    stripedColor = "#f6f8fa",
    highlightColor = "red",
    cellPadding = "8px 12px",
    style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"),
    searchInputStyle = list(width = "100%")
  ))
```

<div class = "col-md-6  col-sm-12">

<center><iframe src="../HtmlWidget/reactable5.html" height="700" width="700" style="border:none;"></iframe></center>

</div>

# Gallery of reactable examples
***
<section class=bg id=portfolio style=padding-top:10px><div class=container><div class=row id=portfolio-items><div class='portfolio-item col-sm-6 col-md-6'><a class=portfolio-link href=../350-introduction-to-reactable.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Custom colors with table rows with reactable<hr><p class=explanation_portfolio>How to customize the colors of rows, borders and when mouse is over</div></div><img alt='Custom colors with table rows with reactable'class='img-fluid imgOfPortfolio'src=../img/graph/350-introduction-to-reactable-2.png></a></div><div class='portfolio-item col-sm-8 col-md-5'><a class=portfolio-link href=../351-column-formatting-in-reactable.html><div class=portfolio-hover><div class=portfolio-hover-content><p>Columns formatting<hr><p class=explanation_portfolio>How to format columns with reactable</div></div><img alt='Column formatting'class='img-fluid imgOfPortfolio'src=../img/graph/351-column-formatting-in-reactable.png></a></div></div></div></section>

<!-- Close container -->
</div>