---
title: "Plot and manipulate geographic features with **sf**"
logo: "sf"
descriptionMeta: "The sf package in R specializes in geospatial visualization, offering streamlined access to geographic features like points, lines, and polygons, and more generally to create maps."
descriptionTop: "The `sf` package in R specializes in **geospatial visualization**, offering streamlined access to **geographic features** like points, lines, and polygons, and more generally to create maps."
documentationLink: "r-spatial.github.io/sf/index.html"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  fig.align='center'
)
```

<div class="container">

# Quick start
***

In just a few lines of code, you can create stunning visualizations with the `sf` package in R. What's more, `sf` lets you create maps from almost **any data format**, such as [shapefiles](../168-load-a-shape-file-into-r.html) or [geojson](../325-background-map-from-geojson-format-in-r.html). You can learn more about geographical data in the [map section](../map.html).

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

This package offers a large variety of functions to manipulate and visualize [geographic data](../map.html). The main steps usually are:

- create a **geospatial object** with `st_read()` or `read_sf()`

- **filter**, **transform**, and **manipulate** the object for your needs

- **plot the object** with `plot()` or with [ggplot2](../ggplot2-package.html)

✍️ **author** &rarr; Edzer Pebesma

📘 **documentation** &rarr; [github](https://r-spatial.github.io/sf/index.html)

⭐️ *more than 1000 stars on github*
</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
plot(nc[1])
```
</div>
</div>

# Installation
***
To get started with `sf`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("sf")
```

# Basic usage 
***

The `sf` package is shipped with the `nc` dataset, which contains the boundaries of North Carolina.

Here's a basic example of how to plot the first feature of the `nc` dataset:

```{r}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
plot(st_geometry(nc), col = "lightblue", border = "black", lwd = 1.2, main = "North Carolina")
```

# Key features
***

## &rarr; Choropleth Maps

The `sf` package allows you to create [choropleth maps](../choropleth-map.html) with ease.

Simply use the `geom_sf()` function with [ggplot2](../ggplot2-package.html) to create a [choropleth map](../choropleth-map.html) of your geographic data.

You can **customize** the map by specifying the `fill` aesthetic and the **color palette** using the `scale_fill_viridis_c()` function

Example:

```{r}
library(sf)
library(ggplot2)

# Read the shapefile
nc = st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)

# Create the plot using ggplot2
ggplot() +
  geom_sf(data = nc, aes(fill = AREA)) +
  scale_fill_viridis_c() +
  ggtitle('Choropleth Map of North Carolina') +
  theme_minimal()
```

<br>

## &rarr; Graticules

Graticules are lines of latitude and longitude that help you orient yourself on a [map](../map.html). You can add graticules to your map using the `plot()` function.

Example:

```{r fig.width=8}
library(maps)
library(sf)
library(ggplot2)
library(patchwork)

# create a map of the USA
usa = st_as_sf(map('usa', plot = FALSE, fill = TRUE))
laea = st_crs("+proj=laea +lat_0=30 +lon_0=-95") # Lambert equal area projection
usa2 <- st_transform(usa, laea)

# create ggplot objects with graticules
p1 = ggplot(data = usa) + 
  geom_sf() + 
  theme_minimal()

p2 = ggplot(data = usa2) + 
  geom_sf() + 
  theme_minimal()

# combine plots with patchwork
p1 + p2
```

<br>

## &rarr; Compute and plot centroids

You can compute and plot the **centroids** of your geographic features using the `st_centroid()` function.

*Caution: Centroids serve merely as an approximation of a polygon's "center" and may not always provide meaningful or accurate insights.*

Example:

```{r}
library(sf)
library(ggplot2)

# Read the shapefile
nc = st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
ncg = st_geometry(nc)
cntrd = st_centroid(ncg)

# Create the plot using ggplot2
ggplot() +
  geom_sf(data = ncg, fill = NA, color = 'grey') +
  geom_sf(data = cntrd, color = 'red', shape = 8, size = 2) +
  ggtitle('Centroids of North Carolina') +
  theme_minimal()
```

<br>

## &rarr; What's more?

While `sf` is not a data visualization package per se, it provides extensive functionalities for handling and [visualizing geographic data](../map.html). This includes:

- **Creating a geospatial object** using `st_read()` or `read_sf()` from virtually any file format.
- **Filtering**, **transforming**, and **manipulating** the geospatial object to suit your specific requirements.
- Enhancing your visualizations by integrating with other packages, such as [ggplot2](../ggplot2-package.html).

Explore the capabilities of `sf` below!


# Gallery of sf examples
***

The gallery is **filled** with examples of how to use the `sf` package to create stunning visualizations. Dive in and discover the possibilities!

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../168-load-a-shape-file-into-r.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Basic world map <hr> <p class=explanation_portfolio>Most simple world map with R </div> </div><img alt="venn diagram" class="img-fluid imgOfPortfolio" src=../img/graph/168-load-a-shape-file-into-r.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../331-basic-cartogram.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Cartogram of Africa <hr> <p class=explanation_portfolio>How to build a cartogram of a specific region </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/331-basic-cartogram4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../331-basic-cartogram.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Cartogram of Africa <hr> <p class=explanation_portfolio>How to build a cartogram of a specific region </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/331-basic-cartogram3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../169-170-basic-manipulation-of-shapefiles.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Label in each sub region <hr> <p class=explanation_portfolio>Discover how to add a label in selected sub regions </div> </div><img alt="pie plot" class="img-fluid imgOfPortfolio" src=../img/graph/169-170-basic-manipulation-of-shapefiles3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../328-hexbin-map-of-the-usa.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hexbin map   <hr> <p class=explanation_portfolio>Learn how to build a hexbin map of the USA </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/328-hexbin-map-of-the-usa2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../328-hexbin-map-of-the-usa.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hexbin + Choropleth <hr> <p class=explanation_portfolio>How to build a hexbin map with a choropleth map </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/328-hexbin-map-of-the-usa3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"> <a class=portfolio-link href=../328-hexbin-map-of-the-usa.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hexbin map <hr> <p class=explanation_portfolio>Advanced example of a hexbin map </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/328-hexbin-map-of-the-usa4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"> <a class=portfolio-link href=../332-hexbin-chloropleth-cartogram.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hexbin + Chloropleth + Cartogram <hr> <p class=explanation_portfolio>How to build a hexbin map with a choropleth map AND a cartogram </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/332-hexbin-chloropleth-cartogram4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../web-choropleth-map-lego-style.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Choropleth map <hr> <p class=explanation_portfolio>Discover how to build a choropleth map with a lego style </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/web-choropleth-map-lego-style.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../web-dorling-cartogram-with-R.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Bubble map   <hr> <p class=explanation_portfolio>Discover to build a bubble map with R </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/web-dorling-cartogram-with-R.png> </a></div> <div class="portfolio-item col-sm-6 col-md-8"> <a class=portfolio-link href=../web-valued-dots-map-bertin.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Dot map   <hr> <p class=explanation_portfolio>France map in the style of Jacques Bertin </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/fromTheWeb/web-valued-dots-map-bertin.png> </a></div> </div> </div> </div> </section>

<!-- Close container -->
</div>
