---
title: "Create Beautiful Thematic Maps with **tmap**"
logo: "tmap"
descriptionMeta: "This post explains how to create thematic maps using the tmap package in R. It provides several reproducible examples with explanation and R code."
descriptionTop: "The `tmap` package in R is designed for creating **thematic maps**, allowing users to visualize spatial data in an intuitive and flexible way.<br/>This post showcases the **key features** of `tmap` and provides a set of **map examples** using the package."
documentationLink: "r-tmap.github.io/tmap/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.align = "center")
```

<div class="container">

# Quick start
***

The `tmap` package in R is designed for creating **thematic maps**, allowing users to visualize spatial data in an intuitive and flexible way.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">

It offers a set of **functions** that make it easy to create complex maps with a consistent and user-friendly syntax, similar to [ggplot2](../ggplot2-package.html). It's built on top of [sf](sf.html) and [leaflet](../179-show-a-map-with-leaflet-r.html).

✍️ **author** &rarr; Martijn Tennekes

📘 **documentation** &rarr; [website](https://r-tmap.github.io/tmap/)

⭐️ *more than 800 stars on github*
</div>

<div class = "col-md-7  col-sm-12">

```{r, echo=FALSE, out.width = "85%", fig.align='center'}
library(tmap)
data("World")
tm_shape(World) +
  tm_polygons("income_grp", palette = "Blues", title = "Income Group") +
  tm_layout(frame = FALSE)
```

</div>
</div>

# Installation
***

To get started with `tmap`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("tmap")
```

# Basic usage 
***

The `tmap` package uses a layered approach to building maps, similar to [ggplot2](../ggplot2-package.html). You start with `tm_shape()` to define the data, then add layers with various `tm_*()` functions.

Here's a basic example:

```{r}
library(tmap)
data("World")

tm_shape(World) +
  tm_polygons("pop_est", style = "quantile", title = "Population")
```

# Key features
***

## &rarr; Draw borders

You can create **minimalist maps** using the `tm_borders()` function:

```{r}
library(tmap)
data("World")

tm_shape(World) +
  tm_borders()
```

<br>


## &rarr; Multiple Layers

You can **add multiple layers** to your map using the `+` operator. The following example will add the `name` of each country with a `size` proportional to the `AREA` column:

```{r}
library(tmap)
data("World")

tm_shape(World) +
  tm_polygons("income_grp", palette = "Blues") +
  tm_text("name", size = "AREA")
```

<br>

## &rarr; Faceted Maps

Use `tm_facets()` to create **small multiples based on a variable**. Here, we use the `continent` column to create a facet:

```{r}
library(tmap)
data("World")

tm_shape(World) +
  tm_polygons("gdp_cap_est", style = "quantile") +
  tm_facets("continent", free.coords = FALSE)
```

<br>


## &rarr; Style Customization

You can apply any custom style to your `tmap` maps. In the following example, we:

- **Color and Style**:
  - Uses a reversed RdYlBu palette for life expectancy
  - Applies the "pretty" style for class intervals
  - Sets white borders with reduced opacity for countries

- **Text**:
  - Overlays country codes (ISO A3) on each country

- **Legend**:
  - Moves the legend outside the map
  - Customizes legend appearance (frame, background color, text size)

- **Layout**:
  - Adds a title with custom positioning and size
  - Sets a background color for the entire plot
  - Adds a frame around the map
  - Customizes inner margins

- **Additional Elements**:
  - Includes a compass and scale bar
  - Adds credits for the data source

- **Overall Style**:
  - Applies the "natural" built-in style as a base

```{r}
library(tmap)
library(dplyr)

# Prepare the data
data("World")
world_data <- World %>%
  filter(!is.na(life_exp)) %>%
  mutate(pop_density = pop_est / area)

# Create the customized map
tm_shape(world_data) +
  tm_polygons("life_exp",
    palette = "-RdYlBu",
    style = "pretty",
    n = 7,
    title = "Life Expectancy",
    popup.vars = c("name", "life_exp", "gdp_cap_est"),
    border.col = "white",
    border.alpha = 0.5
  ) +
  tm_text("iso_a3", size = "AREA", col = "black", fontface = "bold", alpha = 0.7) +
  tm_layout(
    title = "Global Life Expectancy and Economic Indicators",
    title.position = c("center", "top"),
    title.size = 1.5,
    bg.color = "#f5f5f5",
    inner.margins = c(0.1, 0.1, 0.1, 0.1),
    frame = TRUE,
    frame.lwd = 2,
    legend.outside = TRUE,
    legend.outside.position = "right",
    legend.frame = TRUE,
    legend.bg.color = "white",
    legend.text.size = 0.7
  ) +
  tm_compass(position = c("left", "top"), size = 2) +
  tm_scale_bar(position = c("left", "bottom"), text.size = 0.6) +
  tm_credits("Data: {tmap} World dataset", position = c("right", "bottom"), size = 0.6) +
  tm_style("natural")
```

<br>

## &rarr; Interactive Maps

You can super easily make all your `tmap` maps switch to interactive mode with `tmap_mode("view")`. In this example, we:

- **Filters** the World dataset to focus on European countries.
- Uses GDP per capita for the **polygon fill color**, with a "jenks" classification method and the "viridis" color palette.
- Adds **popup information** for country name, population, and GDP per capita.
- Includes country borders with **reduced opacity**.
- Adds **bubbles representing population size** and colored by life expectancy.
- Customizes the layout with a **title** and places the **legend outside the map**.
- Includes a **scale bar** and compass for better orientation.

```{r}
library(tmap)
library(dplyr)
data("World")

europe <- World %>%
  filter(continent == "Europe")

tmap_mode("view")

my_tmap <- tm_shape(europe) +
  tm_polygons("gdp_cap_est",
    style = "jenks",
    palette = "viridis",
    title = "GDP per capita",
    popup.vars = c("name", "pop_est", "gdp_cap_est"),
    id = "name"
  ) +
  tm_borders(alpha = 0.5) +
  tm_bubbles(
    size = "pop_est",
    col = "life_exp",
    scale = 0.5,
    title.size = "Population",
    title.col = "Life Expectancy",
    popup.vars = TRUE
  ) +
  tm_layout(
    title = "European Countries: GDP, Population, and Life Expectancy",
    legend.outside = TRUE,
    legend.outside.position = "right"
  ) +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_compass(position = c("right", "top"))

tmap_save(my_tmap, filename = "../HtmlWidget/my_interactive_map.html")
```

<center>
  <iframe src="../HtmlWidget/my_interactive_map.html" height="600" width="600" style="border:none;"></iframe>
</center>

_Note: You can return to static display mode by running:`tmap_mode("plot")`_

<br>

# Gallery of tmap examples
***

If you want to learn more about `tmap`, it's highly suggested to have a look at:

- this [introduction to tmap](../416-introduction-to-tmap.html)
- [advanced features of tmap](../417-advanced-tmap-usage.html)

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../417-advanced-tmap-usage.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Facetted map with tmap <hr> <p class=explanation_portfolio>Learn how to create facetted map with tmap axis. </div> </div><img alt="Facetted map with tmap" class="img-fluid imgOfPortfolio" src=../img/graph/417-advanced-tmap-usage-1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../417-advanced-tmap-usage.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Combine layout in tmap <hr> <p class=explanation_portfolio>Learn how to combine choropleth and bubble map with tmap axis. </div> </div><img alt="advanced layout with tmap" class="img-fluid imgOfPortfolio" src=../img/graph/417-advanced-tmap-usage-2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../417-advanced-tmap-usage.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Change projection <hr> <p class=explanation_portfolio>How to change your map projection using tmap axis. </div> </div><img alt="change map projection with tmap" class="img-fluid imgOfPortfolio" src=../img/graph/417-advanced-tmap-usage-3.png> </a></div> </div> </div> </div> </section>

<!-- Close container -->
</div>