---
title: "Create beautiful color maps with Viridis"
logo: "Viridis"
descriptionMeta: "The viridis package in R significantly enhances data visualization by offering a collection of color maps that are perceptually uniform in color and brightness. Crafted with the needs of colorblind users in mind, this package features color scales that maintain clarity and consistency even when viewed in grayscale."
descriptionTop: "The `viridis` package in R significantly enhances data visualization by offering a collection of **color maps that are perceptually uniform** in color and brightness. Crafted with the needs of colorblind users in mind, this package features color scales that **maintain clarity and consistency** even when viewed in grayscale."
documentationLink: "sjmgarnier.github.io/viridis/"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template-pkg-post.html
      toc: TRUE
      toc_float: TRUE
      toc_depth: 2
      df_print: "paged"
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  fig.align='center'
)
```

<div class="container">

# Quick start
***
The `viridis` package in R is an extension of the [ggplot2](../ggplot2-package.html) package, designed to simplify the process of creating visually appealing color maps. It offers a set of color maps that are **perceptually uniform** in color and brightness, making them ideal for use in data visualization.

<div class = "row">

<div class = "col-md-5  col-sm-12 align-self-center">


✍️ **author** &rarr; Simon Garnier

📘 **documentation** &rarr; [Github](https://sjmgarnier.github.io/viridis/)

⭐️ *more than 250 stars on github*

</div>

<div class = "col-md-7  col-sm-12">
```{r, echo=FALSE, out.width = "75%", fig.align='center'}
library(viridis)
library(ggplot2)
ggplot(data.frame(x = rnorm(10000), y = rnorm(10000)), aes(x = x, y = y)) +
  geom_hex() +
  scale_fill_viridis(option='C') +
  theme_minimal()
```
</div>
</div>

# Installation
***
To get started with `viridis`, you can install it directly from CRAN using the `install.packages` function:

```{r eval=FALSE}
install.packages("viridis")
```

# Basic usage 
***

The `viridis` package is very easy to use. Start by loading it with `library(viridis)`, and then you can use the **color palettes** in your plots!

```{r}
library(viridis)
colors = viridis(3)

# basic plot
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 6) +
  scale_color_manual(values = colors) +
  theme_minimal()
```

# Key features
***

## &rarr; Combine ggplot2 and Viridis

The `viridis` package is an extension of the [ggplot2](../ggplot2-package.html) package, which means you can use it to customize the color of your plots.

It provides `scale_color_viridis()` and `scale_fill_viridis()` functions that allow you to use the `viridis` color palettes in your plots.

Example:

```{r}
# load libraries
library(viridis)
library(ggplot2)

# create dataframe
dsub <- subset(diamonds, x > 5 & x < 6 & y > 5 & y < 6)
dsub$diff <- with(dsub, sqrt(abs(x - y)) * sign(x - y))

# create plot with Turbo color palette
ggplot(dsub, aes(x, y, colour = diff)) +
  geom_point() +
  scale_color_viridis(option='turbo') +
  theme_minimal()
```

<br>

## &rarr; Beautiful filled.contour

The `filled.contour()` function is a base R function that creates a filled contour plot. You can use the `color.palette` argument to specify a color palette from the `viridis` package.

Example:

```{r}
library(viridis)
x <- y <- seq(-8*pi, 8*pi, len = 40)
r <- sqrt(outer(x^2, y^2, "+"))
filled.contour(
  cos(r^2)*exp(-r/(2*pi)), 
  axes=FALSE,
  color.palette=magma,
  asp=1
)
```

<br>

## &rarr; Use Viridis with lattice

The `viridis` package can be used with the `lattice` package to create beautiful plots. You can use the `colorkey` argument to specify a color palette from the `viridis` package.

Example:

```{r}
library(viridis)
library(lattice)
levelplot(volcano, col.regions = mako(100))
```

<br>

# Gallery of Viridis examples
***

The gallery is **filled** with examples that showcase the versatility of the `viridis` package. Each example is designed to help you understand how to use **custom colors** in your plots.

<section class=bg id=portfolio style=padding-top:10px> <div class=container> <div class=row id=portfolio-items> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../27-levelplot-with-lattice.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Level plot <hr> <p class=explanation_portfolio>Learn how to build a level plot with lattice </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/27-levelplot-with-lattice2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../48-grouped-barplot-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Grouped barplot <hr> <p class=explanation_portfolio>Discover how to build a grouped barplot with viridis colors </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/48-grouped-barplot-with-ggplot24.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../48-grouped-barplot-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Grouped barplot <hr> <p class=explanation_portfolio>Learn how to build a grouped barplot with viridis colors </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/48-grouped-barplot-with-ggplot23.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../79-levelplot-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Grouped barplot <hr> <p class=explanation_portfolio>Discover how to build a grouped barplot with viridis colors </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/79-levelplot-with-ggplot22.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../89-box-and-scatter-plot-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Scatter plot <hr> <p class=explanation_portfolio>How to build a scatter plot with viridis colors </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/89-box-and-scatter-plot-with-ggplot2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../135-stacked-density-graph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked density graph <hr> <p class=explanation_portfolio>How to build a stacked density graph </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/135-stacked-density-graph1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../135-stacked-density-graph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked density graph <hr> <p class=explanation_portfolio>Discover how to build a stacked density graph </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/135-stacked-density-graph2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../135-stacked-density-graph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked density graph <hr> <p class=explanation_portfolio>How to build a stacked density graph </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/135-stacked-density-graph3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"><a class=portfolio-link href=../135-stacked-density-graph.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked density graph <hr> <p class=explanation_portfolio>Discover how to build a stacked density graph </div> </div><img alt="heatmap" class="img-fluid imgOfPortfolio" src=../img/graph/135-stacked-density-graph4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"> <a class=portfolio-link href=../328-hexbin-map-of-the-usa.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Hexbin map <hr> <p class=explanation_portfolio>Discover how to build a hexbin map of the USA </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/328-hexbin-map-of-the-usa4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-6"> <a class=portfolio-link href=../332-hexbin-chloropleth-cartogram.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Chloropleth cartogram <hr> <p class=explanation_portfolio>How to build a hexbin chloropleth cartogram </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/332-hexbin-chloropleth-cartogram4.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../223-faceting-with-ggplot2.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Facetting <hr> <p class=explanation_portfolio>Find out how to create a small multiple of histograms </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/223-faceting-with-ggplot23.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../264-control-ggplot2-boxplot-colors.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Boxplot <hr> <p class=explanation_portfolio>How to control boxplot colors </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/264-control-ggplot2-boxplot-colors.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../283-the-hourly-heatmap.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Heatmap <hr> <p class=explanation_portfolio>Discover how to use heatmap for time series data </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/283-the-hourly-heatmap.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../294-basic-ridgeline-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Ridgeline plot <hr> <p class=explanation_portfolio>How to build a basic ridgeline plot </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/294-basic-ridgeline-plot2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../294-basic-ridgeline-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Ridgeline plot <hr> <p class=explanation_portfolio>Find out how to customize a ridgeline plot </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/294-basic-ridgeline-plot3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../299-circular-stacked-barplot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular stacked barplot <hr> <p class=explanation_portfolio>How to build a circular stacked barplot </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/299-circular-stacked-barplotSmall.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../306-custom-circle-packing-with-one-level.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular packing <hr> <p class=explanation_portfolio>How to build a custom circle packing with one level </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/306-custom-circle-packing-with-one-level.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../314-custom-circle-packing-with-several-levels.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular packing <hr> <p class=explanation_portfolio>How to build a custom circle packing with several levels </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/314-custom-circle-packing-with-several-levels1.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../314-custom-circle-packing-with-several-levels.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Circular packing <hr> <p class=explanation_portfolio>Discover how to build a custom circle packing with several levels </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/314-custom-circle-packing-with-several-levels2.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../320-the-basis-of-bubble-plot.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Bubble plot <hr> <p class=explanation_portfolio>How to build a bubble plot with ggplot2 </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/320-the-basis-of-bubble-plot3.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../341-stacked-barplot-with-negative-values.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Stacked barplot <hr> <p class=explanation_portfolio>How to build a stacked barplot with negative values </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/341-stacked-barplot-with-negative-values.png> </a></div> <div class="portfolio-item col-sm-6 col-md-4"> <a class=portfolio-link href=../histogram_several_group.html> <div class=portfolio-hover> <div class=portfolio-hover-content> <p>Histogram <hr> <p class=explanation_portfolio>Discover how to build a histogram with several groups </div> </div><img alt="" class="img-fluid imgOfPortfolio" src=../img/graph/histogram_several_group.png> </a></div> </div> </div> </div> </section>

<!-- Close container -->
</div>
